### 2021-01-03 outputs/paper/paper.Rmd
Create example repo

---
title: "TBD"
subtitle: "TBD"
author: "TBD"
thanks: "Code and data are available at: LINK."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
output:
  bookdown::pdf_document2:
toc: FALSE
bibliography: references.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(palmerpenguins)
```

# Introduction

# Data

Our data is of penguins (Figure \@ref(fig:bills)).

```{r bills, fig.cap="Bills of penguins", echo = FALSE}
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = FALSE) +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk more about it.

Also bills and their average (Figure \@ref(fig:billssssss)). (Notice how you can change the height and width so they don't take the whole page?)

```{r billssssss, fig.cap="More bills of penguins", echo = FALSE, fig.width=8, fig.height=4}

# This needs to be about the random data tha tI sasmpled?
ggplot(penguins, aes(x = island, fill = species)) +
  geom_bar(alpha = 0.8) +
  scale_fill_manual(values = c("darkorange","purple","cyan4"),
                    guide = FALSE) +
  theme_minimal() +
  facet_wrap(~species, ncol = 1) +
  coord_flip()
```

Talk way more about it.

# Model

\begin{equation}
Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}  (\#eq:bayes)
\end{equation}

Equation \@ref(eq:bayes) seems useful, eh?

Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR]. We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful.

# Results

# Discussion

## First discussion point

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

# Appendix {-}

\newpage

# References

### 2021-01-04 scripts/01-data_cleaning.R
Add example literature and a example script

@@ -1,0 +1,36 @@
+#### Preamble ####
+# Purpose: Clean the survey data downloaded from [...UPDATE ME!!!!!]
+# Author: Rohan Alexander [CHANGE THIS TO YOUR NAME!!!!]
+# Data: 3 January 2021
+# Contact: rohan.alexander@utoronto.ca [PROBABLY CHANGE THIS ALSO!!!!]
+# License: MIT
+# Pre-requisites: 
+# - Need to have downloaded the ACS data and saved it to inputs/data
+# - Don't forget to gitignore it!
+# - Change these to yours
+# Any other information needed?
+
+
+#### Workspace setup ####
+# Use R Projects, not setwd().
+library(haven)
+library(tidyverse)
+# Read in the raw data. 
+raw_data <- readr::read_csv("inputs/data/raw_data.csv"
+                     )
+# Just keep some variables that may be of interest (change 
+# this depending on your interests)
+names(raw_data)
+
+reduced_data <- 
+  raw_data %>% 
+  select(first_col, 
+         second_col)
+rm(raw_data)
+         
+
+#### What's next? ####
+
+
+
+         

### 2021-03-11 outputs/paper/paper.Rmd
Update paper so that appendix numbering works
Add cross-references example

@@
 ---
-title: "TBD"
-subtitle: "TBD"
-author: "TBD"
+title: "My title"
+subtitle: "My subtitle if needed"
+author: 
+  - First author
+  - Another author
 thanks: "Code and data are available at: LINK."
@@
 output:
-  bookdown::pdf_document2:
+  bookdown::pdf_document2
 toc: FALSE
@@
 # Introduction
+
+You can and should cross-reference sections and sub-sections.
+For instance, Section \@ref(data).
+R Markdown automatically makes the sections lower case and adds a dash to spaces to generate labels, for instance, Section \@ref(first-discussion-point).
 
@@
 
+We can use maths by including latex between dollar signs, for instance $\theta$.
+
 # Results
@@
 
+\appendix
+
 # Appendix {-}
+
+# Additional details
 

### 2022-02-06 pp1.Rmd
Add files via upload

---
title: "**Classification Of Toronto Neighbourhoods Safety Based On Crime Volume**"
author: "**Matthew Ma**"
date: "**2022/2/2**"
output: pdf_document
---

\begin{center}
Abstract
\end{center}

```{r, include=FALSE}
library(opendatatoronto)
library(dplyr)

# get package
package <- show_package("fc4d95a6-591f-411f-af17-327e6c5d03c7")
package

# get all resources for this package
resources <- list_package_resources("fc4d95a6-591f-411f-af17-327e6c5d03c7")

# identify datastore resources; by default, Toronto Open Data sets datastore resource format to CSV for non-geospatial and GeoJSON for geospatial resources
datastore_resources <- filter(resources, tolower(format) %in% c('csv', 'geojson'))

# load the first datastore resource as a sample
data <- filter(datastore_resources, row_number()==1) %>% get_resource()
data
```

Crime rate and crime quantity are essential data to measure the security level and livable degree of a particular region and an important indicator to mark the security degree of a city. For Toronto, crime data is an indicator that the police, the government and the general public care about and look at. In this report, I will use Open Data Toronto and Environics Analytics data to analyze robbery, assault, and armed crime in approximately 140 comparable communities in metropolitan Toronto, Felony data such as auto theft and average crime rate per 100,000 population(Toronto Police Service, 2021). Researchers can intuitively compare the crime rate and the total number of crimes in communities with different geographical locations and people in Toronto based on this data set.

# **Introduction**

Based on people's impressions and relevant information, Toronto is regarded as a developed, friendly, livable and multicultural city. It is also one of the most popular and comfortable city in Canada. According to the report, Toronto has one of the lowest crime rates in Canada, compared with other big cities such as Montreal, Calgary, Castlestone and Vancouver(Macleans, 2019). Due to the impact of globalization, local laws and regulations and legalization of drugs, the crime rate in Toronto has constantly been fluctuating. Generally speaking, the crime rate in Toronto has remained in a stable state(Macleans, 2019).
In the past few years, the news has frequently revealed various types of crimes such as armed robbery, car theft and shooting in Toronto (Roccam,2022). These crimes are usually related to the division of regions. Even within The Toronto area, there are significant differences in the crime rate of the different areas. As a result, people in Toronto have various concerns about the safety of living in other areas. As people generally prefer to live in a safer area, housing prices in different areas have also started to change. In addition, the distribution of limited police resources by the government and the police also needs to be allocated and calculated based on the crime data in different areas.
Nowadays, crime data has become a fundamental reference index for real estate developers, police, government and ordinary residents.
In this report, we used community crime data provided by the Toronto government and related data companies to find serious crime data for 140 different Toronto neighbourhoods from 2014 to 2020, including eight felonies including armed crime, assault, theft, and murder(Toronto Police Service, 2021). For the summary of the data, people in different industries can use these data to study the future development moving ideas, but also to police deployment, resource allocation provides a meaningful reference.
The website provides accurate data on eight crimes per year from 2014 to 2020, which is very complex(Toronto Police Service, 2021). Because of this, data is challenging to study and categorize intuitively. We had to refine each piece of data to complete the study.
In general, we have an idea for this study. We firstly find the total number of crimes of different types every year in the unit of year to intuitively obtain the data of the total number of crimes in each region from 2014 to 2020 so that we can compare the security of each community.
After that, we classsify the security level of each reigion and we use the crime number within different community to make compares.

# **Data**

## **Source**:

The data used in this report are published as summary data published in Open Data Toronto by Toronto Police Services. The content includes the statistics of 8 serious crime rates and crime quantity in 140 different areas of Toronto during the seven years from 2014 to 2020(Toronto Police Service, 2021). This information was last updated in May 2021. The information has been updated on an annual basis since its release, with data for the past 12 months republished every May and accessible to all users through the web. Under this page, the publisher also uploaded the set of data CSV files to provide relevant research for the data developer(Toronto Police Service, 2021). I downloaded the CSV file of this group of data under the relevant link of this web page and used R for further analysis.

## **Method**:
The website provides accurate data on eight types of crimes per year from 2014 to 2020, which is very complex(Toronto Police Service, 2021). Because of this, data is challenging to study and categorize intuitively. We had to refine each piece of data to complete the study.
In general, we an ideas for this study, which is to study the total number of crimes of different types every year in the unit of year to intuitively obtain the data of the total number of crimes in each region from 2014 to 2020 so that we can compare the security of each region.
The original data contains the detailed records of 8 different crimes in 140 blocks. Due to the enormous data and the significant difference in the number of crimes in other communities, it is difficult to analyze the crimes in different regions directly. We must first simplify this set of data.
Based on the core purpose of our study is to tell people the safety degree of different communities, we first extracted the sum of crime in other areas of the 140 neighbourhoods in Toronto from 2014 to 2020 to get the sum of corruption in each room and the total crime in each region each year.
So to make the data more meaningful as well as the representative, we find out each the average of the crime in different regions, the maximum and the minimum, with the mean value of the community to do a benchmark for the unit and arranged from big to small, fundamental value more than can be interpreted as the relative risk of communities, the following primary value of community can be understood as the relative safety of the community.

```{r, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(tibble)
library(dplyr)
crime <- read.csv('neighbourhood-crime-rates.csv')
crime <- mutate(crime, Total_2014 = Assault_2014 + AutoTheft_2014 + 
         BreakAndEnter_2014 + Robbery_2014 + TheftOver_2014 + Homicide_2014 + 
         Shootings_2014)
crime <- mutate(crime, Total_2015 = Assault_2015 + AutoTheft_2015 + 
         BreakAndEnter_2015 + Robbery_2015 + TheftOver_2015 + Homicide_2015 + 
         Shootings_2015)
crime <- mutate(crime, Total_2016 = Assault_2016 + AutoTheft_2016 + 
         BreakAndEnter_2016 + Robbery_2016 + TheftOver_2016 + Homicide_2016 + 
         Shootings_2016)
crime <- mutate(crime, Total_2017 = Assault_2017 + AutoTheft_2017 + 
         BreakAndEnter_2017 + Robbery_2017 + TheftOver_2017 + Homicide_2017 + 
         Shootings_2017)
crime <- mutate(crime, Total_2018 = Assault_2018 + AutoTheft_2018 + 
         BreakAndEnter_2018 + Robbery_2018 + TheftOver_2018 + Homicide_2018 + 
         Shootings_2018)
crime <- mutate(crime, Total_2019 = Assault_2019 + AutoTheft_2019 + 
         BreakAndEnter_2019 + Robbery_2019 + TheftOver_2019 + Homicide_2019 + 
         Shootings_2019)
crime <- mutate(crime, Total_2020 = Assault_2020 + AutoTheft_2020 + 
         BreakAndEnter_2020 + Robbery_2020 + TheftOver_2020 + Homicide_2020 + 
         Shootings_2020)
HoodAndYear <- select(crime, OBJECTID, Neighbourhood, Total_2014, Total_2015, Total_2016,
                      Total_2017, Total_2018, Total_2019, Total_2020)
HoodAndYear <- mutate(HoodAndYear, Total = Total_2014 + Total_2015 + Total_2016 +
                      Total_2017 + Total_2018 + Total_2019 + Total_2020)
mmm <- summarise(HoodAndYear, totalMax = max(Total), totalMin = min(Total),
          totalMean = mean(Total))
totalCrime <- select(HoodAndYear, Neighbourhood, Total)
boxplot(totalCrime$Total)
summary(totalCrime)

```
# **Data Characteristics**
```{r}
summary(totalCrime)
```

Through the above induction and calculation, we got a data table and some key data as shown in the above image.

  -The minimum value of this set of data is 382

  -The value of Q1 is 873.2

  -The median value of this set of data is 1334

  -The mean of this group of data is mean=1750

  -Q3 of this set of data =2088.5

  -The maximum value of this set of data =8997

  At the same time, we also divided the safety degree of different areas because the baseline value (the average value) corresponding to the community is medium. Forty-four communities above it are relatively dangerous areas. The three Communities with the highest crime rate are Waterfront Communities, Church-Young Corridor, and Bay Street Corridor
. The 96 neighbourhoods below the average are relatively safe, with Lambton Baby Point, Young-St. Clair and Maple Leaf have the lowest crime rates.

In addition, for further analysis, we also made a box plot for this group of data.
We can observe the visual ratio of safe and dangerous communities in more detail through this figure.
As we gained more information, we were able to further generalize the safety rating of the community, statistically, from the maximum to Q3 of the community.

```{r}
boxplot(totalCrime$Total)
```

Communities were classified as high risk, communities between Q3 and average were classified as relatively dangerous, communities between average and Q1 were classified as relatively safe, and communities between Q1 and minimum were classified as safest.

Through this induction, we can see that there are 35 most dangerous areas, including Waterfront Communities, Church-Young Corridor and the other regions.

There are 10 communities at risk, including Humber Summit, Parkwoods-Donalda, and Newtonbrook West.

There are 60 relatively safe communities, including Don Valley Village, Hummvermede and Triniti-Bellwoods.

35 very safe neighbourhoods, including Young-St.Clair, Guildwood, and Forest Hill South.
Moreover, in order to observe these values, we also made a bar graph for this group of data, and we can see that a higher bar represents a higher crime rate.

```{r}
barplot(totalCrime$Total)
```

# **Discussion**

According to our study of this group of data, summary and observation of images, of the 140 communities in Toronto, nearly 67.9% are relatively safe, and 32.1% are fairly dangerous.
However, whether by observing Boxplot or directly comparing the difference between Q3 and Max, we can see that the difference between Q3 and Max is enormous, which, combined with the above information, represents a very uneven distribution of the number of crimes in the 35 more dangerous neighbourhoods from Q3 to Max. Even though all were classified as very dangerous, the number of crimes concentrated in some group regions was significantly higher than in other groups. Communities that are more at risk have a much smaller total, about 7 percent.
In addition, we can see that the vast majority of the Toronto area is relatively safe, 35 communities are classified as very safe communities, and the gap between Q1 and the minimum value is relatively small. It can be inferred from this information that these communities as a whole still maintain a relatively safe atmosphere. The difference between mean and Q1 in the 60 safer communities is also relatively small, proving that most communities' crime rate is fairly average with slight fluctuation.

# **The Increasing Trend Of Crime Rate**

First, let's reclassify the data. Instead, this time we summed up the crime levels of all neighbourhood in years to get the total crime levels of each neighbourhood for each year of the period and then arranged the method identified the 95 neighbourhoods with the lowest crime levels as safer neighbourhoods.
So to figure out how much crime there was per year in all of these neighbourhoods, we sum up all the data and get the total number of crimes per year in all of these 140 neighbourhoods.
To find the sum of crime per year for the 95 safer neighbourhoods, we summed up the total crime per year for the relative safe neighbourhoods.

```{r}
HoodAndYear <- arrange(HoodAndYear, Total)
bb <- filter(HoodAndYear)
bbbb <- summarise(bb, t2014 = sum(Total_2014), t2015 = sum(Total_2015), t2016 = sum(Total_2016), t2017 = sum(Total_2017), t2018 = sum(Total_2018), t2019 = sum(Total_2019), t2020 = sum(Total_2020))
bbbb <- data.frame(Value = unlist(bbbb), Year = factor(2014:2020))
plot(Value ~ Year, bbbb)
```

As shown in the figure above, in general, the total crime rate of all different types in Toronto increased with a relatively small increase from 2014 to 2019, especially from 2014 to 2017, with no more than 3% of the previous year each time. However, the data entered a relatively rapid growth from 2017 to 2019, and the year with the highest increase was 2018-2019, when the crime increased by 6.65%, more than twice as much as before. From 2019 to 2020, the crime rate fell sharply to roughly the same level as in 2017, decreasing 4,915 crimes, the most significant fluctuation since the data was compiled in 2014 and 2020.
We set the range of the vertical axis to 32,000 to 40,000. In this range, we can see that the increase from 2017 to 2019 was remarkably rapid and then declined in 2019-2020.

```{r}
HoodAndYear <- arrange(HoodAndYear, Total)
hh <- filter(HoodAndYear, row_number() <= 95)
ssss <- summarise(hh, t2014 = sum(Total_2014), t2015 = sum(Total_2015), t2016 = sum(Total_2016), t2017 = sum(Total_2017), t2018 = sum(Total_2018), t2019 = sum(Total_2019), t2020 = sum(Total_2020))
ssss_l <- data.frame(Value = unlist(ssss), Year = factor(2014:2020))
plot(Value ~ Year, ssss_l)
```

As shown in the chart above, for the 95 safer neighbourhoods, we can see very little change from 2014 to 2016, and the values within that range peaked in 2015. In addition, the data changed very little from 2016 to 2017. It can be said that the data hardly changed in the four years, and the most representative decrease was only 11 cases from 2015 to 2016. However, from 2017 to 2019, there was a very clear increase in crime, especially from 2018 to 2019, when the increase reached 1,251 cases, 113 times that of 2015 to 16. There was another significant drop in crime between 2019 and 2020.

By comparing the two groups of data, it is not difficult to find that from 2014 to 2017, the growth rate of 95 safer communities is significantly lower than that of the overall 140 communities. The former fluctuates at a relatively stable value while the latter grows. However, from 2017 to 2019, both data groups showed a rapid growth trend and then began to decline in a relatively fast direction from 2019 to 2020.

# Conclusion

To conclude, by looking at the data provided by open-data-toronto, we found that of the 140 Toronto neighbourhoods, 35 had high crime rates, 10 had high crime rates, and 95 had relatively safe areas, and there has been a clear upward trend in total crime in any community in every year except 2019-20 (Toronto Police Service, 2021). Through this report, people can have a clearer understanding of the safety level of each community so that they can further understand the safety of residents in Toronto. In addition, people of different majors can also use this data set to do other meaningful things.

In addition, based on this set of data, and data generated on the second page, we also observed that the amount of crime in 140 communities in Toronto and the maximum and average or Q3 the value of the difference is tremendous, it can confirm the whole community in Toronto crime quantity gap is very big, the majority of crime are concentrated in specific neighbourhoods, This can provide straightforward and effective information for people looking at this data set.
On the other hand, even though some areas in Toronto have a high crime rate, 67% of the communities are classified as relatively safe communities. The average crime level is far from that of a small number of dangerous communities. For the 95 safer neighbourhoods, crime increased significantly in only two of the seven years of data and remained stable or declining in the rest of the year.
So far, we can conclude that the vast majority of citizens can rest assured in most communities in Toronto, and the overall safety risks are relatively good.

# Reference

Toronto Police Services. (2021, May 6). Open data dataset. City of Toronto Open Data Portal. Retrieved February 2, 2022, from https://open.toronto.ca/dataset/neighbourhood-crime-rates/

Macleans. (2019, November 19). Open data dataset. Canada’s Most Dangerous Places 2020. Retrieved February 3, 2022, from https://www.macleans.ca/canadas-most-dangerous-places-2020/

Rocca, R. (2022, February 4). Man charged in connection with fatal hit-and-run in Toronto: Police - toronto. Global News. Retrieved February 5, 2022, from https://globalnews.ca/news/8596710/weston-road-hit-and-run-suspect-id-charge/

### 2022-02-12 inputs/data/datasheet_template.Rmd
Add datasheets to starter folder

---
title: "My title"
subtitle: "My subtitle if needed"
author: 
  - First author
  - Another author
thanks: "Code and data are available at: LINK."
date: "`r format(Sys.time(), '%d %B %Y')`"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
output:
  bookdown::pdf_document2
toc: FALSE
bibliography: references_datasheet.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Extract of the questions from @gebru2021datasheets

**Motivation**

1. *For what purpose was the dataset created? Was there a specific task in mind? Was there a specific gap that needed to be filled? Please provide a description.*
    - The dataset was created to enable analysis of Australian politicians. We were unable to find a publicly available dataset in a structured format that had the biographical and political information on Australian politicians that was needed for modelling.
2. *Who created the dataset (for example, which team, research group) and on behalf of which entity (for example, company, institution, organization)?*
    - TBD
3. *Who funded the creation of the dataset? If there is an associated grant, please provide the name of the grantor and the grant name and number.*
    - TBD
4. *Any other comments?*
    - TBD

**Composition**

1. *What do the instances that comprise the dataset represent (for example, documents, photos, people, countries)? Are there multiple types of instances (for example, movies, users, and ratings; people and interactions between them; nodes and edges)? Please provide a description.*
	- TBD
2. *How many instances are there in total (of each type, if appropriate)?*
	- TBD
3. *Does the dataset contain all possible instances or is it a sample (not necessarily random) of instances from a larger set? If the dataset is a sample, then what is the larger set? Is the sample representative of the larger set (for example, geographic coverage)? If so, please describe how this representativeness was validated/verified. If it is not representative of the larger set, please describe why not (for example, to cover a more diverse range of instances, because instances were withheld or unavailable).*
	- TBD
4. *What data does each instance consist of? "Raw" data (for example, unprocessed text or images) or features? In either case, please provide a description.*
	- TBD
5. *Is there a label or target associated with each instance? If so, please provide a description.*
	- TBD
6. *Is any information missing from individual instances? If so, please provide a description, explaining why this information is missing (for example, because it was unavailable). This does not include intentionally removed information, but might include, for example, redacted text.*
	- TBD
7. *Are relationships between individual instances made explicit (for example, users' movie ratings, social network links)? If so, please describe how these relationships are made explicit.*
	- TBD
8. *Are there recommended data splits (for example, training, development/validation, testing)? If so, please provide a description of these splits, explaining the rationale behind them.*
	- TBD
9. *Are there any errors, sources of noise, or redundancies in the dataset? If so, please provide a description.*
	- TBD
10. *Is the dataset self-contained, or does it link to or otherwise rely on external resources (for example, websites, tweets, other datasets)? If it links to or relies on external resources, a) are there guarantees that they will exist, and remain constant, over time; b) are there official archival versions of the complete dataset (that is, including the external resources as they existed at the time the dataset was created); c) are there any restrictions (for example, licenses, fees) associated with any of the external resources that might apply to a dataset consumer? Please provide descriptions of all external resources and any restrictions associated with them, as well as links or other access points, as appropriate.*
	- TBD
11. *Does the dataset contain data that might be considered confidential (for example, data that is protected by legal privilege or by doctor-patient confidentiality, data that includes the content of individuals' non-public communications)? If so, please provide a description.*
	- TBD
12. *Does the dataset contain data that, if viewed directly, might be offensive, insulting, threatening, or might otherwise cause anxiety? If so, please describe why.*
	- TBD
13. *Does the dataset identify any sub-populations (for example, by age, gender)? If so, please describe how these subpopulations are identified and provide a description of their respective distributions within the dataset.*
	- TBD
14. *Is it possible to identify individuals (that is, one or more natural persons), either directly or indirectly (that is, in combination with other data) from the dataset? If so, please describe how.*
	- TBD
15. *Does the dataset contain data that might be considered sensitive in any way (for example, data that reveals race or ethnic origins, sexual orientations, religious beliefs, political opinions or union memberships, or locations; financial or health data; biometric or genetic data; forms of government identification, such as social security numbers; criminal history)? If so, please provide a description.*
	- TBD
16. *Any other comments?*
	- TBD

**Collection process**

1. *How was the data associated with each instance acquired? Was the data directly observable (for example, raw text, movie ratings), reported by subjects (for example, survey responses), or indirectly inferred/derived from other data (for example, part-of-speech tags, model-based guesses for age or language)? If the data was reported by subjects or indirectly inferred/derived from other data, was the data validated/verified? If so, please describe how.*
	- TBD
2. *What mechanisms or procedures were used to collect the data (for example, hardware apparatuses or sensors, manual human curation, software programs, software APIs)? How were these mechanisms or procedures validated?*
	- TBD
3. *If the dataset is a sample from a larger set, what was the sampling strategy (for example, deterministic, probabilistic with specific sampling probabilities)?*
	- TBD
4. *Who was involved in the data collection process (for example, students, crowdworkers, contractors) and how were they compensated (for example, how much were crowdworkers paid)?*
	- TBD
5. *Over what timeframe was the data collected? Does this timeframe match the creation timeframe of the data associated with the instances (for example, recent crawl of old news articles)? If not, please describe the timeframe in which the data associated with the instances was created.*
	- TBD
6. *Were any ethical review processes conducted (for example, by an institutional review board)? If so, please provide a description of these review processes, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- TBD
7. *Did you collect the data from the individuals in question directly, or obtain it via third parties or other sources (for example, websites)?*
	- TBD
8. *Were the individuals in question notified about the data collection? If so, please describe (or show with screenshots or other information) how notice was provided, and provide a link or other access point to, or otherwise reproduce, the exact language of the notification itself.*
	- TBD
9. *Did the individuals in question consent to the collection and use of their data? If so, please describe (or show with screenshots or other information) how consent was requested and provided, and provide a link or other access point to, or otherwise reproduce, the exact language to which the individuals consented.*
	- TBD
10. *If consent was obtained, were the consenting individuals provided with a mechanism to revoke their consent in the future or for certain uses? If so, please provide a description, as well as a link or other access point to the mechanism (if appropriate).*
	- TBD
11. *Has an analysis of the potential impact of the dataset and its use on data subjects (for example, a data protection impact analysis) been conducted? If so, please provide a description of this analysis, including the outcomes, as well as a link or other access point to any supporting documentation.*
	- TBD
12. *Any other comments?*
	- TBD

**Preprocessing/cleaning/labeling**

1. *Was any preprocessing/cleaning/labeling of the data done (for example, discretization or bucketing, tokenization, part-of-speech tagging, SIFT feature extraction, removal of instances, processing of missing values)? If so, please provide a description. If not, you may skip the remaining questions in this section.*
	- TBD
2. *Was the "raw" data saved in addition to the preprocessed/cleaned/labeled data (for example, to support unanticipated future uses)? If so, please provide a link or other access point to the "raw" data.*
	- TBD
3. *Is the software that was used to preprocess/clean/label the data available? If so, please provide a link or other access point.*
	- TBD
4. *Any other comments?*
	- TBD

**Uses**

1. *Has the dataset been used for any tasks already? If so, please provide a description.*
	- TBD
2. *Is there a repository that links to any or all papers or systems that use the dataset? If so, please provide a link or other access point.*
	- TBD
3. *What (other) tasks could the dataset be used for?*
	- TBD
4. *Is there anything about the composition of the dataset or the way it was collected and preprocessed/cleaned/labeled that might impact future uses? For example, is there anything that a dataset consumer might need to know to avoid uses that could result in unfair treatment of individuals or groups (for example, stereotyping, quality of service issues) or other risks or harms (for example, legal risks, financial harms)? If so, please provide a description. Is there anything a dataset consumer could do to mitigate these risks or harms?*
	- TBD
5. *Are there tasks for which the dataset should not be used? If so, please provide a description.*
	- TBD
6. *Any other comments?*
	- TBD

**Distribution**

1. *Will the dataset be distributed to third parties outside of the entity (for example, company, institution, organization) on behalf of which the dataset was created? If so, please provide a description.*
	- TBD
2. *How will the dataset be distributed (for example, tarball on website, API, GitHub)? Does the dataset have a digital object identifier (DOI)?*
	- TBD
3. *When will the dataset be distributed?*
	- TBD
4. *Will the dataset be distributed under a copyright or other intellectual property (IP) license, and/or under applicable terms of use (ToU)? If so, please describe this license and/ or ToU, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms or ToU, as well as any fees associated with these restrictions.*
	- TBD
5. *Have any third parties imposed IP-based or other restrictions on the data associated with the instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any relevant licensing terms, as well as any fees associated with these restrictions.*
	- TBD
6. *Do any export controls or other regulatory restrictions apply to the dataset or to individual instances? If so, please describe these restrictions, and provide a link or other access point to, or otherwise reproduce, any supporting documentation.*
	- TBD
7. *Any other comments?*
	- TBD

**Maintenance**

1. *Who will be supporting/hosting/maintaining the dataset?*
	- TBD
2. *How can the owner/curator/manager of the dataset be contacted (for example, email address)?*
	- TBD
3. *Is there an erratum? If so, please provide a link or other access point.*
	- TBD
4. *Will the dataset be updated (for example, to correct labeling errors, add new instances, delete instances)? If so, please describe how often, by whom, and how updates will be communicated to dataset consumers (for example, mailing list, GitHub)?*
	- TBD
5. *If the dataset relates to people, are there applicable limits on the retention of the data associated with the instances (for example, were the individuals in question told that their data would be retained for a fixed period of time and then deleted)? If so, please describe these limits and explain how they will be enforced.*
	- TBD
6. *Will older versions of the dataset continue to be supported/hosted/maintained? If so, please describe how. If not, please describe how its obsolescence will be communicated to dataset consumers.*
	- TBD
7. *If others want to extend/augment/build on/contribute to the dataset, is there a mechanism for them to do so? If so, please provide a description. Will these contributions be validated/verified? If so, please describe how. If not, why not? Is there a process for communicating/distributing these contributions to dataset consumers? If so, please provide a description.*
	- TBD
8. *Any other comments?*
	- TBD

\newpage

# References

### 2022-03-21 STA304 Paper 3-Group_1/304 pape3.Rmd
uploading paper3

---
title: " **The Analyzation of Worked Hours and Demographic In The Perspective of Education and Personal Income For United States In 2018**"
author: '**Haoxuan Cai, Matthew Ma, Zhihao Chen**'
date: "**20/03/2022**"
output: pdf_document
bibliography: reference.bib
nocite: '@*'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(tidyverse)
library(dplyr)
library(broom)
library(arm)
library(knitr)
library(magrittr) # needs to be run every time you start R and want to use %>%

```

# **Abstract**

A person's weekly working hours are closely related to his/her living standard, happiness degree and many other factors. In a typical developed country with a relatively different social status like the United States, it is of great representative significance for government, economy and social research to count the weekly working hours of the people. In this report, according to the General Social Survey(GSS) provided by the Us National Opinion Research Center, statistical information with both quality and quantity to systematically summarize, summarize and further study the average working hours of Native Americans in 2018. An analysis report of working hours and demographic distribution will be made from the perspective of education level and income. In addition, an appendix will be included in this report. A Social survey and fully reproducible method have helped other researchers understand the data in greater detail.

# **Introduction and Hypothesis**

In 2018, with the increase of tariffs on the import of a variety of metal products, restrictions on investment from other countries, the Sino-US trade war and many other events related to the world economy and the local economy of the United States, the social structure and unemployment rate in the United States began. It was not only the most normal year people lived before the COVID-19 pandemic but also the most representative year for the-U.S. President Donald Trump's policies towards the U.S domestic economy.

Under the influence of the above situation, numerous jobs and employment rates in America have changed dramatically. Since the government's policy is proposed to allocate more jobs opportunity for the local United States of American citizens and guarantee more U.S citizens to get paid, many other nation people who live in the U.S are facing the problem of layoffs or higher working pressure. Even though many people got the jobs they needed, many others were unemployed. As a result, many people do not have enough money to support their living expenses due to reducing working hours and salaries, so they have to obtain more income through part-time jobs. On the other hand, as more jobs are available for local U.S citizens, the working hours will also change, and some people will face more work pressure that they did not have before, leading to a lower quality of life.

Therefore, it is imperative for the American people, government and society to analyze the total amount of people working in a certain period. The most intuitive measure is the number of hours worked in a week.

However, collecting data for this metric is difficult because working hours often fluctuate significantly based on different occupations, personal family circumstances, and working cycles, making it challenging to estimate. For example, for self-employed people, entrepreneurs, social service workers, and people with multiple occupations simultaneously, the most intuitive and effective statistical method is to conduct a questionnaire survey.

Attributed to the well-developed Social system in the United States, the General Social Survey (GSS), a sampling Survey method, emerged in 1972, which provides rich questionnaire data for the society. All sampling questions and data statistics were provided and integrated by the National Opinion Research Center. The vast majority of General Social Survey interview models are based on personal interviews, the respondents are all adults in the United States, and the questions studied are all high-quality scientific or Social science-related data with statistical significance. In the past few decades, the General Social Survey has made outstanding contributions to the Social sciences in the United States. It has effectively helped the government, statisticians, economists, sociologists, and a series of other related professions collect information about the adult population, including mental health assessment, happiness, income, race relations, Consumer habits, sexual recognition issues, etc. In addition, except for the first two years, the design of the General Social Survey is a full probability sample design. Therefore, compared with other similar Social questionnaires, General Social Survey has absolute advantages in output quality, output scale and output proportion.

Thus, General Social Survey made a survey on the weekly working hours of people in the United States in 2018 mentioned above and collected the results, which provided highly reliable, scientific and enormous data for the public and relevant researchers to study. By collecting and downloading data from General Social Survey official website, more than 2300 individuals were given feedback on weekly working hours.

# Data

## Data Summary

From 1972 to the current times, the GSS (General Social Survey) has been recording social transformations and making the research on the growth complexity of American society. The General Social Survey always provide a reliable source of data to help researchers, students, and journalists to conduct the analysis and explain trends in American behaviors, demographics, and opinions. In addition, GSS offers a convenient way to analyze, explore, extract and share.

Since GSS supplies the neat and direct dataset for all years, with tools that help you explore and save variables and analyze data online. In fact, we are pretty interested in working hours of people in America in 2018. Hence. We aim to explore the number of hours people, working at all jobs, last week on average. We select 4 variables, which are closely associated with the working hours and the relative topic.

### Similar dataset

As a matter of fact, we can obtain numerous datasets in terms of our topic. But still we select to utilize the dataset created by GSS. Reasons are as followed. First of all, GSS is kind of official institution for the government, which is no doubt more reliable when compared with other datasets supported by some private local institutions. Meanwhile, the data resources are more stable and strict. statistically, GSS has higher rate of receiving the respondents than any other institutions. Besides, normally, any dataset provided by GSS has more than amount of 10,000 data, which is quite a lot in this area. Above all, we determine to select this stable, strict and reliable resources, given there are a lot more dataset regarding our topic.

In this study, we will clean and summarize the data obtained on the official website of GSS in 2018 and primarily analyze the relationship between the proportion of different working hours and the proportion of the population from the data distribution perspective. On top of this, we will use the auxiliary information provided by GSS to find out the proportion of different working hours and the reasons behind it from the perspective of education level and income and make some hypotheses to assist our experiment. Since the total number of respondents is more than 10,000, and each person has different working hours, we will divide the collected receipts into 10-hour units for subsequent analysis. In addition, we have developed a reproducible method for data extraction and analysis and summary. In addition, we have also introduced data obtaining methods in detail and developed a similar survey to further analyze the average weekly working hours of Americans in 2018.

# Hypothsis

Before the experiment, we made an hypothesis. From the perspective of macro society, we believed that even if there were long hours in some particular industries, for the vast majority of population, the working hours of the whole society would appear in a relatively average normal distribution. We hypothesize a nearly perfect bell curve graph should be generated during the research. Nevertheless, we think that in the U.S, one of the most developed countries in the world, people do not need to work too hard for living. Thus, the we deduct the number of people who do less work is greater than people who works more, and there may have a right-skewed normal distribution.

Since the highly educated group is always a minority in society, we hypothesize that the less educated group will be the majority in any working time.

In addition, considering the national conditions of the United States, we believe that no matter in which time-length interval, the average income will be the mainstream of the region, and extremely low or high income will not appear in a particular interval.

## Data cleaning

```{r, include=FALSE}
library(haven)
use_data <- read_dta("GSS2018.dta 2")
head(use_data)
```

For sustaining our further analysis, we have to clean and process the data. To start with, we select four variables, describe the working hours of America and relative means of variable, to establish the new dataset, which are respectively `hrs1`, `earnrs`, `nateducy` and `finrela`. Then we utilize the $names$ function to change their names with more understanding meaning.

Since the response variable working hours is numeric, which make us difficult to do explantory variable analysis. Hence, we first utilize the $tidyverse$ function to transform the numeric hours to categorical. Meanwhile, we set up the criterion to rewrite the working hours. Specifically, we utilize "1" means 0-10 hours, "2" means 11-20 hours and so on, until "9" means 80-90 hours.

In addition, at the first beginning, we find that the dataset has numerous missing values, specifically as NA. In fact, we cannot make any constructable analysis, assume we cannot eliminate the existence of missing values. Hence, after we establish the new dataset, we simply apply the $na.omit$ function to operate the new dataset. The final result is what we can utilie to conduct the further analysis.

```{r, include=FALSE}
hours <- use_data$hrs1
earn <- use_data$earnrs
education <- use_data$nateducy
income <- use_data$finrela
app_data <- data.frame(hours, earn, education, income)
app_data <- na.omit(app_data)
hours <- as.numeric(hours)
attach(app_data)
```

We would exemplify the new dataset. 
```{r, echo=FALSE}
app_data <- app_data %>% 
  mutate(hours = case_when(
    hours <= 10 ~ "1",
    hours <= 20 & hours >10 ~ "2",
    hours <= 30 & hours >20 ~ "3",
    hours <= 40 & hours >30 ~ "4",
    hours <= 50 & hours >40 ~ "5",
    hours <= 60 & hours >50 ~ "6",
    hours <= 70 & hours >60 ~ "7",
    hours <= 80 & hours >70 ~ "8",
    hours <= 90 & hours >80 ~ "9",
    TRUE ~ "NA"
  ))
kable(head(app_data))
```

### Specific variable

`hours`: This variable answers the question in the survey: How many hours did you work last week, at all jobs. It has 9 options, "1" means 0-10 hours and "9" means 80-90 hours in total.

`earn`: This variable answers the question How many persons in the family (including yourself) earned any money last year (2017) from any job or employment. It is a numerical and consistent variable, which describes the number of people working in the family.

`education`: This variable answers the question about the satisfaction in terms of people's attitudes towards the domestic education supported by the government. It has three options. "1" represents too little, while "3" represents very good.

`income`: This variable answers the questions that compared with American families in general, would you say your family income is far below average, below average, average, above average, or far above average. The above is those five options.

```{r, echo=FALSE}
ggplot(app_data, aes(x=hours)) + geom_bar(color = "grey", fill = "orange") + ggtitle("Figure 1 - Bar plot of working hours")
```
According to figure1, we utilize the bar plot to describe the characteristics of `hours`. Apparently, it is symmetric and follows the normal distribution. We can judge that our response variable is quite good to be analysed. In addition, the majority of people just working approximately 4 hours per day on average and working at 8 hours has the least people.

```{r, echo=FALSE}
app_data <- app_data %>%  filter(education <= 5)
app_data$education <- as.factor(app_data$education)
bar_1 <- ggplot(data = app_data, aes(x = hours, fill = education)) + 
  geom_bar() + 
  theme(text = element_text(size = 8)) + 
  labs(x = "hours",
       y = "Number",
       fill = "education",
       title = "Figure 2 - Hours vs education") 
bar_1
```

As shown in the Figure2, It is the bar plot of working hours with the layer of education. The result is that whatever the people in the level of society, they all share the same point that the investment into the education is not enough. And the government needs to take care of this problem and take the measures to handle, given this common consensus.

```{r, echo=FALSE}
app_data$income <- as.factor(app_data$income)
bar_2 <- ggplot(data=as.data.frame(app_data), mapping=aes(x= hours, fill= income))+
  geom_bar(stat="count",width=0.5,position='stack')+
  scale_fill_manual(values=c('#999999','#E69F00','#00BCF2', '#00188F', "red"))+
  geom_text(stat='count',aes(label=..count..), color="white", size=3.5, position=position_stack(0.5))+
  theme_minimal() + ggtitle("Figure 3 - hours vs income")+theme(plot.title=element_text(hjust=0.5))
bar_2
```

As shown in the Figure3, it describes the relationship between the hours and income. The outcome is that majority of people is still with the average income, whatever how long they work. In fact, the working hours is closely related to the people's lifestyles. Analyzing the hours can help us get a more confound understanding of people common lives.

# **Result**

Based on the information provided from the U.S. GSS on the number of hours worked per week in 2018 and the collated data, we found that the number of hours worked has a normal distribution with the population base.

According to figure1, we see a slightly right-skewed bell curve, and the left and right sides of the picture are basically symmetric. The x-axis presents the total working hours in one week from 0 to 10 hours to 90hours; there are nine sections in total, each of which is 10hours long. On the X-axis, the left-most and right-most sections are 0-10hours and 80-90hours, respectively. The Y-axis represents the corresponding numbers of people.

#The relationship between working hours and demographic

The people who work 30-40 hours are the most, accounting for about 40% of the total number. The second and third most are those who work 40-50 hours and 50-60 hours a week respectively, and the sum of the two indicators also accounts for nearly 30%. Based on these two points, we found that almost 70% of the population works from 30-to 60 hours a week, which is also the main reason for the normal distribution of the bell curve. On the other hand, those who work 70-80hours a week are the least. The second and third lowest hours are 0-10 hours and 80-90hours respectively, located on the left and right side of the image, respectively. The sum of the second and third highest values only reaches x% of the maximum value, that indicates it is a unimodal normal distribution. The other three data points are from the highest to the lowest for those working 20 to 30 hours a week, 10 to 20 hours, and 60 to 70 hours a week.

According to the above points, the three values corresponding to the total working hours and the population are respectively in the middle of the x-axis, and the highest values are the smallest and the leftmost of the three values, which is a characteristic of the right-skewed normal distribution. Based on this point, the three smallest values are located at the right and left sides of the image, respectively, which further verifies that this group of data appears in the above distribution pattern. In addition, the three intermediate values are in the middle of the highest and lowest values, respectively, and the highest values in these three groups are on the left side of the highest values, while the other two values are on the right side, which ultimately proves that this set of data followed by right-skewed distribution. Lastly, the extreme highest value is much greater than other high values, therefore, we conclude this data shows a unimodal right-skewed normal distribution, and it also verified that the hypothesis we made before the study was correct.

#Education level’s proportion in every interval

Figure 2 is developed based on figure 1, introducing information about the education level in the different working length intervals. The education level is stratified into three leading groups, respectively represented in red, green, and blue, indicating education level from low to high. The majority of students with education level 1 take up the highest proportion Education Level 2 and 3 are very few.
There is no exact interval in which the highly educated proportion of education is dominated by the highly educated. The overall level of education is almost the same regardless of the range. As can be seen intuitively from Figure 2, except for the 70-80 hours interval, the education level distribution of basically every working hour shows that education level1 is the largest, Education Level2 Next, education Level3 minimum. This fits very well with our hypothesis.
In terms of proportion, education level2 accounts very few in the interval from 70-80hours and 80-90hours. The proportion of education Level3 is very less in the interval of 0 to 20 hours and 70 to 80 hours. The interval of working hours from 70-80 hours does not contain anyone from education level 2.

#The Income proportion

We classified income levels into five levels, respectively far below average, below average, average, above average, or far above average. The above five options correspond to the numbers 1 through 5 to observe with higher efficiency. 
No matter which interval works for a long time, it can be intuitively found that the middle class's average income is the majority. The average income accounts for 46.1% of the total, nearly equal to the sum of the other four. The above Average income and below Average income also account for 25.6% and 21.6%, respectively. These three together account for around 92% of the total. Far Above Average is the least in all the income, accounting for only 3%.
The maximum portion of average income appears from the interval of 40-50 hours, but the highest proportion of that was found in 0-10 hours; in this range, average income accounted for as high as 62.5%. The interval contains the most proportion of far below average income is 10-20 hours at17%. Most of the above-average income appear from 30-60 hours, but in the interval of 80-90 hours, it reaches its highest proportion at 35.7%. The highest proportion of below average income is in 20-30 hours, at nearly 28%.

Besides that, we observed that the proportion of far below average income gradually decrease with the increase of working hours, and after the working hours exceed 60-70 hours a week, it disappears.

# **Discussion**

Based on the information above, most of our hypothesis matches with the real situation after analyzing the result from this data set.

The result is roughly consistent with the hypothesis proposed before our experiment for the distribution type. Yet, although a normal distribution follows this data set, it still has a right-skewed trend. By comparing an ideal normal distribution type, that indicates the majority of people tend to work less. The increasing amplitude from 20-30 hours a week to 30-40 hours a week is exceptionally high, and then there exist a drastically decrease from 30-40 hours a week to 40-50 hours a week. We can infer from this that this is the overwhelming majority of Americans' favourite mode of work in 2018. Furthermore, in an ideal normal distribution situation, the number of people should gradually decrease when the working hours increasingly approaching to a high numbers. However, from 70-80 hours a week to 80-90 hours a week, the number has gone up, which should not happen; this could be an error attributed to the sampling size being not big enough. The problem that caused the people who worked long hours to be too small to have statistical significance is also because of the sampling size. Although the sampling size exceeds ten thousand, it still cannot represent the entire population in the U.S in 2018.

The education level does not show too much influence on this research. The education level remains the same in any interval. Particularly in working hours form 70-80 hours a week, there has no information about it, that is caused by the data sampling size. However, the overall relationship between education and working hours/population matches our hypothesis.

The income is stratified at any interval. On the whole, the variation in each interval is slight. This result is consistent with our hypothesis to a great extent. The only difference is that there is no far below the average income for people who work 60-90 hours a week. We can see that the people who have a far below average income do not exist in any working time from 60-90 hours a week, although this is reasonable because people should not earn too less after doing many works, there still should have some outliers these intervals, which is due to the sampling size. Moreover, although the overall proportion in each interval does not change much, statistically speaking, except the two extreme values on the left and right are too few samples. The shorter hours worked, the more blow average income and far blow average income exist. Reversely, the more hours worked, the more above average income and far above average income appear. We did not expect that in our hypothesis.

Lastly, we can conclude that the entire distribution trend between population and working hours followed by a normal distribution. The education level remains the same in any working time length from 0 to 90 hours a week, and the total number of hours a person works per week does not significantly affect the overall income structure.

\newpage

# **Appendix**

**Supplementary Survey**

Survey’s brief Introduction
In order to make the investigation more effective, we generated a supplementary survey about the working hours as well.

The survey is designed on Google Form; users can easily use it to fill in their information without restriction. There are 11 questions in total, and we intend to use this survey to acquire basic information about one’s working time length in a given period, education level, income situation, attitude toward his/her work and some other relevant information.

The URL Link below is the detail of the Survey; users can directly go to the Survey by clicking this URL.

***Survey's URL :***

https://forms.gle/uoGnHiNWHqYoGTqd6

\newpage

# Reference

### 2023-01-21 outputs/paper/paper.qmd
Change to Quarto

--- 2021-03-11 outputs/paper/paper.Rmd
@@
 thanks: "Code and data are available at: LINK."
-date: "`r format(Sys.time(), '%d %B %Y')`"
+date: "`r Sys.time()`"
+date-format: "D MMMM YYYY"
 abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
-output:
-  bookdown::pdf_document2
-toc: FALSE
+format: pdf
 bibliography: references.bib
@@
 
-```{r setup, include=FALSE}
+```{r}
+#| include: FALSE
+
 knitr::opts_chunk$set(echo = TRUE)
@@
 For instance, Section \@ref(data).
-R Markdown automatically makes the sections lower case and adds a dash to spaces to generate labels, for instance, Section \@ref(first-discussion-point).
+Quarto automatically makes the sections lower case and adds a dash to spaces to generate labels, for instance, Section \@ref(first-discussion-point).
 
@@
 
-Our data is of penguins (Figure \@ref(fig:bills)).
+Our data is of penguins (@fig-bills).
 
-```{r bills, fig.cap="Bills of penguins", echo = FALSE}
+```{r}
+#| label: fig-bills
+#| fig-cap: Bills of penguins
+#| echo: false
+
 ggplot(penguins, aes(x = island, fill = species)) +
@@
   scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = FALSE) +
+                    guide = "none") +
   theme_minimal() +
@@
 
-Also bills and their average (Figure \@ref(fig:billssssss)). (Notice how you can change the height and width so they don't take the whole page?)
+Also bills and their average (@fig-billssssss).
+(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work.)
 
-```{r billssssss, fig.cap="More bills of penguins", echo = FALSE, fig.width=8, fig.height=4}
+```{r}
+#| label: fig-billssssss
+#| fig-cap: More bills of penguins
+#| echo: false
+#| fig-width: 8
+#| fig-height: 4
 
-# This needs to be about the random data tha tI sasmpled?
 ggplot(penguins, aes(x = island, fill = species)) +
@@
 
-\begin{equation}
-Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}  (\#eq:bayes)
-\end{equation}
+$$
+Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}
+$$ {#eq-bayes}
 
-Equation \@ref(eq:bayes) seems useful, eh?
+@eq-bayes seems useful, eh?
 

### 2023-02-11 outputs/paper/paper.qmd
Add more example scripts and add cross-references to quorto

@@
 format: pdf
+number-sections: true
 bibliography: references.bib
@@
 ```{r}
-#| include: FALSE
-
-knitr::opts_chunk$set(echo = TRUE)
+#| include: false
+#| warning: false
+#| message: false
 
@@
 
-You can and should cross-reference sections and sub-sections.
-For instance, Section \@ref(data).
-Quarto automatically makes the sections lower case and adds a dash to spaces to generate labels, for instance, Section \@ref(first-discussion-point).
+You can and should cross-reference sections and sub-sections. For instance, @sec-data and @sec-first-point.
 
-# Data
+# Data {#sec-data}
 
@@
 
-## First discussion point
+## First discussion point {#sec-first-point}
 

### 2023-02-11 scripts/00-simulate_data.R

@@ -1,0 +1,19 @@
+#### Preamble ####
+# Purpose: Simulates... [...UPDATE THIS...]
+# Author: Rohan Alexander [...UPDATE THIS...]
+# Data: 11 February 2023 [...UPDATE THIS...]
+# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# License: MIT
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
+
+
+#### Workspace setup ####
+library(tidyverse)
+# [...UPDATE THIS...]
+
+#### Simulate data ####
+# [...ADD CODE HERE...]
+
+
+

### 2023-02-11 scripts/01-data_cleaning.R

@@ -1,36 +1,0 @@
-#### Preamble ####
-# Purpose: Clean the survey data downloaded from [...UPDATE ME!!!!!]
-# Author: Rohan Alexander [CHANGE THIS TO YOUR NAME!!!!]
-# Data: 3 January 2021
-# Contact: rohan.alexander@utoronto.ca [PROBABLY CHANGE THIS ALSO!!!!]
-# License: MIT
-# Pre-requisites: 
-# - Need to have downloaded the ACS data and saved it to inputs/data
-# - Don't forget to gitignore it!
-# - Change these to yours
-# Any other information needed?
-
-
-#### Workspace setup ####
-# Use R Projects, not setwd().
-library(haven)
-library(tidyverse)
-# Read in the raw data. 
-raw_data <- readr::read_csv("inputs/data/raw_data.csv"
-                     )
-# Just keep some variables that may be of interest (change 
-# this depending on your interests)
-names(raw_data)
-
-reduced_data <- 
-  raw_data %>% 
-  select(first_col, 
-         second_col)
-rm(raw_data)
-         
-
-#### What's next? ####
-
-
-
-         

### 2023-02-11 scripts/01-download_data.R

@@ -1,0 +1,26 @@
+#### Preamble ####
+# Purpose: Downloads and saves the data from [...UPDATE THIS...]
+# Author: Rohan Alexander [...UPDATE THIS...]
+# Data: 11 February 2023 [...UPDATE THIS...]
+# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# License: MIT
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
+
+
+#### Workspace setup ####
+library(opendatatoronto)
+library(tidyverse)
+# [...UPDATE THIS...]
+
+#### Download data ####
+# [...ADD CODE HERE TO DOWNLOAD...]
+
+
+
+#### Save data ####
+# [...UPDATE THIS...]
+# change the_raw_data to whatever name you assigned when you downloaded it.
+write_csv(the_raw_data, "inputs/data/raw_data.csv") 
+
+         

### 2023-02-11 scripts/02-data_cleaning.R

@@ -1,0 +1,26 @@
+#### Preamble ####
+# Purpose: Cleans.... [...UPDATE THIS...]
+# Author: Rohan Alexander [...UPDATE THIS...]
+# Data: 11 February 2023 [...UPDATE THIS...]
+# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# License: MIT
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
+
+
+#### Workspace setup ####
+library(tidyverse)
+# [...UPDATE THIS...]
+
+#### Clean data ####
+# [...UPDATE THIS...]
+raw_data <- read_csv("inputs/data/raw_data.csv")
+
+
+
+
+#### Save data ####
+# [...UPDATE THIS...]
+# change cleaned_data to whatever name you end up with at the end of cleaning
+write_csv(cleaned_data, "outputs/data/cleaned_data.csv")
+

### 2023-02-11 scripts/04-test_data.R

@@ -1,0 +1,15 @@
+#### Preamble ####
+# Purpose: Tests... [...UPDATE THIS...]
+# Author: Rohan Alexander [...UPDATE THIS...]
+# Data: 11 February 2023 [...UPDATE THIS...]
+# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# License: MIT
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
+
+
+#### Workspace setup ####
+library(tidyverse)
+# [...UPDATE THIS...]
+
+#### Test data ####

### 2023-04-05 scripts/00-simulate_data.R
Update 00-simulate_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Simulates... [...UPDATE THIS...]
 # Author: Rohan Alexander [...UPDATE THIS...]
-# Data: 11 February 2023 [...UPDATE THIS...]
+# Date: 11 February 2023 [...UPDATE THIS...]
 # Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
 # License: MIT
 # Pre-requisites: [...UPDATE THIS...]

### 2023-04-05 scripts/01-download_data.R
Update 01-download_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Downloads and saves the data from [...UPDATE THIS...]
 # Author: Rohan Alexander [...UPDATE THIS...]
-# Data: 11 February 2023 [...UPDATE THIS...]
+# Date: 11 February 2023 [...UPDATE THIS...]
 # Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
 # License: MIT
 # Pre-requisites: [...UPDATE THIS...]

### 2023-04-05 scripts/02-data_cleaning.R
Update 02-data_cleaning.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Cleans.... [...UPDATE THIS...]
 # Author: Rohan Alexander [...UPDATE THIS...]
-# Data: 11 February 2023 [...UPDATE THIS...]
+# Date: 11 February 2023 [...UPDATE THIS...]
 # Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
 # License: MIT
 # Pre-requisites: [...UPDATE THIS...]

### 2023-04-05 scripts/04-test_data.R
Update 04-test_data.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Tests... [...UPDATE THIS...]
 # Author: Rohan Alexander [...UPDATE THIS...]
-# Data: 11 February 2023 [...UPDATE THIS...]
+# Date: 11 February 2023 [...UPDATE THIS...]
 # Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
 # License: MIT
 # Pre-requisites: [...UPDATE THIS...]

### 2023-04-06 outputs/paper/paper.qmd
Update with a model example

@@
 
-You can and should cross-reference sections and sub-sections. For instance, @sec-data and @sec-first-point.
+You can and should cross-reference sections and sub-sections.
+
+The remainder of this paper is structured as follows. @sec-data....
 
@@
 
-Our data is of penguins (@fig-bills).
+Some of our data is of penguins (@fig-bills), from @palmerpenguins.
 
@@
 
-Also bills and their average (@fig-billssssss).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work.)
+And also planes (@fig-planes).
+(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
 
 ```{r}
-#| label: fig-billssssss
-#| fig-cap: More bills of penguins
+#| label: fig-planes
+#| fig-cap: Relationship between wing length and width
 #| echo: false
-#| fig-width: 8
-#| fig-height: 4
+#| warning: false
+#| message: false
 
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
+analysis_data <- read_csv(here::here("outputs/data/analysis_data.csv"))
+
+analysis_data |> 
+  ggplot(aes(x = width, y = length)) +
+  geom_point(alpha = 0.8) +
   theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
+  labs(x = "Wing width (mm)",
+       y = "Wing length (mm)")
 ```
@@
 
-$$
-Pr(\theta | y) = \frac{Pr(y | \theta) Pr(\theta)}{Pr(y)}
-$$ {#eq-bayes}
+The goal of our modelling strategy is twofold. Firstly,...
 
-@eq-bayes seems useful, eh?
+Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
 
-Here's a dumb example of how to use some references: In paper we run our analysis in `R` [@citeR].
-We also use the `tidyverse` which was written by @thereferencecanbewhatever If we were interested in baseball data then @citeLahman could be useful.
+## Model set-up
+
+Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+
+\begin{align} 
+y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
+\mu_i &= \alpha + \beta_i + \gamma_i\\
+\alpha &\sim \mbox{Normal}(0, 2.5) \\
+\beta &\sim \mbox{Normal}(0, 2.5) \\
+\gamma &\sim \mbox{Normal}(0, 2.5) \\
+\sigma &\sim \mbox{Exponential}(1)
+\end{align}
+
+We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+
+### Model justification
+
+We expect a positive relationship between the size of the wings and time spent aloft. In particular...
 
@@
 # Results
+
+Our results are summarized in @tbl-modelresults.
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+library(rstanarm)
+
+first_model <-
+  readRDS(file = here::here("outputs/models/first_model.rds"))
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-modelresults
+#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
+#| warning: false
+
+modelsummary::modelsummary(
+  list(
+    "First model" = first_model
+  ),
+  statistic = "mad",
+  fmt = 2
+)
+```
 
@@
 
-# Additional details
+# Additional data details
+
+# Model details {#sec-model-details}
+
+## Posterior predictive check
+
+In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
+
+In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
+
+```{r}
+#| eval: true
+#| echo: false
+#| message: false
+#| warning: false
+#| label: fig-ppcheckandposteriorvsprior
+#| layout-ncol: 2
+#| fig-cap: "Examining how the model fits, and is affected by, the data"
+#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
+
+pp_check(first_model) +
+  theme_classic() +
+  theme(legend.position = "bottom")
+
+posterior_vs_prior(first_model) +
+  theme_minimal() +
+  scale_color_brewer(palette = "Set1") +
+  theme(legend.position = "bottom") +
+  coord_flip()
+```
+
+## Diagnostics
+
+@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
+
+@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
+
+```{r}
+#| echo: false
+#| eval: true
+#| message: false
+#| warning: false
+#| label: fig-stanareyouokay
+#| fig-cap: "Checking the convergence of the MCMC algorithm"
+#| fig-subcap: ["Trace plot", "Rhat plot"]
+#| layout-ncol: 2
+
+plot(first_model, "trace")
+
+plot(first_model, "rhat")
+```
 

### 2023-04-06 scripts/02-data_cleaning.R

@@ -1,26 +1,44 @@
 #### Preamble ####
-# Purpose: Cleans.... [...UPDATE THIS...]
+# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
 # Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
+# Date: 6 April 2023 [...UPDATE THIS...]
 # Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
 # License: MIT
 # Pre-requisites: [...UPDATE THIS...]
 # Any other information needed? [...UPDATE THIS...]
 
-
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
 
 #### Clean data ####
-# [...UPDATE THIS...]
-raw_data <- read_csv("inputs/data/raw_data.csv")
+raw_data <- read_csv("inputs/data/plane_data.csv")
 
-
-
+cleaned_data <-
+  raw_data |>
+  janitor::clean_names() |>
+  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
+  filter(wing_width_mm != "caw") |>
+  mutate(
+    flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
+                                   "1.35",
+                                   flying_time_sec_first_timer)
+  ) |>
+  mutate(wing_width_mm = if_else(wing_width_mm == "490",
+                                 "49",
+                                 wing_width_mm)) |>
+  mutate(wing_width_mm = if_else(wing_width_mm == "6",
+                                 "60",
+                                 wing_width_mm)) |>
+  mutate(
+    wing_width_mm = as.numeric(wing_width_mm),
+    wing_length_mm = as.numeric(wing_length_mm),
+    flying_time_sec_first_timer = as.numeric(flying_time_sec_first_timer)
+  ) |>
+  rename(flying_time = flying_time_sec_first_timer,
+         width = wing_width_mm,
+         length = wing_length_mm
+         ) |> 
+  tidyr::drop_na()
 
 #### Save data ####
-# [...UPDATE THIS...]
-# change cleaned_data to whatever name you end up with at the end of cleaning
-write_csv(cleaned_data, "outputs/data/cleaned_data.csv")
-
+write_csv(cleaned_data, "outputs/data/analysis_data.csv")

### 2023-04-06 scripts/04-model.R

@@ -1,0 +1,37 @@
+#### Preamble ####
+# Purpose: Models... [...UPDATE THIS...]
+# Author: Rohan Alexander [...UPDATE THIS...]
+# Date: 11 February 2023 [...UPDATE THIS...]
+# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# License: MIT
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(rstanarm)
+
+#### Read data ####
+analysis_data <- read_csv("outputs/data/analysis_data.csv")
+
+### Model data ####
+first_model <-
+  stan_glm(
+    formula = flying_time ~ length + width,
+    data = analysis_data,
+    family = gaussian(),
+    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
+    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
+    prior_aux = exponential(rate = 1, autoscale = TRUE),
+    seed = 853
+  )
+
+
+#### Save model ####
+saveRDS(
+  first_model,
+  file = "outputs/models/first_model.rds"
+)
+
+

### 2024-01-09 outputs/paper/paper.qmd
Make tweaks for 2024

@@
 thanks: "Code and data are available at: LINK."
-date: "`r Sys.time()`"
-date-format: "D MMMM YYYY"
+date: today
+date-format: long
 abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."

### 2024-02-07 scripts/04-model.R
update structure especially the data folder

@@ -13,7 +13,7 @@ library(tidyverse)
 library(rstanarm)
 
 #### Read data ####
-analysis_data <- read_csv("outputs/data/analysis_data.csv")
+analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
 
 ### Model data ####
 first_model <-

### 2024-02-07 scripts/99-replications.R

@@ -1,0 +1,19 @@
+#### Preamble ####
+# Purpose: Replicated graphs from... [...UPDATE THIS...]
+# Author: Rohan Alexander [...UPDATE THIS...]
+# Date: 11 February 2023 [...UPDATE THIS...]
+# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# License: MIT
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
+
+
+#### Workspace setup ####
+library(tidyverse)
+# [...UPDATE THIS...]
+
+#### Load data ####
+# [...ADD CODE HERE...]
+
+
+

### 2024-02-07 scripts/04-model.R
Tweak the model bit

@@ -31,7 +31,7 @@ first_model <-
 #### Save model ####
 saveRDS(
   first_model,
-  file = "outputs/models/first_model.rds"
+  file = "models/first_model.rds"
 )
 
 

### 2024-03-29 other/datasheet/datasheet_template.qmd
Move datasheet and add checklist to README

--- 2022-02-12 inputs/data/datasheet_template.Rmd
@@
 ---
-title: "My title"
+title: "Datasheet for 'A dataset'"
 subtitle: "My subtitle if needed"
@@
 thanks: "Code and data are available at: LINK."
-date: "`r format(Sys.time(), '%d %B %Y')`"
+date: "`r Sys.time()`"
+date-format: "D MMMM YYYY"
 abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
-output:
-  bookdown::pdf_document2
-toc: FALSE
-bibliography: references_datasheet.bib
+format: pdf
+number-sections: true
+bibliography: references.bib
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
-```
-
-Extract of the questions from @gebru2021datasheets
+Extract of the questions from @gebru2021datasheets.
 

### 2024-07-12 paper/paper.qmd
Tweak the model bit
more setup, added literature

--- 2024-01-09 outputs/paper/paper.qmd
@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "[Result, but poetic]: Does LLM Use Improve Data Science Education?"
+subtitle: "Evidence from a Canadian Undergraduate Statistics Course"
 author: 
-  - First author
-  - Another author
-thanks: "Code and data are available at: LINK."
+  - Rohan Alexander
+  - Luca Carnegie
+thanks: "Code and data are available at: https://github.com/lcarnegie/llms-achievement."
 date: today
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "Blah Blah Blah Blah"
 format: pdf
@@
 library(tidyverse)
-library(palmerpenguins)
 ```
@@
 
-You can and should cross-reference sections and sub-sections.
+You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.
 
@@
 # Data {#sec-data}
-
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
-
-```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
-#| echo: false
-
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
-```
-
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
-
-analysis_data <- read_csv(here::here("outputs/data/analysis_data.csv"))
-
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
-```
-
-Talk way more about it.
 
@@
 first_model <-
-  readRDS(file = here::here("outputs/models/first_model.rds"))
+  readRDS(file = here::here("models/first_model.rds"))
 ```

### 2024-07-15 paper/paper.qmd
added intro + lit review

@@
 
-You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.
+The rapid advancement of Large Language Models (LLMs) like ChatGPT has had massive effects in all areas of society, with some increases in productivity within some knowledge-based industries.
+Recent papers have demonstrated evidence of some productivity increases, in industries ranging from consulting to computer programming.
 
-The remainder of this paper is structured as follows. @sec-data....
+On the lower end of productivity increases, Ben-Michael et al.
+(2024) conducted a randomized control trial with judges in Wisconsin to assess the impact of AI recommendations on bail hearing decisions.
+The study found that providing AI recommendations did not significantly improve judges' decision-making accuracy.
+On the other hand, Peng et al.
+(2023) investigated the significantly positive impact of GitHub Copilot, an LLM-powered coding assistant, on programmer productivity.
+They conducted an experiment with 95 professional programmers recruited through Upwork.
+The study found that programmers with access to Copilot completed a standardized programming task 55.8% faster than the control group.
+In particular, less experienced programmers appeared to benefit more from the tool.
+
+A study by Dell'Acqua et al.
+(2023) examined the impact of AI on management consulting tasks through a field experiment with 758 Boston Consulting Group consultants.
+They found that AI significantly increased performance across various business tasks, with speed increasing by over 25% and quality by more than 40%.
+However, the study revealed a "jagged technological frontier" where AI excelled in some tasks but struggled with others.
+Notably, AI usage appeared to level performance differences across ability levels, with lower-performing consultants benefiting the most, similar to that of Peng et al.
+
+We can see in the current literature that LLMs have had varying levels of impact across different knowledge-intensive industries, with more computer-intensive tasks like programming receiving the highest performance boosts.
+One area sorely lacking quantitative evidence of performance increases is within higher education, an area where scrutiny of the impact of LLMs and other AI tools has been common.
+Cahill and McCabe (2023) conducted a survey of undergraduate political science students, revealing widespread use of AI tools like ChatGPT.
+However, their study also uncovered a significant gap in students' perceived competence in using these tools effectively for academic work, though it remains to be seen if this gap is smaller in numerical-based fields like statistics and data science.
+
+Given the rapid integration of AI tools in educational and professional settings, there is an urgent need to understand how these technologies affect student learning and performance, particularly in fields like data science and statistics.
+This paper aims to investigate the relationship between LLM usage and academic performance, using evidence from an exit survey of a third-year undergraduate statistics course at the University of Toronto.
+By examining [how students interact with AI tools and how this usage correlates with their grades, we seek to understand the potential benefits and pitfalls of AI integration in data science education].
+
+The remainder of this paper is structured as follows: @sec-data analyzes our survey data; @sec-model models our unstructured data to find fun and cool things; @sec-results lists the results; @sec-discussion discusses the implications of these findings for data science education and for future research and practice in this rapidly evolving field.
 
@@
 
-# Model
+# Model {#sec-model}
 
@@
 
-# Results
+# Results {#sec-results}
 
@@
 
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-library(rstanarm)
-
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
-```
-
-# Discussion
+# Discussion {#sec-discussion}
 
@@
 
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
+In ... we implement a posterior predictive check. This shows...
 
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-```
+In ... we compare the posterior with the prior. This shows...
 
@@
 
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
+... is a trace plot. It shows... This suggests...
 
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
-```
+... is a Rhat plot. It shows... This suggests...
 

### 2024-07-15 paper/paper.qmd
updated paper with info on data science student challenges

@@
 
-The rapid advancement of Large Language Models (LLMs) like ChatGPT has had massive effects in all areas of society, with some increases in productivity within some knowledge-based industries.
+The rapid advancement of Large Language Models (LLMs) like ChatGPT has had massive effects in all areas of society, with notable increases in productivity within some knowledge-based industries.
 Recent papers have demonstrated evidence of some productivity increases, in industries ranging from consulting to computer programming.
 
-On the lower end of productivity increases, Ben-Michael et al.
-(2024) conducted a randomized control trial with judges in Wisconsin to assess the impact of AI recommendations on bail hearing decisions.
+On the lower end of productivity increases, @Ben_2024 conducted a randomized control trial with judges in Wisconsin to assess the impact of AI recommendations on bail hearing decisions.
 The study found that providing AI recommendations did not significantly improve judges' decision-making accuracy.
-On the other hand, Peng et al.
-(2023) investigated the significantly positive impact of GitHub Copilot, an LLM-powered coding assistant, on programmer productivity.
+On the other hand, @Peng_2023 investigated the significantly positive impact of GitHub Copilot, an LLM-powered coding assistant, on programmer productivity.
 They conducted an experiment with 95 professional programmers recruited through Upwork.
@@
 
-A study by Dell'Acqua et al.
-(2023) examined the impact of AI on management consulting tasks through a field experiment with 758 Boston Consulting Group consultants.
+Examining the middle-ground, a study by @DellAcqua_2023 examined the impact of AI on management consulting tasks through a field experiment with 758 Boston Consulting Group consultants.
 They found that AI significantly increased performance across various business tasks, with speed increasing by over 25% and quality by more than 40%.
@@
 One area sorely lacking quantitative evidence of performance increases is within higher education, an area where scrutiny of the impact of LLMs and other AI tools has been common.
-Cahill and McCabe (2023) conducted a survey of undergraduate political science students, revealing widespread use of AI tools like ChatGPT.
-However, their study also uncovered a significant gap in students' perceived competence in using these tools effectively for academic work, though it remains to be seen if this gap is smaller in numerical-based fields like statistics and data science.
+A survey of undergraduate political science students was conducted by @Cahill_2024, revealing widespread use of AI tools like ChatGPT.
+However, their study also uncovered a significant gap in students' self-perceived competence in using these tools effectively for academic work.
 
-Given the rapid integration of AI tools in educational and professional settings, there is an urgent need to understand how these technologies affect student learning and performance, particularly in fields like data science and statistics.
-This paper aims to investigate the relationship between LLM usage and academic performance, using evidence from an exit survey of a third-year undergraduate statistics course at the University of Toronto.
-By examining [how students interact with AI tools and how this usage correlates with their grades, we seek to understand the potential benefits and pitfalls of AI integration in data science education].
+Student challenges while learning Data Science in particular were explored by @kross2019.
+By polling instructor-practitioners of Data Science in academia and industry, they found several key challenges that data science students face.
+The largest challenges they faced were diverse computing backgrounds, integrating various technical workflows, software setup, finding relevant datasets, as well as dealing with the inherent uncertainty of the data analysis process.
+
+Given the rapid integration of AI tools in educational and professional settings, there is a need to understand how these technologies and students' attitudes toward LLMs affect their  learning and performance.
+This paper aims to investigate the relationship between academic performance, LLM usage, and students' perceptions of LLMs, using evidence from an exit survey of a third-year undergraduate statistics course at the University of Toronto.
+By examining how students interact with and perceive LLM tools and how this usage correlates with their grades, we seek to understand the potential benefits and pitfalls of AI integration in data science education.
 

### 2024-07-16 scripts/01-download_data.R
imported data

@@ -1,26 +1,21 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Downloads and saves the data from Excel file.
+# Author: Luca Carnegie
+# Date: 16 July 2024
+# Contact: luca.carnegie@mail.utoronto.ca or rohan.alexander@utoronto.ca 
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: None
 
 
 #### Workspace setup ####
-library(opendatatoronto)
 library(tidyverse)
-# [...UPDATE THIS...]
+library(readr)
 
 #### Download data ####
-# [...ADD CODE HERE TO DOWNLOAD...]
-
+sta302_survey <- read_excel("data/raw_data/raw_STA302_postcourse_survey_w24.xlsx")
 
 
 #### Save data ####
-# [...UPDATE THIS...]
-# change the_raw_data to whatever name you assigned when you downloaded it.
-write_csv(the_raw_data, "inputs/data/raw_data.csv") 
+write_csv(sta302_survey, "data/raw_data/raw_STA302_postcourse_survey_w24.csv") 
 
          

### 2024-07-16 scripts/02-data_cleaning.R
started cleaning data

@@ -1,44 +1,41 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Purpose: Cleans the raw survey data recored by Rohan at the end of STA302
+# Author: Luca Carnegie
+# Date: 16 July 2024
+# Contact: luca.carnegie@mail.utoronto.ca or rohan.alexander@utoronto.ca
+# Pre-requisites: 01-download_data.R
 
 #### Workspace setup ####
 library(tidyverse)
 
 #### Clean data ####
-raw_data <- read_csv("inputs/data/plane_data.csv")
+raw_data <- read_csv("data/raw_data/raw_STA302_postcourse_survey_w24.csv")
 
-cleaned_data <-
-  raw_data |>
-  janitor::clean_names() |>
-  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-  filter(wing_width_mm != "caw") |>
-  mutate(
-    flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
-                                   "1.35",
-                                   flying_time_sec_first_timer)
-  ) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "490",
-                                 "49",
-                                 wing_width_mm)) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "6",
-                                 "60",
-                                 wing_width_mm)) |>
-  mutate(
-    wing_width_mm = as.numeric(wing_width_mm),
-    wing_length_mm = as.numeric(wing_length_mm),
-    flying_time_sec_first_timer = as.numeric(flying_time_sec_first_timer)
-  ) |>
-  rename(flying_time = flying_time_sec_first_timer,
-         width = wing_width_mm,
-         length = wing_length_mm
-         ) |> 
-  tidyr::drop_na()
+raw_data <- as_tibble(raw_data)
+
+# Drop unnecessary columns
+
+raw_data <- raw_data |>
+  select(-c(Email, 
+            Name, 
+            `Last modified time`, 
+            `What is your full name on Quercus?`, #anonymized
[…]

### 2024-07-18 scripts/02-data_cleaning.R
more cleaning

@@ -7,28 +7,78 @@
 
 #### Workspace setup ####
 library(tidyverse)
+library(janitor)
+
 
 #### Clean data ####
 raw_data <- read_csv("data/raw_data/raw_STA302_postcourse_survey_w24.csv")
 
 raw_data <- as_tibble(raw_data)
 
+# Clean up names of columns
+
+raw_data <- raw_data |> 
+  clean_names()
+
+raw_data <- raw_data |> 
+  rename(
+    consent = after_carefully_reading_the_informed_consent_document_please_indicate_below_whether_you_consent_to_have_your_anonymized_responses_included_in_the_research_study,
+    student_name = what_is_your_full_name_on_quercus,
+    student_id = what_is_your_student_id,
+    year = what_year_are_you,
+    program_of_study = what_is_your_specialization, 
+    major = what_is_are_your_major_s,
+    minor = what_is_are_your_minor_s,
+    gpa = what_is_your_gpa, 
+    ai_familiarity = how_familiar_are_you_with_using_generative_ai_tools_such_as_open_a_is_chat_gpt_or_equivalents, 
+    used_ai_any_reason = have_you_used_any_generative_ai_tools_such_as_open_a_is_chat_gpt_or_equivalents_for_any_reason_personal_or_educational, 
+    ai_use_cases = if_you_have_used_generative_ai_tools_such_as_open_a_is_chat_gpt_or_equivalents_in_what_ways_have_you_used_it_select_all_that_apply, 
+    ai_schoolwork_appropriate = to_what_extent_do_you_think_using_generative_ai_tools_such_as_chat_gpt_by_open_ai_or_equivalents_is_ethical_and_appropriate_for_schoolwork, 
+    ai_schoolwork_appropriate_comments = please_elaborate_on_your_answer_above,
+    used_ai_in_sta302 = did_you_use_any_generative_ai_tools_such_as_open_a_is_chat_gpt_or_equivalents_for_sta302, 
+    ai_helpful_weekly_quiz = weekly_quiz,
+    ai_helpful_weekly_mini_essay = weekly_mini_essay,
+    ai_helpful_papers_generating_ideas = papers_generating_ideas,
+    ai_helpful_papers_writing_code = papers_writing_code,
+    ai_helpful_papers_writing_content = papers_writing_content,
+    ai_helpful_papers_improving_content = papers_improving_content,
[…]

### 2024-07-18 paper/paper.qmd
cleaned some more, made a visualization

@@
 library(tidyverse)
+```
+
+```{r}
+# Import data (CHANGE WHEN FULLY CLEAN!)
+raw_data <- read_csv("data/raw_data/raw_STA302_postcourse_survey_w24.csv")
 ```
@@
 # Data {#sec-data}
+
+```{r}
+#Histogram of GPA
+
+raw_data |>
+  ggplot(aes(x = gpa)) +
+  geom_histogram(binwidth = 0.1, fill = "skyblue", color = "black") +
+  labs(title = "Distribution of GPAs in the Survey Data",
+       x = "GPA",
+       y = "Frequency") +
+  theme_minimal()
+
+```
 

### 2024-07-18 scripts/02-data_cleaning.R

@@ -52,6 +52,8 @@ raw_data <- raw_data |>
 
 
 
+
+
 # Drop unnecessary columns
 
 raw_data <- raw_data |>

### 2024-07-19 paper/paper.qmd
finished code to clean program names; finished code for ai_use_cases; visualized GPA and year distribution; data clean!

@@
 
-```{r}
+```{r fig-gpa-hist}
 #Histogram of GPA
@@
   theme_minimal()
+
+```
+
+```{r fig-year-bar}
+#Bar plot of year
+raw_data |> ggplot(aes(x = year)) +
+  geom_bar(fill = "skyblue", color = "black") +
+  labs(title = "Distribution of Students by Year",
+       x = "Year",
+       y = "Frequency") +
+  theme_minimal()
+
+```
+
+```{r fig-writingease-bar}
+#Bar plot of writing ease
+raw_data |> ggplot(aes(x = writing_is_easy_for_me)) +
+  geom_bar(fill = "skyblue", color = "black") +
+  labs(title = "Distribution of Responses to AI Schoolwork Appropriateness",
+       x = "Response",
+       y = "Frequency") +
+  theme_minimal()  # <- fix the labels as they encroach. NO it broke.
 

### 2024-07-19 scripts/02-data_cleaning.R

@@ -8,6 +8,7 @@
 #### Workspace setup ####
 library(tidyverse)
 library(janitor)
+library(stringr)
 
 
 #### Clean data ####
@@ -50,43 +51,121 @@ raw_data <- raw_data |>
 raw_data <- raw_data |>
   filter(consent == "Yes, I authorize the use of the data collected about me for the STA302 course survey to be used. I will be compensated 1% of my course grade for completing the survey.")
 
-
-
-
-
 # Drop unnecessary columns
-
 raw_data <- raw_data |>
-  select(-c(consent, 
+  select(-c(email,
+            name, 
+            last_modified_time, 
+            consent, 
             student_name, 
             student_id, 
             major, #none submitted
             minor, #none submitted
-            
             ))
 
 
-# Clean specialization column (text data oh god)
+# Standardize program names 
 
-#column_clean <-
+#Description: Go through each row. If you find a program that matchs a standard 
+#             program, then put it in a string, do this until all matches are 
+#             found, after which you put a string with the programs separated by 
+#             semi-colons. 
+
+# Define the mapping to standard names. 
+standard_programs <- list(
+  "Statistics" = c("Statistics", "Stats", "Sta", "sta", "Stat", "statistics", "STAT", "STATS", "STATISTICS"),
+  "Economics" = c("Economics", "Economiccs", "Econ", "Eco", "eco", "economics", "ECONOMICS", "ECON", "ECO", "Focus in Data Analytics", "Focus in data analytics"),
+  "Computer Science" = c("Computer science", "Computer Science", "CS", "cs", "computer science", "COMPUTER SCIENCE", "CompSci", "comp sci", "COMP SCI"),
+  "Education" = c("Education", "education", "EDUCATION", "Edu", "edu", "EDU"),
+  "Mathematics" = c("Mathematics", "Math", "math", "Mathematical Applications in Economics and Finance", "mathematical applications in econ and finance", "MATH", "MATHEMATICS", "Mathematical applications in economics and finance"),
[…]

### 2024-07-19 scripts/02-data_cleaning.R
finished cleaning code; visualized GPA and year distributions

@@ -16,7 +16,7 @@ raw_data <- read_csv("data/raw_data/raw_STA302_postcourse_survey_w24.csv")
 
 raw_data <- as_tibble(raw_data)
 
-# Clean up names of columns
+# CLEAN UP COLUMN NAMES
 
 raw_data <- raw_data |> 
   clean_names()
@@ -47,11 +47,14 @@ raw_data <- raw_data |>
     other_comments = optional_any_other_comments
   )
 
-# Drop rows for students who did not consent
+# DROP ROWS FOR UNCONSENTING STUDENT OR TA
 raw_data <- raw_data |>
-  filter(consent == "Yes, I authorize the use of the data collected about me for the STA302 course survey to be used. I will be compensated 1% of my course grade for completing the survey.")
+  filter(
+    consent == "Yes, I authorize the use of the data collected about me for the STA302 course survey to be used. I will be compensated 1% of my course grade for completing the survey.", 
+    year != "TA"
+    )
 
-# Drop unnecessary columns
+# DROP UNECESSARY OR UNSPECIFIED COLUMNS
 raw_data <- raw_data |>
   select(-c(email,
             name, 
@@ -63,6 +66,13 @@ raw_data <- raw_data |>
             minor, #none submitted
             ))
 
+
+# CLEAN YEARS
+raw_data <- raw_data |>
+  mutate(
+    year = str_sub(year, 1, 3)
+  )
+  
 
 # Standardize program names 
 
@@ -83,7 +93,7 @@ standard_programs <- list(
   "Cognitive Science" = c("Cognitive Science", "Cog Sci", "Cogsci", "cognitive science", "COGNITIVE SCIENCE", "COG SCI", "cogsci", "Cogsci", "cog sci"),
   "Engineering" = c("Engineering Science", "engineering science", "ENGINEERING SCIENCE", "Engineering", "engineering", "ENGINEERING"),
   "Life Sciences" = c("Life Sciences", "Life science", "Life Science", "Physiology", "life sciences", "life science", "LIFE SCIENCES", "LIFE SCIENCE", "physiology", "PHYSIOLOGY"),
-  "Physics" = c("Chemical Physics", "chemical physics", "CHEMICAL PHYSICS", "Physics", "physics", "PHYSICS"),
+  "Physics" = c("Chemical Physics Specialist", "chemical physics", "CHEMICAL PHYSICS", "Physics", "physics", "PHYSICS"),
[…]

### 2024-07-19 paper/paper.qmd
finished some initial visualizations

--- 2024-07-15 paper/paper.qmd
@@
 bibliography: references.bib
+nocite: |
+  @*
 ---
 
-```{r}
+```{r setup}
+## Workspace Setup ##
 #| include: false
@@
 library(tidyverse)
+library(knitr)
+library(kableExtra)
+library(lintr)
+
+# Import data
+survey_data <- read_csv("../data/analysis_data/clean_STA302_postcourse_survey_w24.csv")
 ```
@@
 # Data {#sec-data}
+
+```{r fig-boilerplate}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: (CHANGE) Boilerplate figure caption 
+#| fig-align: center
+
+# survey_data |>
+#   ggplot(program_counts, aes(x = "", y = n, fill = program_of_study)) +
+#   #put geom type here (put before compilation) +
+#   labs(title = "Title!",
+#        x = "X label",
+#        y = "Y label") +
+#   theme_minimal() +
+#   scale_fill_discrete(name="Legend Title") +
+#   theme(plot.title.position = "plot",
+#         plot.title = element_text(hjust = 0.5)
+#         )
+
+```
+
+```{r tbl-boilerplate}
+#| message: false
+#| echo: false
+#| warning: false
+#| tbl-cap: (CHANGE) Boilerplate table caption 
+#| fig-align: center
+
+#Put code below this line
+
+```
+
+```{r fig-gpa-hist}
+#| message: false
+#| echo: false
+#| warning: false
+#| tbl-cap: Distribution of Student GPA
+#| fig-align: center
+
+#Histogram of GPA
+
+survey_data |>
+    ggplot(aes(x = gpa)) +
+    geom_histogram(binwidth = 0.1, fill = "skyblue", color = "black") +
+    labs(title = "Distribution of Student GPA",
+         x = "GPA",
+         y = "Frequency") +
+    theme_minimal() +
+    theme(plot.title.position = "plot",
+          plot.title = element_text(hjust = 0.5))
+
+```
+
+```{r fig-year-bar}
+#| message: false
+#| echo: false
+#| warning: false
+#| tbl-cap: Distribution of Students by Year
+#| fig-align: center
+
+#Bar plot of year
+survey_data |> ggplot(aes(x = year)) +
+    geom_bar(fill = "skyblue", color = "black") +
+    labs(title = "Distribution of Students by Year",
+         x = "Year",
+         y = "Frequency") +
+    theme_minimal() +
+    theme(plot.title.position = "plot",
+          plot.title = element_text(hjust = 0.5))
+
+```
+
+```{r fig-programs-pie}
+#| message: false
+#| echo: false
+#| warning: false
+#| tbl-cap: Student Self-Perception of Writing Ease
+#| fig-align: center
+
+#Pie chart of program of study
+
+program_counts <- survey_data |> 
+  count(program_of_study)
+
+ggplot(program_counts, aes(x = "", y = n, fill = program_of_study)) +
+  geom_bar(stat = "identity", width = 1) +
+  coord_polar("y") +
+  labs(title = "Distribution of Unique Programs",
+       x = NULL,
+       y = NULL) +
+  theme_minimal() +
+  scale_fill_discrete(name="Program") +
+  theme(axis.text.x = element_blank(),  # Remove x-axis text
+        axis.title.x = element_blank(), # Remove x-axis title
+        plot.title = element_text(hjust = 1.35)
+        )
+
+```
+
+```{r fig-writingease-bar}
+#| message: false
+#| echo: false
+#| warning: false
+#| tbl-cap: Student Self-Perception of Writing Ease
+#| fig-align: center
+
+#Bar plot of writing ease
+survey_data |> ggplot(aes(x = writing_is_easy_for_me)) +
+  geom_bar(fill = "skyblue", color = "black") +
+  labs(title = "Agreement to the Statement: \"Writing is Easy for Me\"",
+       x = "Response",
+       y = "Frequency") +
+  theme_minimal() +
+  theme(plot.title.position = "plot",
+        plot.title = element_text(hjust = 0.5))
+
+```
 

### 2024-07-19 scripts/02-data_cleaning.R

@@ -68,6 +68,7 @@ raw_data <- raw_data |>
 
 
 # CLEAN YEARS
+
 raw_data <- raw_data |>
   mutate(
     year = str_sub(year, 1, 3)
@@ -205,4 +206,5 @@ raw_data <- raw_data |>
 
 
 #### Save data ####
+cleaned_data <- raw_data
 write_csv(cleaned_data, "data/analysis_data/clean_STA302_postcourse_survey_w24.csv")

### 2024-07-24 paper/paper.qmd
finished multibar self-perception visualizations

@@
 bibliography: references.bib
-nocite: |
+nocite: | #references.bib entries at end, regardless of intext citation
   @* 
@@
 library(lintr)
+library(tidyr)
 
@@
 
+# Change PEY to 4th year
+survey_data <- survey_data |>
+    mutate(
+      year = case_when(
+        year == "PEY" ~ "4th",
+        year == "5th" ~ "5th+",
+        year == "6th" ~ "5th+",
+        TRUE ~ year
+      )
+    )
+
 #Bar plot of year
@@
 
-```{r fig-writingease-bar}
+```{r fig-response-time-hist}
 #| message: false
@@
 #| warning: false
-#| tbl-cap: Student Self-Perception of Writing Ease
+#| fig-cap: Distribution of Survey Response Times
 #| fig-align: center
 
-#Bar plot of writing ease
-survey_data |> ggplot(aes(x = writing_is_easy_for_me)) +
-  geom_bar(fill = "skyblue", color = "black") +
-  labs(title = "Agreement to the Statement: \"Writing is Easy for Me\"",
+#Trim values to histogram
+
+plot_data <- survey_data |>
+    filter(mins_complete < 1000)
+
+plot_data |>
+    ggplot(aes(x = mins_complete)) +
+    geom_histogram(binwidth = 0.1, fill = "skyblue", color = "black") +
+    labs(title = "Distribution of Survey Response Times",
+         x = "Time (Mins)",
+         y = "Frequency") +
+    theme_minimal() +
+    theme(plot.title.position = "plot",
+          plot.title = element_text(hjust = 0.5))
+
+```
+
+```{r fig-writing-multibar-1}
+#| message: false
+#| echo: false
+#| warning: false
+#| tbl-cap: Student Self-Perceptions on Writing, Part 1
+#| fig-align: center
+
+# Create a multi-bar plot for writing perceptions grouped by response
+
+#Select first three statements
+writing_perceptions <- survey_data |>
+  select(
+    writing_is_easy_for_me, 
+    i_like_to_write, 
+    i_believe_it_is_important_to_be_a_good_writer
+  )
+
+writing_perceptions_long <- writing_perceptions |>
+  pivot_longer(cols = everything(), names_to = "Statements", values_to = "Response") |>
+  mutate(
+    Statements = case_when(
+      Statements == "writing_is_easy_for_me" ~ "\"Writing is easy for me\"",
+      Statements == "i_like_to_write" ~ "\"I like to write\"",
+      Statements == "i_believe_it_is_important_to_be_a_good_writer" ~ "\"I believe it is important\n to be a good writer\"",
+      TRUE ~ Statements
+    ), 
+    Response = case_when(
+      Response == "This is a lot like me" ~ "A lot like me",
+      Response == "This is very different to me" ~ "Very different from me",
+      Response == "This somewhat describes me" ~ "Somewhat describes me",
+    )
+  ) |>
+  filter(!is.na(Response))
+
+#Taken from: https://stackoverflow.com/questions/52297978/decrease-overal-legend-size-elements-and-text
+addSmallLegend <- function(myPlot, pointSize = 0.5, textSize = 3, spaceLegend = 0.1) {
+    myPlot +
+        guides(shape = guide_legend(override.aes = list(size = pointSize)),
+               color = guide_legend(override.aes = list(size = pointSize))) +
+        theme(legend.title = element_text(size = textSize), 
+              legend.text  = element_text(size = textSize),
+              legend.key.size = unit(spaceLegend, "lines"))
+}
+
+# Make the plot
+multi_bar <- writing_perceptions_long |>
+  ggplot(aes(x = Response, fill = Statements)) +
+  geom_bar(position = "dodge", stat = "count") +
+  labs(title = "Student Self-Perceptions on Writing",
        x = "Response",
@@
   theme(plot.title.position = "plot",
-        plot.title = element_text(hjust = 0.5))
+        plot.title = element_text(hjust = 0.5),
+        legend.position= "bottom", 
+      )
+
+multi_bar <- addSmallLegend(multi_bar, pointSize = 0.5, textSize = 10, spaceLegend = 0.8)
+
+multi_bar
+
+# TODO: Fix label, Fix Legend title, Fix legibility of Response
+
+```
+
+```{r fig-writing-multibar-2}
+#| message: false
+#| echo: false
+#| warning: false
+#| tbl-cap: Student Self-Perceptions on Writing, Part 2
+#| fig-align: center
+
+# Create a multi-bar plot for writing perceptions grouped by response
+
+# Select last three statements
+writing_perceptions <- survey_data |>
+  select(
+    when_i_edit_it_is_easy_for_me_to_catch_my_mistakes, 
+    i_feel_confident_sharing_my_writing, 
+    i_am_confident_in_my_overall_writing_ability
+  )
+
+writing_perceptions_long <- writing_perceptions |>
+  pivot_longer(cols = everything(), names_to = "Statements", values_to = "Response") |>
+  mutate(
+      Statements = case_when(
+      Statements == "when_i_edit_it_is_easy_for_me_to_catch_my_mistakes" ~ "\"When I edit, it is easy to\n catch my own mistakes\"",
+      Statements == "i_feel_confident_sharing_my_writing" ~ "\"I feel confident\n sharing my writing\"",
+      Statements == "i_am_confident_in_my_overall_writing_ability" ~ "\"I am confident in\n my own writing ability\"",
+      TRUE ~ Statements
+    ), 
+    Response = case_when(
+      Response == "This is a lot like me" ~ "A lot like me",
+      Response == "This is very different to me" ~ "Very different from me",
+      Response == "This somewhat describes me" ~ "Somewhat describes me",
+    )
+  ) |>
+  filter(!is.na(Response))
+
+#Taken from: https://stackoverflow.com/questions/52297978/decrease-overal-legend-size-elements-and-text
+addSmallLegend <- function(myPlot, pointSize = 0.5, textSize = 3, spaceLegend = 0.1) {
+    myPlot +
+        guides(shape = guide_legend(override.aes = list(size = pointSize)),
+               color = guide_legend(override.aes = list(size = pointSize))) +
+        theme(legend.title = element_text(size = textSize), 
+              legend.text  = element_text(size = textSize),
+              legend.key.size = unit(spaceLegend, "lines"))
+}
+
+# Make the plot
+multi_bar <- writing_perceptions_long |>
+  ggplot(aes(x = Response, fill = Statements)) +
+  geom_bar(position = "dodge", stat = "count") +
+  labs(title = "Student Self-Perceptions on Writing",
+       x = "Response",
+       y = "Frequency") +
+  theme_minimal() +
+  theme(plot.title.position = "plot",
+        plot.title = element_text(hjust = 0.5),
+        legend.position= "bottom", 
+      )
+
+multi_bar <- addSmallLegend(multi_bar, pointSize = 0.5, textSize = 10, spaceLegend = 0.8)
+
+multi_bar
+```
+
+```{r fig-coding-multibar-1}
+#| message: false
+#| echo: false
+#| warning: false
+#| tbl-cap: Student Self-Perceptions on Coding, Part 1
+#| fig-align: center
+
+# Create a multi-bar plot for writing perceptions grouped by response
+
+#Select first three statements
+writing_perceptions <- survey_data |>
+  select(
+    coding_is_easy_for_me,
+    i_like_to_code,
+    i_believe_it_is_important_to_be_a_good_coder
+  )
+
+writing_perceptions_long <- writing_perceptions |>
+  pivot_longer(cols = everything(), names_to = "Statements", values_to = "Response") |>
+  mutate(
+    Statements = case_when(
+      Statements == "coding_is_easy_for_me" ~ "\"Coding is easy for me\"",
+      Statements == "i_like_to_code" ~ "\"I like to code\"",
+      Statements == "i_believe_it_is_important_to_be_a_good_coder" ~ "\"I believe it is important\n to be a good coder\"",
+      TRUE ~ Statements
+    ), 
+    Response = case_when(
+      Response == "This is a lot like me" ~ "A lot like me",
+      Response == "This is very different to me" ~ "Very different from me",
+      Response == "This somewhat describes me" ~ "Somewhat describes me",
+    )
+  ) |>
+  filter(!is.na(Response))
+
+#Taken from: https://stackoverflow.com/questions/52297978/decrease-overal-legend-size-elements-and-text
+addSmallLegend <- function(myPlot, pointSize = 0.5, textSize = 3, spaceLegend = 0.1) {
+    myPlot +
+        guides(shape = guide_legend(override.aes = list(size = pointSize)),
+               color = guide_legend(override.aes = list(size = pointSize))) +
+        theme(legend.title = element_text(size = textSize), 
+              legend.text  = element_text(size = textSize),
+              legend.key.size = unit(spaceLegend, "lines"))
+}
+
+# Make the plot
+multi_bar <- writing_perceptions_long |>
+  ggplot(aes(x = Response, fill = Statements)) +
+  geom_bar(position = "dodge", stat = "count") +
+  labs(title = "Student Self-Perceptions of Coding",
+       x = "Response",
+       y = "Frequency") +
+  theme_minimal() +
+  theme(plot.title.position = "plot",
+        plot.title = element_text(hjust = 0.5),
+        legend.position= "bottom", 
+      )
+
+multi_bar <- addSmallLegend(multi_bar, pointSize = 0.5, textSize = 10, spaceLegend = 0.8)
+
+multi_bar
+
+```
+
+```{r fig-coding-multibar-2}
+#| message: false
+#| echo: false
+#| warning: false
+#| tbl-cap: Student Self-Perceptions on Coding, Part 2
+#| fig-align: center
+
+# Create a multi-bar plot for writing perceptions grouped by response
+
+#Select first three statements
+writing_perceptions <- survey_data |>
+  select(
+    when_i_check_my_code_it_is_easy_for_me_to_catch_my_mistakes, 
+    i_feel_confident_sharing_my_code,
+    i_am_confident_in_my_overall_coding_ability
+  )
+
+writing_perceptions_long <- writing_perceptions |>
+  pivot_longer(cols = everything(), names_to = "Statements", values_to = "Response") |>
+  mutate(
+    Statements = case_when(
+      Statements == "when_i_check_my_code_it_is_easy_for_me_to_catch_my_mistakes" ~ "\"When I check my code, it is\n easy for me to catch my mistakes\"",
+      Statements == "i_feel_confident_sharing_my_code" ~ "\"I feel confident sharing my code\"",
+      Statements == "i_am_confident_in_my_overall_coding_ability" ~ "\"I am confident in\n my overall coding ability\"",
+      TRUE ~ Statements
+    ), 
+    Response = case_when(
+      Response == "This is a lot like me" ~ "A lot like me",
+      Response == "This is very different to me" ~ "Very different from me",
+      Response == "This somewhat describes me" ~ "Somewhat describes me",
+    )
+  ) |>
+  filter(!is.na(Response))
+
+#Taken from: https://stackoverflow.com/questions/52297978/decrease-overal-legend-size-elements-and-text
+addSmallLegend <- function(myPlot, pointSize = 0.5, textSize = 3, spaceLegend = 0.1) {
+    myPlot +
+        guides(shape = guide_legend(override.aes = list(size = pointSize)),
+               color = guide_legend(override.aes = list(size = pointSize))) +
+        theme(legend.title = element_text(size = textSize), 
+              legend.text  = element_text(size = textSize),
+              legend.key.size = unit(spaceLegend, "lines"))
+}
+
+# Make the plot
+multi_bar <- writing_perceptions_long |>
+  ggplot(aes(x = Response, fill = Statements)) +
+  geom_bar(position = "dodge", stat = "count") +
+  labs(title = "Student Self-Perceptions of Coding",
+       x = "Response",
+       y = "Frequency") +
+  theme_minimal() +
+  theme(plot.title.position = "plot",
+        plot.title = element_text(hjust = 0.5),
+        legend.position= "bottom", 
+      )
+
+multi_bar <- addSmallLegend(multi_bar, pointSize = 0.4, textSize = 8.5, spaceLegend = 0.8)
+
+multi_bar
 

### 2024-07-24 scripts/02-data_cleaning.R

@@ -54,6 +54,8 @@ raw_data <- raw_data |>
     year != "TA"
     )
 
+
+
 # DROP UNECESSARY OR UNSPECIFIED COLUMNS
 raw_data <- raw_data |>
   select(-c(email,
@@ -62,10 +64,14 @@ raw_data <- raw_data |>
             consent, 
             student_name, 
             student_id, 
-            major, #none submitted
-            minor, #none submitted
             ))
 
+# CALCULATE TIME TO COMPLETE SURVEY
+
+raw_data <- raw_data |>
+  mutate(
+    mins_complete = round(as.numeric(difftime(as.POSIXct(completion_time), as.POSIXct(start_time), units = "mins")), 2)
+  )
 
 # CLEAN YEARS
 

### 2024-07-25 paper/paper.qmd
Finished writing perceptions and usage type visualization

@@
 
+#Maybe a table for the years instead?
+
 ```{r fig-year-bar}
@@
 #| warning: false
-#| tbl-cap: Student Self-Perception of Writing Ease
+#| tbl-cap: Student Distribution by Program of Study
 #| fig-align: center
@@
   coord_polar("y") +
-  labs(title = "Distribution of Unique Programs",
+  labs(title = "Student Distribution by Program of Study",
        x = NULL,
@@
         plot.title = element_text(hjust = 1.35)
-        )
+        ) +
+  geom_text(aes(label = n), 
+            position = position_stack(vjust = 0.5))
 
@@
 
-# Create a multi-bar plot for writing perceptions grouped by response
+# Create a multi-bar plot for coding self-perceptions grouped by response
 
@@
 multi_bar
+
+```
+
+```{r tbl-ai-familiarity}
+#| message: false
+#| echo: false
+#| warning: false
+#| tbl-cap: Familiarity with AI of students
+#| fig-align: center
+
+table_data <- survey_data |>
+    select(ai_familiarity)|>
+    mutate(
+      ai_familiarity = case_when(
+        ai_familiarity == "Limited to some liberal arts, such as history courses" ~ "Somewhat familiar",
+        TRUE ~ ai_familiarity
+      )
+    ) |>
+    count(ai_familiarity) |>
+    pivot_wider(
+      names_from = "ai_familiarity",
+      values_from = "n"
+    )
+
+kable(table_data) |>
+  kable_styling(bootstrap_options = "striped", full_width = F) |>
+  column_spec(1, bold = T)
+
+```
+
+```{r fig-ai_use-multibar}
+#| message: false
+#| echo: false
+#| warning: false
+#| tbl-cap: AI Use
+#| fig-align: center
+
+# Create a multi-bar plot for AI use/perceptions grouped by response
+
+#Select first three statements
+writing_perceptions <- survey_data |>
+  select(
+    used_ai_any_reason, 
+    used_ai_in_sta302
+  )
+
+writing_perceptions_long <- writing_perceptions |>
+  pivot_longer(cols = everything(), names_to = "Statements", values_to = "Response") |>
+  mutate(
+    Statements = case_when(
+      Statements == "used_ai_any_reason" ~ "Used AI for any reason",
+      Statements == "used_ai_in_sta302" ~ "Used AI in Class",
+      TRUE ~ Statements
+    ), 
+  ) |>
+  filter(!is.na(Response)) |>
+  filter(Response == "No" | Response == "Yes")
+
+#Taken from: https://stackoverflow.com/questions/52297978/decrease-overal-legend-size-elements-and-text
+addSmallLegend <- function(myPlot, pointSize = 0.5, textSize = 3, spaceLegend = 0.1) {
+    myPlot +
+        guides(shape = guide_legend(override.aes = list(size = pointSize)),
+               color = guide_legend(override.aes = list(size = pointSize))) +
+        theme(legend.title = element_text(size = textSize), 
+              legend.text  = element_text(size = textSize),
+              legend.key.size = unit(spaceLegend, "lines"))
+}
+
+# Make the plot
+multi_bar <- writing_perceptions_long |>
+  ggplot(aes(x = Response, fill = Statements)) +
+  geom_bar(position = "dodge", stat = "count") +
+  labs(title = "Student Self-Perceptions of Coding",
+       x = "Response",
+       y = "Frequency") +
+  theme_minimal() +
+  theme(plot.title.position = "plot",
+        plot.title = element_text(hjust = 0.5),
+      )
+
+multi_bar <- addSmallLegend(multi_bar, pointSize = 0.5, textSize = 10, spaceLegend = 0.8)
+
+multi_bar
+
+```
+
+```{r fig-ai-appropriate-pie}
+#| message: false
+#| echo: false
+#| warning: false
+#| tbl-cap: Comparison of Whether AI Use is Appropriate in School
+#| fig-align: center
+
+#Pie chart of program of study
+
+appropriate <- survey_data |> 
+  select(ai_schoolwork_appropriate) |>
+  mutate(
+    ai_schoolwork_appropriate = case_when(
+      ai_schoolwork_appropriate == "Appropriate" ~ "Appropriate",
+      ai_schoolwork_appropriate == "Inappropriate" ~ "Inappropriate",
+      TRUE ~ "It depends"
+    )
+  ) |>
+  count(ai_schoolwork_appropriate)
+
+appropriate
+
+ggplot(appropriate, aes(x = "", y = n, fill = ai_schoolwork_appropriate)) +
+  geom_bar(stat = "identity", width = 1) +
+  coord_polar(theta = "y") +
+  labs(title = "Do students think AI use is appropriate in class?",
+       x = NULL,
+       y = NULL) +
+  theme_minimal() +
+  scale_fill_discrete(name="Response") +
+  theme(plot.title.position = "plot",
+        plot.title = element_text(hjust = 0.5),
+        axis.text.x = element_blank(),  # Remove x-axis text
+        axis.title.x = element_blank(),
+      ) +
+  geom_text(aes(label = n), 
+            position = position_stack(vjust = 0.5))
+
+```
+
+```{r fig-AI-usage-type-comp}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: (CHANGE) Boilerplate figure caption 
+#| fig-align: center
+
+survey_data |>
+    select(
+      not_used,
+      used_technical_q,
+      used_conversation,
+      used_general_knowledge,
+      used_solving_hmk,
+      used_check_soln,
+      used_quick_q,
+      used_expl_concepts,
+      used_writing_content,
+      used_writing_code
+    )|>
+    pivot_longer(cols = everything(), names_to = "Statements", values_to = "Response") |>
+    filter(Response == 1) |>
+    count(Statements) |>
+    mutate(Statements = case_when(
+      Statements == "not_used" ~ "Not Used",
+      Statements == "used_technical_q" ~ "Technical Questions",
+      Statements == "used_conversation" ~ "Fun Conversation",
+      Statements == "used_general_knowledge" ~ "General Knowledge",
+      Statements == "used_solving_hmk" ~ "Entirely Solving Homework",
+      Statements == "used_check_soln" ~ "Checking my Solutions",
+      Statements == "used_quick_q" ~ "Quick Questions",
+      Statements == "used_expl_concepts" ~ "Explaining Concepts",
+      Statements == "used_writing_content" ~ "Writing Paper Content",
+      Statements == "used_writing_code" ~ "Writing Paper Code",
+      TRUE ~ Statements
+    ))|>
+    ggplot(aes(x = Statements, y = reorder(Statements, n))) +
+    geom_bar(fill="grey50") +
+    labs(title="How Students Used ChatGPT", x="Count", y="Use Type") +
+    theme(axis.text.x = element_text(angle = 30, hjust = 0.5, vjust = 0.5),
+          panel.background = element_blank(),
+          axis.line.y=element_line(color="black"),
+          axis.ticks.y=element_blank(),
+          panel.grid.major.x=element_line(color="grey80"),
+          panel.grid.minor=element_blank(),
+          plot.title.position = "plot",
+          plot.title = element_text(hjust = 0.5))
+
+bbb
 

### 2024-07-29 paper/paper.qmd
improved introduction and literature review.

@@
 date-format: long
-abstract: "Blah Blah Blah Blah"
+abstract: "An abstract should eventually go here!!"
 format: pdf
@@
 
-```{r setup}
+```{r setup, include=FALSE}
+
 ## Workspace Setup ##
-#| include: false
-#| warning: false
-#| message: false
 
@@
 # Import data
-survey_data <- read_csv("../data/analysis_data/clean_STA302_postcourse_survey_w24.csv")
+survey_data <- read_csv("../data/analysis_data/clean_STA302_postcourse_survey_w24.csv", show_col_types = FALSE)
 ```
@@
 
-The rapid advancement of Large Language Models (LLMs) like ChatGPT has had massive effects in all areas of society, with notable increases in productivity within some knowledge-based industries.
-Recent papers have demonstrated evidence of some productivity increases, in industries ranging from consulting to computer programming.
+The wide release of ChatGPT and other Large Language Models (LLMs) has rapidly transformed various parts of society, in particular the field of education.
+As these tools are increasingly accessible, their adoption and use in academic settings has created excitement and apprehension among educators and students alike.
+LLMs offer unprecedented support for students in tasks ranging from writing assistance to support in complex problem-solving, with the potential to enhance academic performance and student outcomes.
+At the same time, their use raises questions about academic integrity, the development of critical thinking skills (particular in undergraduates), as well as questions about their implications for effective learning overall.
 
-On the lower end of productivity increases, @Ben_2024 conducted a randomized control trial with judges in Wisconsin to assess the impact of AI recommendations on bail hearing decisions.
-The study found that providing AI recommendations did not significantly improve judges' decision-making accuracy.
-On the other hand, @Peng_2023 investigated the significantly positive impact of GitHub Copilot, an LLM-powered coding assistant, on programmer productivity.
-They conducted an experiment with 95 professional programmers recruited through Upwork.
-The study found that programmers with access to Copilot completed a standardized programming task 55.8% faster than the control group.
-In particular, less experienced programmers appeared to benefit more from the tool.
+The potential for LLMs to positively affect students' academic performance in data science can be clearly inferred from already-demonstrated effects on adjacent professional fields.
+The tasks involved in a data science workflow can be generally broken down into two key competencies.
+The first is programming, which is done when cleaning, analyzing, visualizing data using languages like R or Python.
+The second is writing, which is primarily done when communicating results.
+There is experimental evidence from adjacent professional fields that LLMs can improve productivity in both of these competencies.
 
-Examining the middle-ground, a study by @DellAcqua_2023 examined the impact of AI on management consulting tasks through a field experiment with 758 Boston Consulting Group consultants.
-They found that AI significantly increased performance across various business tasks, with speed increasing by over 25% and quality by more than 40%.
-However, the study revealed a "jagged technological frontier" where AI excelled in some tasks but struggled with others.
-Notably, AI usage appeared to level performance differences across ability levels, with lower-performing consultants benefiting the most, similar to that of Peng et al.
+@Peng_2023 shows the potential impact of LLMs on students' computer programming skills, through their uncovering of the significantly positive impacts of GitHub Copilot (an LLM-powered programming assistant) on the productivity of computer programmers.
+In an experiment involving 95 freelance programmers, they found that that programmers with access to Copilot completed a standardized programming task 55.8% faster than the control group.
+Crucially, it was found that programmers with less experience saw the greatest improvements in productivity.
+Given that programming is a core component of data science practice, this suggests that these benefits would carry over to students in data science courses as well.
 
-We can see in the current literature that LLMs have had varying levels of impact across different knowledge-intensive industries, with more computer-intensive tasks like programming receiving the highest performance boosts.
-One area sorely lacking quantitative evidence of performance increases is within higher education, an area where scrutiny of the impact of LLMs and other AI tools has been common.
-A survey of undergraduate political science students was conducted by @Cahill_2024, revealing widespread use of AI tools like ChatGPT.
-However, their study also uncovered a significant gap in students' self-perceived competence in using these tools effectively for academic work.
+@DellAcqua_2023, while primarily about management consulting tasks, provides evidence that LLMs can significantly improve writing productivity.
+In a field experiment involving Consultants recruited from the Boston Consulting Group (BCG), they found that the use of GPT-4 in the experimental group led to a 25% increase in delivery speed of business tasks (most involving some writing), as well as a 40% increase in human-rated performance on those tasks.
+Similar to computer programming, these productivity increases were most pronounced for those with below average performance, with their output increasing by 43%.
+The authors expressed particular optimism for LLMs' potential to generally expedite  knowledge-work-related tasks, such as generating new ideas and persuasive and informative writing, which has direct application to data science education through the writing of more engaging reports and documentation.
+That said, not all AI tasks yielded the same quality of results.
 
-Student challenges while learning Data Science in particular were explored by @kross2019.
-By polling instructor-practitioners of Data Science in academia and industry, they found several key challenges that data science students face.
-The largest challenges they faced were diverse computing backgrounds, integrating various technical workflows, software setup, finding relevant datasets, as well as dealing with the inherent uncertainty of the data analysis process.
+The implications of this evidence appear promising.
+However, literature in domains both adjacent to and distant from statistics and data science suggest much more varied opinions on how LLMs can impact student performance, much less specifically in data science and statistics coursework.
 
-Given the rapid integration of AI tools in educational and professional settings, there is a need to understand how these technologies and students' attitudes toward LLMs affect their  learning and performance.
-This paper aims to investigate the relationship between academic performance, LLM usage, and students' perceptions of LLMs, using evidence from an exit survey of a third-year undergraduate statistics course at the University of Toronto.
-By examining how students interact with and perceive LLM tools and how this usage correlates with their grades, we seek to understand the potential benefits and pitfalls of AI integration in data science education.
+Among other consequences of LLM use, @Valenzuela_2024 argue that LLMs lead to a loss of serendipity (which leads to less original work) and de-skilling (primarily with respect to programming ability).
+These particular outcomes could negatively affect students' effective learning of data science.
+On the other hand, @Ellis_Slade_2023 take a more optimistic perspective, taking the popular stance of comparing ChatGPT to previously controversial learning technologies like calculators.
+They argue that LLMs are just another technology that will impact Statistics and Data Science education like the calculator did.
+However, they take a broader perspective on the issue and more specific inquiry could be done in how effective LLMs are at improving student outcomes within data science education in particular.
 
-The remainder of this paper is structured as follows: @sec-data analyzes our survey data; @sec-model models our unstructured data to find fun and cool things; @sec-results lists the results; @sec-discussion discusses the implications of these findings for data science education and for future research and practice in this rapidly evolving field.
+Speaking generally about data science education, @Tu_2024 acknowledge that for students, LLMs can streamline many parts of a data science workflow - with that in mind, they suggest that budding data scientists should shift their self-perspectives from being an analyst to being a manager responsible for strategic oversight.
+They emphasize that in both education and practice, LLMs and human intelligence should play complementary roles.
+
+The simultaneous excitement and caution the literature expressed toward using LLMs in educational settings was further corroborated by empirical studies on K-12 and university students as well.
+
+At the high school level, @Lazar_2023 conducted a informal survey of secondary school teachers and students on their opinions of ChatGPT, they found that while LLMs could help spark creativity, provide academic support when teachers were unavailable, and model certain types of writing well, teachers were also cognizant of LLMs potential to limit students' learning in certain ways through overreliance.
+Beyond academic integrity concerns, teachers had similar concerns about de-skilling and an overall loss of agency in writing and critical thinking.
+
+@Cahill_2024 surveyed undergraduate Political Science students on their attitudes and self-perceived usage of AI tools (which included LLMs like ChatGPT).
+They found that the use of ChatGPT (among other machine-learning-powered software) was widespread.
+However, they also found that many students lacked the confidence in using AI for academic purposes - in particular, only 11% 'strongly agreed' that they know how to use AI to improve their writing.
+Like educators,  students have nuanced views on appropriate AI use.
+In particular, respondents found that using it to writing whole papers as inappropriate, while using it for basic tasks like general assistance, writing feedback and basic data visualization was perceived more appropriately.
+Interestingly, first-generation college students were found to be more likely to use AI in their work, particularly in writing papers and helping with assignments.
+Their findings suggest that LLMs and other AI tools could be an equalizer for disadvantaged students.
+Moreover, though only political science students were surveyed, statistics and data science students would reasonably have  general attitudes that are similar.
+
+As we have seen, the existing inquiry by educators has explored the general qualitative usage and student perceptions on these tools.
+Though informative, there is still a key lack of evidence of how students' academic performance can be affected through allowing them to use LLMs in classwork.
+This paper aims to fill this gap by quantitatively investigating the relationship between students' grades and measures of student LLM usage and their attitudes toward LLMs in general, using evidence from a third-year undergraduate statistics course at the University of Toronto.
+Through examining how students interact with and perceive LLM tools and how these variables translate into student outcomes, the effects of LLM integration in data science education can be more precisely determined.
+
+The remainder of this paper is structured as follows: @sec-data visualizes and analyzes survey data collected from students at the end of the course; @sec-model models our unstructured data to approximate a relationship between grades and usage and attitudes; @sec-results lists the results; @sec-discussion discusses the implications of these findings for data science education and future research and practice in this rapidly evolving field.
 
 # Data {#sec-data}
+
+## Background, etc.
+
+Put the context of the data here (look at the)
 
@@
     theme(plot.title.position = "plot",
-          plot.title = element_text(hjust = 0.5))
+          plot.title = element_text(hjust = 0.6))
 
@@
 
-#Maybe a table for the years instead?
+Maybe a table for the years instead?
+
+Get inspired by Cahill and McCabe's paper for visualizations etc.
 
@@
 
+```{r fig-ai-use-multibar}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: Hello
+#| fig-align: center
+
+# Create a multi-bar plot for AI use/perceptions grouped by response
+
+# THROWS ERROR: FIX
+
+#Select first three statements
+writing_perceptions <- survey_data |>
+  select(
+    used_ai_any_reason, 
+    used_ai_in_sta302
+  )
+
+writing_perceptions_long <- writing_perceptions |>
+  pivot_longer(cols = everything(), names_to = "Statements", values_to = "Response") |>
+  mutate(
+    Statements = case_when(
+      Statements == "used_ai_any_reason" ~ "Used AI for any reason",
+      Statements == "used_ai_in_sta302" ~ "Used AI in Class",
+      TRUE ~ Statements
+    ), 
+  ) |>
+  filter(!is.na(Response)) |>
+  filter(Response == "No" | Response == "Yes")
+
+#Taken from: https://stackoverflow.com/questions/52297978/decrease-overal-legend-size-elements-and-text
+addSmallLegend <- function(myPlot, pointSize = 0.5, textSize = 3, spaceLegend = 0.1) {
+    myPlot +
+        guides(shape = guide_legend(override.aes = list(size = pointSize)),
+               color = guide_legend(override.aes = list(size = pointSize))) +
+        theme(legend.title = element_text(size = textSize), 
+              legend.text  = element_text(size = textSize),
+              legend.key.size = unit(spaceLegend, "lines"))
+}
+
+# Make the plot
+multi_bar <- writing_perceptions_long |>
+  ggplot(aes(x = Response, fill = Statements)) +
+  geom_bar(position = "dodge", stat = "count") +
+  labs(title = "Student Self-Perceptions of Coding",
+       x = "Response",
+       y = "Frequency") +
+  theme_minimal() +
+  theme(plot.title.position = "plot",
+        plot.title = element_text(hjust = 0.5),
+      )
+
+multi_bar <- addSmallLegend(multi_bar, pointSize = 0.5, textSize = 10, spaceLegend = 0.8)
+
+multi_bar
+
+```
+
 ```{r fig-coding-multibar-1}
@@
 
-```{r fig-ai_use-multibar}
-#| message: false
-#| echo: false
-#| warning: false
-#| tbl-cap: AI Use
-#| fig-align: center
-
-# Create a multi-bar plot for AI use/perceptions grouped by response
-
-#Select first three statements
-writing_perceptions <- survey_data |>
-  select(
-    used_ai_any_reason, 
-    used_ai_in_sta302
-  )
-
-writing_perceptions_long <- writing_perceptions |>
-  pivot_longer(cols = everything(), names_to = "Statements", values_to = "Response") |>
-  mutate(
-    Statements = case_when(
-      Statements == "used_ai_any_reason" ~ "Used AI for any reason",
-      Statements == "used_ai_in_sta302" ~ "Used AI in Class",
-      TRUE ~ Statements
-    ), 
-  ) |>
-  filter(!is.na(Response)) |>
-  filter(Response == "No" | Response == "Yes")
-
-#Taken from: https://stackoverflow.com/questions/52297978/decrease-overal-legend-size-elements-and-text
-addSmallLegend <- function(myPlot, pointSize = 0.5, textSize = 3, spaceLegend = 0.1) {
-    myPlot +
-        guides(shape = guide_legend(override.aes = list(size = pointSize)),
-               color = guide_legend(override.aes = list(size = pointSize))) +
-        theme(legend.title = element_text(size = textSize), 
-              legend.text  = element_text(size = textSize),
-              legend.key.size = unit(spaceLegend, "lines"))
-}
-
-# Make the plot
-multi_bar <- writing_perceptions_long |>
-  ggplot(aes(x = Response, fill = Statements)) +
-  geom_bar(position = "dodge", stat = "count") +
-  labs(title = "Student Self-Perceptions of Coding",
-       x = "Response",
-       y = "Frequency") +
-  theme_minimal() +
-  theme(plot.title.position = "plot",
-        plot.title = element_text(hjust = 0.5),
-      )
-
-multi_bar <- addSmallLegend(multi_bar, pointSize = 0.5, textSize = 10, spaceLegend = 0.8)
-
-multi_bar
-
-```
-
 ```{r fig-ai-appropriate-pie}
@@
   count(ai_schoolwork_appropriate)
-
-appropriate
 
@@
 #| warning: false
-#| fig-cap: (CHANGE) Boilerplate figure caption 
+#| fig-cap: How Students Use ChatGPT
 #| fig-align: center
 
-survey_data |>
+# Create a bar plot of how students used ChatGPT - FIX: sort from largest to smallest
+
+type_data <- survey_data |>
               select(
@@
                 TRUE ~ Statements
-    ))|>
-    ggplot(aes(x = Statements, y = reorder(Statements, n))) +
-    geom_bar(fill="grey50") +
+              )) |>
+              arrange(desc(n))
+
+type_data |>
+    ggplot(aes(x = n, y = Statements)) +
+    geom_bar(stat="identity", fill="grey50") +
     labs(title="How Students Used ChatGPT", x="Count", y="Use Type") +
@@
           plot.title = element_text(hjust = 0.5))
-
-bbb
 

### 2024-07-30 scripts/02-data_cleaning.R
fixed cleaning code

@@ -9,6 +9,7 @@
 library(tidyverse)
 library(janitor)
 library(stringr)
+library(lintr)
 
 
 #### Clean data ####
@@ -18,53 +19,34 @@ raw_data <- as_tibble(raw_data)
 
 # CLEAN UP COLUMN NAMES
 
-raw_data <- raw_data |> 
+raw_data <- raw_data |>
   clean_names()
 
-raw_data <- raw_data |> 
+raw_data <- raw_data |>
   rename(
     consent = after_carefully_reading_the_informed_consent_document_please_indicate_below_whether_you_consent_to_have_your_anonymized_responses_included_in_the_research_study,
     student_name = what_is_your_full_name_on_quercus,
     student_id = what_is_your_student_id,
     year = what_year_are_you,
-    program_of_study = what_is_your_specialization, 
+    program_of_study = what_is_your_specialization,
     major = what_is_are_your_major_s,
     minor = what_is_are_your_minor_s,
-    gpa = what_is_your_gpa, 
-    ai_familiarity = how_familiar_are_you_with_using_generative_ai_tools_such_as_open_a_is_chat_gpt_or_equivalents, 
-    used_ai_any_reason = have_you_used_any_generative_ai_tools_such_as_open_a_is_chat_gpt_or_equivalents_for_any_reason_personal_or_educational, 
-    ai_use_cases = if_you_have_used_generative_ai_tools_such_as_open_a_is_chat_gpt_or_equivalents_in_what_ways_have_you_used_it_select_all_that_apply, 
-    ai_schoolwork_appropriate = to_what_extent_do_you_think_using_generative_ai_tools_such_as_chat_gpt_by_open_ai_or_equivalents_is_ethical_and_appropriate_for_schoolwork, 
+    gpa = what_is_your_gpa,
+    ai_familiarity = how_familiar_are_you_with_using_generative_ai_tools_such_as_open_a_is_chat_gpt_or_equivalents,
+    used_ai_any_reason = have_you_used_any_generative_ai_tools_such_as_open_a_is_chat_gpt_or_equivalents_for_any_reason_personal_or_educational,
+    ai_use_cases = if_you_have_used_generative_ai_tools_such_as_open_a_is_chat_gpt_or_equivalents_in_what_ways_have_you_used_it_select_all_that_apply,
+    ai_schoolwork_appropriate = to_what_extent_do_you_think_using_generative_ai_tools_such_as_chat_gpt_by_open_ai_or_equivalents_is_ethical_and_appropriate_for_schoolwork,
[…]

### 2024-07-30 paper/paper.qmd
Reapply "wrote more paper"
This reverts commit a40be7522ec315e46785a1949a32fac620411096.

@@
 
-@Peng_2023 shows the potential impact of LLMs on students' computer programming skills, through their uncovering of the significantly positive impacts of GitHub Copilot (an LLM-powered programming assistant) on the productivity of computer programmers.
+@Peng_2023 show the potential impact of LLMs on students' computer programming skills, through their uncovering of the significantly positive impacts of GitHub Copilot (an LLM-powered programming assistant) on the productivity of computer programmers.
 In an experiment involving 95 freelance programmers, they found that that programmers with access to Copilot completed a standardized programming task 55.8% faster than the control group.
 Crucially, it was found that programmers with less experience saw the greatest improvements in productivity.
-Given that programming is a core component of data science practice, this suggests that these benefits would carry over to students in data science courses as well.
+Given that programming is a core component of data science practice, this suggests that these benefits could reasonably carry over to students in data science courses as well.
 
@@
 Similar to computer programming, these productivity increases were most pronounced for those with below average performance, with their output increasing by 43%.
-The authors expressed particular optimism for LLMs' potential to generally expedite  knowledge-work-related tasks, such as generating new ideas and persuasive and informative writing, which has direct application to data science education through the writing of more engaging reports and documentation.
-That said, not all AI tasks yielded the same quality of results.
+Though not all tasks done using AI saw the same productivity improvements, the authors expressed particular optimism for LLMs' potential to generally expedite menial knowledge-work tasks.
+These included tasks such as generating new ideas and creating persuasive and informative writing - both examples directly apply to data science education through their analogues of coming up with project ideas and writing more engaging reports.
+As a whole, they show that LLMs could equally augment the quality of student writing and course projects in data science as well.
 
-The implications of this evidence appear promising.
-However, literature in domains both adjacent to and distant from statistics and data science suggest much more varied opinions on how LLMs can impact student performance, much less specifically in data science and statistics coursework.
+Clearly, the implications of these case studies appear promising.
+However, literature from educators both adjacent to and distant from statistics and data science suggest much more varied perspectives on how LLMs can impact student performance.
 
-Among other consequences of LLM use, @Valenzuela_2024 argue that LLMs lead to a loss of serendipity (which leads to less original work) and de-skilling (primarily with respect to programming ability).
+@Valenzuela_2024 argue that LLMs lead to a loss of serendipity (which leads to less original work) and de-skilling (primarily with respect to programming ability), among other consequences.
 These particular outcomes could negatively affect students' effective learning of data science.
@@
 
-Speaking generally about data science education, @Tu_2024 acknowledge that for students, LLMs can streamline many parts of a data science workflow - with that in mind, they suggest that budding data scientists should shift their self-perspectives from being an analyst to being a manager responsible for strategic oversight.
-They emphasize that in both education and practice, LLMs and human intelligence should play complementary roles.
+In a similar attitude, @Tu_2024 acknowledge that for students, LLMs can streamline many parts of a data science workflow - with that in mind, they suggest that data scientists in training should shift their self-perspectives from primarily being an analyst to primarily being a manager responsible for strategic oversight of the analysis.
+Crucially they emphasize that in both education and practice, LLMs and human intelligence should play complementary roles.
 
-The simultaneous excitement and caution the literature expressed toward using LLMs in educational settings was further corroborated by empirical studies on K-12 and university students as well.
+The simultaneous caution and interest the literature expressed toward using LLMs in educational settings was further corroborated by empirical studies on K-12 and university students as well.
 
 At the high school level, @Lazar_2023 conducted a informal survey of secondary school teachers and students on their opinions of ChatGPT, they found that while LLMs could help spark creativity, provide academic support when teachers were unavailable, and model certain types of writing well, teachers were also cognizant of LLMs potential to limit students' learning in certain ways through overreliance.
-Beyond academic integrity concerns, teachers had similar concerns about de-skilling and an overall loss of agency in writing and critical thinking.
+Beyond academic integrity concerns, teachers had similar concerns to @Valenzuela_2024 about de-skilling and an overall loss of agency in writing and critical thinking.
 
-@Cahill_2024 surveyed undergraduate Political Science students on their attitudes and self-perceived usage of AI tools (which included LLMs like ChatGPT).
+@Cahill_2024 surveyed undergraduate Political Science students on their attitudes toward and usage of AI tools.
 They found that the use of ChatGPT (among other machine-learning-powered software) was widespread.
@@
 Their findings suggest that LLMs and other AI tools could be an equalizer for disadvantaged students.
-Moreover, though only political science students were surveyed, statistics and data science students would reasonably have  general attitudes that are similar.
+Rhough only political science students were surveyed, statistics and data science students could reasonably have general attitudes that are similar.
 
@@
 
-The remainder of this paper is structured as follows: @sec-data visualizes and analyzes survey data collected from students at the end of the course; @sec-model models our unstructured data to approximate a relationship between grades and usage and attitudes; @sec-results lists the results; @sec-discussion discusses the implications of these findings for data science education and future research and practice in this rapidly evolving field.
+The remainder of this paper is structured as follows: @sec-data visualizes and analyzes survey data and coursework from students; @sec-model models the unstructured data to approximate a relationship between grades and usage/attitudes; @sec-results describes and analyzes the model's results; @sec-discussion discusses the implications of the findings for data science education and future research and practice in this rapidly evolving field.
 
@@
 
-## Background, etc.
+## Background and Collection
 
-Put the context of the data here (look at the)
+To investigate students' usage and attitudes toward LLM use and how they related to their academic performance, a dataset containing their usage/attitudes, coursework, and academic performance was constructed.
+Data was sampled from the cohort of students taking STA302 - Methods of Data Analysis I, taught by one of the investigators in the Winter 2024 semester at the University of Toronto.
+By virtue of pre-requisites needed, this restricted the data collected to be only on upper-year undergraduate students.
+
+Data was collected from students through an optional end-of-course survey.
+Whether or not they consented to their data being used in this investigation, all respondents received a +1% increase in their final course grade for their participation.
+All consenting responses were then cross-referenced to their course grade, as well as the GitHub account they used to complete their course research papers.
+The responses were finally anonymized by removing any personal references to the students themselves.
+
+## Cleaning and Variables of Interest
+
+All the data was cleaned using R [@citeR] and it's tidyverse [@tidyverse], janitor [@janitor] and stringr [@stringr], then tested for issues using the testthat package [@testthat].
+This led to a final dataset containing 121 responses across (var_num) variables.
+
+These variables were:
+
+- [list of variables]
+
+## Visualization
+The data was then visualized using the ggplot2 [@ggplot2] package.
 
@@
 
-#Pie chart of program of study
+#Pie chart of student opinion on appropriateness of AI use in class.
 
@@
   select(ai_schoolwork_appropriate) |>
-  mutate(
-    ai_schoolwork_appropriate = case_when(
-      ai_schoolwork_appropriate == "Appropriate" ~ "Appropriate",
-      ai_schoolwork_appropriate == "Inappropriate" ~ "Inappropriate",
-      TRUE ~ "It depends"
-    )
-  ) |>
   count(ai_schoolwork_appropriate)
@@
               filter(Response == 1) |>
-              count(Statements) |>
               mutate(Statements = case_when(
@@
               )) |>
+              count(Statements) |>
               arrange(desc(n))
@@
 ```
+
+## Sentiment Analysis
 

### 2024-08-01 paper/paper.qmd
Wrote more paper

@@
 ## Visualization
-The data was then visualized using the ggplot2 [@ggplot2] package.
 
-```{r fig-boilerplate}
+The student survey data was then visualized and analyzed using the ggplot2 [@ggplot2] package.
+
+### General Information
+
+First, some general information about our sample of students was derived through visualizing the basic data about them.
+
+```{r fig-response-time-hist}
 #| message: false
@@
 #| warning: false
-#| fig-cap: (CHANGE) Boilerplate figure caption 
+#| fig-cap: Distribution of Survey Response Times
 #| fig-align: center
 
-# survey_data |>
-#   ggplot(program_counts, aes(x = "", y = n, fill = program_of_study)) +
-#   #put geom type here (put before compilation) +
-#   labs(title = "Title!",
-#        x = "X label",
-#        y = "Y label") +
-#   theme_minimal() +
-#   scale_fill_discrete(name="Legend Title") +
-#   theme(plot.title.position = "plot",
-#         plot.title = element_text(hjust = 0.5)
-#         )
+#Trim values to histogram
+
+plot_data <- survey_data |>
+    filter(mins_complete < 1000)
+
+plot_data |>
+    ggplot(aes(x = mins_complete)) +
+    geom_histogram(binwidth = 2, fill = "skyblue", color = "black") +
+    labs(title = "Distribution of Survey Response Times",
+         x = "Time (Mins)",
+         y = "Frequency") +
+    theme_minimal() +
+    theme(plot.title.position = "plot",
+          plot.title = element_text(hjust = 0.5))
 
 ```
-
-```{r tbl-boilerplate}
-#| message: false
-#| echo: false
-#| warning: false
-#| tbl-cap: (CHANGE) Boilerplate table caption 
-#| fig-align: center
-
-#Put code below this line
-
-```
+@fig-response-time-hist shows the distribution of response times to the end-of-course survey.
+Respondents who consented generally took between 0-40 mins to answer all of the questions, with most response times ranging between 2-10 minutes.
+There was one consenting respondent who took 4308 minutes, whose observation was dropped from the visualization.
+This suggests that students generally took a meaningful amount of time to engage with the survey.
 
@@
 
-Maybe a table for the years instead?
+@fig-gpa-hist shows a wide distribution of student GPAs, with the majority clustering around a B (3.0/4.0) average.
+One factor possibly affecting the range is the course being a required course for programs in the Statistics, Mathematics and Computer Science departments.
+Though programs in these departments generally attract high-achieving students, some may still struggle.
+Another factor is that these GPAs were self-reported, which introduces some ambiguity - students may have provided either their cumulative GPA or their most recent term's GPA, or could have misreported it entirely.
+Despite these two factors, the diversity in reported grades suggests the course attracted students across a broad spectrum of academic performance levels.
 
@@
 
-```{r fig-year-bar}
+```{r tbl-year-dist}
 #| message: false
@@
 
-#Bar plot of year
-survey_data |> ggplot(aes(x = year)) +
-    geom_bar(fill = "skyblue", color = "black") +
-    labs(title = "Distribution of Students by Year",
-         x = "Year",
-         y = "Frequency") +
-    theme_minimal() +
-    theme(plot.title.position = "plot",
-          plot.title = element_text(hjust = 0.5))
+year_count <- survey_data |>
+  count(year) |>
+  pivot_wider(names_from = year, values_from = n, values_fill = list(n = 0))
+
+# Display the original wide format table
+kable(year_count)
 
 ```
+
+We can also see in from @tbl-year-dist that students from a range of years took the course.
+That said, the majority of students were in their 3rd or 4th year of study.
+This makes sense, given the course's prerequisite of general statistics, which is a two-semester sequence typically done in second year statistics.
+This also explains the lack of any first year students in the course as well.
 
@@
         axis.title.x = element_blank(), # Remove x-axis title
-        plot.title = element_text(hjust = 1.35)
+        plot.title = element_text(hjust = 1.6)
         ) +
@@
 
-```{r fig-response-time-hist}
+Programs of study [did something], as we can see in @fig-programs-pie...
+
+### Student Attitudes
+
+```{r fig-ai-familiarity}
 #| message: false
@@
 #| warning: false
-#| fig-cap: Distribution of Survey Response Times
+#| fig-cap: Student Familiarity with AI
 #| fig-align: center
 
-#Trim values to histogram
+table_data <- survey_data |>
+    select(ai_familiarity)|>
+    mutate(
+      ai_familiarity = case_when(
+        ai_familiarity == "Limited to some liberal arts, such as history courses" ~ "Somewhat familiar",
+        TRUE ~ ai_familiarity
+      )
+    ) |>
+    count(ai_familiarity)
 
-plot_data <- survey_data |>
-    filter(mins_complete < 1000)
+table_data |>
+  ggplot(aes(x = "", y = n, fill = ai_familiarity)) +
+  geom_bar(stat = "identity", width = 1) +
+  coord_polar("y") +
+  labs(title = "Student Familiarity with AI",
+       x = NULL,
+       y = NULL) +
+  theme_minimal() +
+  scale_fill_discrete(name="Program") +
+  theme(axis.text.x = element_blank(),  # Remove x-axis text
+        axis.title.x = element_blank(), # Remove x-axis title
+        plot.title = element_text(hjust = 1.35)
+        ) +
+  geom_text(aes(label = n), 
+            position = position_stack(vjust = 0.5)
+            )
+```
 
-plot_data |>
-    ggplot(aes(x = mins_complete)) +
-    geom_histogram(binwidth = 0.1, fill = "skyblue", color = "black") +
-    labs(title = "Distribution of Survey Response Times",
-         x = "Time (Mins)",
-         y = "Frequency") +
-    theme_minimal() +
-    theme(plot.title.position = "plot",
-          plot.title = element_text(hjust = 0.5))
+Just as @Cahill_2024 found, the vast majority of students taking the course were at least "somewhat familiar" with ChatGPT and other LLMs.
+This overrepresentation of familiarity also primarily driven by the fact that the course is primarily for students in Statistics, Computer Science, Economics, Mathematics and other quantitative disciplines, which could likely lead to a skewing of interest/affinity for computers.
 
-```
+Students were then asked about their attitudes and self-perception on writing and coding.
 
@@
 
-```{r fig-ai-use-multibar}
-#| message: false
-#| echo: false
-#| warning: false
-#| fig-cap: Hello
-#| fig-align: center
-
-# Create a multi-bar plot for AI use/perceptions grouped by response
-
-# THROWS ERROR: FIX
-
-#Select first three statements
-writing_perceptions <- survey_data |>
-  select(
-    used_ai_any_reason, 
-    used_ai_in_sta302
-  )
-
-writing_perceptions_long <- writing_perceptions |>
-  pivot_longer(cols = everything(), names_to = "Statements", values_to = "Response") |>
-  mutate(
-    Statements = case_when(
-      Statements == "used_ai_any_reason" ~ "Used AI for any reason",
-      Statements == "used_ai_in_sta302" ~ "Used AI in Class",
-      TRUE ~ Statements
-    ), 
-  ) |>
-  filter(!is.na(Response)) |>
-  filter(Response == "No" | Response == "Yes")
-
-#Taken from: https://stackoverflow.com/questions/52297978/decrease-overal-legend-size-elements-and-text
-addSmallLegend <- function(myPlot, pointSize = 0.5, textSize = 3, spaceLegend = 0.1) {
-    myPlot +
-        guides(shape = guide_legend(override.aes = list(size = pointSize)),
-               color = guide_legend(override.aes = list(size = pointSize))) +
-        theme(legend.title = element_text(size = textSize), 
-              legend.text  = element_text(size = textSize),
-              legend.key.size = unit(spaceLegend, "lines"))
-}
-
-# Make the plot
-multi_bar <- writing_perceptions_long |>
-  ggplot(aes(x = Response, fill = Statements)) +
-  geom_bar(position = "dodge", stat = "count") +
-  labs(title = "Student Self-Perceptions of Coding",
-       x = "Response",
-       y = "Frequency") +
-  theme_minimal() +
-  theme(plot.title.position = "plot",
-        plot.title = element_text(hjust = 0.5),
-      )
-
-multi_bar <- addSmallLegend(multi_bar, pointSize = 0.5, textSize = 10, spaceLegend = 0.8)
-
-multi_bar
-
-```
+Students had varied attitudes toward their writing and their writing abilities in general, as we can see from @fig-writing-multibar-1 and fig-writing-multibar-2. ...
 
@@
 
-```{r tbl-ai-familiarity}
-#| message: false
-#| echo: false
-#| warning: false
-#| tbl-cap: Familiarity with AI of students
-#| fig-align: center
-
-table_data <- survey_data |>
-    select(ai_familiarity)|>
-    mutate(
-      ai_familiarity = case_when(
-        ai_familiarity == "Limited to some liberal arts, such as history courses" ~ "Somewhat familiar",
-        TRUE ~ ai_familiarity
-      )
-    ) |>
-    count(ai_familiarity) |>
-    pivot_wider(
-      names_from = "ai_familiarity",
-      values_from = "n"
-    )
-
-kable(table_data) |>
-  kable_styling(bootstrap_options = "striped", full_width = F) |>
-  column_spec(1, bold = T)
-
-```
+Student self-perceptions about their coding abilities, shown in @fig-coding-multibar-1 and @fig-coding-multibar-2 were also very interesting. ...
 
@@
 
+### Student Usage
+
+```{r fig-ai-use-multibar}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: Hello
+#| fig-align: center
+
+# Create a multi-bar plot for AI use/perceptions grouped by response
+
+# THROWS ERROR: FIX
+
+#Select first three statements
+writing_perceptions <- survey_data |>
+  select(
+    used_ai_any_reason, 
+    used_ai_in_sta302
+  )
+
+writing_perceptions_long <- writing_perceptions |>
+  pivot_longer(cols = everything(), names_to = "Statements", values_to = "Response") |>
+  mutate(
+    Statements = case_when(
+      Statements == "used_ai_any_reason" ~ "Used AI for any reason",
+      Statements == "used_ai_in_sta302" ~ "Used AI in Class",
+      TRUE ~ Statements
+    ), 
+  ) |>
+  filter(!is.na(Response)) |>
+  filter(Response == "No" | Response == "Yes")
+
+#Taken from: https://stackoverflow.com/questions/52297978/decrease-overal-legend-size-elements-and-text
+addSmallLegend <- function(myPlot, pointSize = 0.5, textSize = 3, spaceLegend = 0.1) {
+    myPlot +
+        guides(shape = guide_legend(override.aes = list(size = pointSize)),
+               color = guide_legend(override.aes = list(size = pointSize))) +
+        theme(legend.title = element_text(size = textSize), 
+              legend.text  = element_text(size = textSize),
+              legend.key.size = unit(spaceLegend, "lines"))
+}
+
+# Make the plot
+multi_bar <- writing_perceptions_long |>
+  ggplot(aes(x = Response, fill = Statements)) +
+  geom_bar(position = "dodge", stat = "count") +
+  labs(title = "Student Self-Perceptions of Coding",
+       x = "Response",
+       y = "Frequency") +
+  theme_minimal() +
+  theme(plot.title.position = "plot",
+        plot.title = element_text(hjust = 0.5),
+      )
+
+multi_bar <- addSmallLegend(multi_bar, pointSize = 0.5, textSize = 10, spaceLegend = 0.8)
+
+multi_bar
+
+```
+
 ```{r fig-AI-usage-type-comp}
@@
               pivot_longer(cols = everything(), names_to = "Statements", values_to = "Response") |>
-              filter(Response == 1) |>
               mutate(Statements = case_when(
@@
               )) |>
+              filter(Response == 1) |>
               count(Statements) |>
@@
 type_data |>
-    ggplot(aes(x = n, y = Statements)) +
+    ggplot(aes(x = n, y = reorder(Statements, n))) +
     geom_bar(stat="identity", fill="grey50") +
     labs(title="How Students Used ChatGPT", x="Count", y="Use Type") +
-    theme(axis.text.x = element_text(angle = 30, hjust = 0.5, vjust = 0.5),
-          panel.background = element_blank(),
-          axis.line.y=element_line(color="black"),
-          axis.ticks.y=element_blank(),
-          panel.grid.major.x=element_line(color="grey80"),
-          panel.grid.minor=element_blank(),
-          plot.title.position = "plot",
-          plot.title = element_text(hjust = 0.5))
+    theme_classic()
 
 ```
+
+To understand the role of AI chatbots in learning, students identified their usage by checking off various pre-defined use cases in the survey.
+@fig-AI-usage-type-comp demonstrates that asking technical questions and explaining concepts were the two top use cases among students.
+Less than half used it to entirely solve their homework, which could be explained by the complex nature of the homework, that involved doing extensive research and writing.
+Also just less than half used it to write paper content, which could suggest that students do not feel confident using it to improve writing, concurring with student attitudes in @Cahill_2024.
 

### 2024-08-01 paper/paper.qmd
wrote more paper

@@
 library(tidyr)
+library(quanteda)
 
@@
 
+TODO: Make a viz of helpfulness in various assignments. Describe the context of the assignment.
+
+```{r fig-AI-helpfulness}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: How Students Use ChatGPT
+#| fig-align: center
+
+# Create a bar plot of how students used ChatGPT - FIX: sort from largest to smallest
+
+type_data <- survey_data |>
+              select(
+                not_used,
+                used_technical_q,
+                used_conversation,
+                used_general_knowledge,
+                used_solving_hmk,
+                used_check_soln,
+                used_quick_q,
+                used_expl_concepts,
+                used_writing_content,
+                used_writing_code
+              )|>
+              pivot_longer(cols = everything(), names_to = "Statements", values_to = "Response") |>
+              mutate(Statements = case_when(
+                Statements == "not_used" ~ "Not Used",
+                Statements == "used_technical_q" ~ "Technical Questions",
+                Statements == "used_conversation" ~ "Fun Conversation",
+                Statements == "used_general_knowledge" ~ "General Knowledge",
+                Statements == "used_solving_hmk" ~ "Entirely Solving Homework",
+                Statements == "used_check_soln" ~ "Checking my Solutions",
+                Statements == "used_quick_q" ~ "Quick Questions",
+                Statements == "used_expl_concepts" ~ "Explaining Concepts",
+                Statements == "used_writing_content" ~ "Writing Paper Content",
+                Statements == "used_writing_code" ~ "Writing Paper Code",
+                TRUE ~ Statements
+              )) |>
+              filter(Response == 1) |>
+              count(Statements) |>
+              arrange(desc(n))
+
+type_data |>
+    ggplot(aes(x = n, y = reorder(Statements, n))) +
+    geom_bar(stat="identity", fill="grey50") +
+    labs(title="How Students Used ChatGPT", x="Count", y="Use Type") +
+    theme_classic()
+
+```
+
 ## Sentiment Analysis
+
+```{r senti-analysis}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: Hello
+#| fig-align: center
+
+```
 

### 2024-08-01 paper/paper.qmd
started sentiment analysis (fun!)

@@
 ```{r setup, include=FALSE}
-
 ## Workspace Setup ##
@@
 library(tidyr)
+library(tidytext)
+library(textdata)
 library(quanteda)
@@
 
-The wide release of ChatGPT and other Large Language Models (LLMs) has rapidly transformed various parts of society, in particular the field of education.
-As these tools are increasingly accessible, their adoption and use in academic settings has created excitement and apprehension among educators and students alike.
+The wide release of ChatGPT and other Large Language Models (LLMs) has rapidly transformed education at all levels and disciplines.
+Their quick adoption in academic settings has created excitement and apprehension among educators and students alike.
 LLMs offer unprecedented support for students in tasks ranging from writing assistance to support in complex problem-solving, with the potential to enhance academic performance and student outcomes.
-At the same time, their use raises questions about academic integrity, the development of critical thinking skills (particular in undergraduates), as well as questions about their implications for effective learning overall.
+At the same time, their use raises important questions about academic integrity, the development of critical thinking skills (particular in undergraduates), as well as questions about their implications for effective learning overall.
 
@@
 As we have seen, the existing inquiry by educators has explored the general qualitative usage and student perceptions on these tools.
-Though informative, there is still a key lack of evidence of how students' academic performance can be affected through allowing them to use LLMs in classwork.
+Though informative, a key question still remains: how can students' academic performance be affected, precisely, through allowing them to use LLMs in classwork?
+
 This paper aims to fill this gap by quantitatively investigating the relationship between students' grades and measures of student LLM usage and their attitudes toward LLMs in general, using evidence from a third-year undergraduate statistics course at the University of Toronto.
@@
 All the data was cleaned using R [@citeR] and it's tidyverse [@tidyverse], janitor [@janitor] and stringr [@stringr], then tested for issues using the testthat package [@testthat].
-This led to a final dataset containing 121 responses across (var_num) variables.
+This led to a final dataset containing 121 responses across 41 variables that corresponded to each survey question.
 
-These variables were:
+## General Information
 
-- [list of variables]
-
-## Visualization
-
-The student survey data was then visualized and analyzed using the ggplot2 [@ggplot2] package.
-
-### General Information
-
-First, some general information about our sample of students was derived through visualizing the basic data about them.
+First, general information about the sample of students was derived first, through visualizing the basic data about them. This was done using the ggplot2 [@ggplot2] package.
 
@@
 
-#Trim values to histogram
+# Trim values to histogram
 
@@
   geom_histogram(binwidth = 2, fill = "skyblue", color = "black") +
-    labs(title = "Distribution of Survey Response Times",
+  labs(
+    title = "Distribution of Survey Response Times",
     x = "Time (Mins)",
-         y = "Frequency") +
+    y = "Frequency"
+  ) +
   theme_minimal() +
-    theme(plot.title.position = "plot",
-          plot.title = element_text(hjust = 0.5))
-
+  theme(
+    plot.title.position = "plot",
+    plot.title = element_text(hjust = 0.5)
+  )
 ```
@@
 
-#Histogram of GPA
+# Histogram of GPA
 
@@
   geom_histogram(binwidth = 0.1, fill = "skyblue", color = "black") +
-    labs(title = "Distribution of Student GPA",
+  labs(
+    title = "Distribution of Student GPA",
     x = "GPA",
-         y = "Frequency") +
+    y = "Frequency"
+  ) +
   theme_minimal() +
-    theme(plot.title.position = "plot",
-          plot.title = element_text(hjust = 0.6))
-
+  theme(
+    plot.title.position = "plot",
+    plot.title = element_text(hjust = 0.6)
+  )
 ```
@@
 kable(year_count)
-
 ```
@@
 
-#Pie chart of program of study
+# Pie chart of program of study
 
@@
   coord_polar("y") +
-  labs(title = "Student Distribution by Program of Study",
+  labs(
+    title = "Student Distribution by Program of Study",
     x = NULL,
-       y = NULL) +
+    y = NULL
+  ) +
   theme_minimal() +
-  scale_fill_discrete(name="Program") +
-  theme(axis.text.x = element_blank(),  # Remove x-axis text
+  scale_fill_discrete(name = "Program") +
+  theme(
+    axis.text.x = element_blank(), # Remove x-axis text
     axis.title.x = element_blank(), # Remove x-axis title
@@
   geom_text(aes(label = n),
-            position = position_stack(vjust = 0.5))
-
+    position = position_stack(vjust = 0.5)
+  )
 ```
 
-Programs of study [did something], as we can see in @fig-programs-pie...
+Programs of study [did something; waiting on major/minor data], as we can see in @fig-programs-pie...
 
-### Student Attitudes
+## Student Attitudes
 
@@
 table_data <- survey_data |>
-    select(ai_familiarity)|>
+  select(ai_familiarity) |>
   mutate(
@@
   coord_polar("y") +
-  labs(title = "Student Familiarity with AI",
+  labs(
+    title = "Student Familiarity with AI",
     x = NULL,
-       y = NULL) +
+    y = NULL
+  ) +
   theme_minimal() +
-  scale_fill_discrete(name="Program") +
-  theme(axis.text.x = element_blank(),  # Remove x-axis text
+  scale_fill_discrete(name = "Program") +
+  theme(
+    axis.text.x = element_blank(), # Remove x-axis text
     axis.title.x = element_blank(), # Remove x-axis title
@@
 
-#Select first three statements
+# Select first three statements
 writing_perceptions <- survey_data |>
@@
 
-#Taken from: https://stackoverflow.com/questions/52297978/decrease-overal-legend-size-elements-and-text
+# Taken from: https://stackoverflow.com/questions/52297978/decrease-overal-legend-size-elements-and-text
 addSmallLegend <- function(myPlot, pointSize = 0.5, textSize = 3, spaceLegend = 0.1) {
   myPlot +
-        guides(shape = guide_legend(override.aes = list(size = pointSize)),
-               color = guide_legend(override.aes = list(size = pointSize))) +
-        theme(legend.title = element_text(size = textSize), 
+    guides(
+      shape = guide_legend(override.aes = list(size = pointSize)),
+      color = guide_legend(override.aes = list(size = pointSize))
+    ) +
+    theme(
+      legend.title = element_text(size = textSize),
       legend.text = element_text(size = textSize),
-              legend.key.size = unit(spaceLegend, "lines"))
+      legend.key.size = unit(spaceLegend, "lines")
+    )
 }
@@
   geom_bar(position = "dodge", stat = "count") +
-  labs(title = "Student Self-Perceptions on Writing",
+  labs(
+    title = "Student Self-Perceptions on Writing",
     x = "Response",
-       y = "Frequency") +
+    y = "Frequency"
+  ) +
   theme_minimal() +
-  theme(plot.title.position = "plot",
+  theme(
+    plot.title.position = "plot",
     plot.title = element_text(hjust = 0.5),
-        legend.position= "bottom", 
+    legend.position = "bottom",
   )
@@
 # TODO: Fix label, Fix Legend title, Fix legibility of Response
-
 ```
@@
   geom_bar(position = "dodge", stat = "count") +
-  labs(title = "Student Self-Perceptions of Coding",
+  labs(
+    title = "Student Self-Perceptions of Coding",
     x = "Response",
-       y = "Frequency") +
+    y = "Frequency"
+  ) +
   theme_minimal() +
-  theme(plot.title.position = "plot",
+  theme(
+    plot.title.position = "plot",
     plot.title = element_text(hjust = 0.5),
-        legend.position= "bottom", 
+    legend.position = "bottom",
   )
@@
 multi_bar
-
 ```
@@
 
-#Pie chart of student opinion on appropriateness of AI use in class.
+# Pie chart of student opinion on appropriateness of AI use in class.
 
@@
   coord_polar(theta = "y") +
-  labs(title = "Do students think AI use is appropriate in class?",
+  labs(
+    title = "Do students think AI use is appropriate in class?",
     x = NULL,
-       y = NULL) +
+    y = NULL
+  ) +
   theme_minimal() +
-  scale_fill_discrete(name="Response") +
-  theme(plot.title.position = "plot",
+  scale_fill_discrete(name = "Response") +
+  theme(
+    plot.title.position = "plot",
     plot.title = element_text(hjust = 0.5),
@@
   geom_text(aes(label = n),
-            position = position_stack(vjust = 0.5))
-
+    position = position_stack(vjust = 0.5)
+  )
 ```
 
-### Student Usage
+## Student Usage
 
@@
   geom_bar(position = "dodge", stat = "count") +
-  labs(title = "Student Self-Perceptions of Coding",
+  labs(
+    title = "Student Self-Perceptions of Coding",
     x = "Response",
-       y = "Frequency") +
+    y = "Frequency"
+  ) +
   theme_minimal() +
-  theme(plot.title.position = "plot",
+  theme(
+    plot.title.position = "plot",
     plot.title = element_text(hjust = 0.5),
@@
     used_writing_code
-              )|>
+  ) |>
   pivot_longer(cols = everything(), names_to = "Statements", values_to = "Response") |>
@@
   ggplot(aes(x = n, y = reorder(Statements, n))) +
-    geom_bar(stat="identity", fill="grey50") +
-    labs(title="How Students Used ChatGPT", x="Count", y="Use Type") +
+  geom_bar(stat = "identity", fill = "grey50") +
+  labs(title = "How Students Used ChatGPT", x = "Count", y = "Use Type") +
   theme_classic()
-
 ```
@@
 
-```{r senti-analysis}
+```{r setup-sentiment}
+#| message: false
+#| echo: false
+#| warning: false
+
+#get positive words from NRC lexicon
+pos_words <- get_sentiments("nrc") |>
+  filter(sentiment = c("anger", "disgust", "fear", "negative", "sadness"))
+
+#get negative words from NRC lexicon
+neg_words <- get_sentiments("nrc") |>
+  filter(sentiment = c("joy", "positive",
+"surprise",
+"trust", 
+"anticipation"))
+```
+
+We then performed sentiment analysis on the free-response comment questions asked to students.
+To perform this text analysis, the tidytext package [@tidytext] was used in conjunction with the NRC word-emotion lexicon compiled by @mohammad.
+Although four lexicon options were available in tidytext, the NRC lexicon was chosen due to the relatively balanced amounts of positive and negative words in the dataset, allowing for balanced classification.
+
+### Appropriateness of AI in Schoolwork
+
+In the survey, students were asked to respond to the question: "Please elaborate on to what extent do you think using generative AI tools such as ChatGPT by OpenAI (or equivalents) is ethical and appropriate for schoolwork?".
+
+```{r setup-ai-appropriate}
+
+# AI Appropriateness Comments
+ai_comments <- survey_data |>
+  select(
+    id,
+    ai_schoolwork_appropriate_comments
+  ) |>
+  mutate(
+    comment = ai_schoolwork_appropriate_comments
+  ) |>
+  select(-c(ai_schoolwork_appropriate_comments))
+
+# put comments into tidytext format
+tidy_ai_comments <- ai_comments |>
+  group_by(id) |>
+  unnest_tokens(word, comment)
+
+```
+
+### Recommendations in STA302 
+```{r}
 #| message: false
@@
 
+recs_comments <- survey_data |>
+  select(
+    id,
+    recs_for_ai_sta302
+  )
 ```

### 2024-08-01 paper/paper.qmd
finished ai_appropriateness sentiment

@@
 library(textdata)
-library(quanteda)
+library(wordcloud)
+library(reshape2)
 
@@
 This led to a final dataset containing 121 responses across 41 variables that corresponded to each survey question.
+
+```{r fig-boilerplate}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: (CHANGE) Boilerplate figure caption 
+#| fig-align: center
+# survey_data |>
+#   ggplot(program_counts, aes(x = "", y = n, fill = program_of_study)) +
+#   #put geom type here (put before compilation) +
+#   labs(title = "Title!",
+#        x = "X label",
+#        y = "Y label") +
+#   theme_minimal() +
+#   scale_fill_discrete(name="Legend Title") +
+#   theme(plot.title.position = "plot",
+#         plot.title = element_text(hjust = 0.5)
+#         )
+```
+
+```{r tbl-boilerplate}
+#| message: false
+#| echo: false
+#| warning: false
+#| tbl-cap: (CHANGE) Boilerplate table caption 
+#| fig-align: center
+#Put code below this line
+```
 
@@
 #| warning: false
-#| fig-cap: How Students Use ChatGPT
+#| fig-cap: CHANGE/REMOVE
 #| fig-align: center
 
-# Create a bar plot of how students used ChatGPT - FIX: sort from largest to smallest
-
-type_data <- survey_data |>
-  select(
-    not_used,
-    used_technical_q,
-    used_conversation,
-    used_general_knowledge,
-    used_solving_hmk,
-    used_check_soln,
-    used_quick_q,
-    used_expl_concepts,
-    used_writing_content,
-    used_writing_code
-  ) |>
-  pivot_longer(cols = everything(), names_to = "Statements", values_to = "Response") |>
-  mutate(Statements = case_when(
-    Statements == "not_used" ~ "Not Used",
-    Statements == "used_technical_q" ~ "Technical Questions",
-    Statements == "used_conversation" ~ "Fun Conversation",
-    Statements == "used_general_knowledge" ~ "General Knowledge",
-    Statements == "used_solving_hmk" ~ "Entirely Solving Homework",
-    Statements == "used_check_soln" ~ "Checking my Solutions",
-    Statements == "used_quick_q" ~ "Quick Questions",
-    Statements == "used_expl_concepts" ~ "Explaining Concepts",
-    Statements == "used_writing_content" ~ "Writing Paper Content",
-    Statements == "used_writing_code" ~ "Writing Paper Code",
-    TRUE ~ Statements
-  )) |>
-  filter(Response == 1) |>
-  count(Statements) |>
-  arrange(desc(n))
-
-type_data |>
-  ggplot(aes(x = n, y = reorder(Statements, n))) +
-  geom_bar(stat = "identity", fill = "grey50") +
-  labs(title = "How Students Used ChatGPT", x = "Count", y = "Use Type") +
-  theme_classic()
 ```
@@
 #get positive words from NRC lexicon
-pos_words <- get_sentiments("nrc") |>
-  filter(sentiment = c("anger", "disgust", "fear", "negative", "sadness"))
+neg_words <- get_sentiments("bing") |>
+  filter(sentiment == "negative")
 
 #get negative words from NRC lexicon
-neg_words <- get_sentiments("nrc") |>
-  filter(sentiment = c("joy", "positive",
-"surprise",
-"trust", 
-"anticipation"))
+pos_words <- get_sentiments("bing") |>
+  filter(sentiment == "positive")
 ```
 
-We then performed sentiment analysis on the free-response comment questions asked to students.
-To perform this text analysis, the tidytext package [@tidytext] was used in conjunction with the NRC word-emotion lexicon compiled by @mohammad.
-Although four lexicon options were available in tidytext, the NRC lexicon was chosen due to the relatively balanced amounts of positive and negative words in the dataset, allowing for balanced classification.
+We then performed sentiment analysis on the free-response comment questions asked to students. To perform this text analysis, the tidytext package [@tidytext] was used.
 
@@
 ```{r setup-ai-appropriate}
+#| message: false
+#| echo: false
+#| warning: false
 
@@
   unnest_tokens(word, comment)
+
+```
+
+```{r tbl-ai-appropriate}
+#| message: false
+#| echo: false
+#| warning: false
+#| tbl-cap: Comparing Incidences of Positive and Negative Words.
+#| tbl-subcap: ["Top 10 Mentioned Positive Words", "Top 10 Mentioned Negative Words"]
+#| layout-ncol: 2
+#| fig-align: center
+
+# Generate positive words table
+tbl_ai_pos <- tidy_ai_comments |>
+  ungroup() |>
+  select(-c("id")) |>
+  inner_join(pos_words) |>
+  count(word, sort = TRUE) |>
+  head(10) |>
+  kable(, format = "markdown")
+
+# Generate negative words table
+tbl_ai_neg <- tidy_ai_comments |>
+  ungroup() |>
+  select(-c("id")) |>
+  inner_join(neg_words) |>
+  count(word, sort = TRUE) |>
+  head(10) |>
+  kable(, format = "markdown")
+
+tbl_ai_pos
+tbl_ai_neg
+
+```
+
+```{r fig-wordcloud-ai-appropriate}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: (CHANGE) Boilerplate figure caption 
+#| fig-align: center
+
+# Load sentiment lexicon
+sentiments <- get_sentiments()
+
+# Prepare data for comparison cloud
+comparison_data <- tidy_ai_comments |>
+  inner_join(sentiments) |>
+  count(word, sentiment, sort = TRUE) |>
+  acast(word ~ sentiment, value.var = "n", fill = 0)
+
+# Generate comparative word cloud
+comparison.cloud(
+  comparison_data,
+  colors = c("darkred", "darkgreen"),
+  max.words = 100,
+  title.size = 1.5,
+  scale = c(3, 0.5)  # Adjust scale for word size
+)
 

### 2024-08-02 paper/paper.qmd
worked more on NLP

@@
 number-sections: true
+table-of-contents: true
 bibliography: references.bib
@@
 library(textdata)
+library(quanteda)
 library(wordcloud)
 library(reshape2)
+library(topicmodels)
 
@@
 
+\newpage
+
 # Introduction
 
-The wide release of ChatGPT and other Large Language Models (LLMs) has rapidly transformed education at all levels and disciplines.
-Their quick adoption in academic settings has created excitement and apprehension among educators and students alike.
-LLMs offer unprecedented support for students in tasks ranging from writing assistance to support in complex problem-solving, with the potential to enhance academic performance and student outcomes.
+The wide release of ChatGPT and other Large Language Models (LLMs) has rapidly transformed education at all levels and disciplines, creating excitement and apprehension among educators and students alike.
+They offer unprecedented support for student learning in tasks ranging from writing assistance to support in complex problem-solving, with the potential to enhance academic performance and student outcomes.
 At the same time, their use raises important questions about academic integrity, the development of critical thinking skills (particular in undergraduates), as well as questions about their implications for effective learning overall.
@@
 All the data was cleaned using R [@citeR] and it's tidyverse [@tidyverse], janitor [@janitor] and stringr [@stringr], then tested for issues using the testthat package [@testthat].
-This led to a final dataset containing 121 responses across 41 variables that corresponded to each survey question.
+This led to a final dataset containing 121 unique responses to 18  survey questions.
 
@@
   ggplot(aes(x = mins_complete)) +
-  geom_histogram(binwidth = 2, fill = "skyblue", color = "black") +
+  geom_histogram(binwidth = 2, fill = "lightblue", color = "black") +
   labs(
@@
 There was one consenting respondent who took 4308 minutes, whose observation was dropped from the visualization.
-This suggests that students generally took a meaningful amount of time to engage with the survey.
+Overall, however, this suggests that students generally took a meaningful amount of time and thought to engage with the survey's questions.
 
@@
 #| warning: false
-#| tbl-cap: Distribution of Student GPA
+#| fig-cap: Distribution of Student GPA
 #| fig-align: center
@@
   ggplot(aes(x = gpa)) +
-  geom_histogram(binwidth = 0.1, fill = "skyblue", color = "black") +
+  geom_histogram(binwidth = 0.1, fill = "lightblue", color = "black") +
   labs(
@@
 
-Get inspired by Cahill and McCabe's paper for visualizations etc.
-
 ```{r tbl-year-dist}
@@
 #| warning: false
-#| tbl-cap: Distribution of Students by Year
+#| tbl-cap: Count of Students by Year
 #| fig-align: center
@@
   labs(
-    title = "Student Distribution by Program of Study",
+    title = "Distribution of Students by Program of Study",
+    subtitle = "Visualizing the proportion of students across different programs",
     x = NULL,
@@
   theme_minimal() +
-  scale_fill_discrete(name = "Program") +
+  scale_fill_discrete(name = "Program of Study") +
   theme(
@@
     axis.title.x = element_blank(), # Remove x-axis title
-    plot.title = element_text(hjust = 1.6)
+    plot.title = element_text(hjust = 0.5),
+    plot.subtitle = element_text(hjust = 0.5)
   ) +
@@
 
+# Custom colors
+familiarity_colors <- c(
+  "Not familiar" = "tomato",
+  "Somewhat familiar" = "lightblue",
+  "Very familiar" = "lightgreen"
+)
+
 table_data |>
@@
   theme_minimal() +
-  scale_fill_discrete(name = "Program") +
+  scale_fill_manual(name = "Familiarity Level", values = familiarity_colors) +
   theme(
@@
     axis.title.x = element_blank(), # Remove x-axis title
-    plot.title = element_text(hjust = 1.35)
+    plot.title = element_text(hjust = 0.5),
+    plot.subtitle = element_text(hjust = 0.5)
   ) +
@@
 Just as @Cahill_2024 found, the vast majority of students taking the course were at least "somewhat familiar" with ChatGPT and other LLMs.
-This overrepresentation of familiarity also primarily driven by the fact that the course is primarily for students in Statistics, Computer Science, Economics, Mathematics and other quantitative disciplines, which could likely lead to a skewing of interest/affinity for computers.
+This overrepresentation of familiarity also primarily driven by the fact that the course is primarily targeted for students in Statistics, Computer Science, Economics, Mathematics and other quantitative disciplines, possibly leading to a sample that is skewed to being more interested in computing.
 
@@
       Statements == "i_like_to_write" ~ "\"I like to write\"",
-      Statements == "i_believe_it_is_important_to_be_a_good_writer" ~ "\"I believe it is important\n to be a good writer\"",
+      Statements == "i_believe_it_is_important_to_be_a_good_writer" ~ "\"I believe it is important\nto be a good writer\"",
       TRUE ~ Statements
@@
 
+# Custom colors
+response_colors <- c(
+  "A lot like me" = "lightgreen",
+  "Somewhat describes me" = "lightblue",
+  "Very different from me" = "tomato"
+)
+
 # Taken from: https://stackoverflow.com/questions/52297978/decrease-overal-legend-size-elements-and-text
@@
       shape = guide_legend(override.aes = list(size = pointSize)),
-      color = guide_legend(override.aes = list(size = pointSize))
+      color = guide_legend(override.aes = list(size = pointSize)),
+      fill = guide_legend(title = "Response", override.aes = list(size = pointSize))
     ) +
@@
 multi_bar <- writing_perceptions_long |>
-  ggplot(aes(x = Response, fill = Statements)) +
+  ggplot(aes(x = Response, fill = Response)) +
   geom_bar(position = "dodge", stat = "count") +
+  facet_wrap(~Statements) +
   labs(
     title = "Student Self-Perceptions on Writing",
+    subtitle = "Responses to statements about writing ease, enjoyment, and importance",
     x = "Response",
-    y = "Frequency"
+    y = "Number of Students"
   ) +
+  scale_fill_manual(values = response_colors, name = "Response") + # Using custom colors
   theme_minimal() +
@@
     plot.title = element_text(hjust = 0.5),
+    plot.subtitle = element_text(hjust = 0.5),
+    axis.text.x = element_blank(),
     legend.position = "bottom",
+    legend.title = element_text(size = 10),
+    legend.text = element_text(size = 8)
   )
 
-multi_bar <- addSmallLegend(multi_bar, pointSize = 0.5, textSize = 10, spaceLegend = 0.8)
+multi_bar <- addSmallLegend(multi_bar, pointSize = 0.5, textSize = 9, spaceLegend = 0.8)
 
 multi_bar
-
-# TODO: Fix label, Fix Legend title, Fix legibility of Response
 ```
@@
     Statements = case_when(
-      Statements == "when_i_edit_it_is_easy_for_me_to_catch_my_mistakes" ~ "\"When I edit, it is easy to\n catch my own mistakes\"",
-      Statements == "i_feel_confident_sharing_my_writing" ~ "\"I feel confident\n sharing my writing\"",
-      Statements == "i_am_confident_in_my_overall_writing_ability" ~ "\"I am confident in\n my own writing ability\"",
+      Statements == "when_i_edit_it_is_easy_for_me_to_catch_my_mistakes" ~ "\"When I edit, it is easy to\ncatch my own mistakes\"",
+      Statements == "i_feel_confident_sharing_my_writing" ~ "\"I feel confident\nsharing my writing\"",
+      Statements == "i_am_confident_in_my_overall_writing_ability" ~ "\"I am confident in\nmy own writing ability\"",
       TRUE ~ Statements
@@
 # Make the plot
-multi_bar <- writing_perceptions_long |>
-  ggplot(aes(x = Response, fill = Statements)) +
+multi_bar2 <- writing_perceptions_long |>
+  ggplot(aes(x = Response, fill = Response)) +
   geom_bar(position = "dodge", stat = "count") +
+  facet_wrap(~Statements) +
   labs(
     title = "Student Self-Perceptions on Writing",
+    subtitle = "Responses to statements about confidence and editing in writing",
     x = "Response",
-    y = "Frequency"
+    y = "Number of Students"
   ) +
+  scale_fill_manual(values = response_colors, name = "Response") + # Using custom colors
   theme_minimal() +
@@
     plot.title = element_text(hjust = 0.5),
+    plot.subtitle = element_text(hjust = 0.5),
+    axis.text.x = element_blank(), # Remove x-axis labels
     legend.position = "bottom",
+    legend.title = element_text(size = 10),
+    legend.text = element_text(size = 8)
   )
 
-multi_bar <- addSmallLegend(multi_bar, pointSize = 0.5, textSize = 10, spaceLegend = 0.8)
+multi_bar2 <- addSmallLegend(multi_bar2, pointSize = 0.5, textSize = 9, spaceLegend = 0.8)
 
-multi_bar
+multi_bar2
+
 ```
@@
 
+# Custom colors
+response_colors <- c(
+  "A lot like me" = "lightgreen",
+  "Somewhat describes me" = "lightblue",
+  "Very different from me" = "tomato"
+)
+
 # Make the plot
@@
     x = "Response",
-    y = "Frequency"
+    y = "Number of Students"
   ) +
@@
     legend.position = "bottom",
+    axis.text.x = element_blank(),
   )
@@
 
-```{r fig-ai-use-multibar}
-#| message: false
-#| echo: false
-#| warning: false
-#| fig-cap: Hello
-#| fig-align: center
-
-# Create a multi-bar plot for AI use/perceptions grouped by response
-
-# THROWS ERROR: FIX
-
-# Select first three statements
-writing_perceptions <- survey_data |>
-  select(
-    used_ai_any_reason,
-    used_ai_in_sta302
-  )
-
-writing_perceptions_long <- writing_perceptions |>
-  pivot_longer(cols = everything(), names_to = "Statements", values_to = "Response") |>
-  mutate(
-    Statements = case_when(
-      Statements == "used_ai_any_reason" ~ "Used AI for any reason",
-      Statements == "used_ai_in_sta302" ~ "Used AI in Class",
-      TRUE ~ Statements
-    ),
-  ) |>
-  filter(!is.na(Response)) |>
-  filter(Response == "No" | Response == "Yes")
-
-# Taken from: https://stackoverflow.com/questions/52297978/decrease-overal-legend-size-elements-and-text
-addSmallLegend <- function(myPlot, pointSize = 0.5, textSize = 3, spaceLegend = 0.1) {
-  myPlot +
-    guides(
-      shape = guide_legend(override.aes = list(size = pointSize)),
-      color = guide_legend(override.aes = list(size = pointSize))
-    ) +
-    theme(
-      legend.title = element_text(size = textSize),
-      legend.text = element_text(size = textSize),
-      legend.key.size = unit(spaceLegend, "lines")
-    )
-}
-
-# Make the plot
-multi_bar <- writing_perceptions_long |>
-  ggplot(aes(x = Response, fill = Statements)) +
-  geom_bar(position = "dodge", stat = "count") +
-  labs(
-    title = "Student Self-Perceptions of Coding",
-    x = "Response",
-    y = "Frequency"
-  ) +
-  theme_minimal() +
-  theme(
-    plot.title.position = "plot",
-    plot.title = element_text(hjust = 0.5),
-  )
-
-multi_bar <- addSmallLegend(multi_bar, pointSize = 0.5, textSize = 10, spaceLegend = 0.8)
-
-multi_bar
-```
-
 ```{r fig-AI-usage-type-comp}
@@
 #| fig-align: center
-
-# Create a bar plot of how students used ChatGPT - FIX: sort from largest to smallest
 
@@
     Statements == "used_solving_hmk" ~ "Entirely Solving Homework",
-    Statements == "used_check_soln" ~ "Checking my Solutions",
+    Statements == "used_check_soln" ~ "Checking Solutions",
     Statements == "used_quick_q" ~ "Quick Questions",
@@
   ggplot(aes(x = n, y = reorder(Statements, n))) +
-  geom_bar(stat = "identity", fill = "grey50") +
-  labs(title = "How Students Used ChatGPT", x = "Count", y = "Use Type") +
+  geom_bar(stat = "identity", fill = "lightblue") +
+  labs(title = "How Students Used ChatGPT", x = "Number of Students", y = "Type of Usage") +
   theme_classic()
@@
 To understand the role of AI chatbots in learning, students identified their usage by checking off various pre-defined use cases in the survey.
-@fig-AI-usage-type-comp demonstrates that asking technical questions and explaining concepts were the two top use cases among students.
-Less than half used it to entirely solve their homework, which could be explained by the complex nature of the homework, that involved doing extensive research and writing.
-Also just less than half used it to write paper content, which could suggest that students do not feel confident using it to improve writing, concurring with student attitudes in @Cahill_2024.
+As we can see in @fig-AI-usage-type-comp, technical questions and explaining concepts were the two top use cases among students.
+More than half of students also used ChatGPT for Quick Questions, General Knowledge, Writing Paper Code, and Checking Solutions.
+Just less than half used it to write paper content, which could suggest that students do not feel confident using it to improve writing, concurring with student attitudes in @Cahill_2024.
 
@@
 
-#get positive words from NRC lexicon
-neg_words <- get_sentiments("bing") |>
+#get positive words from tidytext
+neg_words <- get_sentiments() |>
   filter(sentiment == "negative")
 
-#get negative words from NRC lexicon
-pos_words <- get_sentiments("bing") |>
+#get negative words from tidytext
+pos_words <- get_sentiments() |>
   filter(sentiment == "positive")
@@
 
-We then performed sentiment analysis on the free-response comment questions asked to students. To perform this text analysis, the tidytext package [@tidytext] was used.
+We then performed basic sentiment analysis on the open-response comment questions asked to students. To perform this text analysis, the tidytext package [@tidytext] was used.
 
@@
 
-In the survey, students were asked to respond to the question: "Please elaborate on to what extent do you think using generative AI tools such as ChatGPT by OpenAI (or equivalents) is ethical and appropriate for schoolwork?".
+In the survey, students were asked to provide an open-ended response to the question: "Please elaborate on to what extent do you think using generative AI tools such as ChatGPT by OpenAI (or equivalents) is ethical and appropriate for schoolwork?".
+These responses were attempted to be understood thematically.
 
@@
   group_by(id) |>
-  unnest_tokens(word, comment)
-
-```
-
-```{r tbl-ai-appropriate}
-#| message: false
-#| echo: false
-#| warning: false
-#| tbl-cap: Comparing Incidences of Positive and Negative Words.
-#| tbl-subcap: ["Top 10 Mentioned Positive Words", "Top 10 Mentioned Negative Words"]
-#| layout-ncol: 2
-#| fig-align: center
-
-# Generate positive words table
-tbl_ai_pos <- tidy_ai_comments |>
-  ungroup() |>
-  select(-c("id")) |>
-  inner_join(pos_words) |>
-  count(word, sort = TRUE) |>
-  head(10) |>
-  kable(, format = "markdown")
-
-# Generate negative words table
-tbl_ai_neg <- tidy_ai_comments |>
-  ungroup() |>
-  select(-c("id")) |>
-  inner_join(neg_words) |>
-  count(word, sort = TRUE) |>
-  head(10) |>
-  kable(, format = "markdown")
-
-tbl_ai_pos
-tbl_ai_neg
+  unnest_tokens(word, comment) |>
+  mutate(word = str_to_lower(string = word))
 
@@
 )
+```
+
+```{r tbl-ai-appropriate}
+#| message: false
+#| echo: false
+#| warning: false
+#| tbl-cap: Comparing Incidences of Positive and Negative Words.
+#| tbl-subcap: ["Top 10 Positive Words", "Top 10 Negative Words"]
+#| layout-ncol: 2
+#| fig-align: center
+
+# Generate positive words table
+tbl_ai_pos <- tidy_ai_comments |>
+  ungroup() |>
+  select(-c("id")) |>
+  inner_join(pos_words) |>
+  count(word, sort = TRUE) |>
+  head(10) |>
+  kable(, format = "markdown")
+
+# Generate negative words table
+tbl_ai_neg <- tidy_ai_comments |>
+  ungroup() |>
+  select(-c("id")) |>
+  inner_join(neg_words) |>
+  count(word, sort = TRUE) |>
+  head(10) |>
+  kable(, format = "markdown")
+
+tbl_ai_pos
+tbl_ai_neg
 
@@
 
-### Recommendations in STA302 
-```{r}
+### AI Implementation Recommendations
+
+```{r recs-setup}
 #| message: false
@@
 #| warning: false
-#| fig-cap: Hello
+
+# Example sentiments data frame
+sentiments <- get_sentiments()
+
+# Prepare data
+recs_comments <- survey_data |>
+  select(id, recs_for_ai_sta302) |>
+  mutate(comment = recs_for_ai_sta302) |>
+  select(-c(recs_for_ai_sta302))
+
+# Tokenize and clean individual words
+tidy_recs_comments <- recs_comments |>
+  ungroup() |>
+  unnest_tokens(word, comment) |>
+  anti_join(stop_words)
+
+# Count word frequencies
+word_counts <- tidy_recs_comments |>
+  count(word, sort = TRUE)
+
+# Bigram analysis
+bigrams <- recs_comments |>
+  ungroup() |>
+  unnest_tokens(bigram, comment, token = "ngrams", n = 2) |>
+  separate(bigram, into = c("word1", "word2"), sep = " ") |>
+  filter(!word1 %in% stop_words$word & !word2 %in% stop_words$word) |>
+  count(word1, word2, sort = TRUE) |>
+  unite(bigram, word1, word2, sep = " ")
+
+# Prepare data for comparison cloud (including both words and bigrams)
+comparison_data_words <- tidy_recs_comments |>
+  inner_join(sentiments, by = "word") |>
+  count(word, sentiment, sort = TRUE)
+
+comparison_data_bigrams <- bigrams |>
+  separate(bigram, into = c("word1", "word2"), sep = " ") |>
+  inner_join(sentiments, by = c("word1" = "word")) |>
+  inner_join(sentiments, by = c("word2" = "word"), suffix = c(".word1", ".word2")) |>
+  unite(bigram, word1, word2, sep = " ") |>
+  count(bigram, sentiment.word1, sort = TRUE) |>
+  rename(sentiment = sentiment.word1)
+
+# Combine both data sets
+combined_data <- bind_rows(
+  comparison_data_words |> rename(term = word),
+  comparison_data_bigrams |> rename(term = bigram)
+) |>
+  arrange(desc(n))
+
+```
+
+```{r recs-wordcloud}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: Word Cloud for Recommendations Comments
 #| fig-align: center
 
-recs_comments <- survey_data |>
-  select(
-    id,
-    recs_for_ai_sta302
+combined_data |>
+  with(
+    wordcloud(
+      words = term,
+      freq = n,
+      scale = c(3, 0.5),
+      min.freq = 1, # Adjust scale as needed
+      max.words = 100,    # Maximum number of words to display
+      colors = "blue"     # Color for the word cloud
+    )
+  )
+
+```
+
+```{r tbl-recs-relpolarity}
+#| message: false
+#| echo: false
+#| warning: false
+#| tbl-cap: Comparing Incidences of Positive and Negative Words in Student.
+#| tbl-subcap: ["Top 10 Positive Words", "Top 10 Negative Words"]
+#| layout-ncol: 2
+#| fig-align: center
+
+# Generate positive words table
+tbl_recs_pos <- tidy_recs_comments |>
+  ungroup() |>
+  select(-c("id")) |>
+  inner_join(pos_words) |>
+  count(word, sort = TRUE) |>
+  head(10) |>
+  kable(, format = "markdown")
+
+# Generate negative words table
+tbl_recs_neg <- tidy_recs_comments |>
+  ungroup() |>
+  select(-c("id")) |>
+  inner_join(neg_words) |>
+  count(word, sort = TRUE) |>
+  head(10) |>
+  kable(, format = "markdown")
+
+tbl_recs_pos
+tbl_recs_neg
+```
+
+```{r word-cloud}
+#| message: false
+#| echo: false
+#| warning: false
+
+# Word cloud based on bigrams
+word_counts |>
+  with(
+    wordcloud(
+      words = word,
+      freq = n,
+      scale = c(3, 0.5),  # Adjust scale as needed
+      max.words = 100,    # Maximum number of words to display
+      colors = "blue"     # Color for the word cloud
+    )
   )

### 2024-08-02 paper/paper.qmd
Further cleaned up visualizations

@@
 
-# Create a multi-bar plot for coding self-perceptions grouped by response
-
-# Select first three statements
-writing_perceptions <- survey_data |>
+# Select first three statements for coding self-perceptions
+coding_perceptions <- survey_data |>
   select(
@@
 
-writing_perceptions_long <- writing_perceptions |>
+# Convert to long format and adjust labels
+coding_perceptions_long <- coding_perceptions |>
   pivot_longer(cols = everything(), names_to = "Statements", values_to = "Response") |>
@@
       Response == "This somewhat describes me" ~ "Somewhat describes me",
+      TRUE ~ Response
     )
@@
   filter(!is.na(Response))
-
-# Taken from: https://stackoverflow.com/questions/52297978/decrease-overal-legend-size-elements-and-text
-addSmallLegend <- function(myPlot, pointSize = 0.5, textSize = 3, spaceLegend = 0.1) {
-  myPlot +
-    guides(
-      shape = guide_legend(override.aes = list(size = pointSize)),
-      color = guide_legend(override.aes = list(size = pointSize))
-    ) +
-    theme(
-      legend.title = element_text(size = textSize),
-      legend.text = element_text(size = textSize),
-      legend.key.size = unit(spaceLegend, "lines")
-    )
-}
 
@@
 
+# Function to adjust legend size
+addSmallLegend <- function(myPlot, pointSize = 0.5, textSize = 3, spaceLegend = 0.1) {
+  myPlot +
+    guides(
+      shape = guide_legend(override.aes = list(size = pointSize)),
+      color = guide_legend(override.aes = list(size = pointSize)),
+      fill = guide_legend(title = "Response", override.aes = list(size = pointSize))
+    ) +
+    theme(
+      legend.title = element_text(size = textSize),
+      legend.text = element_text(size = textSize),
+      legend.key.size = unit(spaceLegend, "lines")
+    )
+}
+
 # Make the plot
-multi_bar <- writing_perceptions_long |>
-  ggplot(aes(x = Response, fill = Statements)) +
+multi_bar <- coding_perceptions_long |>
+  ggplot(aes(x = Response, fill = Response)) +
   geom_bar(position = "dodge", stat = "count") +
+  facet_wrap(~Statements) +
   labs(
     title = "Student Self-Perceptions of Coding",
+    subtitle = "Responses to statements about coding ease, enjoyment, and importance",
     x = "Response",
@@
   ) +
+  scale_fill_manual(values = response_colors, name = "Response") + # Using custom colors
   theme_minimal() +
@@
     plot.title = element_text(hjust = 0.5),
+    plot.subtitle = element_text(hjust = 0.5),
+    axis.text.x = element_blank(),
     legend.position = "bottom",
-    axis.text.x = element_blank(),
-  )
-
-multi_bar <- addSmallLegend(multi_bar, pointSize = 0.5, textSize = 10, spaceLegend = 0.8)
-
+    legend.title = element_text(size = 10),
+    legend.text = element_text(size = 8)
+  )
+
+multi_bar <- addSmallLegend(multi_bar, pointSize = 0.5, textSize = 9, spaceLegend = 0.8)
+
+# Display the plot
 multi_bar
@@
 
-# Create a multi-bar plot for coding self-perceptions grouped by response
-
-# Select first three statements
-writing_perceptions <- survey_data |>
+# Select the statements for coding self-perceptions
+coding_perceptions <- survey_data |>
   select(
@@
       Statements == "when_i_check_my_code_it_is_easy_for_me_to_catch_my_mistakes" ~ "\"When I check my code, it is\n easy for me to catch my mistakes\"",
-      Statements == "i_feel_confident_sharing_my_code" ~ "\"I feel confident sharing my code\"",
+      Statements == "i_feel_confident_sharing_my_code" ~ "\"I feel confident\n sharing my code\"",
       Statements == "i_am_confident_in_my_overall_coding_ability" ~ "\"I am confident in\n my overall coding ability\"",
@@
 
-# Taken from: https://stackoverflow.com/questions/52297978/decrease-overal-legend-size-elements-and-text
+# Custom colors
+response_colors <- c(
+  "A lot like me" = "lightgreen",
+  "Somewhat describes me" = "lightblue",
+  "Very different from me" = "tomato"
+)
+
+# Function to adjust legend size
 addSmallLegend <- function(myPlot, pointSize = 0.5, textSize = 3, spaceLegend = 0.1) {
@@
 # Make the plot
-multi_bar <- writing_perceptions_long |>
-  ggplot(aes(x = Response, fill = Statements)) +
+multi_bar <- coding_perceptions_long |>
+  ggplot(aes(x = Response, fill = Response)) +
   geom_bar(position = "dodge", stat = "count") +
+  facet_wrap(~Statements) +
   labs(
     title = "Student Self-Perceptions of Coding",
+    subtitle = "Responses to statements about coding confidence and ease",
     x = "Response",
-    y = "Frequency"
-  ) +
+    y = "Number of Students"
+  ) +
+  scale_fill_manual(values = response_colors, name = "Response") + # Using custom colors
   theme_minimal() +
@@
     plot.title = element_text(hjust = 0.5),
+    plot.subtitle = element_text(hjust = 0.5),
+    axis.text.x = element_blank(),
     legend.position = "bottom",
+    legend.title = element_text(size = 10),
+    legend.text = element_text(size = 8)
   )
@@
 
+# Display the plot
 multi_bar
@@
   theme_minimal() +
-  scale_fill_discrete(name = "Response") +
+  scale_fill_manual(
+      name = "Response", 
+      values = c("lightgreen", "tomato", "lightblue")) +
   theme(
@@
 ```
+
+### Task-Specific Helpfulness
 
@@
 To understand the role of AI chatbots in learning, students identified their usage by checking off various pre-defined use cases in the survey.
-As we can see in @fig-AI-usage-type-comp, technical questions and explaining concepts were the two top use cases among students.
+@fig-AI-usage-type-comp shows that technical questions and explaining concepts were the two top use cases among students.
 More than half of students also used ChatGPT for Quick Questions, General Knowledge, Writing Paper Code, and Checking Solutions.
-Just less than half used it to write paper content, which could suggest that students do not feel confident using it to improve writing, concurring with student attitudes in @Cahill_2024.
+Just less than half used it to write paper content, which could suggest that students do not feel confident using it to improve writing, concurring with student attitudes expressed in @Cahill_2024.
 
@@
 
+Based on the word cloud and the relative counts of
+
 ### AI Implementation Recommendations
@@
       scale = c(3, 0.5),
-      min.freq = 1, # Adjust scale as needed
+      min.freq = 1, # Min number of occurences
       max.words = 100,    # Maximum number of words to display
-      colors = "blue"     # Color for the word cloud
+      colors = "skyblue4"     # Color for the word cloud
     )
@@
       freq = n,
-      scale = c(3, 0.5),  # Adjust scale as needed
+      scale = c(3, 0.5),
+      min.freq = 1,# Adjust scale as needed
       max.words = 100,    # Maximum number of words to display

### 2024-08-02 paper/paper.qmd
finished data section commentary

@@
 
-# Introduction
+# Introduction and Literature Review
 
@@
 
-```{r fig-boilerplate}
-#| message: false
-#| echo: false
-#| warning: false
-#| fig-cap: (CHANGE) Boilerplate figure caption 
-#| fig-align: center
-# survey_data |>
-#   ggplot(program_counts, aes(x = "", y = n, fill = program_of_study)) +
-#   #put geom type here (put before compilation) +
-#   labs(title = "Title!",
-#        x = "X label",
-#        y = "Y label") +
-#   theme_minimal() +
-#   scale_fill_discrete(name="Legend Title") +
-#   theme(plot.title.position = "plot",
-#         plot.title = element_text(hjust = 0.5)
-#         )
-```
-
-```{r tbl-boilerplate}
-#| message: false
-#| echo: false
-#| warning: false
-#| tbl-cap: (CHANGE) Boilerplate table caption 
-#| fig-align: center
-#Put code below this line
-```
-
 ## General Information
@@
 We can also see in from @tbl-year-dist that students from a range of years took the course.
-That said, the majority of students were in their 3rd or 4th year of study.
-This makes sense, given the course's prerequisite of general statistics, which is a two-semester sequence typically done in second year statistics.
-This also explains the lack of any first year students in the course as well.
+That said, the majority of students were more senior undergraduates in their 3rd or 4th year of study.
+This makes sense, given the course's prerequisite of general statistics, which is a two-course sequence typically completed by students in their second year.
 
@@
 Just as @Cahill_2024 found, the vast majority of students taking the course were at least "somewhat familiar" with ChatGPT and other LLMs.
-This overrepresentation of familiarity also primarily driven by the fact that the course is primarily targeted for students in Statistics, Computer Science, Economics, Mathematics and other quantitative disciplines, possibly leading to a sample that is skewed to being more interested in computing.
+However, this vast over-representation of familiarity could be driven by the fact that the course is primarily targeted for students in Statistics, Computer Science, Economics, Mathematics and other quantitative disciplines, leading to a sample that is skewed to being more interested in computing in general.
 
 Students were then asked about their attitudes and self-perception on writing and coding.
+First, students had varied attitudes toward their writing and their writing abilities in general, as seen in @fig-writing-multibar-1 and fig-writing-multibar-2.
 
@@
 ```
+We can see in @fig-writing-multibar-1 that most students believe it is important to be a good writing, though most are either indifferent or do not really like to write.
+In parallel way, students also do not find writing to be particularly easy, which could feed into the relative antipathy toward writing in general.
 
@@
 ```
-
-Students had varied attitudes toward their writing and their writing abilities in general, as we can see from @fig-writing-multibar-1 and fig-writing-multibar-2. ...
+Mirroring this, @fig-writing-multibar-2 shows that most students in the class were at least somewhat confident in their own writing abilities, but a significant contingent felt otherwise.
+Interestingly, although few students felt that they were confident writing ability, more felt that they were able to catch their mistakes, which could indicate a disconnect between how students perceive their work and how the work was actually evaluated.
+
+Students' self-perceptions about their coding abilities, shown in @fig-coding-multibar-1 and @fig-coding-multibar-2, also shed light on how students perceived themselves with respect to coursework, as R programming was a significant component of the course.
 
@@
 ```
+@fig-coding-multibar-1 shows that student self-perceptions regarding coding proficiency and importance varied.
+There was strong consensus on the perceived importance of coding skills, with a majority of students strongly affirming this belief.
+However, students' self-assessed coding abilities and enjoyment are more heterogeneous, with a substantial proportion reporting moderate rather than high levels of ease and enjoyment in coding tasks.
+These findings suggest a complex interplay between students' recognition of coding's significance and their personal experiences with programming activities.
 
@@
 
-Student self-perceptions about their coding abilities, shown in @fig-coding-multibar-1 and @fig-coding-multibar-2 were also very interesting. ...
+@fig-coding-multibar-2 reveals a predominant moderate level of confidence among students in their overall coding ability, willingness to share code, and capacity to identify errors.
+Notably, students express slightly higher confidence in detecting their own coding mistakes compared to general coding ability or code sharing.
+These patterns suggest that while students have developed some coding self-efficacy, there is still considerable potential for enhancing their perceived competence and comfort across various coding-related activities.
 
@@
 
-### Task-Specific Helpfulness
+Though students' self-perceptions around writing and coding are varied, there was strong consensus that the use of student AI tools is appropriate within an academic setting.
+One thing of note is that most who responded "It depends" generally found artificial intelligence tools to be appropriate, though with certain guidelines and rules governing their use.
 
@@
 
-To understand the role of AI chatbots in learning, students identified their usage by checking off various pre-defined use cases in the survey.
-@fig-AI-usage-type-comp shows that technical questions and explaining concepts were the two top use cases among students.
-More than half of students also used ChatGPT for Quick Questions, General Knowledge, Writing Paper Code, and Checking Solutions.
+To understand the role of AI chatbots in learning, students more granularly identified their usage by checking off various pre-defined use cases in the survey.
+@fig-AI-usage-type-comp shows that technical questions and explaining concepts were the two top use cases among students in the course.
+More than half of students also used ChatGPT for quick questions, general knowledge, writing paper code, and checking solutions.
 Just less than half used it to write paper content, which could suggest that students do not feel confident using it to improve writing, concurring with student attitudes expressed in @Cahill_2024.
 
-TODO: Make a viz of helpfulness in various assignments. Describe the context of the assignment.
-
-```{r fig-AI-helpfulness}
+### Task-Specific Attitudes
+
+Make a grid of all the tasks, with
+
+Students were also asked to rate the helpfulness of LLMs on various tasks assigned during the course on a 4 point scale of "Did not use" to "Very Helpful".
+To simplify the presentation, responses were grouped into two main categories: "Less Helpful" and "More Helpful." The "Less Helpful" category combines responses where students found the AI either "Not helpful" or did not use it for the task, while the "More Helpful" category includes responses where the AI was considered "Somewhat helpful" or "Very helpfu
+
+```{r fig-AI-helpfulness-grid}
 #| message: false
@@
 
-```
+helpful <- survey_data |>
+  select(
+    ai_helpful_weekly_quiz,
+    ai_helpful_weekly_mini_essay,
+    ai_helpful_papers_generating_ideas,
+    ai_helpful_papers_writing_code, 
+    ai_helpful_papers_writing_content, 
+    ai_helpful_papers_improving_content
+  )
+
+# Convert data to long format
+helpful_long <- helpful |>
+  pivot_longer(cols = everything(), names_to = "task", values_to = "response")
+
+# Simplify response categories and rename tasks
+helpful_long <- helpful_long |>
+  mutate(
+    response_simplified = case_when(
+      response %in% c("Not helpful", "I did not use generative AI for this component") ~ "Less Helpful",
+      response %in% c("Somewhat helpful", "Very helpful") ~ "More Helpful",
+      TRUE ~ "Unknown"  # Handle unexpected values
+    ),
+    task = case_when(
+      task == "ai_helpful_weekly_quiz" ~ "Weekly Quiz",
+      task == "ai_helpful_weekly_mini_essay" ~ "Weekly Mini Essay",
+      task == "ai_helpful_papers_generating_ideas" ~ "Generating Ideas",
+      task == "ai_helpful_papers_writing_code" ~ "Writing Code",
+      task == "ai_helpful_papers_writing_content" ~ "Writing Content",
+      task == "ai_helpful_papers_improving_content" ~ "Improving Content",
+      TRUE ~ "Other"  # Handle any unexpected task names
+    )
+  )
+
+# Count the number of responses for each simplified category and task
+helpful_summary <- helpful_long |>
+  group_by(task, response_simplified) |>
+  summarise(count = n(), .groups = 'drop')
+
+custom_colors <- c(
+  "Less Helpful" = "tomato",
+  "More Helpful" = "lightgreen"
+)
+
+# Create grouped bar charts
+grid_bar_chart <- helpful_summary |>
+  ggplot(aes(x = response_simplified, y = count, fill = response_simplified)) +
+  geom_bar(stat = "identity") +
+  facet_wrap(~ task, scales = "free_y") +
+  labs(title = "Helpfulness of LLMs for Different Tasks", x = "", y = "Number of Students") +
+  theme_minimal() +
+  coord_flip() + 
+  scale_fill_manual(name = "Level", values = custom_colors) +
+  theme(
+    plot.title.position = "plot",
+    plot.title = element_text(hjust = 0.5),
+    plot.subtitle = element_text(hjust = 0.5),
+    axis.text.y = element_blank(),
+    legend.position = "bottom",
+    legend.title = element_text(size = 10),
+    legend.text = element_text(size = 8)
+  )
+
+grid_bar_chart <- addSmallLegend(grid_bar_chart, pointSize = 0.5, textSize = 9, spaceLegend = 0.5)
+
+grid_bar_chart
+```
+
+@fig-AI-helpfulness-grid shows insights about how students used LLMs within the course.
+While most tasks were roughly split between students finding LLMs helpful or not, it was clear that students found them most helpful in generating code.
+In the context of the course, this meant generating R code for transforming, analyzing, and visualizing data.
+To a lesser extent, students also found LLMs to be helpful in improving the existing writing they had, while at the same time not favouring it for writing content from scratch.
 
@@
 
-In the survey, students were asked to provide an open-ended response to the question: "Please elaborate on to what extent do you think using generative AI tools such as ChatGPT by OpenAI (or equivalents) is ethical and appropriate for schoolwork?".
-These responses were attempted to be understood thematically.
+In the survey, students were asked to provide an open-ended response to the question: "Do you think using generative AI tools such as ChatGPT by OpenAI (or equivalents) is ethical and appropriate for schoolwork?".
 

### 2024-08-04 paper/paper.qmd
First round

--- 2024-08-02 paper/paper.qmd
@@
 ---
-title: "[Result, but poetic]: Does LLM Use Improve Data Science Education?"
-subtitle: "Evidence from a Canadian Undergraduate Statistics Course"
+title: "Self-reported LLM usage and results on a data science project: Evidence from a Canadian undergraduate data science course"
 author: 
@@
   - Luca Carnegie
-thanks: "Code and data are available at: https://github.com/lcarnegie/llms-achievement."
+thanks: "Code and data are available at: https://github.com/lcarnegie/llms-achievement.
+We thank Nathalie Moon for developing and sharing her survey questions.
+This research is currently under review by the University of Toronto's University Research Ethics Board.
+Comments can be sent to: rohan.alexander@utoronto.ca."
 date: today
 date-format: long
-abstract: "An abstract should eventually go here!!"
+abstract: "To help understand the effect of Large Language Models (LLMs) on data science practice we examine the extent to which LLM usage is correlated with the mark that a student gets on a final paper in a classroom data science setting.
+We find no clear relationship between more extensive LLM usage and the student's mark.
+Despite the classroom setting used for evaluation, the particular activity of interest reflects the work done by professional data scientists.
+Our finding suggests the need for more extensive work evaluating how LLMs can be integrated into the data science workflow in a way that provides value."
 format: pdf
 number-sections: true
-table-of-contents: true
+table-of-contents: false
 bibliography: references.bib
-nocite: | #references.bib entries at end, regardless of intext citation
-  @* 
 ---
 
-```{r setup, include=FALSE}
-## Workspace Setup ##
+```{r setup}
+#| include: false
 
+library(knitr)
+library(quanteda)
+library(readxl)
+library(reshape2)
+library(tidytext)
 library(tidyverse)
-library(knitr)
-library(kableExtra)
-library(lintr)
-library(tidyr)
-library(tidytext)
-library(textdata)
-library(quanteda)
-library(wordcloud)
-library(reshape2)
-library(topicmodels)
 
-# Import data
+# Old data (REMOVE ONCE THE TEXT ANALYSIS BIT IS UPDATED)
 survey_data <- read_csv("../data/analysis_data/clean_STA302_postcourse_survey_w24.csv", show_col_types = FALSE)
+
+# Read in data
+all_data <- read_excel(here::here("data/raw_data/merged_data.xlsx"), sheet = "Sheet1") |>
+  janitor::clean_names()
 ```
-
-\newpage
 
@@
 
-The wide release of ChatGPT and other Large Language Models (LLMs) has rapidly transformed education at all levels and disciplines, creating excitement and apprehension among educators and students alike.
-They offer unprecedented support for student learning in tasks ranging from writing assistance to support in complex problem-solving, with the potential to enhance academic performance and student outcomes.
-At the same time, their use raises important questions about academic integrity, the development of critical thinking skills (particular in undergraduates), as well as questions about their implications for effective learning overall.
+There are many aspects of trustworthy data science, including education, culture, and workflow to name a few.
+One aspect of special importance is the tools that we use to do data science: the computers, programming languages, settings, and environments.
+The wide release of user-friendly Large Language Models (LLMs), especially OpenAI's ChatGPT, is a rare instance of a new tool being made available.
 
-The potential for LLMs to positively affect students' academic performance in data science can be clearly inferred from already-demonstrated effects on adjacent professional fields.
-The tasks involved in a data science workflow can be generally broken down into two key competencies.
-The first is programming, which is done when cleaning, analyzing, visualizing data using languages like R or Python.
+Like all new tools, LLMs have created both excitement and apprehension.
+ChatGPT's public release on November 30, 2022, brought LLMs into the mainstream conversation.
+By now many people, especially educators and students, have some experience with LLMs, in both personal and professional contexts.
+
+In the context of teaching statistics and data science LLMs could be useful for many tasks.
+For instance, there is considerable interest in the potential of chatbots and personalized tutors **[ADD CITE]**.
+ChatGPT has also been the catalyst for many interesting and important conversations around academic integrity **[ADD CITE]**, the development of critical thinking skills **[ADD CITE]**, and what effective learning looks like **[ADD CITE]**.
+
+In this paper we are interested in better understanding LLMs as a tool for producing trustworthy data science.
+We study how they were used by students in an upper-year undergraduate data science course and whether students who used LLMs tended to have higher scores than those who did not.
+
+The tasks involved in a trustworthy data science workflow can be generally broken down into a number of key competencies [@Adhikari2021Interleaving; @Gibbs2021Building].
+The potential for LLMs to positively affect students' academic performance in data science can be clearly inferred from already-demonstrated effects in a number of those competencies on adjacent professional fields.
+The first is programming, which is done when cleaning, analyzing, visualizing data often using languages like R or Python.
 The second is writing, which is primarily done when communicating results.
-There is experimental evidence from adjacent professional fields that LLMs can improve productivity in both of these competencies.
 
-@Peng_2023 show the potential impact of LLMs on students' computer programming skills, through their uncovering of the significantly positive impacts of GitHub Copilot (an LLM-powered programming assistant) on the productivity of computer programmers.
-In an experiment involving 95 freelance programmers, they found that that programmers with access to Copilot completed a standardized programming task 55.8% faster than the control group.
-Crucially, it was found that programmers with less experience saw the greatest improvements in productivity.
-Given that programming is a core component of data science practice, this suggests that these benefits could reasonably carry over to students in data science courses as well.
+In terms of general programming, @Peng_2023 found a positive impact of GitHub Copilot (an LLM-powered programming assistant) on productivity.
+Specifically, in an experiment involving 95 freelance programmers, they found that that programmers with access to GitHub Copilot completed a standardized programming task 56% faster than the control group.
+Programmers with less experience saw the greatest improvements in productivity.
 
-@DellAcqua_2023, while primarily about management consulting tasks, provides evidence that LLMs can significantly improve writing productivity.
-In a field experiment involving Consultants recruited from the Boston Consulting Group (BCG), they found that the use of GPT-4 in the experimental group led to a 25% increase in delivery speed of business tasks (most involving some writing), as well as a 40% increase in human-rated performance on those tasks.
+@DellAcqua_2023 focus on management consulting tasks and provide evidence that LLMs can improve writing productivity.
+In a field experiment involving consultants from Boston Consulting Group they found that the use of OpenAI's GPT-4 led to a 25% increase in delivery speed of business tasks (most involving some writing), as well as a 40% increase in human-rated performance on those tasks.
 Similar to computer programming, these productivity increases were most pronounced for those with below average performance, with their output increasing by 43%.
-Though not all tasks done using AI saw the same productivity improvements, the authors expressed particular optimism for LLMs' potential to generally expedite menial knowledge-work tasks.
-These included tasks such as generating new ideas and creating persuasive and informative writing - both examples directly apply to data science education through their analogues of coming up with project ideas and writing more engaging reports.
-As a whole, they show that LLMs could equally augment the quality of student writing and course projects in data science as well.
 
-Clearly, the implications of these case studies appear promising.
-However, literature from educators both adjacent to and distant from statistics and data science suggest much more varied perspectives on how LLMs can impact student performance.
+On the other hand, @Valenzuela_2024 argue that LLMs lead to a loss of serendipity (which leads to less original work) and de-skilling (primarily with respect to programming ability), among other consequences.
+These outcomes could negatively affect students' effective learning of data science.
 
-@Valenzuela_2024 argue that LLMs lead to a loss of serendipity (which leads to less original work) and de-skilling (primarily with respect to programming ability), among other consequences.
-These particular outcomes could negatively affect students' effective learning of data science.
-On the other hand, @Ellis_Slade_2023 take a more optimistic perspective, taking the popular stance of comparing ChatGPT to previously controversial learning technologies like calculators.
-They argue that LLMs are just another technology that will impact Statistics and Data Science education like the calculator did.
-However, they take a broader perspective on the issue and more specific inquiry could be done in how effective LLMs are at improving student outcomes within data science education in particular.
+@Ellis_Slade_2023 take a more optimistic perspective and argue that LLMs are just another technology that will impact statistics and data science education.
+Similarly, @Tu_2024 acknowledge that LLMs can streamline many parts of a data science workflow.
+With that in mind, they suggest that data-scientists-in-training should shift their self-perspectives from primarily being an analyst to primarily being a manager responsible for strategic oversight of the analysis.
 
-In a similar attitude, @Tu_2024 acknowledge that for students, LLMs can streamline many parts of a data science workflow - with that in mind, they suggest that data scientists in training should shift their self-perspectives from primarily being an analyst to primarily being a manager responsible for strategic oversight of the analysis.
-Crucially they emphasize that in both education and practice, LLMs and human intelligence should play complementary roles.
-
-The simultaneous caution and interest the literature expressed toward using LLMs in educational settings was further corroborated by empirical studies on K-12 and university students as well.
-
-At the high school level, @Lazar_2023 conducted a informal survey of secondary school teachers and students on their opinions of ChatGPT, they found that while LLMs could help spark creativity, provide academic support when teachers were unavailable, and model certain types of writing well, teachers were also cognizant of LLMs potential to limit students' learning in certain ways through overreliance.
+At the high school level @Lazar_2023 conducted a informal survey of secondary school teachers and students on their opinions of ChatGPT, and found that while LLMs could: help creativity, provide academic support when teachers were unavailable, and model certain types of writing well; teachers were also aware of LLMs potential to limit students' learning in certain ways through over-reliance.
 Beyond academic integrity concerns, teachers had similar concerns to @Valenzuela_2024 about de-skilling and an overall loss of agency in writing and critical thinking.
 
-@Cahill_2024 surveyed undergraduate Political Science students on their attitudes toward and usage of AI tools.
-They found that the use of ChatGPT (among other machine-learning-powered software) was widespread.
-However, they also found that many students lacked the confidence in using AI for academic purposes - in particular, only 11% 'strongly agreed' that they know how to use AI to improve their writing.
-Like educators,  students have nuanced views on appropriate AI use.
-In particular, respondents found that using it to writing whole papers as inappropriate, while using it for basic tasks like general assistance, writing feedback and basic data visualization was perceived more appropriately.
-Interestingly, first-generation college students were found to be more likely to use AI in their work, particularly in writing papers and helping with assignments.
-Their findings suggest that LLMs and other AI tools could be an equalizer for disadvantaged students.
-Rhough only political science students were surveyed, statistics and data science students could reasonably have general attitudes that are similar.
+@Cahill_2024 surveyed undergraduate political science students on their attitudes toward and usage of AI tools.
+They found that the use of ChatGPT was widespread.
+However, they also found that many students lacked confidence in using AI for academic purposes.
+In particular, only 11% 'strongly agreed' that they know how to use AI to improve their writing.
+Students had nuanced views on appropriate AI use.
+In particular, respondents found that using it to write whole papers as inappropriate, while using it for basic tasks like general assistance, writing feedback and basic data visualization was perceived more appropriately.
 
-As we have seen, the existing inquiry by educators has explored the general qualitative usage and student perceptions on these tools.
-Though informative, a key question still remains: how can students' academic performance be affected, precisely, through allowing them to use LLMs in classwork?
+To understand the current state of LLMs as a tool for trustworthy data science, this paper focuses on the association between student academic performance and their LLM usage.
+Specifically, we examine the relationship between students' grades and self-reported measures of student LLM usage, as well as student attitudes toward LLMs in general.
+This is based on final papers and a survey, conducted in a third-year undergraduate data science course at the University of Toronto.
+By examining how students interact with and perceive LLMs as tools, and how these variables translate into student outcomes, the effects of LLM integration in data science can be more precisely determined and recommendations made for future development.
 
-This paper aims to fill this gap by quantitatively investigating the relationship between students' grades and measures of student LLM usage and their attitudes toward LLMs in general, using evidence from a third-year undergraduate statistics course at the University of Toronto.
-Through examining how students interact with and perceive LLM tools and how these variables translate into student outcomes, the effects of LLM integration in data science education can be more precisely determined.
-
-The remainder of this paper is structured as follows: @sec-data visualizes and analyzes survey data and coursework from students; @sec-model models the unstructured data to approximate a relationship between grades and usage/attitudes; @sec-results describes and analyzes the model's results; @sec-discussion discusses the implications of the findings for data science education and future research and practice in this rapidly evolving field.
+The remainder of this paper is structured as follows: @sec-data visualizes and analyzes survey data and coursework from students.
+@sec-model specifies a model used to investigate the relationship.
+@sec-results describes and analyzes the model's results.
+@sec-discussion discusses the implications of the findings for data science education and future research and practice at the intersection of LLMs and trustworthy data science.
 
@@
 
-## Background and Collection
+## Background
 
-To investigate students' usage and attitudes toward LLM use and how they related to their academic performance, a dataset containing their usage/attitudes, coursework, and academic performance was constructed.
-Data was sampled from the cohort of students taking STA302 - Methods of Data Analysis I, taught by one of the investigators in the Winter 2024 semester at the University of Toronto.
-By virtue of pre-requisites needed, this restricted the data collected to be only on upper-year undergraduate students.
+To investigate students' usage and attitudes towards LLMs and how they related to their academic performance, a dataset containing their usage/attitudes, coursework, and academic performance was constructed.
+This was based on three components:
+
+1. an optional survey; 
+2. self-reported LLM usage; and 
+3. student marks on the final paper.
+
+All data are from the cohort of students taking STA302 "Methods of Data Analysis I" in the Winter 2024 semester at the University of Toronto.
+This course had 275 students initially enrolled which, reflecting a normal rate of attrition for undergraduate statistics courses at the University of Toronto, reduced to 154 students by the end of the semester.
+Assessment was heavily based on three papers submitted over the course of the 12 week semester.
+
+The student marks that we analyze are based only on the final paper, which is done individually.
+By this stage, uninterested students have typically dropped the course, and students are familiar with course expectations.
+A typical paper submission is 10-20 pages, and requires students to conduct original research to answer a research question of interest to them.
+It reflects the skills typically used by a professional data scientist.
+Students are expected to develop a research question of interest to them, identify or collect data to answer the question, conduct statistical analysis, and write a short paper.
+Examples of final papers (shared with consent) include: @hannahyu; @emilysu; and @bennyrochwerg.
+
+By the time they are working on their final paper, students have submitted and received feedback on two previous papers with similar requirements and rubrics to that of the final paper.
+Each paper has the same basic structure and expectations.
+Before the final paper is due students have received feedback on all their previous work in the class (including their past papers) and there is an, optional, two-day period of peer review.
+
+By virtue of pre-requisites of this course the typical student is an upper-year undergraduate.
+Coding and writing are major parts of the course.
+Students are welcome to use R or Python, but the majority code in R because that is the programming language currently mostly taught in pre-requisite courses.
+All writing must be in English.
+The aim of the course for students to create a public portfolio of work they can use to apply for jobs.
+
+Throughout the semester students were encouraged to use LLMs.
+Formal instruction was provided twice during the semester, with a masterclass taught by a computer science faculty member on the ethics of using LLMs (see @Horton2024 for details), and another masterclass taught by a TA on writing with LLMs.
 
 Data was collected from students through an optional end-of-course survey.
-Whether or not they consented to their data being used in this investigation, all respondents received a +1% increase in their final course grade for their participation.
-All consenting responses were then cross-referenced to their course grade, as well as the GitHub account they used to complete their course research papers.
-The responses were finally anonymized by removing any personal references to the students themselves.
+[Appendix -@sec-survey-details] provides all the questions asked in the survey.
+Whether or not they consented to their data being used, all respondents received a 1% increase in their final course grade for their participation.
+Consenting responses were then matched to their final paper mark, as well as the GitHub repository for their final paper.
+The responses were anonymized by removing any personal references to the students themselves including names, emails, student numbers, and GitHub links.
 
-## Cleaning and Variables of Interest
+All data cleaning and analysis was done using the `R` statistical programming language [@citeR], and especially uses the `tidyverse` [@tidyverse], `janitor` [@janitor], `reshape2` [@reshape], and `readxl` [@readxl] libraries.
 
-All the data was cleaned using R [@citeR] and it's tidyverse [@tidyverse], janitor [@janitor] and stringr [@stringr], then tested for issues using the testthat package [@testthat].
-This led to a final dataset containing 121 unique responses to 18  survey questions.
+## Survey data
 
-```{r fig-boilerplate}
+There were 146 responses to the survey.
+Of these, 119 respondents provided authorization for their data to be collected and used.
+Four of those respondents submitted the survey twice, and after removing their second response, 115 responses remained.
+Of those, 15 respondents did not include a statement on LLM usage in the README of the GitHub repository of their final paper, leaving 100 responses that were of use and were merged based on student name.
+
+All but one respondent completed the survey within 45 minutes (@fig-response-time-and-gpa-1).
+That one respondent took more than 4,000 minutes to complete the survey, suggesting they took a break while filling it out.
+That respondent is included in the analysis dataset, but without that respondent, the average time to complete the survey was 11 minutes, and the standard deviation was 8 minutes.
+
+```{r fig-response-time-and-gpa}
 #| message: false
@@
 #| warning: false
-#| fig-cap: (CHANGE) Boilerplate figure caption 
-#| fig-align: center
-# survey_data |>
-#   ggplot(program_counts, aes(x = "", y = n, fill = program_of_study)) +
-#   #put geom type here (put before compilation) +
-#   labs(title = "Title!",
-#        x = "X label",
-#        y = "Y label") +
-#   theme_minimal() +
-#   scale_fill_discrete(name="Legend Title") +
-#   theme(plot.title.position = "plot",
-#         plot.title = element_text(hjust = 0.5)
-#         )
+#| fig-cap: Distribution of respondents' survey response times and GPA
+#| fig-subcap: ["Distribution of survey response times","\"What is your GPA?\""]
+#| layout-ncol: 1
+
+all_data |>
+  mutate(mins_complete = completion_time - start_time) |> 
+  filter(mins_complete < 1000) |> # Remove one student who took 4,000 minutes to complete
+  # summarize(mean = mean(mins_complete),
+            # standard_dev = sd(mins_complete))
+  ggplot(aes(x = mins_complete)) +
+  geom_histogram(binwidth = 1) +
+  labs(
+    x = "Time to complete survey (minutes)",
+    y = "Frequency"
+  ) +
+  theme_minimal()
+
+all_data |>
+  # summarize(mean = mean(what_is_your_gpa, na.rm = TRUE),
+  #           standard_dev = sd(what_is_your_gpa, na.rm = TRUE))
+  ggplot(aes(x = what_is_your_gpa)) +
+  geom_histogram(binwidth = 0.1) +
+  labs(
+    x = "Self-reported GPA",
+    y = "Frequency"
+  ) +
+  theme_minimal()
 ```
 
-```{r tbl-boilerplate}
-#| message: false
-#| echo: false
-#| warning: false
-#| tbl-cap: (CHANGE) Boilerplate table caption 
-#| fig-align: center
-#Put code below this line
-```
+There is a wide distribution of self-reported GPAs (@fig-response-time-and-gpa-2).
+Seven of the 100 respondents included in the analysis dataset did not report their GPA.
+The majority of responses cluster around a B (3.0/4.0), and the average is 3.06 with a standard deviation of 0.55.
+One factor that may affect the range is that the course is required for programs in the Statistics, Mathematics and Computer Science Departments.
+Self-reported GPAs introduce the possibility of reporting bias.
+For instance, respondents may have provided their cumulative GPA, their most recent term's GPA, or could have misreported it entirely.
 
-## General Information
-
-First, general information about the sample of students was derived first, through visualizing the basic data about them. This was done using the ggplot2 [@ggplot2] package.
-
-```{r fig-response-time-hist}
-#| message: false
-#| echo: false
-#| warning: false
-#| fig-cap: Distribution of Survey Response Times
-#| fig-align: center
-
-# Trim values to histogram
-
-plot_data <- survey_data |>
-  filter(mins_complete < 1000)
-
-plot_data |>
-  ggplot(aes(x = mins_complete)) +
-  geom_histogram(binwidth = 2, fill = "lightblue", color = "black") +
-  labs(
-    title = "Distribution of Survey Response Times",
-    x = "Time (Mins)",
-    y = "Frequency"
-  ) +
-  theme_minimal() +
-  theme(
-    plot.title.position = "plot",
-    plot.title = element_text(hjust = 0.5)
-  )
-```
-@fig-response-time-hist shows the distribution of response times to the end-of-course survey.
-Respondents who consented generally took between 0-40 mins to answer all of the questions, with most response times ranging between 2-10 minutes.
-There was one consenting respondent who took 4308 minutes, whose observation was dropped from the visualization.
-Overall, however, this suggests that students generally took a meaningful amount of time and thought to engage with the survey's questions.
-
-```{r fig-gpa-hist}
-#| message: false
-#| echo: false
-#| warning: false
-#| fig-cap: Distribution of Student GPA
-#| fig-align: center
-
-# Histogram of GPA
-
-survey_data |>
-  ggplot(aes(x = gpa)) +
-  geom_histogram(binwidth = 0.1, fill = "lightblue", color = "black") +
-  labs(
-    title = "Distribution of Student GPA",
-    x = "GPA",
-    y = "Frequency"
-  ) +
-  theme_minimal() +
-  theme(
-    plot.title.position = "plot",
-    plot.title = element_text(hjust = 0.6)
-  )
-```
-
-@fig-gpa-hist shows a wide distribution of student GPAs, with the majority clustering around a B (3.0/4.0) average.
-One factor possibly affecting the range is the course being a required course for programs in the Statistics, Mathematics and Computer Science departments.
-Though programs in these departments generally attract high-achieving students, some may still struggle.
-Another factor is that these GPAs were self-reported, which introduces some ambiguity - students may have provided either their cumulative GPA or their most recent term's GPA, or could have misreported it entirely.
-Despite these two factors, the diversity in reported grades suggests the course attracted students across a broad spectrum of academic performance levels.
+Students from a range of years took the course, however the majority of respondents were in their 3rd or 4th year of study (@tbl-year-dist).
+The course's prerequisite of general statistics, which is a two-course sequence typically completed by students in their second year, would make it difficult to take this course earlier than 3rd year.
 
@@
 #| warning: false
-#| tbl-cap: Count of Students by Year
-#| fig-align: center
+#| tbl-cap: \"What year are you?\"
 
-# Change PEY to 4th year
-survey_data <- survey_data |>
+all_data |> 
   mutate(
-    year = case_when(
-      year == "PEY" ~ "4th",
-      year == "5th" ~ "5th+",
-      year == "6th" ~ "5th+",
-      TRUE ~ year
+    what_year_are_you = case_when(
+      what_year_are_you == "PEY" ~ "4th", # Change PEY to 4th year
+      what_year_are_you == "5th" ~ "5th or over",
+      what_year_are_you == "6th" ~ "5th or over",
+      TRUE ~ what_year_are_you
+    )
+  ) |> 
+  count(what_year_are_you) |>
+  pivot_wider(names_from = what_year_are_you, 
+              values_from = n, 
+              values_fill = list(n = 0)) |> 
+  kable()
+```
+
+Respondents had a varied self-perception of their coding and writing abilities (@fig-selfperception).
+Most respondents believed it is important to be good at writing, but many are either indifferent or do not like to write (@fig-selfperception-1).
+Respondents also do not find writing to be particularly easy, which could be associated with the reported relative antipathy toward writing.
+Most respondents were at least somewhat confident in their own writing abilities, but a substantial contingent felt otherwise.
+Although few respondents felt that they were confident in their writing ability, more felt that they were able to catch their mistakes, which could indicate a disconnect between how respondents perceive their work and how the work was evaluated.
+
+```{r fig-selfperception}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: Self-perception of coding and writing abilities
+#| fig-subcap: ["\"Please rate how much each statement describes you, on a scale from 'This is very different to me' to 'This is a lot like me'\"", "\"Please rate how much each statement describes you, on a scale from 'This is very different to me' to 'This is a lot like me'\""]
+
+# Writing
+writing_perceptions_long <- 
+  all_data |>
+  select(
+    writing_is_easy_for_me, 
+    i_like_to_write, 
+    i_believe_it_is_important_to_be_a_good_writer,
+    when_i_edit_it_is_easy_for_me_to_catch_my_mistakes,
+    i_feel_confident_sharing_my_writing,
+    i_am_confident_in_my_overall_writing_ability
+  ) |> 
+  pivot_longer(cols = everything(), names_to = "statements", values_to = "response") |>
+  mutate(
+    statements = case_when(
+      statements == "writing_is_easy_for_me" ~ "\"Writing is easy for me.\"",
+      statements == "i_like_to_write" ~ "\"I like to write.\"",
+      statements == "i_believe_it_is_important_to_be_a_good_writer" ~ "\"I believe it is important\nto be a good writer.\"",
+      statements == "when_i_edit_it_is_easy_for_me_to_catch_my_mistakes" ~ "\"When I edit it is easy for\nme to catch my mistakes.\"",
+      statements == "i_feel_confident_sharing_my_writing" ~ "\"I feel confident\nsharing my writing.\"",
+      statements == "i_am_confident_in_my_overall_writing_ability" ~ "\"I am confident in my\noverall writing ability.\"",
+      TRUE ~ statements
+    ),
+    response = case_when(
+      response == "\"This is a lot like me\"" ~ "A lot like me",
+      response == "\"This is very different to me\"" ~ "Very different from me",
+      response == "\"This somewhat describes me\"" ~ "Somewhat describes me",
+      TRUE ~ response
     )
@@
 
-year_count <- survey_data |>
-  count(year) |>
-  pivot_wider(names_from = year, values_from = n, values_fill = list(n = 0))
+writing_perceptions_long |>
+  ggplot(aes(x = response, fill = response)) +
+  geom_bar(position = "stack", stat = "count") +
+  facet_wrap(vars(statements)) +
+  labs(
+    x = "Response",
+    y = "Number of repondents",
+    fill = "Response:"
+  ) +
+  scale_fill_brewer(palette = "Set1") +
+  theme_minimal() +
+  theme(
+    axis.text.x = element_blank(),
+    legend.position = "bottom"
+  )
 
-# Display the original wide format table
-kable(year_count)
+# Coding
+coding_perceptions_long <- 
+  all_data |>
+  select(
+    coding_is_easy_for_me,
+    i_like_to_code,
+    i_believe_it_is_important_to_be_a_good_coder,
+    when_i_check_my_code_it_is_easy_for_me_to_catch_my_mistakes,
+    i_feel_confident_sharing_my_code,
+    i_am_confident_in_my_overall_coding_ability
+  ) |> 
+  pivot_longer(cols = everything(), names_to = "statements", values_to = "response") |>
+  mutate(
+    statements = case_when(
+      statements == "coding_is_easy_for_me" ~ "\"Coding is easy for me\"",
+      statements == "i_like_to_code" ~ "\"I like to code\"",
+      statements == "i_believe_it_is_important_to_be_a_good_coder" ~ "\"I believe it is important\n to be a good coder\"",
+      statements == "when_i_check_my_code_it_is_easy_for_me_to_catch_my_mistakes" ~ "\"When I check my code, it is\n easy for me to catch my\nmistakes\"",
+      statements == "i_feel_confident_sharing_my_code" ~ "\"I feel confident\n sharing my code\"",
+      statements == "i_am_confident_in_my_overall_coding_ability" ~ "\"I am confident in\n my overall coding ability\"",
+      TRUE ~ statements
+    ),
+    response = case_when(
+      response == "\"This is a lot like me\"" ~ "A lot like me",
+      response == "\"This is very different to me\"" ~ "Very different from me",
+      response == "\"This somewhat describes me\"" ~ "Somewhat describes me",
+      TRUE ~ response
+    )
+  )
+
+coding_perceptions_long |>
+  ggplot(aes(x = response, fill = response)) +
+  geom_bar(position = "stack", stat = "count") +
+  facet_wrap(vars(statements)) +
+  labs(
+    x = "Response",
+    y = "Number of repondents",
+    fill = "Response:"
+  ) +
+  scale_fill_brewer(palette = "Set1") +
+  theme_minimal() +
+  theme(
+    axis.text.x = element_blank(),
+    legend.position = "bottom"
+  )
 ```
 
-We can also see in from @tbl-year-dist that students from a range of years took the course.
-That said, the majority of students were in their 3rd or 4th year of study.
-This makes sense, given the course's prerequisite of general statistics, which is a two-semester sequence typically done in second year statistics.
-This also explains the lack of any first year students in the course as well.
+Respondent self-perceptions regarding coding proficiency and importance varied (@fig-selfperception-2).
+There was strong consensus on the perceived importance of coding skills, with a majority of respondents strongly affirming this belief.
+However, respondents' self-assessed coding abilities and enjoyment are more heterogeneous, with a substantial proportion reporting moderate rather than high levels of ease and enjoyment in coding tasks.
+These findings suggest a complicated relationship between respondents' recognition of coding's importance and their personal experience with coding.
 
-```{r fig-programs-pie}
+Overall there was a moderate level of confidence among respondents in their overall coding ability, willingness to share code, and capacity to identify errors (@fig-selfperception-2).
+Notably, respondents express slightly higher confidence in detecting their own coding mistakes compared to general coding ability or code sharing.
+These patterns suggest that while respondents have developed some coding self-efficacy, there is still considerable potential for enhancing their perceived competence and comfort across various coding-related activities.
+
+The majority of respondents were at least "somewhat familiar" with generative AI such as ChatGPT (@tbl-ai-familiarity-usage-1).
+Though respondents' self-perceptions around writing and coding are varied, there was strong consensus that the use of generative AI tools is appropriate within an academic setting (@tbl-ai-familiarity-usage-2).
+Most respondents who responded "It depends" generally found artificial intelligence tools to be appropriate, though with certain guidelines and rules governing their use.
+
+```{r tbl-ai-familiarity-usage}
 #| message: false
@@
 #| warning: false
-#| tbl-cap: Student Distribution by Program of Study
-#| fig-align: center
+#| tbl-cap: Familiarity with, and appropriateness of, generative AI
+#| tbl-subcap: ["'How familiar are you with using generative AI tools such as OpenAI's ChatGPT or equivalents?'","'To what extent do you think using generative AI tools such as ChatGPT by OpenAI (or equivalents) is ethical and appropriate for schoolwork?'"]
 
-# Pie chart of program of study
-
-program_counts <- survey_data |>
-  count(program_of_study)
-
-ggplot(program_counts, aes(x = "", y = n, fill = program_of_study)) +
-  geom_bar(stat = "identity", width = 1) +
-  coord_polar("y") +
-  labs(
-    title = "Distribution of Students by Program of Study",
-    subtitle = "Visualizing the proportion of students across different programs",
-    x = NULL,
-    y = NULL
-  ) +
-  theme_minimal() +
-  scale_fill_discrete(name = "Program of Study") +
-  theme(
-    axis.text.x = element_blank(), # Remove x-axis text
-    axis.title.x = element_blank(), # Remove x-axis title
-    plot.title = element_text(hjust = 0.5),
-    plot.subtitle = element_text(hjust = 0.5)
-  ) +
-  geom_text(aes(label = n),
-    position = position_stack(vjust = 0.5)
-  )
-```
-
-Programs of study [did something; waiting on major/minor data], as we can see in @fig-programs-pie...
-
-## Student Attitudes
-
-```{r fig-ai-familiarity}
-#| message: false
-#| echo: false
-#| warning: false
-#| fig-cap: Student Familiarity with AI
-#| fig-align: center
-
-table_data <- survey_data |>
-  select(ai_familiarity) |>
+all_data |>
+  rename(ai_familiarity = how_familiar_are_you_with_using_generative_ai_tools_such_as_open_a_is_chat_gpt_or_equivalents) |> 
   mutate(
@@
   ) |>
-  count(ai_familiarity)
+  count(ai_familiarity) |> 
+  kable(col.names = c("Extent of AI familiarity", "Number"))
 
-# Custom colors
-familiarity_colors <- c(
-  "Not familiar" = "tomato",
-  "Somewhat familiar" = "lightblue",
-  "Very familiar" = "lightgreen"
-)
-
-table_data |>
-  ggplot(aes(x = "", y = n, fill = ai_familiarity)) +
-  geom_bar(stat = "identity", width = 1) +
-  coord_polar("y") +
-  labs(
-    title = "Student Familiarity with AI",
-    x = NULL,
-    y = NULL
-  ) +
-  theme_minimal() +
-  scale_fill_manual(name = "Familiarity Level", values = familiarity_colors) +
-  theme(
-    axis.text.x = element_blank(), # Remove x-axis text
-    axis.title.x = element_blank(), # Remove x-axis title
-    plot.title = element_text(hjust = 0.5),
-    plot.subtitle = element_text(hjust = 0.5)
-  ) +
-  geom_text(aes(label = n),
-    position = position_stack(vjust = 0.5)
-  )
+all_data |>
+  rename(ai_schoolwork_appropriate = to_what_extent_do_you_think_using_generative_ai_tools_such_as_chat_gpt_by_open_ai_or_equivalents_is_ethical_and_appropriate_for_schoolwork) |> 
+  mutate(
+    ai_schoolwork_appropriate = case_when(
+      ai_schoolwork_appropriate == "Depends of what the main goal of schoolwork is" ~ "It depends",
+      ai_schoolwork_appropriate == "I believe student can only use it if they are allowed to" ~ "It depends",
+      ai_schoolwork_appropriate == "a controllable amount" ~ "It depends",
+      ai_schoolwork_appropriate == "appropriate depending on the context" ~ "It depends",
+      ai_schoolwork_appropriate == "appropriate with rules for usage" ~ "It depends",
+      ai_schoolwork_appropriate == "depends" ~ "It depends",
+      ai_schoolwork_appropriate == "depends on how it is used" ~ "It depends",
+      ai_schoolwork_appropriate == "half and half" ~ "It depends",
+      ai_schoolwork_appropriate == "unsure" ~ "It depends",
+      TRUE ~ ai_schoolwork_appropriate
+      )) |> 
+  count(ai_schoolwork_appropriate) |> 
+  kable(col.names = c("Ethical and appropriate for school?", "Number"))
 ```
 
-Just as @Cahill_2024 found, the vast majority of students taking the course were at least "somewhat familiar" with ChatGPT and other LLMs.
-This overrepresentation of familiarity also primarily driven by the fact that the course is primarily targeted for students in Statistics, Computer Science, Economics, Mathematics and other quantitative disciplines, possibly leading to a sample that is skewed to being more interested in computing.
+To understand the role of LLMs in learning, students more granularly identified their usage by checking off various pre-defined use cases in the survey (@fig-AI-use-and-usefulness-1).
+Technical questions and explaining concepts were the two top use cases among students in the course.
+More than half of students also used LLMs for quick questions, general knowledge, writing paper code, and checking solutions.
+Just less than half used it to write paper content, which could suggest that students do not feel confident using it to improve writing.
 
-Students were then asked about their attitudes and self-perception on writing and coding.
-
-```{r fig-writing-multibar-1}
+```{r fig-AI-use-and-usefulness}
 #| message: false
@@
 #| warning: false
-#| tbl-cap: Student Self-Perceptions on Writing, Part 1
-#| fig-align: center
+#| fig-cap: "Use and usefulness of generative AI"
+#| fig-subcap: ["\"If you have used generative AI tools such as OpenAI's ChatGPT or equivalents, in what ways have you used it (select all that apply)?\"", "\"How helpful did you find generative AI tools such as ChatGPT by OpenAI (or equivalents) for each component of STA302?\""]
 
-# Create a multi-bar plot for writing perceptions grouped by response
-
-# Select first three statements
-writing_perceptions <- survey_data |>
-  select(
-    writing_is_easy_for_me,
-    i_like_to_write,
-    i_believe_it_is_important_to_be_a_good_writer
-  )
-
-writing_perceptions_long <- writing_perceptions |>
-  pivot_longer(cols = everything(), names_to = "Statements", values_to = "Response") |>
+# Use
+all_data |>
+  rename(how_have_you_used = if_you_have_used_generative_ai_tools_such_as_open_a_is_chat_gpt_or_equivalents_in_what_ways_have_you_used_it_select_all_that_apply) |> 
+  select(how_have_you_used) |> 
+  separate_rows(how_have_you_used, sep = ";") |> 
   mutate(
-    Statements = case_when(
-      Statements == "writing_is_easy_for_me" ~ "\"Writing is easy for me\"",
-      Statements == "i_like_to_write" ~ "\"I like to write\"",
-      Statements == "i_believe_it_is_important_to_be_a_good_writer" ~ "\"I believe it is important\nto be a good writer\"",
-      TRUE ~ Statements
-    ),
-    Response = case_when(
-      Response == "This is a lot like me" ~ "A lot like me",
-      Response == "This is very different to me" ~ "Very different from me",
-      Response == "This somewhat describes me" ~ "Somewhat describes me",
+    how_have_you_used = case_when(
+      how_have_you_used == "check grammar mistake" ~ "Other",
+      how_have_you_used == "debugging code" ~ "Other",
+      how_have_you_used == "mental health support" ~ "Other",
+      how_have_you_used == "Making practice questions for me with increasing difficulty" ~ "Other",
+      how_have_you_used == "Generating test questions" ~ "Other",
+      TRUE ~ how_have_you_used
     )
   ) |> 
-  filter(!is.na(Response))
+  count(how_have_you_used) |>
+  filter(how_have_you_used != "") |> 
+  arrange(desc(n)) |> 
+  ggplot(aes(x = n, y = reorder(how_have_you_used, n))) +
+  geom_bar(stat = "identity") +
+  labs(x = "Number of respondents", y = "Type of usage") +
+  theme_minimal()
 
-# Custom colors
-response_colors <- c(
-  "A lot like me" = "lightgreen",
-  "Somewhat describes me" = "lightblue",
-  "Very different from me" = "tomato"
-)
-
-# Taken from: https://stackoverflow.com/questions/52297978/decrease-overal-legend-size-elements-and-text
-addSmallLegend <- function(myPlot, pointSize = 0.5, textSize = 3, spaceLegend = 0.1) {
-  myPlot +
-    guides(
-      shape = guide_legend(override.aes = list(size = pointSize)),
-      color = guide_legend(override.aes = list(size = pointSize)),
-      fill = guide_legend(title = "Response", override.aes = list(size = pointSize))
-    ) +
-    theme(
-      legend.title = element_text(size = textSize),
-      legend.text = element_text(size = textSize),
-      legend.key.size = unit(spaceLegend, "lines")
-    )
-}
-
-# Make the plot
-multi_bar <- writing_perceptions_long |>
-  ggplot(aes(x = Response, fill = Response)) +
-  geom_bar(position = "dodge", stat = "count") +
-  facet_wrap(~Statements) +
-  labs(
-    title = "Student Self-Perceptions on Writing",
-    subtitle = "Responses to statements about writing ease, enjoyment, and importance",
-    x = "Response",
-    y = "Number of Students"
-  ) +
-  scale_fill_manual(values = response_colors, name = "Response") + # Using custom colors
-  theme_minimal() +
-  theme(
-    plot.title.position = "plot",
-    plot.title = element_text(hjust = 0.5),
-    plot.subtitle = element_text(hjust = 0.5),
-    axis.text.x = element_blank(),
-    legend.position = "bottom",
-    legend.title = element_text(size = 10),
-    legend.text = element_text(size = 8)
-  )
-
-multi_bar <- addSmallLegend(multi_bar, pointSize = 0.5, textSize = 9, spaceLegend = 0.8)
-
-multi_bar
-```
-
-```{r fig-writing-multibar-2}
-#| message: false
-#| echo: false
-#| warning: false
-#| tbl-cap: Student Self-Perceptions on Writing, Part 2
-#| fig-align: center
-
-# Create a multi-bar plot for writing perceptions grouped by response
-
-# Select last three statements
-writing_perceptions <- survey_data |>
+# Usefulness
+all_data |>
   select(
-    when_i_edit_it_is_easy_for_me_to_catch_my_mistakes,
-    i_feel_confident_sharing_my_writing,
-    i_am_confident_in_my_overall_writing_ability
-  )
-
-writing_perceptions_long <- writing_perceptions |>
-  pivot_longer(cols = everything(), names_to = "Statements", values_to = "Response") |>
+    weekly_quiz,
+    weekly_mini_essay,
+    papers_generating_ideas,
+    papers_writing_code,
+    papers_writing_content,
+    papers_improving_content
+  ) |>
+  pivot_longer(cols = everything(),
+               names_to = "task",
+               values_to = "response") |>
   mutate(
-    Statements = case_when(
-      Statements == "when_i_edit_it_is_easy_for_me_to_catch_my_mistakes" ~ "\"When I edit, it is easy to\ncatch my own mistakes\"",
-      Statements == "i_feel_confident_sharing_my_writing" ~ "\"I feel confident\nsharing my writing\"",
-      Statements == "i_am_confident_in_my_overall_writing_ability" ~ "\"I am confident in\nmy own writing ability\"",
-      TRUE ~ Statements
+    # Simplify response categories and rename tasks
+    response_simplified = case_when(
+      response %in% c("\"Not helpful\"", "\"I did not use generative AI for this component\"") ~ "Less Helpful",
+      response %in% c("\"Somewhat helpful\"", "\"Very helpful\"") ~ "More Helpful",
+      TRUE ~ "Unknown"  # Handle unexpected values
     ),
-    Response = case_when(
-      Response == "This is a lot like me" ~ "A lot like me",
-      Response == "This is very different to me" ~ "Very different from me",
-      Response == "This somewhat describes me" ~ "Somewhat describes me",
+    task = case_when(
+      task == "weekly_quiz" ~ "Weekly Quiz",
+      task == "weekly_mini_essay" ~ "Weekly Mini Essay",
+      task == "papers_generating_ideas" ~ "Generating Ideas",
+      task == "papers_writing_code" ~ "Writing Code",
+      task == "papers_writing_content" ~ "Writing Content",
+      task == "papers_improving_content" ~ "Improving Content",
+      TRUE ~ "Other"  # Handle any unexpected task names
     )
   ) |> 
-  filter(!is.na(Response))
-
-# Custom colors
-response_colors <- c(
-  "A lot like me" = "lightgreen",
-  "Somewhat describes me" = "lightblue",
-  "Very different from me" = "tomato"
+  group_by(task, response_simplified) |>
+  summarise(count = n(), .groups = 'drop') |> 
+  ggplot(aes(x = response_simplified, y = count, fill = response_simplified)) +
+  geom_bar(stat = "identity") +
+  facet_wrap(~ task, scales = "free_y") +
+  labs(y = "Number of respondents",
+       x = "",
+       fill = "Response type: ") +
+  theme_minimal() +
+  coord_flip() + 
+  scale_fill_brewer(palette = "Set1") +
+  theme(
+    axis.text.y = element_blank(),
+    legend.position = "bottom",
   )
-
-# Taken from: https://stackoverflow.com/questions/52297978/decrease-overal-legend-size-elements-and-text
-addSmallLegend <- function(myPlot, pointSize = 0.5, textSize = 3, spaceLegend = 0.1) {
-  myPlot +
-    guides(
-      shape = guide_legend(override.aes = list(size = pointSize)),
-      color = guide_legend(override.aes = list(size = pointSize)),
-      fill = guide_legend(title = "Response", override.aes = list(size = pointSize))
-    ) +
-    theme(
-      legend.title = element_text(size = textSize),
-      legend.text = element_text(size = textSize),
-      legend.key.size = unit(spaceLegend, "lines")
-    )
-}
-
-# Make the plot
-multi_bar2 <- writing_perceptions_long |>
-  ggplot(aes(x = Response, fill = Response)) +
-  geom_bar(position = "dodge", stat = "count") +
-  facet_wrap(~Statements) +
-  labs(
-    title = "Student Self-Perceptions on Writing",
-    subtitle = "Responses to statements about confidence and editing in writing",
-    x = "Response",
-    y = "Number of Students"
-  ) +
-  scale_fill_manual(values = response_colors, name = "Response") + # Using custom colors
-  theme_minimal() +
-  theme(
-    plot.title.position = "plot",
-    plot.title = element_text(hjust = 0.5),
-    plot.subtitle = element_text(hjust = 0.5),
-    axis.text.x = element_blank(), # Remove x-axis labels
-    legend.position = "bottom",
-    legend.title = element_text(size = 10),
-    legend.text = element_text(size = 8)
-  )
-
-multi_bar2 <- addSmallLegend(multi_bar2, pointSize = 0.5, textSize = 9, spaceLegend = 0.8)
-
-multi_bar2
 
@@
 
-Students had varied attitudes toward their writing and their writing abilities in general, as we can see from @fig-writing-multibar-1 and fig-writing-multibar-2. ...
+Respondents were also asked to rate the helpfulness of LLMs on various tasks assigned during the course on a 4-point scale of "Did not use" to "Very Helpful" (@fig-AI-use-and-usefulness-2).
+To simplify the presentation, responses were grouped into two main categories: "Less Helpful" and "More Helpful." The "Less Helpful" category combines responses where students found the AI either "Not helpful" or did not use it for the task, while the "More Helpful" category includes responses where the AI was considered "Somewhat helpful" or "Very helpful".
 
-```{r fig-coding-multibar-1}
-#| message: false
-#| echo: false
-#| warning: false
-#| tbl-cap: Student Self-Perceptions on Coding, Part 1
-#| fig-align: center
+Respondents differed in terms of how they used LLMs in the course (@fig-AI-use-and-usefulness-2).
+While most tasks were roughly split between respondents finding LLMs helpful or not, respondents found them most helpful in generating code.
+In the context of the course, this meant generating R code for transforming, analyzing, and visualizing data.
+To a lesser extent, respondents also found LLMs to be helpful in improving the existing writing they had, while at the same time not favouring it for writing content from scratch.
 
-# Create a multi-bar plot for coding self-perceptions grouped by response
-
-# Select first three statements
-writing_perceptions <- survey_data |>
-  select(
-    coding_is_easy_for_me,
-    i_like_to_code,
-    i_believe_it_is_important_to_be_a_good_coder
-  )
-
-writing_perceptions_long <- writing_perceptions |>
-  pivot_longer(cols = everything(), names_to = "Statements", values_to = "Response") |>
-  mutate(
-    Statements = case_when(
-      Statements == "coding_is_easy_for_me" ~ "\"Coding is easy for me\"",
-      Statements == "i_like_to_code" ~ "\"I like to code\"",
-      Statements == "i_believe_it_is_important_to_be_a_good_coder" ~ "\"I believe it is important\n to be a good coder\"",
-      TRUE ~ Statements
-    ),
-    Response = case_when(
-      Response == "This is a lot like me" ~ "A lot like me",
-      Response == "This is very different to me" ~ "Very different from me",
-      Response == "This somewhat describes me" ~ "Somewhat describes me",
-    )
-  ) |>
-  filter(!is.na(Response))
-
-# Taken from: https://stackoverflow.com/questions/52297978/decrease-overal-legend-size-elements-and-text
-addSmallLegend <- function(myPlot, pointSize = 0.5, textSize = 3, spaceLegend = 0.1) {
-  myPlot +
-    guides(
-      shape = guide_legend(override.aes = list(size = pointSize)),
-      color = guide_legend(override.aes = list(size = pointSize))
-    ) +
-    theme(
-      legend.title = element_text(size = textSize),
-      legend.text = element_text(size = textSize),
-      legend.key.size = unit(spaceLegend, "lines")
-    )
-}
-
-# Custom colors
-response_colors <- c(
-  "A lot like me" = "lightgreen",
-  "Somewhat describes me" = "lightblue",
-  "Very different from me" = "tomato"
-)
-
-# Make the plot
-multi_bar <- writing_perceptions_long |>
-  ggplot(aes(x = Response, fill = Statements)) +
-  geom_bar(position = "dodge", stat = "count") +
-  labs(
-    title = "Student Self-Perceptions of Coding",
-    x = "Response",
-    y = "Number of Students"
-  ) +
-  theme_minimal() +
-  theme(
-    plot.title.position = "plot",
-    plot.title = element_text(hjust = 0.5),
-    legend.position = "bottom",
-    axis.text.x = element_blank(),
-  )
-
-multi_bar <- addSmallLegend(multi_bar, pointSize = 0.5, textSize = 10, spaceLegend = 0.8)
-
-multi_bar
-```
-
-```{r fig-coding-multibar-2}
-#| message: false
-#| echo: false
-#| warning: false
-#| tbl-cap: Student Self-Perceptions on Coding, Part 2
-#| fig-align: center
-
-# Create a multi-bar plot for coding self-perceptions grouped by response
-
-# Select first three statements
-writing_perceptions <- survey_data |>
-  select(
-    when_i_check_my_code_it_is_easy_for_me_to_catch_my_mistakes,
-    i_feel_confident_sharing_my_code,
-    i_am_confident_in_my_overall_coding_ability
-  )
-
-writing_perceptions_long <- writing_perceptions |>
-  pivot_longer(cols = everything(), names_to = "Statements", values_to = "Response") |>
-  mutate(
-    Statements = case_when(
-      Statements == "when_i_check_my_code_it_is_easy_for_me_to_catch_my_mistakes" ~ "\"When I check my code, it is\n easy for me to catch my mistakes\"",
-      Statements == "i_feel_confident_sharing_my_code" ~ "\"I feel confident sharing my code\"",
-      Statements == "i_am_confident_in_my_overall_coding_ability" ~ "\"I am confident in\n my overall coding ability\"",
-      TRUE ~ Statements
-    ),
-    Response = case_when(
-      Response == "This is a lot like me" ~ "A lot like me",
-      Response == "This is very different to me" ~ "Very different from me",
-      Response == "This somewhat describes me" ~ "Somewhat describes me",
-    )
-  ) |>
-  filter(!is.na(Response))
-
-# Taken from: https://stackoverflow.com/questions/52297978/decrease-overal-legend-size-elements-and-text
-addSmallLegend <- function(myPlot, pointSize = 0.5, textSize = 3, spaceLegend = 0.1) {
-  myPlot +
-    guides(
-      shape = guide_legend(override.aes = list(size = pointSize)),
-      color = guide_legend(override.aes = list(size = pointSize))
-    ) +
-    theme(
-      legend.title = element_text(size = textSize),
-      legend.text = element_text(size = textSize),
-      legend.key.size = unit(spaceLegend, "lines")
-    )
-}
-
-# Make the plot
-multi_bar <- writing_perceptions_long |>
-  ggplot(aes(x = Response, fill = Statements)) +
-  geom_bar(position = "dodge", stat = "count") +
-  labs(
-    title = "Student Self-Perceptions of Coding",
-    x = "Response",
-    y = "Frequency"
-  ) +
-  theme_minimal() +
-  theme(
-    plot.title.position = "plot",
-    plot.title = element_text(hjust = 0.5),
-    legend.position = "bottom",
-  )
-
-multi_bar <- addSmallLegend(multi_bar, pointSize = 0.4, textSize = 8.5, spaceLegend = 0.8)
-
-multi_bar
-```
-
-Student self-perceptions about their coding abilities, shown in @fig-coding-multibar-1 and @fig-coding-multibar-2 were also very interesting. ...
-
-```{r fig-ai-appropriate-pie}
-#| message: false
-#| echo: false
-#| warning: false
-#| tbl-cap: Comparison of Whether AI Use is Appropriate in School
-#| fig-align: center
-
-# Pie chart of student opinion on appropriateness of AI use in class.
-
-appropriate <- survey_data |>
-  select(ai_schoolwork_appropriate) |>
-  count(ai_schoolwork_appropriate)
-
-ggplot(appropriate, aes(x = "", y = n, fill = ai_schoolwork_appropriate)) +
-  geom_bar(stat = "identity", width = 1) +
-  coord_polar(theta = "y") +
-  labs(
-    title = "Do students think AI use is appropriate in class?",
-    x = NULL,
-    y = NULL
-  ) +
-  theme_minimal() +
-  scale_fill_discrete(name = "Response") +
-  theme(
-    plot.title.position = "plot",
-    plot.title = element_text(hjust = 0.5),
-    axis.text.x = element_blank(), # Remove x-axis text
-    axis.title.x = element_blank(),
-  ) +
-  geom_text(aes(label = n),
-    position = position_stack(vjust = 0.5)
-  )
-```
-
-## Student Usage
-
-```{r fig-AI-usage-type-comp}
-#| message: false
-#| echo: false
-#| warning: false
-#| fig-cap: How Students Use ChatGPT
-#| fig-align: center
-
-type_data <- survey_data |>
-  select(
-    not_used,
-    used_technical_q,
-    used_conversation,
-    used_general_knowledge,
-    used_solving_hmk,
-    used_check_soln,
-    used_quick_q,
-    used_expl_concepts,
-    used_writing_content,
-    used_writing_code
-  ) |>
-  pivot_longer(cols = everything(), names_to = "Statements", values_to = "Response") |>
-  mutate(Statements = case_when(
-    Statements == "not_used" ~ "Not Used",
-    Statements == "used_technical_q" ~ "Technical Questions",
-    Statements == "used_conversation" ~ "Fun Conversation",
-    Statements == "used_general_knowledge" ~ "General Knowledge",
-    Statements == "used_solving_hmk" ~ "Entirely Solving Homework",
-    Statements == "used_check_soln" ~ "Checking Solutions",
-    Statements == "used_quick_q" ~ "Quick Questions",
-    Statements == "used_expl_concepts" ~ "Explaining Concepts",
-    Statements == "used_writing_content" ~ "Writing Paper Content",
-    Statements == "used_writing_code" ~ "Writing Paper Code",
-    TRUE ~ Statements
-  )) |>
-  filter(Response == 1) |>
-  count(Statements) |>
-  arrange(desc(n))
-
-type_data |>
-  ggplot(aes(x = n, y = reorder(Statements, n))) +
-  geom_bar(stat = "identity", fill = "lightblue") +
-  labs(title = "How Students Used ChatGPT", x = "Number of Students", y = "Type of Usage") +
-  theme_classic()
-```
-
-To understand the role of AI chatbots in learning, students identified their usage by checking off various pre-defined use cases in the survey.
-As we can see in @fig-AI-usage-type-comp, technical questions and explaining concepts were the two top use cases among students.
-More than half of students also used ChatGPT for Quick Questions, General Knowledge, Writing Paper Code, and Checking Solutions.
-Just less than half used it to write paper content, which could suggest that students do not feel confident using it to improve writing, concurring with student attitudes in @Cahill_2024.
-
-TODO: Make a viz of helpfulness in various assignments. Describe the context of the assignment.
-
-```{r fig-AI-helpfulness}
-#| message: false
-#| echo: false
-#| warning: false
-#| fig-cap: CHANGE/REMOVE
-#| fig-align: center
-
-```
-
-## Sentiment Analysis
+## Open-ended survey responses
 
@@
 
-We then performed basic sentiment analysis on the open-response comment questions asked to students. To perform this text analysis, the tidytext package [@tidytext] was used.
-
-### Appropriateness of AI in Schoolwork
-
-In the survey, students were asked to provide an open-ended response to the question: "Please elaborate on to what extent do you think using generative AI tools such as ChatGPT by OpenAI (or equivalents) is ethical and appropriate for schoolwork?".
-These responses were attempted to be understood thematically.
+One question asked students to elaborate on whether they thought generative AI tools such as ChatGPT were ethical and appropriate for schoolwork.
+This was an open-response question.
+We summarized the responses to that question using text analysis and the `tidytext` package [@tidytext].
 
@@
 ai_comments <- survey_data |>
-  select(
-    id,
-    ai_schoolwork_appropriate_comments
-  ) |>
-  mutate(
-    comment = ai_schoolwork_appropriate_comments
-  ) |>
+  select(id, ai_schoolwork_appropriate_comments) |>
+  mutate(comment = ai_schoolwork_appropriate_comments) |>
   select(-c(ai_schoolwork_appropriate_comments))
@@
   mutate(word = str_to_lower(string = word))
-
-```
-
-```{r fig-wordcloud-ai-appropriate}
-#| message: false
-#| echo: false
-#| warning: false
-#| fig-cap: (CHANGE) Boilerplate figure caption 
-#| fig-align: center
 
@@
   acast(word ~ sentiment, value.var = "n", fill = 0)
-
-# Generate comparative word cloud
-comparison.cloud(
-  comparison_data,
-  colors = c("darkred", "darkgreen"),
-  max.words = 100,
-  title.size = 1.5,
-  scale = c(3, 0.5)  # Adjust scale for word size
-)
-```
-
-```{r tbl-ai-appropriate}
-#| message: false
-#| echo: false
-#| warning: false
-#| tbl-cap: Comparing Incidences of Positive and Negative Words.
-#| tbl-subcap: ["Top 10 Positive Words", "Top 10 Negative Words"]
-#| layout-ncol: 2
-#| fig-align: center
-
-# Generate positive words table
-tbl_ai_pos <- tidy_ai_comments |>
-  ungroup() |>
-  select(-c("id")) |>
-  inner_join(pos_words) |>
-  count(word, sort = TRUE) |>
-  head(10) |>
-  kable(, format = "markdown")
-
-# Generate negative words table
-tbl_ai_neg <- tidy_ai_comments |>
-  ungroup() |>
-  select(-c("id")) |>
-  inner_join(neg_words) |>
-  count(word, sort = TRUE) |>
-  head(10) |>
-  kable(, format = "markdown")
-
-tbl_ai_pos
-tbl_ai_neg
-
-```
-
-### AI Implementation Recommendations
-
-```{r recs-setup}
-#| message: false
-#| echo: false
-#| warning: false
 
@@
   separate(bigram, into = c("word1", "word2"), sep = " ") |>
-  filter(!word1 %in% stop_words$word & !word2 %in% stop_words$word) |>
+  filter(!word1 %in% stop_words$word &
+           !word2 %in% stop_words$word) |>
   count(word1, word2, sort = TRUE) |>
@@
   inner_join(sentiments, by = c("word1" = "word")) |>
-  inner_join(sentiments, by = c("word2" = "word"), suffix = c(".word1", ".word2")) |>
+  inner_join(sentiments,
+             by = c("word2" = "word"),
+             suffix = c(".word1", ".word2")) |>
   unite(bigram, word1, word2, sep = " ") |>
@@
   comparison_data_words |> rename(term = word),
-  comparison_data_bigrams |> rename(term = bigram)
-) |>
+  comparison_data_bigrams |> rename(term = bigram)) |>
   arrange(desc(n))
-
-```
-
-```{r recs-wordcloud}
-#| message: false
-#| echo: false
-#| warning: false
-#| fig-cap: Word Cloud for Recommendations Comments
-#| fig-align: center
-
-combined_data |>
-  with(
-    wordcloud(
-      words = term,
-      freq = n,
-      scale = c(3, 0.5),
-      min.freq = 1, # Adjust scale as needed
-      max.words = 100,    # Maximum number of words to display
-      colors = "blue"     # Color for the word cloud
-    )
-  )
-
 ```
@@
 # Generate positive words table
-tbl_recs_pos <- tidy_recs_comments |>
+tidy_recs_comments |>
   ungroup() |>
@@
 # Generate negative words table
-tbl_recs_neg <- tidy_recs_comments |>
+tidy_recs_comments |>
   ungroup() |>
@@
   kable(, format = "markdown") 
-
-tbl_recs_pos
-tbl_recs_neg
 ```
 
-```{r word-cloud}
+## LLM usage and final paper marks
+
+Two other components were merged with the survey responses: self-reported LLM usage on the final paper, and final paper mark.
+
+Students were encouraged to use LLMs to complete their papers.
+Each paper required the students to disclose their usage through a statement in the GitHub repository README for the paper.
+Even students who did not use generative AI at all were required to state this in the README.
+For students who did use generative AI, there was an additional requirement that they save the logs of their usage in a txt file which was also included in their GitHub repository.
+
+Those README statements were gathered and parsed using OpenAI's ChatGPT 4o model (as at 26 July).
+The following prompt was used: 'The following statement is about to what extent LLMs were used by a student.
+Please characterize it as one of: "None", "Minimal", "Somewhat", "Extensive", "Unsure".
+Respond with only one of those options.' All classifications were then manually checked by hand for reasonableness.
+
+We find a varied extent of self-reported LLM usage (@tbl-llm-usage).
+41 respondents were classified as having made extensive use of LLMs, while 28 were classified as having made somewhat use.
+31 respondents were classified as having made minimal or no use of LLMs.
+In the analysis dataset we combine those two classifications because only 8 respondents were classified as having minimal usage.
+
+```{r tbl-llm-usage}
 #| message: false
@@
 #| warning: false
+#| tbl-cap: Self-reported LLM usage in final paper
 
-# Word cloud based on bigrams
-word_counts |>
-  with(
-    wordcloud(
-      words = word,
-      freq = n,
-      scale = c(3, 0.5),  # Adjust scale as needed
-      max.words = 100,    # Maximum number of words to display
-      colors = "blue"     # Color for the word cloud
+all_data |>
+  count(llm_usage) |> 
+  kable(col.names = c("Self-reported LLM usage", "Number"))
+```
+
+The third, and final, component is the mark, in percentages, on the final paper (@fig-mark).
+The overall mean was 78% and standard deviation was 17 percentage points.
+However there was considerable differences by the extent of LLM usage.
+For extensive use, the mean was 82% and the standard deviation was 15 percentage points.
+For somewhat usage, the mean was 77% and the standard deviation was 16 percentage points.
+For minimal usage, the mean was 74% and the standard deviation was 23 percentage points.
+Finally for no usage, the mean was 75% and the standard deviation was 19 percentage points.
+
+```{r fig-mark}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: Mark on final paper, in percentages, and LLM usage in final paper
+
+# all_data |>
+#   group_by(llm_usage) |> 
+#   summarize(mean = mean(mark) |> round(2),
+#             sd = sd(mark) |> round(2))
+# 
+# all_data |>
+#   summarize(mean = mean(mark) |> round(2),
+#             sd = sd(mark) |> round(2))
+
+all_data |>
+  ggplot(mapping = aes(x = mark, color =)) +
+  geom_histogram(binwidth = 0.01) +
+  scale_x_continuous(labels = scales::label_percent()) +
+  labs(x = "Final paper mark (percentage)", y = "Count") +
+  facet_wrap(vars(llm_usage),
+             ncol = 2,
+             labeller = labeller(
+               llm_usage = c(
+                 "Extensive" = "Extensive LLM usage",
+                 "Minimal" = "Minimal LLM usage",
+                 "None" = "No LLM usage",
+                 "Somewhat" = "Some LLM usage"
                )
-  )
+             )) +
+  theme_minimal()
 ```
@@
 
-The goal of our modelling strategy is twofold. Firstly,...
+The goal of our modelling strategy is to better understand how respondents result on their final paper is associated with their self-reported LLM usage.
+The result on the final paper is a proportion and some respondents got full marks and so we use zero-one-inflated beta regression.
+Here we briefly describe the model that we use, which follows @kurztotherescue.
+Diagnostics are included in [Appendix -@sec-model-details].
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
-
-## Model set-up
-
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+Define $y_i$ as the percentage received on the final paper. Then $\beta_1$ is the characterization of self-reported LLM usage and $\beta_2$ is self-reported GPA.
 
 \begin{align}
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
+y_i &\sim \mbox{Beta}(\mu_i, \phi) \\
+logit(\mu_i) &= \beta_0 + \beta_1 \times \mbox{LLM usage}_i + \beta_2 \times \mbox{GPA}_i\\
+\beta_0, \beta_1, \beta_2  &\sim \mbox{Normal}(0, 1) \\
+\phi &\sim \mbox{Gamma}(4, 0.1)
 \end{align}
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+We estimate the model in `R` [@citeR] using `brms` [@brms].
 
-### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
+The expected relationship between LLM usage and final mark is unclear.
+It may be that stronger students used LLMs in a more sophisticated way or that they did not need to use them.
+However, the relationship between GPA and final mark is expected to be positive.
 
@@
 
-Our results are summarized in @tbl-modelresults.
+Our results are summarized in @tbl-model-results and @fig-model-results. We especially draw on `modelsummary` [@modelsummary].
+
+```{r tbl-model-results}
+#| message: false
+#| echo: false
+#| warning: false
+#| tbl-cap: Coefficient estimates and mean absolute deviation (MAD)
+
+fit1 <-
+  readRDS(file = here::here("models/fit1brms.rds"))
+fit2 <-
+  readRDS(file = here::here("models/fit2brms.rds"))
+
+modelsummary::modelsummary(
+  list(
+    "Base model" = fit1,
+    "Including self-reported GPA" = fit2
+  ),
+  statistic = "mad",
+  fmt = 2
+)
+```
+
+```{r fig-model-results}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: Coefficient estimates and 90 per cent credibility intervals
+
+modelsummary::modelplot(
+  list(
+    "Base model" = fit1,
+    "Including self-reported GPA" = fit2
+  ),
+  conf_level = 0.9,
+  coef_omit = "phi") +
+  labs(x= "") +
+  scale_color_brewer(palette = "Set1")
+```
+
+Interestingly, we find a slightly negative effect, relative to self-reported extensive use, of using LLMs less (@fig-model-results).
+As expected, self-reported GPA is positively associated with final mark.
+
+As always with regression, average estimates over the full dataset.
+It may be that the results are different for different tranches of respondents.
+For instance, looking at the 28 students who received an A+ for the final paper, 13 of them had extensive LLM usage.
 
@@
 ## First discussion point {#sec-first-point}
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
 
@@
 
-Weaknesses and next steps should also be included.
+We don't know whether it made them faster.
 
@@
 
-# Additional data details
+# Survey questions {#sec-survey-details}
+
+1. After carefully reading the informed consent document, please indicate below whether you consent to have your anonymized responses included in the research study?
+    - Yes, I authorize the use of the data collected about me for the STA302 course survey to be used. I will be compensated 1% of my course grade for completing the survey.
+    - No I do not want my data included in the research study, but I want to complete the survey. I will be compensated 1% of my course grade for completing the survey.
+    - I do not want to complete this survey. I realize that I am forfeiting the corresponding course credit.
+2. What is your full name on Quercus?
+3. What is your Student ID?
+4. What year are you?
+5. What is your specialization?
+6. What is/are your major/s?
+7. What is/are your minor/s?
+8. What is your GPA?
+9.
+Please rate how much each statement describes you, on a scale from "This is very different to me" to "This is a lot like me" ["This is very different to me"; "This somewhat describes me"; "This is a lot like me"]
+    - Writing is easy for me
+    - I like to write
+    - I believe it is important to be a good writer.
+    - When I edit it is easy for me to catch my mistakes.
+    - I feel confident sharing my writing.
+    - I am confident in my overall writing ability.
+10.
+Please rate how much each statement describes you, on a scale from "This is very different to me" to "This is a lot like me".
+When answering, please consider whichever programming language you are most familiar with.
+["This is very different to me"; "This somewhat describes me"; "This is a lot like me"]
+    - Coding is easy for me
+    - I like to code
+    - I believe it is important to be a good coder.
+    - When I check my code it is easy for me to catch my mistakes.
+    - I feel confident sharing my code.
+    - I am confident in my overall coding ability.
+11. How familiar are you with using generative AI tools such as OpenAI's ChatGPT or equivalents?
+    - Very familiar
+    - Somewhat familiar
+    - Not familiar
+    - Other
+12. Have you used any generative AI tools such as OpenAI's ChatGPT or equivalents for any reason (personal or educational)?
+    - Yes
+    - No
+    - Other
+13. If you have used generative AI tools such as OpenAI's ChatGPT or equivalents, in what ways have you used it (select all that apply)?
+    - Asking technical questions
+    - Carrying on a conversation out of curiosity
+    - Asking general knowledge questions
+    - Solving homework
+    - Checking solutions
+    - Asking quick questions when stuck
+    - Explaining concepts
+    - Writing essays or paragraphs
+    - Writing code
+    - Never used it
+    - Other
+14. To what extent do you think using generative AI tools such as ChatGPT by OpenAI (or equivalents) is ethical and appropriate for schoolwork?
+    - Appropriate
+    - Inappropriate
+    - Other
+15. Please elaborate on your answer above.
+16. Did you use any generative AI tools such as OpenAI's ChatGPT or equivalents for STA302?
+    - Yes
+    - No
+    - Other
+17.
+How helpful did you find generative AI tools such as ChatGPT by OpenAI (or equivalents) for each component of STA302?
+["Not helpful"; "Somewhat helpful"; "Very helpful"; "I did not use generative AI for this component"]
+    - Weekly quiz
+    - Weekly mini-essay
+    - Papers: Generating ideas
+    - Papers: Writing code
+    - Papers: Writing content
+    - Papers: Improving content
+18. What is your recommendation for how generative AI tools such as ChatGPT by OpenAI (or equivalents) should be used in the course in future?
+19. (Optional) Any other comments?
+
+\newpage
 
@@
 
-In ... we implement a posterior predictive check. This shows...
+```{r fig-posteriorchecks}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: Posterior predictive checking
+#| fig-subcap: ["Base model","Including self-reported GPA"]
+#| layout-ncol: 2
 
-In ... we compare the posterior with the prior. This shows...
+bayesplot::pp_check(fit1) +
+  theme_classic() +
+  theme(legend.position = "bottom")
+
+bayesplot::pp_check(fit2) +
+  theme_classic() +
+  theme(legend.position = "bottom")
+```
+
+```{r tbl-loo}
+#| message: false
+#| echo: false
+#| warning: false
+#| tbl-cap: Model comparison
+
+brms::loo_compare(loo::loo(fit1, cores = 2), loo::loo(fit2, cores = 2))
+```
+
+\newpage
 
@@
 
-... is a trace plot. It shows... This suggests...
+```{r fig-basemodeldiagnostics}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: Base model diagnostics
 
-... is a Rhat plot. It shows... This suggests...
+plot(fit1)
+```
+
+```{r fig-gpamodeldiagnostics}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: Model including self-reported GPA diagnostics
+
+plot(fit2)
+```
 

### 2024-08-04 scripts/04-model.R

@@ -1,37 +1,73 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Models
+# Author: Rohan Alexander
+# Date: 4 August 2024
+# Contact: rohan.alexander@utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: -
+# Any other information needed? -
 
 
 #### Workspace setup ####
 library(tidyverse)
 library(rstanarm)
+library(brms)
+library(readxl)
+
 
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+all_data <-
+  read_excel(here::here("data/raw_data/merged_data.xlsx"), sheet = "Sheet1") |>
+  janitor::clean_names() |>
+  mutate(mark = if_else(mark == 1, 0.99, mark)) |> # I'm a monster
+  mutate(llm_usage =
+           if_else(llm_usage %in% c("Minimal", "None"), "None or minimal", llm_usage)) |>
+  filter(!is.na(what_is_your_gpa))
+
 
 ### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
+# Using rstanarm
+fit1rstanarm <-
+  stan_betareg(mark ~ llm_usage,
+               data = all_data,
+               link = "logit",
+               seed = 853)
+
+fit2rstanarm <-
+  stan_betareg(
+    mark ~ llm_usage + what_is_your_gpa,
+    data = all_data,
+    link = "logit",
     seed = 853
   )
 
+# Using brms
+fit1brms <- brm(
+  data = all_data,
+  family = Beta,
+  mark ~ llm_usage,
+  prior = prior(normal(0, 1), class = b) +
[…]

### 2024-08-06 paper/paper.qmd
Almost finished first draft

@@
 We thank Nathalie Moon for developing and sharing her survey questions.
+We thank Tiffany Timbers and attendees at JSM 2024 for helpful suggestions.
 This research is currently under review by the University of Toronto's University Research Ethics Board.
@@
 abstract: "To help understand the effect of Large Language Models (LLMs) on data science practice we examine the extent to which LLM usage is correlated with the mark that a student gets on a final paper in a classroom data science setting.
-We find no clear relationship between more extensive LLM usage and the student's mark.
+We find some very mild evidence from this observational study that LLM usage may be associated with better scores, but our main conclusion is that there is no clear relationship between more extensive LLM usage and the student's mark.
 Despite the classroom setting used for evaluation, the particular activity of interest reflects the work done by professional data scientists.
@@
 
-# Old data (REMOVE ONCE THE TEXT ANALYSIS BIT IS UPDATED)
-survey_data <- read_csv("../data/analysis_data/clean_STA302_postcourse_survey_w24.csv", show_col_types = FALSE)
-
 # Read in data
-all_data <- read_excel(here::here("data/raw_data/merged_data.xlsx"), sheet = "Sheet1") |>
+all_data <-
+  read_excel(here::here("data/merged_data.xlsx"), sheet = "Sheet1") |>
   janitor::clean_names()
@@
 Each paper has the same basic structure and expectations.
-Before the final paper is due students have received feedback on all their previous work in the class (including their past papers) and there is an, optional, two-day period of peer review.
+Before the final paper is due students have received feedback on all their previous work in the class (including their past papers) and there is an optional two-day period of peer review.
 
@@
   geom_histogram(binwidth = 1) +
-  labs(
-    x = "Time to complete survey (minutes)",
-    y = "Frequency"
-  ) +
+  labs(x = "Time to complete survey (minutes)", y = "Frequency") +
   theme_minimal()
@@
   geom_histogram(binwidth = 0.1) +
-  labs(
-    x = "Self-reported GPA",
-    y = "Frequency"
-  ) +
+  labs(x = "Self-reported GPA", y = "Frequency") +
   theme_minimal()
@@
     what_year_are_you = case_when(
-      what_year_are_you == "PEY" ~ "4th", # Change PEY to 4th year
+      what_year_are_you == "PEY" ~ "4th",
+      # Change PEY to 4th year
       what_year_are_you == "5th" ~ "5th or over",
@@
   count(what_year_are_you) |>
-  pivot_wider(names_from = what_year_are_you, 
+  pivot_wider(
+    names_from = what_year_are_you,
     values_from = n,
-              values_fill = list(n = 0)) |> 
+    values_fill = list(n = 0)
+  ) |>
   kable()
@@
   ) |>
-  pivot_longer(cols = everything(), names_to = "statements", values_to = "response") |>
+  pivot_longer(cols = everything(),
+               names_to = "statements",
+               values_to = "response") |>
   mutate(
@@
   facet_wrap(vars(statements)) +
-  labs(
-    x = "Response",
-    y = "Number of repondents",
-    fill = "Response:"
-  ) +
+  labs(x = "Response", y = "Number of repondents", fill = "Response:") +
   scale_fill_brewer(palette = "Set1") +
   theme_minimal() +
-  theme(
-    axis.text.x = element_blank(),
-    legend.position = "bottom"
-  )
+  theme(axis.text.x = element_blank(), legend.position = "bottom")
 
@@
   facet_wrap(vars(statements)) +
-  labs(
-    x = "Response",
-    y = "Number of repondents",
-    fill = "Response:"
-  ) +
+  labs(x = "Response", y = "Number of repondents", fill = "Response:") +
   scale_fill_brewer(palette = "Set1") +
   theme_minimal() +
-  theme(
-    axis.text.x = element_blank(),
-    legend.position = "bottom"
-  )
+  theme(axis.text.x = element_blank(), legend.position = "bottom")
 ```
@@
       TRUE ~ ai_schoolwork_appropriate
-      )) |> 
+    )
+  ) |>
   count(ai_schoolwork_appropriate) |>
@@
 
-To understand the role of LLMs in learning, students more granularly identified their usage by checking off various pre-defined use cases in the survey (@fig-AI-use-and-usefulness-1).
+To understand the role of LLMs in learning, students identified their usage in a more granular way by checking off various pre-defined use cases in the survey (@fig-AI-use-and-usefulness-1).
 Technical questions and explaining concepts were the two top use cases among students in the course.
@@
     response_simplified = case_when(
-      response %in% c("\"Not helpful\"", "\"I did not use generative AI for this component\"") ~ "Less Helpful",
+      response %in% c(
+        "\"Not helpful\"",
+        "\"I did not use generative AI for this component\""
+      ) ~ "Less Helpful",
       response %in% c("\"Somewhat helpful\"", "\"Very helpful\"") ~ "More Helpful",
@@
   facet_wrap(~ task, scales = "free_y") +
-  labs(y = "Number of respondents",
-       x = "",
-       fill = "Response type: ") +
+  labs(y = "Number of respondents", x = "", fill = "Response type: ") +
   theme_minimal() +
@@
   scale_fill_brewer(palette = "Set1") +
-  theme(
-    axis.text.y = element_blank(),
-    legend.position = "bottom",
-  )
-
+  theme(axis.text.y = element_blank(), legend.position = "bottom")
 ```
@@
 
-## Open-ended survey responses
+Finally, one question asked students to elaborate on whether they thought generative AI tools such as ChatGPT were ethical and appropriate for schoolwork.
+This was an open-response question.
+To provide a sense of the responses, we used Anthropic's Claude 3.5 Sonnet model (as at 5 August 2024) to summarize the comments and it provided:
 
-```{r setup-sentiment}
-#| message: false
-#| echo: false
-#| warning: false
-
-#get positive words from tidytext
-neg_words <- get_sentiments() |>
-  filter(sentiment == "negative")
-
-#get negative words from tidytext
-pos_words <- get_sentiments() |>
-  filter(sentiment == "positive")
-```
-
-One question asked students to elaborate on whether they thought generative AI tools such as ChatGPT were ethical and appropriate for schoolwork.
-This was an open-response question.
-We summarized the responses to that question using text analysis and the `tidytext` package [@tidytext].
-
-```{r setup-ai-appropriate}
-#| message: false
-#| echo: false
-#| warning: false
-
-# AI Appropriateness Comments
-ai_comments <- survey_data |>
-  select(id, ai_schoolwork_appropriate_comments) |>
-  mutate(comment = ai_schoolwork_appropriate_comments) |>
-  select(-c(ai_schoolwork_appropriate_comments))
-
-# put comments into tidytext format
-tidy_ai_comments <- ai_comments |>
-  group_by(id) |>
-  unnest_tokens(word, comment) |>
-  mutate(word = str_to_lower(string = word))
-
-# Load sentiment lexicon
-sentiments <- get_sentiments()
-
-# Prepare data for comparison cloud
-comparison_data <- tidy_ai_comments |>
-  inner_join(sentiments) |>
-  count(word, sentiment, sort = TRUE) |>
-  acast(word ~ sentiment, value.var = "n", fill = 0)
-
-# Example sentiments data frame
-sentiments <- get_sentiments()
-
-# Prepare data
-recs_comments <- survey_data |>
-  select(id, recs_for_ai_sta302) |>
-  mutate(comment = recs_for_ai_sta302) |>
-  select(-c(recs_for_ai_sta302))
-
-# Tokenize and clean individual words
-tidy_recs_comments <- recs_comments |>
-  ungroup() |>
-  unnest_tokens(word, comment) |>
-  anti_join(stop_words)
-
-# Count word frequencies
-word_counts <- tidy_recs_comments |>
-  count(word, sort = TRUE)
-
-# Bigram analysis
-bigrams <- recs_comments |>
-  ungroup() |>
-  unnest_tokens(bigram, comment, token = "ngrams", n = 2) |>
-  separate(bigram, into = c("word1", "word2"), sep = " ") |>
-  filter(!word1 %in% stop_words$word &
-           !word2 %in% stop_words$word) |>
-  count(word1, word2, sort = TRUE) |>
-  unite(bigram, word1, word2, sep = " ")
-
-# Prepare data for comparison cloud (including both words and bigrams)
-comparison_data_words <- tidy_recs_comments |>
-  inner_join(sentiments, by = "word") |>
-  count(word, sentiment, sort = TRUE)
-
-comparison_data_bigrams <- bigrams |>
-  separate(bigram, into = c("word1", "word2"), sep = " ") |>
-  inner_join(sentiments, by = c("word1" = "word")) |>
-  inner_join(sentiments,
-             by = c("word2" = "word"),
-             suffix = c(".word1", ".word2")) |>
-  unite(bigram, word1, word2, sep = " ") |>
-  count(bigram, sentiment.word1, sort = TRUE) |>
-  rename(sentiment = sentiment.word1)
-
-# Combine both data sets
-combined_data <- bind_rows(
-  comparison_data_words |> rename(term = word),
-  comparison_data_bigrams |> rename(term = bigram)) |>
-  arrange(desc(n))
-```
-
-```{r tbl-recs-relpolarity}
-#| message: false
-#| echo: false
-#| warning: false
-#| tbl-cap: Comparing Incidences of Positive and Negative Words in Student.
-#| tbl-subcap: ["Top 10 Positive Words", "Top 10 Negative Words"]
-#| layout-ncol: 2
-#| fig-align: center
-
-# Generate positive words table
-tidy_recs_comments |>
-  ungroup() |>
-  select(-c("id")) |>
-  inner_join(pos_words) |>
-  count(word, sort = TRUE) |>
-  head(10) |>
-  kable(, format = "markdown")
-
-# Generate negative words table
-tidy_recs_comments |>
-  ungroup() |>
-  select(-c("id")) |>
-  inner_join(neg_words) |>
-  count(word, sort = TRUE) |>
-  head(10) |>
-  kable(, format = "markdown") 
-```
+> Many students view AI as a helpful supplementary tool, comparing it to resources like Google or calculators.
+They believe it can aid in understanding concepts, debugging code, brainstorming ideas, and saving time on routine tasks.
+However, there's a consensus that AI should not be used to complete entire assignments or replace original thinking.
+Students emphasize the importance of using AI ethically, citing it when appropriate, and not relying on it exclusively.
+Some argue that learning to use AI effectively is a valuable skill for future careers.
+Concerns raised include the potential for plagiarism, the risk of hindering critical thinking skills, and the possibility of receiving incorrect information.
+Overall, most students support the responsible use of AI in education, with proper guidelines and transparency, while recognizing the need to maintain academic integrity and develop independent learning skills.
 
@@
 all_data |>
-  ggplot(mapping = aes(x = mark, color =)) +
+  ggplot(mapping = aes(x = mark, color = )) +
   geom_histogram(binwidth = 0.01) +
@@
 
-As always with regression, average estimates over the full dataset.
-It may be that the results are different for different tranches of respondents.
-For instance, looking at the 28 students who received an A+ for the final paper, 13 of them had extensive LLM usage.
-
 # Discussion {#sec-discussion}
 
-## First discussion point {#sec-first-point}
+There is some very mild evidence from this observational study that LLM usage may be associated with better scores.
+However, this should mostly be treated as a direction for future research.
+More broadly, our work has three major take-aways: 1) LLMs have a major role to play in enhancing the trustworthiness of data science.
+2) Making general purpose chatbots more opinionated and knowledgeable about data science would considerably help.
+3) Alternative interaction approaches, beyond chatbots, could be especially useful.
 
-## Second discussion point
+## Impact of LLMs on trustworthy data science
 
-## Third discussion point
+## General purpose compared with opinionated specific chatbots
+
+## On the need for LLM-based infrastructure
 
@@
 
-We don't know whether it made them faster.
+There are a number of substantial weaknesses of this study.
+The foundational one is that we used observational data, much of which was self-reported.
+There may be selection bias present in terms of who used LLMs, who reported their LLM usage truthfully, and even who remained in the class.
+A different design, specifically a randomized controlled trial (RCT) would deal with many of these issues, although likely at some cost.
+
+Regression reports average estimates over the full dataset.
+However, many earlier studies found distributional effects, with low-performing individuals benefiting more than high-performing individuals.
+Again, a change in design toward an RCT could enable the exploration of this question.
+Stratification of our dataset would result in small sample size, but nonetheless our data does provide some limited suggestive evidence that this effect may be present in data science.
+For instance, looking at the 28 students who received an A+ for the final paper, 13 of them had extensive LLM usage, whereas looking at the same number of worst performing students finds that only six of them had extensive LLM usage.
+
+Along these lines, we have considered LLM usage for code and writing as equivalent, but they should actually have different impacts depending on student backgrounds.
+Distinguishing between native and second-language English speakers, and then focusing on differential LLM usage, could add a great deal to the analysis.
+
+Finally, we only considered one outcome measure, namely grade on the final paper.
+Each paper required a considerable amount of time for the student to produce.
+If an LLM was found to reduce the time taken to produce a paper, without any reduction in quality, then that would be a similarly useful outcome.
+
+Despite these shortcomings, our work clearly identifies a need for further research examining how LLMs can be used to develop a more trustworthy data science.
 
@@
 
+We use `bayesplot` [@bayesplot] and `loo` [@loo] help with posterior predictive checks and model diagnostics.
+
 ## Posterior predictive check
@@
 #| fig-cap: Posterior predictive checking
-#| fig-subcap: ["Base model","Including self-reported GPA"]
+#| fig-subcap: ["Base model", "Including self-reported GPA"]
 #| layout-ncol: 2
@@
 ```{r fig-basemodeldiagnostics}
+#| fig-height: 7
 #| message: false
@@
 ```{r fig-gpamodeldiagnostics}
+#| fig-height: 7
 #| message: false

### 2024-08-06 scripts/01-download_data.R

@@ -1,21 +1,0 @@
-#### Preamble ####
-# Purpose: Downloads and saves the data from Excel file.
-# Author: Luca Carnegie
-# Date: 16 July 2024
-# Contact: luca.carnegie@mail.utoronto.ca or rohan.alexander@utoronto.ca 
-# License: MIT
-# Pre-requisites: None
-
-
-#### Workspace setup ####
-library(tidyverse)
-library(readr)
-
-#### Download data ####
-sta302_survey <- read_excel("data/raw_data/raw_STA302_postcourse_survey_w24.xlsx")
-
-
-#### Save data ####
-write_csv(sta302_survey, "data/raw_data/raw_STA302_postcourse_survey_w24.csv") 
-
-         

### 2024-08-06 scripts/03-model.R

@@ -17,7 +17,7 @@ library(readxl)
 
 #### Read data ####
 all_data <-
-  read_excel(here::here("data/raw_data/merged_data.xlsx"), sheet = "Sheet1") |>
+  read_excel(here::here("data/merged_data.xlsx"), sheet = "Sheet1") |>
   janitor::clean_names() |>
   mutate(mark = if_else(mark == 1, 0.99, mark)) |> # I'm a monster
   mutate(llm_usage =

### 2024-08-06 scripts/05-replications.R

@@ -1,19 +1,0 @@
-#### Preamble ####
-# Purpose: Replicated graphs from... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
-
-#### Workspace setup ####
-library(tidyverse)
-# [...UPDATE THIS...]
-
-#### Load data ####
-# [...ADD CODE HERE...]
-
-
-

### 2024-08-06 paper/paper.qmd
found sources and started fixing up written bits

@@
   - Luca Carnegie
+  - Nathalie Moon
 thanks: "Code and data are available at: https://github.com/lcarnegie/llms-achievement.
@@
 library(tidyverse)
+library(tinytex)
 
@@
 There are many aspects of trustworthy data science, including education, culture, and workflow to name a few.
-One aspect of special importance is the tools that we use to do data science: the computers, programming languages, settings, and environments.
-The wide release of user-friendly Large Language Models (LLMs), especially OpenAI's ChatGPT, is a rare instance of a new tool being made available.
+One aspect of special importance is the tools that we use to do data science: the computers, programming languages, and environments.
+The wide release of user-friendly Large Language Models (LLMs), particularly OpenAI's ChatGPT, is a rare instance of a new and powerful tool being made available to add to a data scientist's toolbelt.
 
-Like all new tools, LLMs have created both excitement and apprehension.
-ChatGPT's public release on November 30, 2022, brought LLMs into the mainstream conversation.
-By now many people, especially educators and students, have some experience with LLMs, in both personal and professional contexts.
+Like all new tools, however, LLMs have created both excitement and apprehension.
+ChatGPT's public release on November 30, 2022, brought LLMs into mainstream conversation, quickly being put to use in both industry and academia.
+By now, many people, especially educators and students, have some experience with LLMs, in both personal and professional contexts.
 
-In the context of teaching statistics and data science LLMs could be useful for many tasks.
-For instance, there is considerable interest in the potential of chatbots and personalized tutors **[ADD CITE]**.
-ChatGPT has also been the catalyst for many interesting and important conversations around academic integrity **[ADD CITE]**, the development of critical thinking skills **[ADD CITE]**, and what effective learning looks like **[ADD CITE]**.
+In the context of teaching statistics and data science, LLMs could be useful for many tasks.
+For instance, there is considerable interest in the potential of chatbots to act as personalized tutors for students [@Winkler_2018].
+ChatGPT has also been the catalyst for many interesting and important conversations around academic integrity [@Okaibedi_2023], the development of critical thinking skills, and what effective learning looks like [@BaidooAnu_2023].
 
@@
 The tasks involved in a trustworthy data science workflow can be generally broken down into a number of key competencies [@Adhikari2021Interleaving; @Gibbs2021Building].
-The potential for LLMs to positively affect students' academic performance in data science can be clearly inferred from already-demonstrated effects in a number of those competencies on adjacent professional fields.
 The first is programming, which is done when cleaning, analyzing, visualizing data often using languages like R or Python.
 The second is writing, which is primarily done when communicating results.
+The potential for LLMs to positively affect students' academic performance in data science can be clearly inferred from already-demonstrated effects in a number of those competencies on adjacent professional fields.
 
-In terms of general programming, @Peng_2023 found a positive impact of GitHub Copilot (an LLM-powered programming assistant) on productivity.
+In terms of computer programming, @Peng_2023 found a positive impact of GitHub Copilot (an LLM-powered programming assistant) on productivity.
 Specifically, in an experiment involving 95 freelance programmers, they found that that programmers with access to GitHub Copilot completed a standardized programming task 56% faster than the control group.
@@
 
-@DellAcqua_2023 focus on management consulting tasks and provide evidence that LLMs can improve writing productivity.
+@DellAcqua_2023, focusing on management consulting tasks, provide evidence that LLMs can improve writing productivity.
 In a field experiment involving consultants from Boston Consulting Group they found that the use of OpenAI's GPT-4 led to a 25% increase in delivery speed of business tasks (most involving some writing), as well as a 40% increase in human-rated performance on those tasks.
@@
 Similarly, @Tu_2024 acknowledge that LLMs can streamline many parts of a data science workflow.
-With that in mind, they suggest that data-scientists-in-training should shift their self-perspectives from primarily being an analyst to primarily being a manager responsible for strategic oversight of the analysis.
+With that in mind, they suggest that data-scientists-in-training should shift their self-perspectives from primarily being an 'analyst' to primarily being a 'product manager' responsible for strategic oversight of the analysis carried out by LLMs.
 
@@
 Specifically, we examine the relationship between students' grades and self-reported measures of student LLM usage, as well as student attitudes toward LLMs in general.
-This is based on final papers and a survey, conducted in a third-year undergraduate data science course at the University of Toronto.
-By examining how students interact with and perceive LLMs as tools, and how these variables translate into student outcomes, the effects of LLM integration in data science can be more precisely determined and recommendations made for future development.
+This is based on students' final papers and a survey, conducted in a third-year undergraduate data science course at the University of Toronto.
+By examining how students interact with and perceive LLMs as tools, and how these variables translate into student outcomes, the effects of LLM integration in data science can be more precisely determined, leading to better recommendations for their future development.
 
@@
 2. self-reported LLM usage; and 
-3. student marks on the final paper.
+3. student marks on their final paper.
 
@@
 Each paper has the same basic structure and expectations.
-Before the final paper is due students have received feedback on all their previous work in the class (including their past papers) and there is an optional two-day period of peer review.
+Before the final paper is due, students have received feedback on all their previous work in the class (including their past papers) and there is an optional two-day period of peer review.
 
@@
 All writing must be in English.
-The aim of the course for students to create a public portfolio of work they can use to apply for jobs.
+The primary aim of the course for students is to create a public portfolio of work they can use to apply for jobs.
 
@@
 
-All data cleaning and analysis was done using the `R` statistical programming language [@citeR], and especially uses the `tidyverse` [@tidyverse], `janitor` [@janitor], `reshape2` [@reshape], and `readxl` [@readxl] libraries.
+All data cleaning and analysis was done using the `R` statistical programming language [@citeR], and especially uses the `tidyverse` [@tidyverse], `janitor` [@janitor], `reshape2` [@reshape], and `readxl` [@readxl] packages.
 
@@
 One factor that may affect the range is that the course is required for programs in the Statistics, Mathematics and Computer Science Departments.
-Self-reported GPAs introduce the possibility of reporting bias.
+Self-reported GPAs also introduce the possibility of reporting bias.
 For instance, respondents may have provided their cumulative GPA, their most recent term's GPA, or could have misreported it entirely.
@@
 Students from a range of years took the course, however the majority of respondents were in their 3rd or 4th year of study (@tbl-year-dist).
-The course's prerequisite of general statistics, which is a two-course sequence typically completed by students in their second year, would make it difficult to take this course earlier than 3rd year.
+The course's prerequisite of general statistics, which is a two-course sequence typically completed by students in their second year, would make it difficult to take this course earlier than their 3rd year.
 
@@
   count(ai_familiarity) |>
-  kable(col.names = c("Extent of AI familiarity", "Number"))
+  kable(col.names = c("Extent of AI familiarity", "Students"))
 
@@
   count(ai_schoolwork_appropriate) |>
-  kable(col.names = c("Ethical and appropriate for school?", "Number"))
+  kable(col.names = c("Ethical & Appropriate for school?", "Students"))
 ```
@@
 There is some very mild evidence from this observational study that LLM usage may be associated with better scores.
-However, this should mostly be treated as a direction for future research.
+However, this work should mostly be treated as a direction for future research in this area.
 More broadly, our work has three major take-aways: 1) LLMs have a major role to play in enhancing the trustworthiness of data science.
-2) Making general purpose chatbots more opinionated and knowledgeable about data science would considerably help.
-3) Alternative interaction approaches, beyond chatbots, could be especially useful.
+2) Making general purpose chatbots more opinionated and knowledgeable about data science would considerably help (i.e.
+fine-tuning them - maybe talk more about the GPT store and the "Telling Stories with Data chatbot") 3) Alternative interaction approaches, beyond a chatbot interface, could be especially useful.
 
@@
 Along these lines, we have considered LLM usage for code and writing as equivalent, but they should actually have different impacts depending on student backgrounds.
-Distinguishing between native and second-language English speakers, and then focusing on differential LLM usage, could add a great deal to the analysis.
+Distinguishing between native and second-language English speakers, and then focusing on differential LLM usage, could have added a great deal of nuance to the analysis.
 

### 2024-08-07 paper/paper.qmd
changed bin width

@@
   - Luca Carnegie
-  - Nathalie Moon
 thanks: "Code and data are available at: https://github.com/lcarnegie/llms-achievement.
@@
 
-The goal of our modelling strategy is to better understand how respondents result on their final paper is associated with their self-reported LLM usage.
-The result on the final paper is a proportion and some respondents got full marks and so we use zero-one-inflated beta regression.
-Here we briefly describe the model that we use, which follows @kurztotherescue.
-Diagnostics are included in [Appendix -@sec-model-details].
+The goal of our modelling strategy is to better understand how a respondents' result on their final paper associates with their self-reported LLM usage.
+Students received their result on the final paper as a proportion (a 0-100% mark), so we use zero-one-inflated beta regression to best fit the distribution in this interval.
+We use this type of regression given the ability of the beta distribution to effectively describe continuous and skewed or heteroskedastic datasets, which we can see in the asymmetrical histogram of the distribution of marks.
+
+```{r final-grade-dist}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: Distribution of grades on Final Paper
+
+all_data |>
+  select(mark) |>
+  ggplot(aes(x = mark)) +
+  geom_histogram(binwidth = 0.01) +
+  labs(x = "Final paper mark (percentage)", y = "Count") +
+  theme_minimal()
+
+```
+
+Transforming the regression with a logit function as the link function ensures that the predicted values stay between the 0-1 range, like in logistic regression.
+This allows the regression model to only return values between the 0-1, respecting the boundaries of the data.
+Finally, since students' grades could be either 0, 100, or in between, the zero-one-inflated beta model design is used to include grades that are 0 and 1 as well.
+
+Here we briefly describe the model that we use, which follows @kurztotherescue. Model Diagnostics are included in [Appendix -@sec-model-details].
 
@@
 
-We estimate the model in `R` [@citeR] using `brms` [@brms].
+We estimate the model in `R` [@citeR] as well, using `brms` [@brms].
 
 The expected relationship between LLM usage and final mark is unclear.
-It may be that stronger students used LLMs in a more sophisticated way or that they did not need to use them.
+It could be that stronger students used LLMs in a more sophisticated way or that they did not need to use them.
 However, the relationship between GPA and final mark is expected to be positive.

### 2024-08-07 paper/paper.qmd
changed bin width

@@
   ggplot(aes(x = mark)) +
-  geom_histogram(binwidth = 0.01) +
+  geom_histogram(binwidth = 0.05) +
   labs(x = "Final paper mark (percentage)", y = "Count") +

### 2024-08-09 paper/paper.qmd
update bibliography and start move toward complete draft

--- 2024-08-06 paper/paper.qmd
@@
 thanks: "Code and data are available at: https://github.com/lcarnegie/llms-achievement.
-We thank Nathalie Moon for developing and sharing her survey questions.
 We thank Tiffany Timbers and attendees at JSM 2024 for helpful suggestions.
-This research is currently under review by the University of Toronto's University Research Ethics Board.
+This research is currently under review by the University of Toronto's Research Ethics Board.
 Comments can be sent to: rohan.alexander@utoronto.ca."
@@
 There are many aspects of trustworthy data science, including education, culture, and workflow to name a few.
-One aspect of special importance is the tools that we use to do data science: the computers, programming languages, and environments.
-The wide release of user-friendly Large Language Models (LLMs), particularly OpenAI's ChatGPT, is a rare instance of a new and powerful tool being made available to add to a data scientist's toolbelt.
+One aspect of special importance is the tools that we use to do data science: the computers, programming languages, and environments among other factors.
+The wide release of user-friendly Large Language Models (LLMs), particularly OpenAI's ChatGPT, is the rare instance of a new tool being added to the data science toolkit.
 
 Like all new tools, however, LLMs have created both excitement and apprehension.
-ChatGPT's public release on November 30, 2022, brought LLMs into mainstream conversation, quickly being put to use in both industry and academia.
+ChatGPT's public release on November 30, 2022, brought LLMs into the mainstream conversation.
+LLMs have quickly been put to use in both industry and academia.
 By now, many people, especially educators and students, have some experience with LLMs, in both personal and professional contexts.
@@
 For instance, there is considerable interest in the potential of chatbots to act as personalized tutors for students [@Winkler_2018].
-ChatGPT has also been the catalyst for many interesting and important conversations around academic integrity [@Okaibedi_2023], the development of critical thinking skills, and what effective learning looks like [@BaidooAnu_2023].
+ChatGPT has also been the catalyst for many interesting and important conversations around academic integrity [@Eke2023], the development of critical thinking skills, and what effective learning looks like [@BaidooAnu_2023].
 
@@
 
-At the high school level @Lazar_2023 conducted a informal survey of secondary school teachers and students on their opinions of ChatGPT, and found that while LLMs could: help creativity, provide academic support when teachers were unavailable, and model certain types of writing well; teachers were also aware of LLMs potential to limit students' learning in certain ways through over-reliance.
+At the high school level @Lazar_2023 conducted an informal survey of secondary school teachers and students on their opinions of ChatGPT, and found that while LLMs could: help creativity, provide academic support when teachers were unavailable, and model certain types of writing well; teachers were also aware of LLMs potential to limit students' learning in certain ways through over-reliance.
 Beyond academic integrity concerns, teachers had similar concerns to @Valenzuela_2024 about de-skilling and an overall loss of agency in writing and critical thinking.
@@
 However, they also found that many students lacked confidence in using AI for academic purposes.
-In particular, only 11% 'strongly agreed' that they know how to use AI to improve their writing.
+In particular, only 11% "strongly agreed" that they know how to use AI to improve their writing.
 Students had nuanced views on appropriate AI use.
@@
       response == "\"This is a lot like me\"" ~ "A lot like me",
-      response == "\"This is very different to me\"" ~ "Very different from me",
+      response == "\"This is very different to me\"" ~ "Very different\nfrom me",
       response == "\"This somewhat describes me\"" ~ "Somewhat describes me",
@@
 
-Those README statements were gathered and parsed using OpenAI's ChatGPT 4o model (as at 26 July).
-The following prompt was used: 'The following statement is about to what extent LLMs were used by a student.
+Those README statements were gathered and parsed using OpenAI's ChatGPT 4o model (as at 26 July). The following prompt was used:
+
+> The following statement is about to what extent LLMs were used by a student.
 Please characterize it as one of: "None", "Minimal", "Somewhat", "Extensive", "Unsure".
-Respond with only one of those options.' All classifications were then manually checked by hand for reasonableness.
+Respond with only one of those options.
+
+All classifications were then manually checked by hand for reasonableness.
 
@@
 
-The third, and final, component is the mark, in percentages, on the final paper (@fig-mark).
+The third, and final, component is the mark, in percentages, on the final paper (@fig-mark-1).
 The overall mean was 78% and standard deviation was 17 percentage points.
-However there was considerable differences by the extent of LLM usage.
-For extensive use, the mean was 82% and the standard deviation was 15 percentage points.
-For somewhat usage, the mean was 77% and the standard deviation was 16 percentage points.
-For minimal usage, the mean was 74% and the standard deviation was 23 percentage points.
-Finally for no usage, the mean was 75% and the standard deviation was 19 percentage points.
+However there was considerable differences by the extent of LLM usage (@tbl-mark; @fig-mark-2).
 
@@
 #| warning: false
-#| fig-cap: Mark on final paper, in percentages, and LLM usage in final paper
-
-# all_data |>
-#   group_by(llm_usage) |>
-#   summarize(mean = mean(mark) |> round(2),
-#             sd = sd(mark) |> round(2))
-#
-# all_data |>
-#   summarize(mean = mean(mark) |> round(2),
-#             sd = sd(mark) |> round(2))
+#| fig-cap: Distribution of marks on final paper (in percentages)
+#| fig-subcap: ["Overall", "By self-reported LLM usage"]
 
 all_data |>
-  ggplot(mapping = aes(x = mark, color = )) +
+  select(mark) |>
+  ggplot(aes(x = mark)) +
+  geom_histogram(binwidth = 0.05) +
+  scale_x_continuous(labels = scales::label_percent()) +
+  labs(x = "Final paper mark (percentage)", y = "Count") +
+  theme_minimal()
+
+all_data |>
+  ggplot(mapping = aes(x = mark)) +
   geom_histogram(binwidth = 0.01) +
@@
 
+```{r tbl-mark}
+#| message: false
+#| echo: false
+#| warning: false
+#| tbl-cap: Self-reported LLM usage in final paper and final paper mark
+
+all_data |>
+  group_by(llm_usage) |>
+  mutate(llm_usage = factor(llm_usage, levels = c(
+    "Extensive", "Somewhat", "Minimal", "None"
+  ))) |>
+  summarize(mean = mean(mark) |> round(2),
+            sd = sd(mark) |> round(2)) |>
+  kable(col.names = c("Self-reported LLM usage", "Mean", "Std dev"))
+```
+
 # Model {#sec-model}
 
-The goal of our modelling strategy is to better understand how respondents result on their final paper is associated with their self-reported LLM usage.
-The result on the final paper is a proportion and some respondents got full marks and so we use zero-one-inflated beta regression.
-Here we briefly describe the model that we use, which follows @kurztotherescue.
-Diagnostics are included in [Appendix -@sec-model-details].
+The goal of our modelling strategy is to better understand how a respondents' result on their final paper associates with their self-reported LLM usage.
+Students received their result on the final paper as a proportion (a 0-100% mark), so we use zero-one-inflated beta regression to best fit the distribution in this interval.
+We use this type of regression given the ability of the beta distribution to effectively describe continuous and skewed or heteroskedastic datasets, which we can see in the asymmetrical histogram of the distribution of marks.
+
+Transforming the regression with a logit function as the link function ensures that the predicted values stay between zero and one, respecting the boundaries of the data.
+Some students got full marks for the paper, so we use the zero-one-inflated beta model.
+
+Here we briefly describe the model that we use, which follows @kurztotherescue. Model diagnostics are included in [Appendix -@sec-model-details].
 
 Define $y_i$ as the percentage received on the final paper. Then $\beta_1$ is the characterization of self-reported LLM usage and $\beta_2$ is self-reported GPA.
-
 \begin{align}
@@
 ```{r fig-gpamodeldiagnostics}
-#| fig-height: 7
+#| fig-height: 6
 #| message: false

### 2024-08-11 paper/paper.qmd
Initial changes & added more recent chatbot references

@@
 In the context of teaching statistics and data science, LLMs could be useful for many tasks.
-For instance, there is considerable interest in the potential of chatbots to act as personalized tutors for students [@Winkler_2018].
+For instance, there is considerable interest in the potential of chatbots to act as personalized tutors for students [fulgencio2024developing; afzal2024tailoring; @Winkler_2018].
 ChatGPT has also been the catalyst for many interesting and important conversations around academic integrity [@Eke2023], the development of critical thinking skills, and what effective learning looks like [@BaidooAnu_2023].
@@
 Students had nuanced views on appropriate AI use.
-In particular, respondents found that using it to write whole papers as inappropriate, while using it for basic tasks like general assistance, writing feedback and basic data visualization was perceived more appropriately.
+In particular, respondents found that using it to write whole papers as inappropriate, while using it for basic tasks like general assistance, writing feedback and basic data visualization was perceived as more appropriate.
 
@@
 That one respondent took more than 4,000 minutes to complete the survey, suggesting they took a break while filling it out.
-That respondent is included in the analysis dataset, but without that respondent, the average time to complete the survey was 11 minutes, and the standard deviation was 8 minutes.
+That respondent is included in the analysis dataset, but without them, the average time to complete the survey was 11 minutes, and the standard deviation was 8 minutes.
 
@@
   ) |>
+   mutate(
+     ai_schoolwork_appropriate = factor(ai_schoolwork_appropriate, levels=c("Appropriate", "It depends", "Inappropriate"))
+   ) |>
   count(ai_schoolwork_appropriate) |>
@@
 all_data |>
+  mutate(llm_usage = factor(llm_usage, levels = c("Extensive", "Somewhat", "Minimal", "None"))) |>
   count(llm_usage) |> 
@@
 
-Define $y_i$ as the percentage received on the final paper. Then $\beta_1$ is the characterization of self-reported LLM usage and $\beta_2$ is self-reported GPA.
+Define $y_i$ as the percentage received on the final paper.
+Then $\beta_1$ and $\beta_2$ are the effect of self-reported LLM usage and self-reported GPA on the final paper grade, respectively.
 \begin{align}
@@
 
-# Appendix {-}
+# Appendix {.unnumbered}
 
@@
 9.
-Please rate how much each statement describes you, on a scale from "This is very different to me" to "This is a lot like me" ["This is very different to me"; "This somewhat describes me"; "This is a lot like me"]
+Please rate how much each statement describes you, on a scale from "This is very different to me" to "This is a lot like me" \["This is very different to me"; "This somewhat describes me"; "This is a lot like me"\]
     -   Writing is easy for me
@@
 When answering, please consider whichever programming language you are most familiar with.
-["This is very different to me"; "This somewhat describes me"; "This is a lot like me"]
+\["This is very different to me"; "This somewhat describes me"; "This is a lot like me"\]
     -   Coding is easy for me
@@
 How helpful did you find generative AI tools such as ChatGPT by OpenAI (or equivalents) for each component of STA302?
-["Not helpful"; "Somewhat helpful"; "Very helpful"; "I did not use generative AI for this component"]
+\["Not helpful"; "Somewhat helpful"; "Very helpful"; "I did not use generative AI for this component"\]
     -   Weekly quiz

### 2024-08-12 paper/paper.qmd
Updated first paragraph of introduction

@@
 
-There are many aspects of trustworthy data science, including education, culture, and workflow to name a few.
-One aspect of special importance is the tools that we use to do data science: the computers, programming languages, and environments among other factors.
-The wide release of user-friendly Large Language Models (LLMs), particularly OpenAI's ChatGPT, is the rare instance of a new tool being added to the data science toolkit.
+Trustworthy data science is the practice of conducting data analysis in a transparent, ethical, and reliable manner.
+These principles are reflected in various ways, through education, culture, and workflow.
+To stay current, data scientists must continually update their knowledge of best practices for transparency and reproducibility, foster a culture which values different perspectives and accountability, and critically think about the broader impact of data-driven decisions.
+The wide release of user-friendly Large Language Models (LLMs), particularly OpenAI’s ChatGPT, has transformed the data science toolkit, offering powerful tools for natural language processing and generation, thus enhancing aspects of data analysis and automation.
 

### 2024-08-12 paper/paper.qmd
Reorder rows in Table 2a from Very familiar to Not familiar

@@
     )
+  ) |>
+  mutate(
+     ai_familiarity = factor(ai_familiarity, levels=c("Very familiar", "Somewhat familiar", "Not familiar"))
      ) |>

### 2024-08-12 paper/paper.qmd
Changes to the introduction

@@
 Trustworthy data science is the practice of conducting data analysis in a transparent, ethical, and reliable manner.
-These principles are reflected in various ways, through education, culture, and workflow.
-To stay current, data scientists must continually update their knowledge of best practices for transparency and reproducibility, foster a culture which values different perspectives and accountability, and critically think about the broader impact of data-driven decisions.
-The wide release of user-friendly Large Language Models (LLMs), particularly OpenAI’s ChatGPT, has transformed the data science toolkit, offering powerful tools for natural language processing and generation, thus enhancing aspects of data analysis and automation.
+These principles are upheld in various ways, through ongoing education, adherence to professional norms and, and implementation of reproducible workflows.
+To stay current, data scientists must continually update their knowledge of best practices for transparency and reproducibility, foster a culture which values different perspectives and accountability, and critically consider the broader impact of data-driven decisions.
+Recent advances, such as the wide release of user-friendly Large Language Models (LLMs), particularly OpenAI’s ChatGPT, have transformed the data science toolkit.
+These powerful tools for natural language processing and generation influence various aspects of data analysis and automation, further emphasizing the need for trustworthy practices.
 
-Like all new tools, however, LLMs have created both excitement and apprehension.
+Like all new tools, LLMs have created both excitement and apprehension.
 ChatGPT's public release on November 30, 2022, brought LLMs into the mainstream conversation.
@@
 In the context of teaching statistics and data science, LLMs could be useful for many tasks.
-For instance, there is considerable interest in the potential of chatbots to act as personalized tutors for students [fulgencio2024developing; afzal2024tailoring; @Winkler_2018].
-ChatGPT has also been the catalyst for many interesting and important conversations around academic integrity [@Eke2023], the development of critical thinking skills, and what effective learning looks like [@BaidooAnu_2023].
+For instance, there is considerable interest in the potential of chatbots to act as personalized tutors for students [@fulgencio2024developing; @afzal2024tailoring].
+ChatGPT has also been the catalyst for many interesting and important conversations around academic integrity [@Eke2023], the development of critical thinking skills [@thiga2024generative], and what effective learning looks like [@BaidooAnu_2023].
 

### 2024-08-12 paper/paper.qmd
Minor tweaks throughout paper. Fix model to be one-inflated. Move all data cleaning from qmd to the data cleaning script.

@@
 This research is currently under review by the University of Toronto's Research Ethics Board.
+Contributions: RA: Conceptualization, Data Curation, Formal Analysis, Funding Acquisition, Investigation, Methodology, Project Administration, Software, Visualization, Writing – Original Draft Preparation, Writing – Review & Editing; LC: Formal Analysis, Visualization, Writing – Original Draft Preparation, Writing – Review & Editing; NM: Conceptualization, Methodology; Writing – Original Draft Preparation, Writing – Review & Editing.
 Comments can be sent to: rohan.alexander@utoronto.ca."
@@
 We find some very mild evidence from this observational study that LLM usage may be associated with better scores, but our main conclusion is that there is no clear relationship between more extensive LLM usage and the student's mark.
-Despite the classroom setting used for evaluation, the particular activity of interest reflects the work done by professional data scientists.
+Despite the classroom setting used for evaluation, the particular activity of interest is similar to the work done by professional data scientists.
 Our finding suggests the need for more extensive work evaluating how LLMs can be integrated into the data science workflow in a way that provides value."
@@
 
+library(arrow)
+library(here)
 library(knitr)
-library(quanteda)
-library(readxl)
-library(reshape2)
-library(tidytext)
 library(tidyverse)
-library(tinytex)
+library(scales)
+library(modelsummary)
 
@@
 all_data <-
-  read_excel(here::here("data/merged_data.xlsx"), sheet = "Sheet1") |>
-  janitor::clean_names()
+  read_parquet(here("data/analysis_data.parquet"))
 ```
@@
 
-Like all new tools, LLMs have created both excitement and apprehension.
+Like all new tools LLMs have created both excitement and apprehension.
 ChatGPT's public release on November 30, 2022, brought LLMs into the mainstream conversation.
@@
 The tasks involved in a trustworthy data science workflow can be generally broken down into a number of key competencies [@Adhikari2021Interleaving; @Gibbs2021Building].
-The first is programming, which is done when cleaning, analyzing, visualizing data often using languages like R or Python.
-The second is writing, which is primarily done when communicating results.
-The potential for LLMs to positively affect students' academic performance in data science can be clearly inferred from already-demonstrated effects in a number of those competencies on adjacent professional fields.
+One is programming, which is done when cleaning, analyzing, and visualizing data often using programming languages like R or Python.
+Another is writing, which is primarily done when communicating results.
+The potential for LLMs to positively affect students' academic performance in data science can be clearly inferred from their already-demonstrated capability in those competencies in adjacent professional fields.
 
 In terms of computer programming, @Peng_2023 found a positive impact of GitHub Copilot (an LLM-powered programming assistant) on productivity.
-Specifically, in an experiment involving 95 freelance programmers, they found that that programmers with access to GitHub Copilot completed a standardized programming task 56% faster than the control group.
+Specifically, in an experiment involving 95 freelance programmers, they found that a treatment group of programmers with access to GitHub Copilot completed a standardized programming task 56% faster than the control group.
 Programmers with less experience saw the greatest improvements in productivity.
@@
 @DellAcqua_2023, focusing on management consulting tasks, provide evidence that LLMs can improve writing productivity.
-In a field experiment involving consultants from Boston Consulting Group they found that the use of OpenAI's GPT-4 led to a 25% increase in delivery speed of business tasks (most involving some writing), as well as a 40% increase in human-rated performance on those tasks.
+In a field experiment involving consultants from Boston Consulting Group they found that the use of OpenAI's GPT-4 led to a 25% increase in delivery speed of business tasks, most of which involved some writing, as well as a 40% increase in human-rated performance on those tasks.
 Similar to computer programming, these productivity increases were most pronounced for those with below average performance, with their output increasing by 43%.
 
-On the other hand, @Valenzuela_2024 argue that LLMs lead to a loss of serendipity (which leads to less original work) and de-skilling (primarily with respect to programming ability), among other consequences.
+On the other hand, @Valenzuela_2024 argue that LLMs lead to a loss of serendipity which may lead to less original work, and potentially de-skilling primarily with respect to programming ability, among other consequences.
 These outcomes could negatively affect students' effective learning of data science.
@@
 Similarly, @Tu_2024 acknowledge that LLMs can streamline many parts of a data science workflow.
-With that in mind, they suggest that data-scientists-in-training should shift their self-perspectives from primarily being an 'analyst' to primarily being a 'product manager' responsible for strategic oversight of the analysis carried out by LLMs.
+With that in mind, they suggest that data-scientists-in-training should shift their self-perspectives from primarily being an "analyst" to primarily being a "product manager" responsible for strategic oversight of the analysis carried out by LLMs.
 
-At the high school level @Lazar_2023 conducted an informal survey of secondary school teachers and students on their opinions of ChatGPT, and found that while LLMs could: help creativity, provide academic support when teachers were unavailable, and model certain types of writing well; teachers were also aware of LLMs potential to limit students' learning in certain ways through over-reliance.
-Beyond academic integrity concerns, teachers had similar concerns to @Valenzuela_2024 about de-skilling and an overall loss of agency in writing and critical thinking.
+At the high school level @Lazar_2023 conducted an informal survey of secondary school teachers and students on their opinions of ChatGPT, and found that while LLMs could: help creativity, provide academic support when teachers were unavailable, and model certain types of writing well.
+Teachers were also concerned about the potential of LLMs to limit students' learning in certain ways through over-reliance.
+Beyond academic integrity concerns, @Lazar_2023 found that teachers had similar concerns to @Valenzuela_2024 about de-skilling and an overall loss of agency in writing and critical thinking.
 
-@Cahill_2024 surveyed undergraduate political science students on their attitudes toward and usage of AI tools.
+@Cahill_2024 surveyed undergraduate political science students on their attitudes toward, and usage of, AI tools.
 They found that the use of ChatGPT was widespread.
@@
 Students had nuanced views on appropriate AI use.
-In particular, respondents found that using it to write whole papers as inappropriate, while using it for basic tasks like general assistance, writing feedback and basic data visualization was perceived as more appropriate.
+In particular, many respondents felt that using it to write whole papers was inappropriate, but using it for basic tasks like general assistance, writing feedback and basic data visualization appropriate.
 
@@
 This is based on students' final papers and a survey, conducted in a third-year undergraduate data science course at the University of Toronto.
-By examining how students interact with and perceive LLMs as tools, and how these variables translate into student outcomes, the effects of LLM integration in data science can be more precisely determined, leading to better recommendations for their future development.
+By examining how students interact with and perceive LLMs as tools, and how these variables translate into student outcomes, current practice with regard to LLM integration in data science can be better understood, leading to better recommendations for their future development.
 
@@
 
-To investigate students' usage and attitudes towards LLMs and how they related to their academic performance, a dataset containing their usage/attitudes, coursework, and academic performance was constructed.
+To investigate students' usage and attitudes towards LLMs and how they related to their academic performance, a dataset containing their usage and attitudes, coursework, and academic performance was constructed.
 This was based on three components:
@@
 
-By virtue of pre-requisites of this course the typical student is an upper-year undergraduate.
+The pre-requisites of this course mean that the typical student is an upper-year undergraduate.
 Coding and writing are major parts of the course.
@@
 All writing must be in English.
-The primary aim of the course for students is to create a public portfolio of work they can use to apply for jobs.
+The primary motivation for having students write three papers as the main assessment for the course is to give them the opportunity to create a public portfolio of work they can use to apply for jobs.
 
 Throughout the semester students were encouraged to use LLMs.
-Formal instruction was provided twice during the semester, with a masterclass taught by a computer science faculty member on the ethics of using LLMs (see @Horton2024 for details), and another masterclass taught by a TA on writing with LLMs.
+Formal instruction was provided twice during the semester.
+The first was a masterclass taught by a computer science faculty member on the ethics of using LLMs (see @Horton2024 for details).
+The second was a masterclass taught by a TA on writing with LLMs.
 
 Data was collected from students through an optional end-of-course survey.
-[Appendix -@sec-survey-details] provides all the questions asked in the survey.
+[Appendix -@sec-survey-details] details the questions asked in the survey.
 Whether or not they consented to their data being used, all respondents received a 1% increase in their final course grade for their participation.
 Consenting responses were then matched to their final paper mark, as well as the GitHub repository for their final paper.
-The responses were anonymized by removing any personal references to the students themselves including names, emails, student numbers, and GitHub links.
+The responses were anonymized by removing any personal references to the students themselves including, names, emails, student numbers, and GitHub links.
 
-All data cleaning and analysis was done using the `R` statistical programming language [@citeR], and especially uses the `tidyverse` [@tidyverse], `janitor` [@janitor], `reshape2` [@reshape], and `readxl` [@readxl] packages.
+Data cleaning and analysis was done using the `R` statistical programming language [@citeR], and the `tidyverse` [@tidyverse], `janitor` [@janitor], `reshape2` [@reshape], and `readxl` [@readxl] packages.
 
@@
 Four of those respondents submitted the survey twice, and after removing their second response, 115 responses remained.
-Of those, 15 respondents did not include a statement on LLM usage in the README of the GitHub repository of their final paper, leaving 100 responses that were of use and were merged based on student name.
+Of those, 15 respondents did not include a statement on LLM usage in the README of the GitHub repository of their final paper, leaving 100 responses.
+Finally 7 of those respondents did not provide a usable GPA response.
+This leaves 93 respondents that were of use and were merged based on student name.
 
@@
 #| warning: false
-#| fig-cap: Distribution of respondents' survey response times and GPA
+#| fig-cap: Distribution of respondents' survey response times and self-reported GPA
 #| fig-subcap: ["Distribution of survey response times","\"What is your GPA?\""]
@@
 all_data |>
-  mutate(mins_complete = completion_time - start_time) |>
-  filter(mins_complete < 1000) |> # Remove one student who took 4,000 minutes to complete
-  # summarize(mean = mean(mins_complete),
-  # standard_dev = sd(mins_complete))
-  ggplot(aes(x = mins_complete)) +
+  # summarize(mean = mean(mins_complete/60),
+  # standard_dev = sd(mins_complete/60))
+  ggplot(aes(x = mins_complete/60)) +
   geom_histogram(binwidth = 1) +
@@
 There is a wide distribution of self-reported GPAs (@fig-response-time-and-gpa-2).
-Seven of the 100 respondents included in the analysis dataset did not report their GPA.
-The majority of responses cluster around a B (3.0/4.0), and the average is 3.06 with a standard deviation of 0.55.
+The majority of responses cluster around a B (3.0 out of 4.0), and the average is 3.06 with a standard deviation of 0.55.
 One factor that may affect the range is that the course is required for programs in the Statistics, Mathematics and Computer Science Departments.
@@
 Students from a range of years took the course, however the majority of respondents were in their 3rd or 4th year of study (@tbl-year-dist).
-The course's prerequisite of general statistics, which is a two-course sequence typically completed by students in their second year, would make it difficult to take this course earlier than their 3rd year.
+The course's prerequisite two-course sequence is typically completed by students in their second year, would make it difficult to take this course earlier than their 3rd year.
 
@@
 all_data |>
-  mutate(
-    what_year_are_you = case_when(
-      what_year_are_you == "PEY" ~ "4th",
-      # Change PEY to 4th year
-      what_year_are_you == "5th" ~ "5th or over",
-      what_year_are_you == "6th" ~ "5th or over",
-      TRUE ~ what_year_are_you
-    )
-  ) |>
   count(what_year_are_you) |>
@@
 Respondents had a varied self-perception of their coding and writing abilities (@fig-selfperception).
-Most respondents believed it is important to be good at writing, but many are either indifferent or do not like to write (@fig-selfperception-1).
+Most respondents believe it is important to be good at writing, but many are either indifferent or do not like to write (@fig-selfperception-1).
 Respondents also do not find writing to be particularly easy, which could be associated with the reported relative antipathy toward writing.
-Most respondents were at least somewhat confident in their own writing abilities, but a substantial contingent felt otherwise.
+Most respondents were at least somewhat confident in their own writing abilities, but a substantial number felt otherwise.
 Although few respondents felt that they were confident in their writing ability, more felt that they were able to catch their mistakes, which could indicate a disconnect between how respondents perceive their work and how the work was evaluated.
@@
 Respondent self-perceptions regarding coding proficiency and importance varied (@fig-selfperception-2).
-There was strong consensus on the perceived importance of coding skills, with a majority of respondents strongly affirming this belief.
+There is strong consensus on the perceived importance of coding skills.
 However, respondents' self-assessed coding abilities and enjoyment are more heterogeneous, with a substantial proportion reporting moderate rather than high levels of ease and enjoyment in coding tasks.
-These findings suggest a complicated relationship between respondents' recognition of coding's importance and their personal experience with coding.
 
@@
 all_data |>
-  rename(ai_familiarity = how_familiar_are_you_with_using_generative_ai_tools_such_as_open_a_is_chat_gpt_or_equivalents) |>
-  mutate(
-    ai_familiarity = case_when(
-      ai_familiarity == "Limited to some liberal arts, such as history courses" ~ "Somewhat familiar",
-      TRUE ~ ai_familiarity
-    )
-  ) |>
-  mutate(
-     ai_familiarity = factor(ai_familiarity, levels=c("Very familiar", "Somewhat familiar", "Not familiar"))
-     ) |>
   count(ai_familiarity) |>
@@
 all_data |>
-  rename(ai_schoolwork_appropriate = to_what_extent_do_you_think_using_generative_ai_tools_such_as_chat_gpt_by_open_ai_or_equivalents_is_ethical_and_appropriate_for_schoolwork) |>
-  mutate(
-    ai_schoolwork_appropriate = case_when(
-      ai_schoolwork_appropriate == "Depends of what the main goal of schoolwork is" ~ "It depends",
-      ai_schoolwork_appropriate == "I believe student can only use it if they are allowed to" ~ "It depends",
-      ai_schoolwork_appropriate == "a controllable amount" ~ "It depends",
-      ai_schoolwork_appropriate == "appropriate depending on the context" ~ "It depends",
-      ai_schoolwork_appropriate == "appropriate with rules for usage" ~ "It depends",
-      ai_schoolwork_appropriate == "depends" ~ "It depends",
-      ai_schoolwork_appropriate == "depends on how it is used" ~ "It depends",
-      ai_schoolwork_appropriate == "half and half" ~ "It depends",
-      ai_schoolwork_appropriate == "unsure" ~ "It depends",
-      TRUE ~ ai_schoolwork_appropriate
-    )
-  ) |>
-   mutate(
-     ai_schoolwork_appropriate = factor(ai_schoolwork_appropriate, levels=c("Appropriate", "It depends", "Inappropriate"))
-   ) |>
   count(ai_schoolwork_appropriate) |>
@@
 
-To understand the role of LLMs in learning, students identified their usage in a more granular way by checking off various pre-defined use cases in the survey (@fig-AI-use-and-usefulness-1).
+To understand the role of LLMs in learning, students identified their usage in a more granular way by selecting various pre-defined use cases in the survey (@fig-AI-use-and-usefulness-1).
 Technical questions and explaining concepts were the two top use cases among students in the course.
@@
 all_data |>
-  rename(how_have_you_used = if_you_have_used_generative_ai_tools_such_as_open_a_is_chat_gpt_or_equivalents_in_what_ways_have_you_used_it_select_all_that_apply) |>
   select(how_have_you_used) |>
@@
 Respondents were also asked to rate the helpfulness of LLMs on various tasks assigned during the course on a 4-point scale of "Did not use" to "Very Helpful" (@fig-AI-use-and-usefulness-2).
-To simplify the presentation, responses were grouped into two main categories: "Less Helpful" and "More Helpful." The "Less Helpful" category combines responses where students found the AI either "Not helpful" or did not use it for the task, while the "More Helpful" category includes responses where the AI was considered "Somewhat helpful" or "Very helpful".
+To simplify the presentation, responses were grouped into two main categories: "Less Helpful" and "More Helpful." The "Less Helpful" category combines responses where students found the AI either "Not helpful" or did not use it for the task, while the "More Helpful" category includes responses where the LLM was considered "Somewhat helpful" or "Very helpful".
 
@@
 While most tasks were roughly split between respondents finding LLMs helpful or not, respondents found them most helpful in generating code.
-In the context of the course, this meant generating R code for transforming, analyzing, and visualizing data.
+In the context of the course, this mostly meant generating R code for transforming, analyzing, and visualizing data.
 To a lesser extent, respondents also found LLMs to be helpful in improving the existing writing they had, while at the same time not favouring it for writing content from scratch.
@@
 Even students who did not use generative AI at all were required to state this in the README.
-For students who did use generative AI, there was an additional requirement that they save the logs of their usage in a txt file which was also included in their GitHub repository.
+For students who did use generative AI, there was an additional requirement, where possible, that they save the logs of their usage in a txt file which was also included in their GitHub repository.
 
@@
 
-All classifications were then manually checked by hand for reasonableness.
+All classifications were then manually checked for reasonableness.
 
@@
 all_data |>
-  mutate(llm_usage = factor(llm_usage, levels = c("Extensive", "Somewhat", "Minimal", "None"))) |>
   count(llm_usage) |> 
@@
   geom_histogram(binwidth = 0.05) +
-  scale_x_continuous(labels = scales::label_percent()) +
+  scale_x_continuous(labels = label_percent()) +
   labs(x = "Final paper mark (percentage)", y = "Count") +
@@
   geom_histogram(binwidth = 0.01) +
-  scale_x_continuous(labels = scales::label_percent()) +
+  scale_x_continuous(labels = label_percent()) +
   labs(x = "Final paper mark (percentage)", y = "Count") +
@@
   group_by(llm_usage) |>
-  mutate(llm_usage = factor(llm_usage, levels = c(
-    "Extensive", "Somewhat", "Minimal", "None"
-  ))) |>
   summarize(mean = mean(mark) |> round(2),
@@
 fit1 <-
-  readRDS(file = here::here("models/fit1brms.rds"))
+  readRDS(file = here("models/fit1brms.rds"))
 fit2 <-
-  readRDS(file = here::here("models/fit2brms.rds"))
+  readRDS(file = here("models/fit2brms.rds"))
 
-modelsummary::modelsummary(
+modelsummary(
   list(
@@
 
-modelsummary::modelplot(
+modelplot(
   list(
@@
   labs(x= "") +
-  scale_color_brewer(palette = "Set1")
+  scale_color_brewer(palette = "Set1") +
+  theme(legend.position = "bottom")
 ```

### 2024-08-12 scripts/01-data_cleaning.R

@@ -1,278 +1,114 @@
 #### Preamble ####
-# Purpose: Cleans the raw survey data recored by Rohan at the end of STA302
-# Author: Luca Carnegie
-# Date: 16 July 2024
-# Contact: luca.carnegie@mail.utoronto.ca or rohan.alexander@utoronto.ca
-# Pre-requisites: 01-download_data.R
+# Purpose: Prepare the data
+# Author: Rohan Alexander
+# Date: 12 August 2024
+# Contact: rohan.alexander@utoronto.ca
+
 
 #### Workspace setup ####
+library(here)
 library(tidyverse)
 library(janitor)
 library(stringr)
-library(lintr)
+library(readxl)
+library(arrow)
+
+
+#### Read in data ####
+all_data <-
+  read_excel(here("data/merged_data.xlsx"), sheet = "Sheet1") |>
+  clean_names()
 
 
 #### Clean data ####
-raw_data <- read_csv("data/raw_data/raw_STA302_postcourse_survey_w24.csv")
+# Add completion time
+all_data <-
+  all_data |>
+  mutate(mins_complete = completion_time - start_time) |>
+  filter(mins_complete < 1000) # Remove one student who took 4,000 minutes to complete
 
-raw_data <- as_tibble(raw_data)
-
-# CLEAN UP COLUMN NAMES
-
-raw_data <- raw_data |>
-  clean_names()
-
-raw_data <- raw_data |>
-  rename(
-    consent = after_carefully_reading_the_informed_consent_document_please_indicate_below_whether_you_consent_to_have_your_anonymized_responses_included_in_the_research_study,
-    student_name = what_is_your_full_name_on_quercus,
-    student_id = what_is_your_student_id,
-    year = what_year_are_you,
-    program_of_study = what_is_your_specialization,
-    major = what_is_are_your_major_s,
-    minor = what_is_are_your_minor_s,
-    gpa = what_is_your_gpa,
-    ai_familiarity = how_familiar_are_you_with_using_generative_ai_tools_such_as_open_a_is_chat_gpt_or_equivalents,
-    used_ai_any_reason = have_you_used_any_generative_ai_tools_such_as_open_a_is_chat_gpt_or_equivalents_for_any_reason_personal_or_educational,
-    ai_use_cases = if_you_have_used_generative_ai_tools_such_as_open_a_is_chat_gpt_or_equivalents_in_what_ways_have_you_used_it_select_all_that_apply,
[…]

### 2024-08-12 scripts/03-model.R

@@ -12,30 +12,31 @@
 library(tidyverse)
 library(rstanarm)
 library(brms)
-library(readxl)
+library(arrow)
 
 
 #### Read data ####
 all_data <-
-  read_excel(here::here("data/merged_data.xlsx"), sheet = "Sheet1") |>
-  janitor::clean_names() |>
-  mutate(mark = if_else(mark == 1, 0.99, mark)) |> # I'm a monster
-  mutate(llm_usage =
-           if_else(llm_usage %in% c("Minimal", "None"), "None or minimal", llm_usage)) |>
-  filter(!is.na(what_is_your_gpa))
+  read_parquet(here::here("data/analysis_data.parquet"))
 
 
-### Model data ####
+#### Model data ####
+#### Simplified data ####
+# Change the 1s to 0.99 so we can just use Beta initially
+all_data <-
+  all_data |>
+  mutate(mark_no_ones = if_else(mark == 1, 0.99, mark))
+
 # Using rstanarm
 fit1rstanarm <-
-  stan_betareg(mark ~ llm_usage,
+  stan_betareg(mark_no_ones ~ llm_usage,
                data = all_data,
                link = "logit",
                seed = 853)
 
 fit2rstanarm <-
   stan_betareg(
-    mark ~ llm_usage + what_is_your_gpa,
+    mark_no_ones ~ llm_usage + what_is_your_gpa,
     data = all_data,
     link = "logit",
     seed = 853
@@ -45,7 +46,7 @@ fit2rstanarm <-
 fit1brms <- brm(
   data = all_data,
   family = Beta,
-  mark ~ llm_usage,
+  mark_no_ones ~ llm_usage,
   prior = prior(normal(0, 1), class = b) +
     prior(gamma(4, 0.1), class = phi),
   cores = 4,
@@ -55,7 +56,7 @@ fit1brms <- brm(
 fit2brms <- brm(
   data = all_data,
   family = Beta,
-  mark ~ llm_usage + what_is_your_gpa,
+  mark_no_ones ~ llm_usage + what_is_your_gpa,
   prior = prior(normal(0, 1), class = b) +
     prior(gamma(4, 0.1), class = phi),
   cores = 4,
@@ -63,7 +64,39 @@ fit2brms <- brm(
 )
 
 
-#### Save model ####
+#### Real data ####
[…]

### 2024-08-13 paper/paper.qmd
Update model and add first pass at discussion and results.

@@
 library(modelsummary)
+library(bayesplot)
 
@@
 Students had nuanced views on appropriate AI use.
-In particular, many respondents felt that using it to write whole papers was inappropriate, but using it for basic tasks like general assistance, writing feedback and basic data visualization appropriate.
+Many respondents felt that using it to write whole papers was inappropriate, but using it for basic tasks like general assistance, writing feedback and basic data visualization appropriate.
 
@@
 
+OLD:
+
 Define $y_i$ as the percentage received on the final paper.
@@
 \end{align}
+
+NEW:
+
+Let $Y_i$ represent the dependent variable "mark" for observation $i$. The model assumes a zero-one inflated beta distribution for $Y_i$, which can be represented as:
+
+$Y_i \sim \text{Zero-One Inflated Beta}(\mu_i, \phi_i, \pi_{zoi_i}, \pi_{coi_i})$
+
+Where:
+
+- $\mu_i$ is the mean of the beta distribution (related to the location parameter).
+- $\phi_i$ is the precision parameter (related to the dispersion of the beta distribution).
+- $\pi_{zoi_i}$ is the probability of $Y_i$ being exactly 0 or 1 (zero-one inflation component).
+- $\pi_{coi_i}$ is the probability of $Y_i$ being in the continuous part of the beta distribution (continuous component inflation).
+
+The specific linear models for each parameter are given by:
+
+1. **Mean (Location) Component:**
+   $logit(\mu_i) = \beta_{0,\mu} + \beta_{1,\mu} \cdot \text{llm\_usage}_i + \beta_{2,\mu} \cdot \text{what\_is\_your\_gpa}_i$
+
+2. **Precision Component:**
+   $\log(\phi_i) = \beta_{0,\phi} + \beta_{1,\phi} \cdot \text{llm\_usage}_i + \beta_{2,\phi} \cdot \text{what\_is\_your\_gpa}_i$
+
+3. **Zero-One Inflation Component:**
+   $logit(\pi_{zoi_i}) = \beta_{0,zoi} + \beta_{1,zoi} \cdot \text{llm\_usage}_i + \beta_{2,zoi} \cdot \text{what\_is\_your\_gpa}_i$
+
+4. **Continuous Outcome Inflation (coi) Component:**
+   $logit(\pi_{coi_i}) = \beta_{0,coi}$
+
+In this model:
+
+- $\beta_{0,\mu}, \beta_{0,\phi}, \beta_{0,zoi},$ and $\beta_{0,coi}$ are the intercepts for the respective components.
+- $\beta_{1,\mu}, \beta_{1,\phi}, \beta_{1,zoi}$ represent the coefficients for the effect of "llm\_usage" on the respective components.
+- $\beta_{2,\mu}, \beta_{2,\phi}, \beta_{2,zoi}$ represent the coefficients for the effect of "what\_is\_your\_gpa" on the respective components. 
+- The term $logit(\cdot)$ denotes the logit link function, commonly used for modeling probabilities.
+
+This notation captures the model's structure as defined in the `brms` code, with each parameter modeled as a function of the predictors "llm\_usage" and "what\_is\_your\_gpa", except for the continuous outcome inflation (coi) component, which is modeled with just an intercept.
 
@@
 fit1 <-
-  readRDS(file = here("models/fit1brms.rds"))
+  readRDS(file = here("models/fit_one_inflated.rds"))
 fit2 <-
-  readRDS(file = here("models/fit2brms.rds"))
+  readRDS(file = here("models/fit_one_inflated_gpa.rds"))
 
@@
 #| warning: false
-#| fig-cap: Coefficient estimates and 90 per cent credibility intervals
+#| fig-cap: Selected coefficient estimates and 90 per cent credibility intervals
 
@@
   conf_level = 0.9,
-  coef_omit = "phi") +
+  drop = c("phi", "b_zoi_Intercept")) +
   labs(x= "") +
@@
 
-Interestingly, we find a slightly negative effect, relative to self-reported extensive use, of using LLMs less (@fig-model-results).
-As expected, self-reported GPA is positively associated with final mark.
+@tbl-model-results and @fig-model-results present the results from two models estimating the association between LLM usage, GPA and final paper mark.
+The first model includes only LLM usage, while the second model also includes self-reported GPA.
+
+LLM usage estimates are in relation to "Extensive" LLM usage.
+In the base model, the coefficients suggest that less usage of LLMs was associated with lower scores.
+The second model retains this slight association, but with slightly smaller coefficients.
+For instance, the negative impact of “b_llm_usageSomewhat” is slightly reduced from -0.35 to -0.29, indicating that part of the effect observed in the base model might be explained by the GPA variable.
+
+Including self-reported GPA in the second model does considerably change the  changes the interpretation of some coefficients.
+For instance, intercept for "b_zoi_Intercept" becomes much more negative in the GPA-included model, suggesting that the GPA significantly affects the zero-one inflation part of the model.
+However, the essential relationship between LLM usage and final mark is retained.
+
+The overall fit of the model improves when GPA is included.
+There is an increase in R-squared from 0.095 to 0.478 and improvements in the ELPD and LOOIC metrics.
+This is also suggested by the LOO comparison between the two models (@tbl-loo in [Appendix -@sec-model-details]).
 
@@
 However, this work should mostly be treated as a direction for future research in this area.
-More broadly, our work has three major take-aways: 1) LLMs have a major role to play in enhancing the trustworthiness of data science.
-2) Making general purpose chatbots more opinionated and knowledgeable about data science would considerably help (i.e.
-fine-tuning them - maybe talk more about the GPT store and the "Telling Stories with Data chatbot") 3) Alternative interaction approaches, beyond a chatbot interface, could be especially useful.
+More broadly, our work has three major take-aways: 1) LLMs could help enhance the trustworthiness of data science, but not without considerable futher development.
+2) Making general purpose chatbots more opinionated and knowledgeable about data science would considerably help.
+3) Alternative interaction approaches, beyond a chatbot interface, could be especially useful.
 
@@
 
+We found that many students did use LLMs to help them complete a data science project closely related to what professional data scientists do.
+A common use case to ask the LLM to write code to make a graph.
+Another was to ask for some example paragraphs.
+
+Data science code tends to be written differently to software engineering code.
+Functions and loops are used less, as there is often less reuse of code.
+The training data of these LLMs is dominated by software engineering code, and so when an LLM is asked to write code for a data science project it tends to produce code with some of those features.
+This is neither good nor bad in and of itself, however it can clash with the style of the code that surrounds it.
+
+LLMs are also being used as an alternative search engine.
+Students often asked, for instance, how to modify a graph to add a vertical line, or similar.
+Here LLMs tended to be especially useful and able to answer questions.
+In the past students may have used StackOverflow or similar to find their answers, but the nature of the chat response likely provided a faster answer.
+
+Finally, there is the question of whether the nature of the data science has fundamentally changed or improved.
+For now, there was no evidence of that.
+It was clear when a student had just used LLM-generated text without modification, and it was often clear when large amounts of code had been written by an LLM.
+But it was not obvious when small amounts of either had been used, or when LLM-generated content had just been used as an initial draft.
+
+Many of these issues were alluded to by @bommasani2021opportunities, in the context of data science, and it is striking that they remain today.
+
 ## General purpose compared with opinionated specific chatbots
 
-## On the need for LLM-based infrastructure
+ChatGPT and equivalents are general-purpose chatbots.
+One can ask them a technical data science question and then in the same session ask for a brownie recipe!
+However this general-purpose nature means that they are not necessarily optimized toward one particular task [@thoppilan2022lamda].
+When used for data science, this means that sometimes the functions or libraries recommended may not exist, code may not run, or out-dated approaches may be used.
+Frontier models make things up and are unnervingly confident, even when wrong [@openai2023gpt4, p.
+59], which can create special issues for learners.
+
+General-purpose chatbots being used for data science are fine for experienced data scientists who have an established sense of when something might be not ideal or even wrong.
+But it is problematic for novices who may not know, can lack the confidence to push back, and possibly, may learn the wrong approach.
+
+Retrieval-augmented generation (RAG) means providing various sources, for instance books or manuals, which the LLM uses to base its response on.
+Fine-tuning means providing the previously trained LLM with a large number of example input and outputs, which again guide the response from the LLM [@raffel2020exploring].
+A chatbot based interface, focused on the needs of data science beginners, could use RAG and/or fine-tuning to considerably improve the quality of its responses.
+The trade-off would be a reduction in the general-purpose usability of the chatbot, but in the context of this use this that trade-off is likely acceptable.
+
+## Changed LLM-based interfaces
+
+LLMs existed before ChatGPT, but it was the launch of ChatGPT that brought them into the mainstream.
+The underlying model was not considerably different to others and the main difference that ChatGPT brought was the chat interaction.
+These are useful, but there is considerable opportunity to develop additional ways of interacting with LLMs.
+This is particularly relevant for enhancing trustworthy data science.
+
+Search-based LLM tools like Elicit and Perplexity.ai differ to ChatGPT in that they focus on search.
+In contrast to Google, which typically provides a list of websites related to a particular search, Perplexity.ai attempts to provide an answer.
+In contrast to ChatGPT, search-based LLM tools focus on answering a question rather than engaging in discussion.
+
+Students in this class were required to use GitHub to host their papers, as well as supporting code and data.
+GitHub Issues and Pull Requests were used for peer review and submission occurred based on links to GitHub repositories.
+Trustworthy data science could be considerably enhanced by establishing something analogous to continuous integration and a suite of tests, but for data science.
+The once-off nature of much data science code would mean that LLMs could be especially useful in this proposed infrastructure stack.
+For instance, when a data scientist committed their code or writing, a suite of static tests could run, for instance to check that the data are being read in, that classes are appropriate, etc.
+But additionally LLMs could be used to check for possible improvements in writing, and propose context-specific improvements in code.
+For instance, to make sure that what is being described in the paper is actually what is happening in the code.
+
+Finally, one exciting area of current development in LLMs is tool use.
+One early example of this was enabling LLMs to use search engines such as Google.
+As LLMs can produce writing, and writing is the input to a Google Search, enabling this tool broadened the types of queries that LLMs could respond effectively to.
+Tool use, focused on data science, exists.
+For instance, LLMs can use R and Python to respond to queries about a dataset.
+However, further development could be useful.
+For instance, when a user wrote code in an IDE to import some data, a tool-enhanced LLM could notice and automatically write code that establishes a Pydantic-based validation model for that data.
+Similarly, consider an LLM in an IDE context where a data scientist wrote out the model, in statistical notation, that they were interested in estimating, and the LLM was able write the model in Stan, run it, save the estimates, and add a summary table to the paper, all automatically and in the background.
+Similar work has occured in other contexts and found to bring substantial benefits [@chen2022evaluating].
 
@@
 
-There are a number of substantial weaknesses of this study.
+There are several substantial weaknesses of this study.
 The foundational one is that we used observational data, much of which was self-reported.
@@
 
-We use `bayesplot` [@bayesplot] and `loo` [@loo] help with posterior predictive checks and model diagnostics.
+We use `bayesplot` [@bayesplot] and `loo` [@loo] conduct posterior predictive checks and evaluate model diagnostics.
 
@@
 
-bayesplot::pp_check(fit1) +
+pp_check(fit1) +
   theme_classic() +
@@
 
-bayesplot::pp_check(fit2) +
+pp_check(fit2) +
   theme_classic() +
@@
 
-brms::loo_compare(loo::loo(fit1, cores = 2), loo::loo(fit2, cores = 2))
+brms::loo_compare(loo::loo(fit1, cores = 2), loo::loo(fit2, cores = 2)) |>
+  kable(digits = 3)
 ```

### 2024-08-14 paper/paper.qmd

@@
 This research is currently under review by the University of Toronto's Research Ethics Board.
-Contributions: RA: Conceptualization, Data Curation, Formal Analysis, Funding Acquisition, Investigation, Methodology, Project Administration, Software, Visualization, Writing – Original Draft Preparation, Writing – Review & Editing; LC: Formal Analysis, Visualization, Writing – Original Draft Preparation, Writing – Review & Editing; NM: Conceptualization, Methodology; Writing – Original Draft Preparation, Writing – Review & Editing.
+Contributions: RA: Conceptualization, Data Curation, Formal Analysis, Funding Acquisition, Investigation, Methodology, Project Administration, Software, Visualization, Writing – Original Draft Preparation, Writing – Review & Editing; LC: Formal Analysis, Visualization, Writing – Original Draft Preparation, Writing – Review & Editing; NM: Conceptualization, Methodology, Writing – Original Draft Preparation, Writing – Review & Editing.
 Comments can be sent to: rohan.alexander@utoronto.ca."
@@
 
+*NM Comments (I'm not familiar with 0-1 inflated beta regression so I'll leave these as comments to be considered)*
+
+- The text refers to a zero-one inflated beta, but the code chunk below is for a one-inflated model (linked to the zero-or-one inflated beta referenced, if I understand correctly).
+I think the text should be updated
+- Is the Kurz (2023) reference correct? It looks like this is for standard beta regression, so perhaps an additional reference to one-inflated beta regression would be helpful (e.g. Ospina 2012)
+- Regarding $\pi_{zoi}$ and $\pi_{coi}$ - based on the descriptions it seems like these should sum to 1, but if that's the case it's not clear to me why we would model both.
+I suspect I'm missing something here, but tweaking the descriptions could help
+- llm_usage is in the model as a categorical variable, do we want to make that explicit in the model descriptions?
+- Related to the above, perhaps it makes more sense to have "None" as the reference level for llm_usage instead of "Extensive"
+- I typically prefer not to have raw variable names in the body of a paper, so if there's a way to do that without being to cumbersome that could be good (perhaps a personal preference though)
+- Similarly, in Table 5 is it possible to use the mathematical notation instead of the variable names?
+- In Figure 5 is it possible to have different symbols for the two models to make it more readable in grayscale?
+- [Related to section 5.1]: There are references to how students used LLMs (e.g.
+to write code/text, as an alternative search engine), and I'm just curious where this information comes from.
+Were the LLM usage statements submitted and analysed, or is this more annecdotal?
+Either way, I think a bit more context about where it came from would help give the reader context.
+
 # Results {#sec-results}
@@
   theme(legend.position = "bottom")
+
+# Can you change the symbol for one of the models in the above plot?
+
 ```
@@
 
-Data science code tends to be written differently to software engineering code.
+Data science code tends to be written differently than software engineering code.
 Functions and loops are used less, as there is often less reuse of code.
@@
 Finally, there is the question of whether the nature of the data science has fundamentally changed or improved.
-For now, there was no evidence of that.
+For now, there is no evidence of that.
 It was clear when a student had just used LLM-generated text without modification, and it was often clear when large amounts of code had been written by an LLM.
@@
 
-Many of these issues were alluded to by @bommasani2021opportunities, in the context of data science, and it is striking that they remain today.
+Many of these issues were alluded to by @bommasani2021opportunities, in the context of data science, and it is striking that they remain relevant today.
 
@@
 Fine-tuning means providing the previously trained LLM with a large number of example input and outputs, which again guide the response from the LLM [@raffel2020exploring].
-A chatbot based interface, focused on the needs of data science beginners, could use RAG and/or fine-tuning to considerably improve the quality of its responses.
-The trade-off would be a reduction in the general-purpose usability of the chatbot, but in the context of this use this that trade-off is likely acceptable.
+A chatbot-based interface, focused on the needs of data science beginners, could use RAG and/or fine-tuning to considerably improve the quality of its responses.
+The trade-off would be a reduction in the general-purpose usability of the chatbot, but in the context of this use, such a trade-off is likely acceptable.
 
@@
 Trustworthy data science could be considerably enhanced by establishing something analogous to continuous integration and a suite of tests, but for data science.
-The once-off nature of much data science code would mean that LLMs could be especially useful in this proposed infrastructure stack.
-For instance, when a data scientist committed their code or writing, a suite of static tests could run, for instance to check that the data are being read in, that classes are appropriate, etc.
+The one-off nature of much data science code would mean that LLMs could be especially useful in this proposed infrastructure stack; for instance, when a data scientist commits their code or writing, a suite of static tests could run, for instance to check that the data are being read in, that classes are appropriate, etc.
 But additionally LLMs could be used to check for possible improvements in writing, and propose context-specific improvements in code.

### 2024-08-14 paper/paper.qmd
mild edits

@@
 Students from a range of years took the course, however the majority of respondents were in their 3rd or 4th year of study (@tbl-year-dist).
-The course's prerequisite two-course sequence is typically completed by students in their second year, would make it difficult to take this course earlier than their 3rd year.
+The course's prerequisite two-course sequence is typically completed by students in their second year, which makes it difficult for students to take this course earlier than their 3rd year.
 
@@
 
+# so that "Very different from me" is Red - better visuals
+set1_palette <- c("#4DAF4A", "#377EB8", "#E41A1C")
+
 writing_perceptions_long |>
@@
   labs(x = "Response", y = "Number of repondents", fill = "Response:") +
-  scale_fill_brewer(palette = "Set1") +
+  scale_fill_manual(values = set1_palette) +
   theme_minimal() +
@@
 More than half of students also used LLMs for quick questions, general knowledge, writing paper code, and checking solutions.
-Just less than half used it to write paper content, which could suggest that students do not feel confident using it to improve writing.
+Just less than half used it to write paper content, which could suggest that students do not feel confident using it to improve writing; this could be similar to the attitudes expressed by students surveyed in @Cahill_2024.
 
@@
 The goal of our modelling strategy is to better understand how a respondents' result on their final paper associates with their self-reported LLM usage.
-Students received their result on the final paper as a proportion (a 0-100% mark), so we use zero-one-inflated beta regression to best fit the distribution in this interval.
-We use this type of regression given the ability of the beta distribution to effectively describe continuous and skewed or heteroskedastic datasets, which we can see in the asymmetrical histogram of the distribution of marks.
+Students received their result on the final paper as a proportion (a 0-100% mark), so we use a Bayesian zero-one-inflated beta regression to best fit the distribution in this interval.
+We use this type of regression given the ability of the beta distribution to effectively describe continuous and skewed  (heteroskedastic) datasets, which we can see in the asymmetrical histogram of the distribution of marks.
 
 Transforming the regression with a logit function as the link function ensures that the predicted values stay between zero and one, respecting the boundaries of the data.
-Some students got full marks for the paper, so we use the zero-one-inflated beta model.
 
-Here we briefly describe the model that we use, which follows @kurztotherescue. Model diagnostics are included in [Appendix -@sec-model-details].
+We briefly describe the model that we use, which follows @kurztotherescue.
+No students got zeroes, though some got full marks (100%) one the paper, so we adapt the regression to one-inflated beta model as described by @heisstotherescue to account for this.
+
+Model diagnostics are included in [Appendix -@sec-model-details].
 
@@
 In the base model, the coefficients suggest that less usage of LLMs was associated with lower scores.
-The second model retains this slight association, but with slightly smaller coefficients.
+The second model retains this association, but with slightly smaller coefficients.
 For instance, the negative impact of “b_llm_usageSomewhat” is slightly reduced from -0.35 to -0.29, indicating that part of the effect observed in the base model might be explained by the GPA variable.

### 2024-09-27 paper/paper.qmd
Update based on 2024 student feedback

--- 2024-08-13 paper/paper.qmd
@@
 ---
-title: "Self-reported LLM usage and results on a data science project: Evidence from a Canadian undergraduate data science course"
+title: "My title"
+subtitle: "My subtitle if needed"
 author: 
-  - Rohan Alexander
-  - Luca Carnegie
-  - Nathalie Moon
-thanks: "Code and data are available at: https://github.com/lcarnegie/llms-achievement.
-We thank Tiffany Timbers and attendees at JSM 2024 for helpful suggestions.
-This research is currently under review by the University of Toronto's Research Ethics Board.
-Contributions: RA: Conceptualization, Data Curation, Formal Analysis, Funding Acquisition, Investigation, Methodology, Project Administration, Software, Visualization, Writing – Original Draft Preparation, Writing – Review & Editing; LC: Formal Analysis, Visualization, Writing – Original Draft Preparation, Writing – Review & Editing; NM: Conceptualization, Methodology; Writing – Original Draft Preparation, Writing – Review & Editing.
-Comments can be sent to: rohan.alexander@utoronto.ca."
+  - First author
+  - Another author
+thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
 date: today
 date-format: long
-abstract: "To help understand the effect of Large Language Models (LLMs) on data science practice we examine the extent to which LLM usage is correlated with the mark that a student gets on a final paper in a classroom data science setting.
-We find some very mild evidence from this observational study that LLM usage may be associated with better scores, but our main conclusion is that there is no clear relationship between more extensive LLM usage and the student's mark.
-Despite the classroom setting used for evaluation, the particular activity of interest is similar to the work done by professional data scientists.
-Our finding suggests the need for more extensive work evaluating how LLMs can be integrated into the data science workflow in a way that provides value."
+abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
 format: pdf
 number-sections: true
-table-of-contents: false
 bibliography: references.bib
@@
 
-```{r setup}
+```{r}
 #| include: false
+#| warning: false
+#| message: false
 
-library(arrow)
-library(here)
-library(knitr)
 library(tidyverse)
-library(scales)
-library(modelsummary)
-library(bayesplot)
-
-# Read in data
-all_data <-
-  read_parquet(here("data/analysis_data.parquet"))
+library(palmerpenguins)
 ```
@@
 
-Trustworthy data science is the practice of conducting data analysis in a transparent, ethical, and reliable manner.
-These principles are upheld in various ways, through ongoing education, adherence to professional norms and, and implementation of reproducible workflows.
-To stay current, data scientists must continually update their knowledge of best practices for transparency and reproducibility, foster a culture which values different perspectives and accountability, and critically consider the broader impact of data-driven decisions.
-Recent advances, such as the wide release of user-friendly Large Language Models (LLMs), particularly OpenAI’s ChatGPT, have transformed the data science toolkit.
-These powerful tools for natural language processing and generation influence various aspects of data analysis and automation, further emphasizing the need for trustworthy practices.
+You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.
 
-Like all new tools LLMs have created both excitement and apprehension.
-ChatGPT's public release on November 30, 2022, brought LLMs into the mainstream conversation.
-LLMs have quickly been put to use in both industry and academia.
-By now, many people, especially educators and students, have some experience with LLMs, in both personal and professional contexts.
-
-In the context of teaching statistics and data science, LLMs could be useful for many tasks.
-For instance, there is considerable interest in the potential of chatbots to act as personalized tutors for students [@fulgencio2024developing; @afzal2024tailoring].
-ChatGPT has also been the catalyst for many interesting and important conversations around academic integrity [@Eke2023], the development of critical thinking skills [@thiga2024generative], and what effective learning looks like [@BaidooAnu_2023].
-
-In this paper we are interested in better understanding LLMs as a tool for producing trustworthy data science.
-We study how they were used by students in an upper-year undergraduate data science course and whether students who used LLMs tended to have higher scores than those who did not.
-
-The tasks involved in a trustworthy data science workflow can be generally broken down into a number of key competencies [@Adhikari2021Interleaving; @Gibbs2021Building].
-One is programming, which is done when cleaning, analyzing, and visualizing data often using programming languages like R or Python.
-Another is writing, which is primarily done when communicating results.
-The potential for LLMs to positively affect students' academic performance in data science can be clearly inferred from their already-demonstrated capability in those competencies in adjacent professional fields.
-
-In terms of computer programming, @Peng_2023 found a positive impact of GitHub Copilot (an LLM-powered programming assistant) on productivity.
-Specifically, in an experiment involving 95 freelance programmers, they found that a treatment group of programmers with access to GitHub Copilot completed a standardized programming task 56% faster than the control group.
-Programmers with less experience saw the greatest improvements in productivity.
-
-@DellAcqua_2023, focusing on management consulting tasks, provide evidence that LLMs can improve writing productivity.
-In a field experiment involving consultants from Boston Consulting Group they found that the use of OpenAI's GPT-4 led to a 25% increase in delivery speed of business tasks, most of which involved some writing, as well as a 40% increase in human-rated performance on those tasks.
-Similar to computer programming, these productivity increases were most pronounced for those with below average performance, with their output increasing by 43%.
-
-On the other hand, @Valenzuela_2024 argue that LLMs lead to a loss of serendipity which may lead to less original work, and potentially de-skilling primarily with respect to programming ability, among other consequences.
-These outcomes could negatively affect students' effective learning of data science.
-
-@Ellis_Slade_2023 take a more optimistic perspective and argue that LLMs are just another technology that will impact statistics and data science education.
-Similarly, @Tu_2024 acknowledge that LLMs can streamline many parts of a data science workflow.
-With that in mind, they suggest that data-scientists-in-training should shift their self-perspectives from primarily being an "analyst" to primarily being a "product manager" responsible for strategic oversight of the analysis carried out by LLMs.
-
-At the high school level @Lazar_2023 conducted an informal survey of secondary school teachers and students on their opinions of ChatGPT, and found that while LLMs could: help creativity, provide academic support when teachers were unavailable, and model certain types of writing well.
-Teachers were also concerned about the potential of LLMs to limit students' learning in certain ways through over-reliance.
-Beyond academic integrity concerns, @Lazar_2023 found that teachers had similar concerns to @Valenzuela_2024 about de-skilling and an overall loss of agency in writing and critical thinking.
-
-@Cahill_2024 surveyed undergraduate political science students on their attitudes toward, and usage of, AI tools.
-They found that the use of ChatGPT was widespread.
-However, they also found that many students lacked confidence in using AI for academic purposes.
-In particular, only 11% "strongly agreed" that they know how to use AI to improve their writing.
-Students had nuanced views on appropriate AI use.
-Many respondents felt that using it to write whole papers was inappropriate, but using it for basic tasks like general assistance, writing feedback and basic data visualization appropriate.
-
-To understand the current state of LLMs as a tool for trustworthy data science, this paper focuses on the association between student academic performance and their LLM usage.
-Specifically, we examine the relationship between students' grades and self-reported measures of student LLM usage, as well as student attitudes toward LLMs in general.
-This is based on students' final papers and a survey, conducted in a third-year undergraduate data science course at the University of Toronto.
-By examining how students interact with and perceive LLMs as tools, and how these variables translate into student outcomes, current practice with regard to LLM integration in data science can be better understood, leading to better recommendations for their future development.
-
-The remainder of this paper is structured as follows: @sec-data visualizes and analyzes survey data and coursework from students.
-@sec-model specifies a model used to investigate the relationship.
-@sec-results describes and analyzes the model's results.
-@sec-discussion discusses the implications of the findings for data science education and future research and practice at the intersection of LLMs and trustworthy data science.
+The remainder of this paper is structured as follows. @sec-data....
 
@@
 
-## Background
+Some of our data is of penguins (@fig-bills), from @palmerpenguins.
 
-To investigate students' usage and attitudes towards LLMs and how they related to their academic performance, a dataset containing their usage and attitudes, coursework, and academic performance was constructed.
-This was based on three components:
+```{r}
+#| label: fig-bills
+#| fig-cap: Bills of penguins
+#| echo: false
 
-1.  an optional survey;
-2.  self-reported LLM usage; and
-3.  student marks on their final paper.
+ggplot(penguins, aes(x = island, fill = species)) +
+  geom_bar(alpha = 0.8) +
+  scale_fill_manual(values = c("darkorange","purple","cyan4"),
+                    guide = "none") +
+  theme_minimal() +
+  facet_wrap(~species, ncol = 1) +
+  coord_flip()
+```
 
-All data are from the cohort of students taking STA302 "Methods of Data Analysis I" in the Winter 2024 semester at the University of Toronto.
-This course had 275 students initially enrolled which, reflecting a normal rate of attrition for undergraduate statistics courses at the University of Toronto, reduced to 154 students by the end of the semester.
-Assessment was heavily based on three papers submitted over the course of the 12 week semester.
+Talk more about it.
 
-The student marks that we analyze are based only on the final paper, which is done individually.
-By this stage, uninterested students have typically dropped the course, and students are familiar with course expectations.
-A typical paper submission is 10-20 pages, and requires students to conduct original research to answer a research question of interest to them.
-It reflects the skills typically used by a professional data scientist.
-Students are expected to develop a research question of interest to them, identify or collect data to answer the question, conduct statistical analysis, and write a short paper.
-Examples of final papers (shared with consent) include: @hannahyu; @emilysu; and @bennyrochwerg.
+And also planes (@fig-planes).
+(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
 
-By the time they are working on their final paper, students have submitted and received feedback on two previous papers with similar requirements and rubrics to that of the final paper.
-Each paper has the same basic structure and expectations.
-Before the final paper is due, students have received feedback on all their previous work in the class (including their past papers) and there is an optional two-day period of peer review.
-
-The pre-requisites of this course mean that the typical student is an upper-year undergraduate.
-Coding and writing are major parts of the course.
-Students are welcome to use R or Python, but the majority code in R because that is the programming language currently mostly taught in pre-requisite courses.
-All writing must be in English.
-The primary motivation for having students write three papers as the main assessment for the course is to give them the opportunity to create a public portfolio of work they can use to apply for jobs.
-
-Throughout the semester students were encouraged to use LLMs.
-Formal instruction was provided twice during the semester.
-The first was a masterclass taught by a computer science faculty member on the ethics of using LLMs (see @Horton2024 for details).
-The second was a masterclass taught by a TA on writing with LLMs.
-
-Data was collected from students through an optional end-of-course survey.
-[Appendix -@sec-survey-details] details the questions asked in the survey.
-Whether or not they consented to their data being used, all respondents received a 1% increase in their final course grade for their participation.
-Consenting responses were then matched to their final paper mark, as well as the GitHub repository for their final paper.
-The responses were anonymized by removing any personal references to the students themselves including, names, emails, student numbers, and GitHub links.
-
-Data cleaning and analysis was done using the `R` statistical programming language [@citeR], and the `tidyverse` [@tidyverse], `janitor` [@janitor], `reshape2` [@reshape], and `readxl` [@readxl] packages.
-
-## Survey data
-
-There were 146 responses to the survey.
-Of these, 119 respondents provided authorization for their data to be collected and used.
-Four of those respondents submitted the survey twice, and after removing their second response, 115 responses remained.
-Of those, 15 respondents did not include a statement on LLM usage in the README of the GitHub repository of their final paper, leaving 100 responses.
-Finally 7 of those respondents did not provide a usable GPA response.
-This leaves 93 respondents that were of use and were merged based on student name.
-
-All but one respondent completed the survey within 45 minutes (@fig-response-time-and-gpa-1).
-That one respondent took more than 4,000 minutes to complete the survey, suggesting they took a break while filling it out.
-That respondent is included in the analysis dataset, but without them, the average time to complete the survey was 11 minutes, and the standard deviation was 8 minutes.
-
-```{r fig-response-time-and-gpa}
-#| message: false
+```{r}
+#| label: fig-planes
+#| fig-cap: Relationship between wing length and width
 #| echo: false
 #| warning: false
-#| fig-cap: Distribution of respondents' survey response times and self-reported GPA
-#| fig-subcap: ["Distribution of survey response times","\"What is your GPA?\""]
-#| layout-ncol: 1
+#| message: false
 
-all_data |>
-  # summarize(mean = mean(mins_complete/60),
-  # standard_dev = sd(mins_complete/60))
-  ggplot(aes(x = mins_complete/60)) +
-  geom_histogram(binwidth = 1) +
-  labs(x = "Time to complete survey (minutes)", y = "Frequency") +
-  theme_minimal()
+analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
 
-all_data |>
-  # summarize(mean = mean(what_is_your_gpa, na.rm = TRUE),
-  #           standard_dev = sd(what_is_your_gpa, na.rm = TRUE))
-  ggplot(aes(x = what_is_your_gpa)) +
-  geom_histogram(binwidth = 0.1) +
-  labs(x = "Self-reported GPA", y = "Frequency") +
-  theme_minimal()
+analysis_data |> 
+  ggplot(aes(x = width, y = length)) +
+  geom_point(alpha = 0.8) +
+  theme_minimal() +
+  labs(x = "Wing width (mm)",
+       y = "Wing length (mm)")
 ```
 
-There is a wide distribution of self-reported GPAs (@fig-response-time-and-gpa-2).
-The majority of responses cluster around a B (3.0 out of 4.0), and the average is 3.06 with a standard deviation of 0.55.
-One factor that may affect the range is that the course is required for programs in the Statistics, Mathematics and Computer Science Departments.
-Self-reported GPAs also introduce the possibility of reporting bias.
-For instance, respondents may have provided their cumulative GPA, their most recent term's GPA, or could have misreported it entirely.
+Talk way more about it.
 
-Students from a range of years took the course, however the majority of respondents were in their 3rd or 4th year of study (@tbl-year-dist).
-The course's prerequisite two-course sequence is typically completed by students in their second year, would make it difficult to take this course earlier than their 3rd year.
+# Model
 
-```{r tbl-year-dist}
+The goal of our modelling strategy is twofold. Firstly,...
+
+Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+
+## Model set-up
+
+Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+
+\begin{align} 
+y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
+\mu_i &= \alpha + \beta_i + \gamma_i\\
+\alpha &\sim \mbox{Normal}(0, 2.5) \\
+\beta &\sim \mbox{Normal}(0, 2.5) \\
+\gamma &\sim \mbox{Normal}(0, 2.5) \\
+\sigma &\sim \mbox{Exponential}(1)
+\end{align}
+
+We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+
+### Model justification
+
+We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+
+We can use maths by including latex between dollar signs, for instance $\theta$.
+
+# Results
+
+Our results are summarized in @tbl-modelresults.
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
 #| message: false
-#| echo: false
-#| warning: false
-#| tbl-cap: \"What year are you?\"
 
-all_data |>
-  count(what_year_are_you) |>
-  pivot_wider(
-    names_from = what_year_are_you,
-    values_from = n,
-    values_fill = list(n = 0)
-  ) |>
-  kable()
+library(rstanarm)
+
+first_model <-
+  readRDS(file = here::here("models/first_model.rds"))
 ```
 
-Respondents had a varied self-perception of their coding and writing abilities (@fig-selfperception).
-Most respondents believe it is important to be good at writing, but many are either indifferent or do not like to write (@fig-selfperception-1).
-Respondents also do not find writing to be particularly easy, which could be associated with the reported relative antipathy toward writing.
-Most respondents were at least somewhat confident in their own writing abilities, but a substantial number felt otherwise.
-Although few respondents felt that they were confident in their writing ability, more felt that they were able to catch their mistakes, which could indicate a disconnect between how respondents perceive their work and how the work was evaluated.
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-modelresults
+#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
+#| warning: false
 
-```{r fig-selfperception}
-#| message: false
-#| echo: false
-#| warning: false
-#| fig-cap: Self-perception of coding and writing abilities
-#| fig-subcap: ["\"Please rate how much each statement describes you, on a scale from 'This is very different to me' to 'This is a lot like me'\"", "\"Please rate how much each statement describes you, on a scale from 'This is very different to me' to 'This is a lot like me'\""]
-
-# Writing
-writing_perceptions_long <-
-  all_data |>
-  select(
-    writing_is_easy_for_me,
-    i_like_to_write,
-    i_believe_it_is_important_to_be_a_good_writer,
-    when_i_edit_it_is_easy_for_me_to_catch_my_mistakes,
-    i_feel_confident_sharing_my_writing,
-    i_am_confident_in_my_overall_writing_ability
-  ) |>
-  pivot_longer(cols = everything(),
-               names_to = "statements",
-               values_to = "response") |>
-  mutate(
-    statements = case_when(
-      statements == "writing_is_easy_for_me" ~ "\"Writing is easy for me.\"",
-      statements == "i_like_to_write" ~ "\"I like to write.\"",
-      statements == "i_believe_it_is_important_to_be_a_good_writer" ~ "\"I believe it is important\nto be a good writer.\"",
-      statements == "when_i_edit_it_is_easy_for_me_to_catch_my_mistakes" ~ "\"When I edit it is easy for\nme to catch my mistakes.\"",
-      statements == "i_feel_confident_sharing_my_writing" ~ "\"I feel confident\nsharing my writing.\"",
-      statements == "i_am_confident_in_my_overall_writing_ability" ~ "\"I am confident in my\noverall writing ability.\"",
-      TRUE ~ statements
-    ),
-    response = case_when(
-      response == "\"This is a lot like me\"" ~ "A lot like me",
-      response == "\"This is very different to me\"" ~ "Very different\nfrom me",
-      response == "\"This somewhat describes me\"" ~ "Somewhat describes me",
-      TRUE ~ response
-    )
-  )
-
-writing_perceptions_long |>
-  ggplot(aes(x = response, fill = response)) +
-  geom_bar(position = "stack", stat = "count") +
-  facet_wrap(vars(statements)) +
-  labs(x = "Response", y = "Number of repondents", fill = "Response:") +
-  scale_fill_brewer(palette = "Set1") +
-  theme_minimal() +
-  theme(axis.text.x = element_blank(), legend.position = "bottom")
-
-# Coding
-coding_perceptions_long <-
-  all_data |>
-  select(
-    coding_is_easy_for_me,
-    i_like_to_code,
-    i_believe_it_is_important_to_be_a_good_coder,
-    when_i_check_my_code_it_is_easy_for_me_to_catch_my_mistakes,
-    i_feel_confident_sharing_my_code,
-    i_am_confident_in_my_overall_coding_ability
-  ) |>
-  pivot_longer(cols = everything(),
-               names_to = "statements",
-               values_to = "response") |>
-  mutate(
-    statements = case_when(
-      statements == "coding_is_easy_for_me" ~ "\"Coding is easy for me\"",
-      statements == "i_like_to_code" ~ "\"I like to code\"",
-      statements == "i_believe_it_is_important_to_be_a_good_coder" ~ "\"I believe it is important\n to be a good coder\"",
-      statements == "when_i_check_my_code_it_is_easy_for_me_to_catch_my_mistakes" ~ "\"When I check my code, it is\n easy for me to catch my\nmistakes\"",
-      statements == "i_feel_confident_sharing_my_code" ~ "\"I feel confident\n sharing my code\"",
-      statements == "i_am_confident_in_my_overall_coding_ability" ~ "\"I am confident in\n my overall coding ability\"",
-      TRUE ~ statements
-    ),
-    response = case_when(
-      response == "\"This is a lot like me\"" ~ "A lot like me",
-      response == "\"This is very different to me\"" ~ "Very different\nfrom me",
-      response == "\"This somewhat describes me\"" ~ "Somewhat describes me",
-      TRUE ~ response
-    )
-  )
-
-coding_perceptions_long |>
-  ggplot(aes(x = response, fill = response)) +
-  geom_bar(position = "stack", stat = "count") +
-  facet_wrap(vars(statements)) +
-  labs(x = "Response", y = "Number of repondents", fill = "Response:") +
-  scale_fill_brewer(palette = "Set1") +
-  theme_minimal() +
-  theme(axis.text.x = element_blank(), legend.position = "bottom")
-```
-
-Respondent self-perceptions regarding coding proficiency and importance varied (@fig-selfperception-2).
-There is strong consensus on the perceived importance of coding skills.
-However, respondents' self-assessed coding abilities and enjoyment are more heterogeneous, with a substantial proportion reporting moderate rather than high levels of ease and enjoyment in coding tasks.
-
-Overall there was a moderate level of confidence among respondents in their overall coding ability, willingness to share code, and capacity to identify errors (@fig-selfperception-2).
-Notably, respondents express slightly higher confidence in detecting their own coding mistakes compared to general coding ability or code sharing.
-These patterns suggest that while respondents have developed some coding self-efficacy, there is still considerable potential for enhancing their perceived competence and comfort across various coding-related activities.
-
-The majority of respondents were at least "somewhat familiar" with generative AI such as ChatGPT (@tbl-ai-familiarity-usage-1).
-Though respondents' self-perceptions around writing and coding are varied, there was strong consensus that the use of generative AI tools is appropriate within an academic setting (@tbl-ai-familiarity-usage-2).
-Most respondents who responded "It depends" generally found artificial intelligence tools to be appropriate, though with certain guidelines and rules governing their use.
-
-```{r tbl-ai-familiarity-usage}
-#| message: false
-#| echo: false
-#| warning: false
-#| tbl-cap: Familiarity with, and appropriateness of, generative AI
-#| tbl-subcap: ["'How familiar are you with using generative AI tools such as OpenAI's ChatGPT or equivalents?'","'To what extent do you think using generative AI tools such as ChatGPT by OpenAI (or equivalents) is ethical and appropriate for schoolwork?'"]
-
-all_data |>
-  count(ai_familiarity) |>
-  kable(col.names = c("Extent of AI familiarity", "Students"))
-
-all_data |>
-  count(ai_schoolwork_appropriate) |>
-  kable(col.names = c("Ethical & Appropriate for school?", "Students"))
-```
-
-To understand the role of LLMs in learning, students identified their usage in a more granular way by selecting various pre-defined use cases in the survey (@fig-AI-use-and-usefulness-1).
-Technical questions and explaining concepts were the two top use cases among students in the course.
-More than half of students also used LLMs for quick questions, general knowledge, writing paper code, and checking solutions.
-Just less than half used it to write paper content, which could suggest that students do not feel confident using it to improve writing.
-
-```{r fig-AI-use-and-usefulness}
-#| message: false
-#| echo: false
-#| warning: false
-#| fig-cap: "Use and usefulness of generative AI"
-#| fig-subcap: ["\"If you have used generative AI tools such as OpenAI's ChatGPT or equivalents, in what ways have you used it (select all that apply)?\"", "\"How helpful did you find generative AI tools such as ChatGPT by OpenAI (or equivalents) for each component of STA302?\""]
-
-# Use
-all_data |>
-  select(how_have_you_used) |>
-  separate_rows(how_have_you_used, sep = ";") |>
-  mutate(
-    how_have_you_used = case_when(
-      how_have_you_used == "check grammar mistake" ~ "Other",
-      how_have_you_used == "debugging code" ~ "Other",
-      how_have_you_used == "mental health support" ~ "Other",
-      how_have_you_used == "Making practice questions for me with increasing difficulty" ~ "Other",
-      how_have_you_used == "Generating test questions" ~ "Other",
-      TRUE ~ how_have_you_used
-    )
-  ) |> 
-  count(how_have_you_used) |>
-  filter(how_have_you_used != "") |>
-  arrange(desc(n)) |>
-  ggplot(aes(x = n, y = reorder(how_have_you_used, n))) +
-  geom_bar(stat = "identity") +
-  labs(x = "Number of respondents", y = "Type of usage") +
-  theme_minimal()
-
-# Usefulness
-all_data |>
-  select(
-    weekly_quiz,
-    weekly_mini_essay,
-    papers_generating_ideas,
-    papers_writing_code,
-    papers_writing_content,
-    papers_improving_content
-  ) |>
-  pivot_longer(cols = everything(),
-               names_to = "task",
-               values_to = "response") |>
-  mutate(
-    # Simplify response categories and rename tasks
-    response_simplified = case_when(
-      response %in% c(
-        "\"Not helpful\"",
-        "\"I did not use generative AI for this component\""
-      ) ~ "Less Helpful",
-      response %in% c("\"Somewhat helpful\"", "\"Very helpful\"") ~ "More Helpful",
-      TRUE ~ "Unknown"  # Handle unexpected values
-    ),
-    task = case_when(
-      task == "weekly_quiz" ~ "Weekly Quiz",
-      task == "weekly_mini_essay" ~ "Weekly Mini Essay",
-      task == "papers_generating_ideas" ~ "Generating Ideas",
-      task == "papers_writing_code" ~ "Writing Code",
-      task == "papers_writing_content" ~ "Writing Content",
-      task == "papers_improving_content" ~ "Improving Content",
-      TRUE ~ "Other"  # Handle any unexpected task names
-    )
-  ) |>
-  group_by(task, response_simplified) |>
-  summarise(count = n(), .groups = 'drop') |>
-  ggplot(aes(x = response_simplified, y = count, fill = response_simplified)) +
-  geom_bar(stat = "identity") +
-  facet_wrap(~ task, scales = "free_y") +
-  labs(y = "Number of respondents", x = "", fill = "Response type: ") +
-  theme_minimal() +
-  coord_flip() +
-  scale_fill_brewer(palette = "Set1") +
-  theme(axis.text.y = element_blank(), legend.position = "bottom")
-```
-
-Respondents were also asked to rate the helpfulness of LLMs on various tasks assigned during the course on a 4-point scale of "Did not use" to "Very Helpful" (@fig-AI-use-and-usefulness-2).
-To simplify the presentation, responses were grouped into two main categories: "Less Helpful" and "More Helpful." The "Less Helpful" category combines responses where students found the AI either "Not helpful" or did not use it for the task, while the "More Helpful" category includes responses where the LLM was considered "Somewhat helpful" or "Very helpful".
-
-Respondents differed in terms of how they used LLMs in the course (@fig-AI-use-and-usefulness-2).
-While most tasks were roughly split between respondents finding LLMs helpful or not, respondents found them most helpful in generating code.
-In the context of the course, this mostly meant generating R code for transforming, analyzing, and visualizing data.
-To a lesser extent, respondents also found LLMs to be helpful in improving the existing writing they had, while at the same time not favouring it for writing content from scratch.
-
-Finally, one question asked students to elaborate on whether they thought generative AI tools such as ChatGPT were ethical and appropriate for schoolwork.
-This was an open-response question.
-To provide a sense of the responses, we used Anthropic's Claude 3.5 Sonnet model (as at 5 August 2024) to summarize the comments and it provided:
-
-> Many students view AI as a helpful supplementary tool, comparing it to resources like Google or calculators.
-They believe it can aid in understanding concepts, debugging code, brainstorming ideas, and saving time on routine tasks.
-However, there's a consensus that AI should not be used to complete entire assignments or replace original thinking.
-Students emphasize the importance of using AI ethically, citing it when appropriate, and not relying on it exclusively.
-Some argue that learning to use AI effectively is a valuable skill for future careers.
-Concerns raised include the potential for plagiarism, the risk of hindering critical thinking skills, and the possibility of receiving incorrect information.
-Overall, most students support the responsible use of AI in education, with proper guidelines and transparency, while recognizing the need to maintain academic integrity and develop independent learning skills.
-
-## LLM usage and final paper marks
-
-Two other components were merged with the survey responses: self-reported LLM usage on the final paper, and final paper mark.
-
-Students were encouraged to use LLMs to complete their papers.
-Each paper required the students to disclose their usage through a statement in the GitHub repository README for the paper.
-Even students who did not use generative AI at all were required to state this in the README.
-For students who did use generative AI, there was an additional requirement, where possible, that they save the logs of their usage in a txt file which was also included in their GitHub repository.
-
-Those README statements were gathered and parsed using OpenAI's ChatGPT 4o model (as at 26 July). The following prompt was used:
-
-> The following statement is about to what extent LLMs were used by a student.
-Please characterize it as one of: "None", "Minimal", "Somewhat", "Extensive", "Unsure".
-Respond with only one of those options.
-
-All classifications were then manually checked for reasonableness.
-
-We find a varied extent of self-reported LLM usage (@tbl-llm-usage).
-41 respondents were classified as having made extensive use of LLMs, while 28 were classified as having made somewhat use.
-31 respondents were classified as having made minimal or no use of LLMs.
-In the analysis dataset we combine those two classifications because only 8 respondents were classified as having minimal usage.
-
-```{r tbl-llm-usage}
-#| message: false
-#| echo: false
-#| warning: false
-#| tbl-cap: Self-reported LLM usage in final paper
-
-all_data |>
-  count(llm_usage) |> 
-  kable(col.names = c("Self-reported LLM usage", "Number"))
-```
-
-The third, and final, component is the mark, in percentages, on the final paper (@fig-mark-1).
-The overall mean was 78% and standard deviation was 17 percentage points.
-However there was considerable differences by the extent of LLM usage (@tbl-mark; @fig-mark-2).
-
-```{r fig-mark}
-#| message: false
-#| echo: false
-#| warning: false
-#| fig-cap: Distribution of marks on final paper (in percentages)
-#| fig-subcap: ["Overall", "By self-reported LLM usage"]
-
-all_data |>
-  select(mark) |>
-  ggplot(aes(x = mark)) +
-  geom_histogram(binwidth = 0.05) +
-  scale_x_continuous(labels = label_percent()) +
-  labs(x = "Final paper mark (percentage)", y = "Count") +
-  theme_minimal()
-
-all_data |>
-  ggplot(mapping = aes(x = mark)) +
-  geom_histogram(binwidth = 0.01) +
-  scale_x_continuous(labels = label_percent()) +
-  labs(x = "Final paper mark (percentage)", y = "Count") +
-  facet_wrap(vars(llm_usage),
-             ncol = 2,
-             labeller = labeller(
-               llm_usage = c(
-                 "Extensive" = "Extensive LLM usage",
-                 "Minimal" = "Minimal LLM usage",
-                 "None" = "No LLM usage",
-                 "Somewhat" = "Some LLM usage"
-               )
-             )) +
-  theme_minimal()
-```
-
-```{r tbl-mark}
-#| message: false
-#| echo: false
-#| warning: false
-#| tbl-cap: Self-reported LLM usage in final paper and final paper mark
-
-all_data |>
-  group_by(llm_usage) |>
-  summarize(mean = mean(mark) |> round(2),
-            sd = sd(mark) |> round(2)) |>
-  kable(col.names = c("Self-reported LLM usage", "Mean", "Std dev"))
-```
-
-# Model {#sec-model}
-
-The goal of our modelling strategy is to better understand how a respondents' result on their final paper associates with their self-reported LLM usage.
-Students received their result on the final paper as a proportion (a 0-100% mark), so we use zero-one-inflated beta regression to best fit the distribution in this interval.
-We use this type of regression given the ability of the beta distribution to effectively describe continuous and skewed or heteroskedastic datasets, which we can see in the asymmetrical histogram of the distribution of marks.
-
-Transforming the regression with a logit function as the link function ensures that the predicted values stay between zero and one, respecting the boundaries of the data.
-Some students got full marks for the paper, so we use the zero-one-inflated beta model.
-
-Here we briefly describe the model that we use, which follows @kurztotherescue. Model diagnostics are included in [Appendix -@sec-model-details].
-
-OLD:
-
-Define $y_i$ as the percentage received on the final paper.
-Then $\beta_1$ and $\beta_2$ are the effect of self-reported LLM usage and self-reported GPA on the final paper grade, respectively.
-\begin{align}
-y_i &\sim \mbox{Beta}(\mu_i, \phi) \\
-logit(\mu_i) &= \beta_0 + \beta_1 \times \mbox{LLM usage}_i + \beta_2 \times \mbox{GPA}_i\\
-\beta_0, \beta_1, \beta_2  &\sim \mbox{Normal}(0, 1) \\
-\phi &\sim \mbox{Gamma}(4, 0.1)
-\end{align}
-
-NEW:
-
-Let $Y_i$ represent the dependent variable "mark" for observation $i$. The model assumes a zero-one inflated beta distribution for $Y_i$, which can be represented as:
-
-$Y_i \sim \text{Zero-One Inflated Beta}(\mu_i, \phi_i, \pi_{zoi_i}, \pi_{coi_i})$
-
-Where:
-
-- $\mu_i$ is the mean of the beta distribution (related to the location parameter).
-- $\phi_i$ is the precision parameter (related to the dispersion of the beta distribution).
-- $\pi_{zoi_i}$ is the probability of $Y_i$ being exactly 0 or 1 (zero-one inflation component).
-- $\pi_{coi_i}$ is the probability of $Y_i$ being in the continuous part of the beta distribution (continuous component inflation).
-
-The specific linear models for each parameter are given by:
-
-1. **Mean (Location) Component:**
-   $logit(\mu_i) = \beta_{0,\mu} + \beta_{1,\mu} \cdot \text{llm\_usage}_i + \beta_{2,\mu} \cdot \text{what\_is\_your\_gpa}_i$
-
-2. **Precision Component:**
-   $\log(\phi_i) = \beta_{0,\phi} + \beta_{1,\phi} \cdot \text{llm\_usage}_i + \beta_{2,\phi} \cdot \text{what\_is\_your\_gpa}_i$
-
-3. **Zero-One Inflation Component:**
-   $logit(\pi_{zoi_i}) = \beta_{0,zoi} + \beta_{1,zoi} \cdot \text{llm\_usage}_i + \beta_{2,zoi} \cdot \text{what\_is\_your\_gpa}_i$
-
-4. **Continuous Outcome Inflation (coi) Component:**
-   $logit(\pi_{coi_i}) = \beta_{0,coi}$
-
-In this model:
-
-- $\beta_{0,\mu}, \beta_{0,\phi}, \beta_{0,zoi},$ and $\beta_{0,coi}$ are the intercepts for the respective components.
-- $\beta_{1,\mu}, \beta_{1,\phi}, \beta_{1,zoi}$ represent the coefficients for the effect of "llm\_usage" on the respective components.
-- $\beta_{2,\mu}, \beta_{2,\phi}, \beta_{2,zoi}$ represent the coefficients for the effect of "what\_is\_your\_gpa" on the respective components. 
-- The term $logit(\cdot)$ denotes the logit link function, commonly used for modeling probabilities.
-
-This notation captures the model's structure as defined in the `brms` code, with each parameter modeled as a function of the predictors "llm\_usage" and "what\_is\_your\_gpa", except for the continuous outcome inflation (coi) component, which is modeled with just an intercept.
-
-We estimate the model in `R` [@citeR] as well, using `brms` [@brms].
-
-The expected relationship between LLM usage and final mark is unclear.
-It could be that stronger students used LLMs in a more sophisticated way or that they did not need to use them.
-However, the relationship between GPA and final mark is expected to be positive.
-
-# Results {#sec-results}
-
-Our results are summarized in @tbl-model-results and @fig-model-results. We especially draw on `modelsummary` [@modelsummary].
-
-```{r tbl-model-results}
-#| message: false
-#| echo: false
-#| warning: false
-#| tbl-cap: Coefficient estimates and mean absolute deviation (MAD)
-
-fit1 <-
-  readRDS(file = here("models/fit_one_inflated.rds"))
-fit2 <-
-  readRDS(file = here("models/fit_one_inflated_gpa.rds"))
-
-modelsummary(
+modelsummary::modelsummary(
   list(
-    "Base model" = fit1,
-    "Including self-reported GPA" = fit2
+    "First model" = first_model
   ),
@@
 
-```{r fig-model-results}
-#| message: false
-#| echo: false
-#| warning: false
-#| fig-cap: Selected coefficient estimates and 90 per cent credibility intervals
+# Discussion
 
-modelplot(
-  list(
-    "Base model" = fit1,
-    "Including self-reported GPA" = fit2
-  ),
-  conf_level = 0.9,
-  drop = c("phi", "b_zoi_Intercept")) +
-  labs(x= "") +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom")
-```
+## First discussion point {#sec-first-point}
 
-@tbl-model-results and @fig-model-results present the results from two models estimating the association between LLM usage, GPA and final paper mark.
-The first model includes only LLM usage, while the second model also includes self-reported GPA.
+If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
 
-LLM usage estimates are in relation to "Extensive" LLM usage.
-In the base model, the coefficients suggest that less usage of LLMs was associated with lower scores.
-The second model retains this slight association, but with slightly smaller coefficients.
-For instance, the negative impact of “b_llm_usageSomewhat” is slightly reduced from -0.35 to -0.29, indicating that part of the effect observed in the base model might be explained by the GPA variable.
+## Second discussion point
 
-Including self-reported GPA in the second model does considerably change the  changes the interpretation of some coefficients.
-For instance, intercept for "b_zoi_Intercept" becomes much more negative in the GPA-included model, suggesting that the GPA significantly affects the zero-one inflation part of the model.
-However, the essential relationship between LLM usage and final mark is retained.
-
-The overall fit of the model improves when GPA is included.
-There is an increase in R-squared from 0.095 to 0.478 and improvements in the ELPD and LOOIC metrics.
-This is also suggested by the LOO comparison between the two models (@tbl-loo in [Appendix -@sec-model-details]).
-
-# Discussion {#sec-discussion}
-
-There is some very mild evidence from this observational study that LLM usage may be associated with better scores.
-However, this work should mostly be treated as a direction for future research in this area.
-More broadly, our work has three major take-aways: 1) LLMs could help enhance the trustworthiness of data science, but not without considerable futher development.
-2) Making general purpose chatbots more opinionated and knowledgeable about data science would considerably help.
-3) Alternative interaction approaches, beyond a chatbot interface, could be especially useful.
-
-## Impact of LLMs on trustworthy data science
-
-We found that many students did use LLMs to help them complete a data science project closely related to what professional data scientists do.
-A common use case to ask the LLM to write code to make a graph.
-Another was to ask for some example paragraphs.
-
-Data science code tends to be written differently to software engineering code.
-Functions and loops are used less, as there is often less reuse of code.
-The training data of these LLMs is dominated by software engineering code, and so when an LLM is asked to write code for a data science project it tends to produce code with some of those features.
-This is neither good nor bad in and of itself, however it can clash with the style of the code that surrounds it.
-
-LLMs are also being used as an alternative search engine.
-Students often asked, for instance, how to modify a graph to add a vertical line, or similar.
-Here LLMs tended to be especially useful and able to answer questions.
-In the past students may have used StackOverflow or similar to find their answers, but the nature of the chat response likely provided a faster answer.
-
-Finally, there is the question of whether the nature of the data science has fundamentally changed or improved.
-For now, there was no evidence of that.
-It was clear when a student had just used LLM-generated text without modification, and it was often clear when large amounts of code had been written by an LLM.
-But it was not obvious when small amounts of either had been used, or when LLM-generated content had just been used as an initial draft.
-
-Many of these issues were alluded to by @bommasani2021opportunities, in the context of data science, and it is striking that they remain today.
-
-## General purpose compared with opinionated specific chatbots
-
-ChatGPT and equivalents are general-purpose chatbots.
-One can ask them a technical data science question and then in the same session ask for a brownie recipe!
-However this general-purpose nature means that they are not necessarily optimized toward one particular task [@thoppilan2022lamda].
-When used for data science, this means that sometimes the functions or libraries recommended may not exist, code may not run, or out-dated approaches may be used.
-Frontier models make things up and are unnervingly confident, even when wrong [@openai2023gpt4, p.
-59], which can create special issues for learners.
-
-General-purpose chatbots being used for data science are fine for experienced data scientists who have an established sense of when something might be not ideal or even wrong.
-But it is problematic for novices who may not know, can lack the confidence to push back, and possibly, may learn the wrong approach.
-
-Retrieval-augmented generation (RAG) means providing various sources, for instance books or manuals, which the LLM uses to base its response on.
-Fine-tuning means providing the previously trained LLM with a large number of example input and outputs, which again guide the response from the LLM [@raffel2020exploring].
-A chatbot based interface, focused on the needs of data science beginners, could use RAG and/or fine-tuning to considerably improve the quality of its responses.
-The trade-off would be a reduction in the general-purpose usability of the chatbot, but in the context of this use this that trade-off is likely acceptable.
-
-## Changed LLM-based interfaces
-
-LLMs existed before ChatGPT, but it was the launch of ChatGPT that brought them into the mainstream.
-The underlying model was not considerably different to others and the main difference that ChatGPT brought was the chat interaction.
-These are useful, but there is considerable opportunity to develop additional ways of interacting with LLMs.
-This is particularly relevant for enhancing trustworthy data science.
-
-Search-based LLM tools like Elicit and Perplexity.ai differ to ChatGPT in that they focus on search.
-In contrast to Google, which typically provides a list of websites related to a particular search, Perplexity.ai attempts to provide an answer.
-In contrast to ChatGPT, search-based LLM tools focus on answering a question rather than engaging in discussion.
-
-Students in this class were required to use GitHub to host their papers, as well as supporting code and data.
-GitHub Issues and Pull Requests were used for peer review and submission occurred based on links to GitHub repositories.
-Trustworthy data science could be considerably enhanced by establishing something analogous to continuous integration and a suite of tests, but for data science.
-The once-off nature of much data science code would mean that LLMs could be especially useful in this proposed infrastructure stack.
-For instance, when a data scientist committed their code or writing, a suite of static tests could run, for instance to check that the data are being read in, that classes are appropriate, etc.
-But additionally LLMs could be used to check for possible improvements in writing, and propose context-specific improvements in code.
-For instance, to make sure that what is being described in the paper is actually what is happening in the code.
-
-Finally, one exciting area of current development in LLMs is tool use.
-One early example of this was enabling LLMs to use search engines such as Google.
-As LLMs can produce writing, and writing is the input to a Google Search, enabling this tool broadened the types of queries that LLMs could respond effectively to.
-Tool use, focused on data science, exists.
-For instance, LLMs can use R and Python to respond to queries about a dataset.
-However, further development could be useful.
-For instance, when a user wrote code in an IDE to import some data, a tool-enhanced LLM could notice and automatically write code that establishes a Pydantic-based validation model for that data.
-Similarly, consider an LLM in an IDE context where a data scientist wrote out the model, in statistical notation, that they were interested in estimating, and the LLM was able write the model in Stan, run it, save the estimates, and add a summary table to the paper, all automatically and in the background.
-Similar work has occured in other contexts and found to bring substantial benefits [@chen2022evaluating].
+## Third discussion point
 
@@
 
-There are several substantial weaknesses of this study.
-The foundational one is that we used observational data, much of which was self-reported.
-There may be selection bias present in terms of who used LLMs, who reported their LLM usage truthfully, and even who remained in the class.
-A different design, specifically a randomized controlled trial (RCT) would deal with many of these issues, although likely at some cost.
-
-Regression reports average estimates over the full dataset.
-However, many earlier studies found distributional effects, with low-performing individuals benefiting more than high-performing individuals.
-Again, a change in design toward an RCT could enable the exploration of this question.
-Stratification of our dataset would result in small sample size, but nonetheless our data does provide some limited suggestive evidence that this effect may be present in data science.
-For instance, looking at the 28 students who received an A+ for the final paper, 13 of them had extensive LLM usage, whereas looking at the same number of worst performing students finds that only six of them had extensive LLM usage.
-
-Along these lines, we have considered LLM usage for code and writing as equivalent, but they should actually have different impacts depending on student backgrounds.
-Distinguishing between native and second-language English speakers, and then focusing on differential LLM usage, could have added a great deal of nuance to the analysis.
-
-Finally, we only considered one outcome measure, namely grade on the final paper.
-Each paper required a considerable amount of time for the student to produce.
-If an LLM was found to reduce the time taken to produce a paper, without any reduction in quality, then that would be a similarly useful outcome.
-
-Despite these shortcomings, our work clearly identifies a need for further research examining how LLMs can be used to develop a more trustworthy data science.
+Weaknesses and next steps should also be included.
 
@@
 
-# Appendix {.unnumbered}
+# Appendix {-}
 
-# Survey questions {#sec-survey-details}
-
-1.  After carefully reading the informed consent document, please indicate below whether you consent to have your anonymized responses included in the research study?
-    -   Yes, I authorize the use of the data collected about me for the STA302 course survey to be used. I will be compensated 1% of my course grade for completing the survey.
-    -   No I do not want my data included in the research study, but I want to complete the survey. I will be compensated 1% of my course grade for completing the survey.
-    -   I do not want to complete this survey. I realize that I am forfeiting the corresponding course credit.
-2.  What is your full name on Quercus?
-3.  What is your Student ID?
-4.  What year are you?
-5.  What is your specialization?
-6.  What is/are your major/s?
-7.  What is/are your minor/s?
-8.  What is your GPA?
-9.
-Please rate how much each statement describes you, on a scale from "This is very different to me" to "This is a lot like me" \["This is very different to me"; "This somewhat describes me"; "This is a lot like me"\]
-    -   Writing is easy for me
-    -   I like to write
-    -   I believe it is important to be a good writer.
-    -   When I edit it is easy for me to catch my mistakes.
-    -   I feel confident sharing my writing.
-    -   I am confident in my overall writing ability.
-10.
-Please rate how much each statement describes you, on a scale from "This is very different to me" to "This is a lot like me".
-When answering, please consider whichever programming language you are most familiar with.
-\["This is very different to me"; "This somewhat describes me"; "This is a lot like me"\]
-    -   Coding is easy for me
-    -   I like to code
-    -   I believe it is important to be a good coder.
-    -   When I check my code it is easy for me to catch my mistakes.
-    -   I feel confident sharing my code.
-    -   I am confident in my overall coding ability.
-11. How familiar are you with using generative AI tools such as OpenAI's ChatGPT or equivalents?
-    -   Very familiar
-    -   Somewhat familiar
-    -   Not familiar
-    -   Other
-12. Have you used any generative AI tools such as OpenAI's ChatGPT or equivalents for any reason (personal or educational)?
-    -   Yes
-    -   No
-    -   Other
-13. If you have used generative AI tools such as OpenAI's ChatGPT or equivalents, in what ways have you used it (select all that apply)?
-    -   Asking technical questions
-    -   Carrying on a conversation out of curiosity
-    -   Asking general knowledge questions
-    -   Solving homework
-    -   Checking solutions
-    -   Asking quick questions when stuck
-    -   Explaining concepts
-    -   Writing essays or paragraphs
-    -   Writing code
-    -   Never used it
-    -   Other
-14. To what extent do you think using generative AI tools such as ChatGPT by OpenAI (or equivalents) is ethical and appropriate for schoolwork?
-    -   Appropriate
-    -   Inappropriate
-    -   Other
-15. Please elaborate on your answer above.
-16. Did you use any generative AI tools such as OpenAI's ChatGPT or equivalents for STA302?
-    -   Yes
-    -   No
-    -   Other
-17.
-How helpful did you find generative AI tools such as ChatGPT by OpenAI (or equivalents) for each component of STA302?
-\["Not helpful"; "Somewhat helpful"; "Very helpful"; "I did not use generative AI for this component"\]
-    -   Weekly quiz
-    -   Weekly mini-essay
-    -   Papers: Generating ideas
-    -   Papers: Writing code
-    -   Papers: Writing content
-    -   Papers: Improving content
-18. What is your recommendation for how generative AI tools such as ChatGPT by OpenAI (or equivalents) should be used in the course in future?
-19. (Optional) Any other comments?
-
-\newpage
+# Additional data details
 
@@
 
-We use `bayesplot` [@bayesplot] and `loo` [@loo] conduct posterior predictive checks and evaluate model diagnostics.
-
 ## Posterior predictive check
 
-```{r fig-posteriorchecks}
+In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
+
+In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
+
+```{r}
+#| eval: true
+#| echo: false
 #| message: false
-#| echo: false
 #| warning: false
-#| fig-cap: Posterior predictive checking
-#| fig-subcap: ["Base model", "Including self-reported GPA"]
+#| label: fig-ppcheckandposteriorvsprior
 #| layout-ncol: 2
+#| fig-cap: "Examining how the model fits, and is affected by, the data"
+#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
 
-pp_check(fit1) +
+pp_check(first_model) +
   theme_classic() +
@@
 
-pp_check(fit2) +
-  theme_classic() +
-  theme(legend.position = "bottom")
+posterior_vs_prior(first_model) +
+  theme_minimal() +
+  scale_color_brewer(palette = "Set1") +
+  theme(legend.position = "bottom") +
+  coord_flip()
 ```
-
-```{r tbl-loo}
-#| message: false
-#| echo: false
-#| warning: false
-#| tbl-cap: Model comparison
-
-brms::loo_compare(loo::loo(fit1, cores = 2), loo::loo(fit2, cores = 2)) |>
-  kable(digits = 3)
-```
-
-\newpage
 
@@
 
-```{r fig-basemodeldiagnostics}
-#| fig-height: 7
+@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
+
+@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
+
+```{r}
+#| echo: false
+#| eval: true
 #| message: false
-#| echo: false
 #| warning: false
-#| fig-cap: Base model diagnostics
+#| label: fig-stanareyouokay
+#| fig-cap: "Checking the convergence of the MCMC algorithm"
+#| fig-subcap: ["Trace plot", "Rhat plot"]
+#| layout-ncol: 2
 
-plot(fit1)
-```
+plot(first_model, "trace")
 
-```{r fig-gpamodeldiagnostics}
-#| fig-height: 6
-#| message: false
-#| echo: false
-#| warning: false
-#| fig-cap: Model including self-reported GPA diagnostics
-
-plot(fit2)
+plot(first_model, "rhat")
 ```

### 2024-09-27 scripts/00-simulate_data.R

@@ -1,19 +1,52 @@
 #### Preamble ####
-# Purpose: Simulates... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Simulates a dataset of Australian electoral divisions, including the 
+  #state and party that won each division.
+# Author: Rohan Alexander
+# Date: 26 September 2024
+# Contact: rohan.alexander@utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: The `tidyverse` package must be installed
+# Any other information needed? Make sure you are in the `starter_folder` rproj
 
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
+set.seed(853)
+
 
 #### Simulate data ####
-# [...ADD CODE HERE...]
+# State names
+states <- c(
+  "New South Wales",
+  "Victoria",
+  "Queensland",
+  "South Australia",
+  "Western Australia",
+  "Tasmania",
+  "Northern Territory",
+  "Australian Capital Territory"
+)
+
+# Political parties
+parties <- c("Labor", "Liberal", "Greens", "National", "Other")
+
+# Create a dataset by randomly assigning states and parties to divisions
+analysis_data <- tibble(
+  division = paste("Division", 1:151),  # Add "Division" to make it a character
+  state = sample(
+    states,
+    size = 151,
+    replace = TRUE,
+    prob = c(0.25, 0.25, 0.15, 0.1, 0.1, 0.1, 0.025, 0.025) # Rough state population distribution
+  ),
+  party = sample(
+    parties,
+    size = 151,
+    replace = TRUE,
+    prob = c(0.40, 0.40, 0.05, 0.1, 0.05) # Rough party distribution
+  )
+)
 
 
-
+#### Save data ####
+write_csv(analysis_data, "data/00-simulated_data/simulated_data.csv")

### 2024-09-27 scripts/01-test_simulated_data.R

@@ -1,0 +1,89 @@
+#### Preamble ####
+# Purpose: Tests the structure and validity of the simulated Australian 
+  #electoral divisions dataset.
+# Author: Rohan Alexander
+# Date: 26 September 2024
+# Contact: rohan.alexander@utoronto.ca
+# License: MIT
+# Pre-requisites: 
+  # - The `tidyverse` package must be installed and loaded
+  # - 00-simulate_data.R must have been run
+# Any other information needed? Make sure you are in the `starter_folder` rproj
+
+
+#### Workspace setup ####
+library(tidyverse)
+
+analysis_data <- read_csv("data/00-simulated_data/simulated_data.csv")
+
+# Test if the data was successfully loaded
+if (exists("analysis_data")) {
+  message("Test Passed: The dataset was successfully loaded.")
+} else {
+  stop("Test Failed: The dataset could not be loaded.")
+}
+
+
+#### Test data ####
+
+# Check if the dataset has 151 rows
+if (nrow(analysis_data) == 151) {
+  message("Test Passed: The dataset has 151 rows.")
+} else {
+  stop("Test Failed: The dataset does not have 151 rows.")
+}
+
+# Check if the dataset has 3 columns
+if (ncol(analysis_data) == 3) {
+  message("Test Passed: The dataset has 3 columns.")
+} else {
+  stop("Test Failed: The dataset does not have 3 columns.")
+}
+
+# Check if all values in the 'division' column are unique
+if (n_distinct(analysis_data$division) == nrow(analysis_data)) {
+  message("Test Passed: All values in 'division' are unique.")
+} else {
+  stop("Test Failed: The 'division' column contains duplicate values.")
+}
+
+# Check if the 'state' column contains only valid Australian state names
+valid_states <- c("New South Wales", "Victoria", "Queensland", "South Australia", 
+                  "Western Australia", "Tasmania", "Northern Territory", 
+                  "Australian Capital Territory")
+
+if (all(analysis_data$state %in% valid_states)) {
+  message("Test Passed: The 'state' column contains only valid Australian state names.")
+} else {
+  stop("Test Failed: The 'state' column contains invalid state names.")
+}
+
+# Check if the 'party' column contains only valid party names
+valid_parties <- c("Labor", "Liberal", "Greens", "National", "Other")
+
+if (all(analysis_data$party %in% valid_parties)) {
+  message("Test Passed: The 'party' column contains only valid party names.")
+} else {
[…]

### 2024-09-27 scripts/03-test_data.R

@@ -1,15 +1,0 @@
-#### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
-# License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
-
-#### Workspace setup ####
-library(tidyverse)
-# [...UPDATE THIS...]
-
-#### Test data ####

### 2024-09-27 scripts/04-test_analysis_data.R

@@ -1,0 +1,69 @@
+#### Preamble ####
+# Purpose: Tests... [...UPDATE THIS...]
+# Author: Rohan Alexander [...UPDATE THIS...]
+# Date: 26 September 2024 [...UPDATE THIS...]
+# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# License: MIT
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(testthat)
+
+data <- read_csv("data/02-analysis_data/analysis_data.csv")
+
+
+#### Test data ####
+# Test that the dataset has 151 rows - there are 151 divisions in Australia
+test_that("dataset has 151 rows", {
+  expect_equal(nrow(analysis_data), 151)
+})
+
+# Test that the dataset has 3 columns
+test_that("dataset has 3 columns", {
+  expect_equal(ncol(analysis_data), 3)
+})
+
+# Test that the 'division' column is character type
+test_that("'division' is character", {
+  expect_type(analysis_data$division, "character")
+})
+
+# Test that the 'party' column is character type
+test_that("'party' is character", {
+  expect_type(analysis_data$party, "character")
+})
+
+# Test that the 'state' column is character type
+test_that("'state' is character", {
+  expect_type(analysis_data$state, "character")
+})
+
+# Test that there are no missing values in the dataset
+test_that("no missing values in dataset", {
+  expect_true(all(!is.na(analysis_data)))
+})
+
+# Test that 'division' contains unique values (no duplicates)
+test_that("'division' column contains unique values", {
+  expect_equal(length(unique(analysis_data$division)), 151)
+})
+
+# Test that 'state' contains only valid Australian state or territory names
+valid_states <- c("New South Wales", "Victoria", "Queensland", "South Australia", "Western Australia", 
+                  "Tasmania", "Northern Territory", "Australian Capital Territory")
+test_that("'state' contains valid Australian state names", {
+  expect_true(all(analysis_data$state %in% valid_states))
+})
+
+# Test that there are no empty strings in 'division', 'party', or 'state' columns
+test_that("no empty strings in 'division', 'party', or 'state' columns", {
+  expect_false(any(analysis_data$division == "" | analysis_data$party == "" | analysis_data$state == ""))
+})
+
+# Test that the 'party' column contains at least 2 unique values
[…]

### 2024-09-27 scripts/00-simulate_data.py
Add python and eda script

@@ -1,0 +1,48 @@
+#### Preamble ####
+# Purpose: Simulates a dataset of Australian electoral divisions, including the 
+  # state and party that won each division.
+# Author: Rohan Alexander
+# Date: 26 September 2024
+# Contact: rohan.alexander@utoronto.ca
+# License: MIT
+# Pre-requisites: 
+  # - `polars` must be installed (pip install polars)
+  # - `numpy` must be installed (pip install numpy)
+
+
+#### Workspace setup ####
+import polars as pl
+import numpy as np
+np.random.seed(853)
+
+
+#### Simulate data ####
+# State names
+states = [
+    "New South Wales", "Victoria", "Queensland", "South Australia", 
+    "Western Australia", "Tasmania", "Northern Territory", 
+    "Australian Capital Territory"
+]
+
+# Political parties
+parties = ["Labor", "Liberal", "Greens", "National", "Other"]
+
+# Probabilities for state and party distribution
+state_probs = [0.25, 0.25, 0.15, 0.1, 0.1, 0.1, 0.025, 0.025]
+party_probs = [0.40, 0.40, 0.05, 0.1, 0.05]
+
+# Generate the data using numpy and polars
+divisions = [f"Division {i}" for i in range(1, 152)]
+states_sampled = np.random.choice(states, size=151, replace=True, p=state_probs)
+parties_sampled = np.random.choice(parties, size=151, replace=True, p=party_probs)
+
+# Create a polars DataFrame
+analysis_data = pl.DataFrame({
+    "division": divisions,
+    "state": states_sampled,
+    "party": parties_sampled
+})
+
+
+#### Save data ####
+analysis_data.write_csv("data/00-simulated_data/simulated_data.csv")

### 2024-09-27 scripts/06-model_data.R

@@ -1,0 +1,37 @@
+#### Preamble ####
+# Purpose: Models... [...UPDATE THIS...]
+# Author: Rohan Alexander [...UPDATE THIS...]
+# Date: 11 February 2023 [...UPDATE THIS...]
+# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# License: MIT
+# Pre-requisites: [...UPDATE THIS...]
+# Any other information needed? [...UPDATE THIS...]
+
+
+#### Workspace setup ####
+library(tidyverse)
+library(rstanarm)
+
+#### Read data ####
+analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+
+### Model data ####
+first_model <-
+  stan_glm(
+    formula = flying_time ~ length + width,
+    data = analysis_data,
+    family = gaussian(),
+    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
+    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
+    prior_aux = exponential(rate = 1, autoscale = TRUE),
+    seed = 853
+  )
+
+
+#### Save model ####
+saveRDS(
+  first_model,
+  file = "models/first_model.rds"
+)
+
+

### 2024-10-02 paper/paper.qmd
Update with more guidance.

@@
 
-You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.
+Overview paragraph
 
-The remainder of this paper is structured as follows. @sec-data....
+Estimand paragraph
+
+Results paragraph
+
+Why it matters paragraph
+
+Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
 
 # Data {#sec-data}
+
+## Overview
+
+We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
+
+Overview text
+
+## Measurement
+
+Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
+
+## Outcome variables
+
+Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
 
@@
 Talk way more about it.
+
+## Predictor variables
+
+Add graphs, tables and text.
+
+Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
 
@@
 
+Please don't use these as sub-heading labels - change them to be what your point actually is.
+
 ## Third discussion point

### 2024-10-19 paper/paper.qmd
Updated pollster intro

@@
 
+## pollster methodology analysis
+
+This survey was conducted by Redfield & Wilton Strategies to assess the voting intentions of eligible voters in key U.S.
+swing states ahead of the 2024 Presidential Election.
+The primary goal of this poll is to provide an accurate and timely snapshot of public opinion in states where electoral outcomes are uncertain and could have a decisive impact on the overall result of the election.
+Swing states, due to their political volatility and diverse voter bases, are critical in determining the balance of power in the U.S.
+electoral system.
+Understanding voter preferences in these states is essential for political analysts, campaigns, and the general public.
+
+### Population of Interest
+The population of interest for this survey consists of all eligible voters residing in major U.S.
+swing states, specifically Arizona, Florida, Georgia, Michigan, North Carolina, and Pennsylvania.
+These states are known for their fluctuating political alignments and are expected to play a crucial role in the upcoming election.
+
+### Sampling Frame
+The population sampled includes eligible voters from Arizona, Florida, Georgia, Michigan, North Carolina, and Pennsylvania. Participants were selected via an online panel.
+
+### Sample
+The sample sizes for each state were as follows:
+
+Arizona: 750 respondents
+
+Florida: 1,350 respondents
+
+Georgia: 927 respondents
+
+Michigan: 970 respondents
+
+North Carolina: 880 respondents
+
+Pennsylvania: 1,070 respondents
+
 # Model details {#sec-model-details}

### 2024-10-22 paper/paper.qmd
Update the paper.qmd

@@
 ---
-title: "My title"
-subtitle: "My subtitle if needed"
+title: "US President Prediction"
+subtitle: "Predict the results of 2024 US President Election"
 author: 
-  - First author
-  - Another author
-thanks: "Code and data are available at: [https://github.com/RohanAlexander/starter_folder](https://github.com/RohanAlexander/starter_folder)."
-date: today
+  - Yun Chu
+  - Felix Li
+  - Wen Han Zhao 
+thanks: "Code and data are available at: [https://github.com/younazhao/US-President-Prediction/tree/main](https://github.com/younazhao/US-President-Prediction/tree/main)."
+date: 22 October 2024
 date-format: long
-abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
+abstract: "In this study, we aim to study the 2024 U.S.
+Presidential Election with the methodology of  polls of polls across various states.
+With the insightful analyses on polls provided by Redfiled & Wilton Strategies, the conducted survey on the population highlights various key point for the prediction of US president.
+Using Generalized Linear Models (GLMs), this report analyzes the accuracy of prediction in president votes through several key variables from pollster.
+The model incorporates variables such as pollscore, methodology, transparency_score and hypothetical.
+The prediction finding highlights the significance of candicate trustworthiness among the voter decision, providing insights for swing voters.
+This analysis offers valuable information into the potential electoral outcomes driving the 2024 US president election.
+"
 format: pdf
@@
 #| message: false
-
 library(tidyverse)
-library(palmerpenguins)
+library(knitr)
+library(kableExtra)
+library(readr)
+library(dplyr)
 ```
@@
 
-We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...
+We use the statistical programming language R [@citeR].... Our data.... Following @tellingstories, we consider...
 
@@
 
-Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
-
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
-
-```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
-#| echo: false
-
-ggplot(penguins, aes(x = island, fill = species)) +
-  geom_bar(alpha = 0.8) +
-  scale_fill_manual(values = c("darkorange","purple","cyan4"),
-                    guide = "none") +
-  theme_minimal() +
-  facet_wrap(~species, ncol = 1) +
-  coord_flip()
-```
-
-Talk more about it.
-
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
-
-analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
-
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
-```
-
 Talk way more about it.
@@
 
-# Model
+# Model {#set-model}
 
@@
 
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix @sec-model-details].
 
@@
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+Define $y_i$ as the percentage of support for a candicate in a poll.
 
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
+${y_i} = \beta_0 + \beta_1 pollscore + \beta_2 methodology + \beta_3transparency score + \beta_4 hypothetical + \epsilon$
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+pollscore: A numeric value representing the score or reliability of the pollster in question (e.g., -1.1).
+"The error and bias we can attribute to a pollster.
+Negative numbers are better.
+Stands for "Predictive Optimization of Latent skill Level in Surveys, Considering Overall Record, Empirically.
+
+methodology: The method used to conduct the poll (e.g., Online Panel).
+
+transparency_score: A score reflecting the pollster’s transparency about their methodology (e.g., 9.0).
+"A grade for how transparent a pollster is, calculated based on how much information it discloses about its polls and weighted by
+by recency. The highest Transparency Score is 10."
+
+hypothetical: Indicates whether the poll is about a hypothetical match-up.
+
+We run the model in R [@citeR].
 
 ### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
-
-# Results
-
-Our results are summarized in @tbl-modelresults.
 
@@
 
-library(rstanarm)
-
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
+# Library setup
+#install.packages("glm")
+library(glmnet)
+library(ggplot2)
 ```
@@
 #| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
+#| label: tbl-model1
+#| tbl-cap: "model 1"
+#| warning: false
+# Get the analysis data
+analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
+
+# Model
+glm_model1 <- glm(pct ~ pollscore + methodology + transparency_score + hypothetical, data = analysis_data, family = gaussian())
+
+# Get the summary
+model_summary <- summary(glm_model1)
+
+# Convert the summary coefficients to a data frame
+summary_table <- as.data.frame(model_summary$coefficients)
+
+# Create a table with a caption and control the size
+kable(summary_table, caption = "Model 1 Summary Table", size = "small")
+```
+
+## Lasso Regularization {#sec-model-lassoregularization}
+Lasso Regularization is performed to see if the selected variable in our model would best predict the results of polls percentage of a candidate.
+If the variable do not align with our actual numbers, we should consider other variables or a interaction variable.
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-lassoregularization
+#| tbl-cap: "lasso regularization"
+#| warning: false
+# Prepare the predictors
+x <- model.matrix(pct ~ pollscore + methodology + transparency_score + hypothetical, data = analysis_data)[, -1]
+
+# Prepare the response variable
+y <- analysis_data$pct
+
+# Lasso regression (alpha = 1)
+lasso_model <- glmnet(x, y, alpha = 1, family = "gaussian")
+
+# Cross-validation for Lasso
+cv_lasso <- cv.glmnet(x, y, alpha = 1, family = "gaussian")
+
+plot(cv_lasso)
+
+# Get the best lambda
+best_lambda_lasso <- cv_lasso$lambda.min
+
+# Refit Lasso with the optimal lambda
+lasso_final_model <- glmnet(x, y, alpha = 1, lambda = best_lambda_lasso, family = "gaussian")
+
+# View Lasso coefficients
+coef.lasso <- coef(lasso_final_model)
+
+# Predict using Lasso
+pred_lasso <- predict(lasso_final_model, s = best_lambda_lasso, newx = x)
+
+```
+
+Computing the Mean Square Error would give us a sense of our lasso model prediction. 
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-msr
+#| tbl-cap: "Mean Squared Error"
+#| warning: false
+# Calculate Mean Squared Error
+mse <- mean((y - pred_lasso)^2)
+cat("Mean Squared Error:", mse, "\n")
+
+# Calculate R-squared
+sst <- sum((y - mean(y))^2)
+sse <- sum((y - pred_lasso)^2)
+r_squared <- 1 - (sse / sst)
+cat("R-squared:", r_squared, "\n")
+
+# Create a data frame for plotting
+results_df <- data.frame(
+  Metric = c("SST", "SSE", "R-squared"),
+  Value = c(sst, sse, r_squared)
+)
+
+kable(results_df, caption = "Mean Squared Error Table")
+```
+
+From this graph, we can see that the cluster points do not perfectly align on the line of the best fit so we should still investigate in other variables. 
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-actualvspredicted
+#| tbl-cap: "actual vs predicted"
 #| warning: false
 
-modelsummary::modelsummary(
-  list(
-    "First model" = first_model
-  ),
-  statistic = "mad",
-  fmt = 2
-)
+# Compare predicted vs actual values
+comparison <- data.frame(Actual = y, Predicted = as.vector(pred_lasso))
+
+kable(head(comparison), caption = "Actual vs Predicted")
+
+# Plot actual vs predicted values
+ggplot(comparison, aes(x = Actual, y = Predicted)) +
+  geom_point() +
+  geom_abline(slope = 1, intercept = 0, color = "red") +
+  labs(title = "Actual vs Predicted Values",
+       x = "Actual Values",
+       y = "Predicted Values") +
+  theme_minimal()
 ```
-
-# Discussion
-
-## First discussion point {#sec-first-point}
-
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
-
-## Second discussion point
-
-Please don't use these as sub-heading labels - change them to be what your point actually is.
-
-## Third discussion point
-
-## Weaknesses and next steps
-
-Weaknesses and next steps should also be included.
 
@@
 
+## Weakness & Strength of the methodology
+
+In terms of strengths, Redfield & Wilton has a great reputation for producing reliable polling data.
+They have employ a mix of online and telephone survey, which add in various resources of collecting their data.
+This approach can help reduce bias.
+In addition, Redfield & Wilton often target swing states, which makes their polls results important to the US president election.
+
+The weakness of their methodology would incorporate a certain potential bias based on their political leaning of their clients or media. This could cause a certain neutrality in their dataset.
+
+Overall, Redfield & Wilton has been considered as a competent, reputable and reliable pollster with his variaty and methodology on polls.
+Even though the pollster contained a potential bias, it has been a reliable resource in prediction of US president election.
+
 # Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-```{r}
-#| eval: true
-#| echo: false
-#| message: false
-#| warning: false
-#| label: fig-ppcheckandposteriorvsprior
-#| layout-ncol: 2
-#| fig-cap: "Examining how the model fits, and is affected by, the data"
-#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
-
-pp_check(first_model) +
-  theme_classic() +
-  theme(legend.position = "bottom")
-
-posterior_vs_prior(first_model) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-```
-
-## Diagnostics
-
-@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...
-
-@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...
-
-```{r}
-#| echo: false
-#| eval: true
-#| message: false
-#| warning: false
-#| label: fig-stanareyouokay
-#| fig-cap: "Checking the convergence of the MCMC algorithm"
-#| fig-subcap: ["Trace plot", "Rhat plot"]
-#| layout-ncol: 2
-
-plot(first_model, "trace")
-
-plot(first_model, "rhat")
-```
 

### 2024-10-22 scripts/02-download_data.R
Update 02-download_data.R

@@ -1,26 +1,24 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Downloads and saves the data from 538 president general 
+# Authors: Yun Chu, Felix Li, and Wen Han Zhao 
+# Date: 22 October 2024
+# Contact: youna.zhao@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: Dataset found from 538 website
+# Any other information needed? None
 
 
 #### Workspace setup ####
-library(opendatatoronto)
+
 library(tidyverse)
-# [...UPDATE THIS...]
+library(readr)
 
 #### Download data ####
-# [...ADD CODE HERE TO DOWNLOAD...]
 
+president_polls <- read_csv("~/Downloads/president_polls.csv")
+data <- president_polls
 
 
 #### Save data ####
-# [...UPDATE THIS...]
-# change the_raw_data to whatever name you assigned when you downloaded it.
-write_csv(the_raw_data, "inputs/data/raw_data.csv") 
 
-         
+write_csv(data, "data/01-raw_data/raw_data.csv") 

### 2024-10-22 scripts/03-clean_data.R
Update 03-clean_data.R

@@ -1,44 +1,20 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Cleans the raw data from 538 by removing missing values
+# Author: Yun Chu, Felix Li, and Wen Han Zhao 
+# Date: 22 October 2024
+# Contact: youna.zhao@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: raw data has been downloaded from the website
+# Any other information needed? None
 
 #### Workspace setup ####
 library(tidyverse)
 
 #### Clean data ####
-raw_data <- read_csv("inputs/data/plane_data.csv")
+raw_data <- read_csv("data/01-raw_data/raw_data.csv")
 
-cleaned_data <-
-  raw_data |>
-  janitor::clean_names() |>
-  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-  filter(wing_width_mm != "caw") |>
-  mutate(
-    flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
-                                   "1.35",
-                                   flying_time_sec_first_timer)
-  ) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "490",
-                                 "49",
-                                 wing_width_mm)) |>
-  mutate(wing_width_mm = if_else(wing_width_mm == "6",
-                                 "60",
-                                 wing_width_mm)) |>
-  mutate(
-    wing_width_mm = as.numeric(wing_width_mm),
-    wing_length_mm = as.numeric(wing_length_mm),
-    flying_time_sec_first_timer = as.numeric(flying_time_sec_first_timer)
-  ) |>
-  rename(flying_time = flying_time_sec_first_timer,
-         width = wing_width_mm,
-         length = wing_length_mm
-         ) |> 
-  tidyr::drop_na()
+# Remove rows with any missing values in the relevant columns
+data_clean <- na.omit(data[, c("pct", "pollscore", "methodology", "transparency_score", "hypothetical")])
 
 #### Save data ####
[…]

### 2024-10-27 scripts/00-simulate_data.R
simuated data and script

@@ -1,52 +1,40 @@
 #### Preamble ####
-# Purpose: Simulates a dataset of Australian electoral divisions, including the 
-  #state and party that won each division.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
+# Purpose: Cleans the raw data from 538 by removing missing values
+# Author: Yun Chu, Felix Li, and Wen Han Zhao 
+# Date: 22 October 2024
+# Contact: youna.zhao@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: The `tidyverse` package must be installed
-# Any other information needed? Make sure you are in the `starter_folder` rproj
+# Pre-requisites: raw data has been downloaded from the website
+# Any other information needed? None
 
 
 #### Workspace setup ####
 library(tidyverse)
+library(dplyr)
 set.seed(853)
 
-
 #### Simulate data ####
-# State names
-states <- c(
-  "New South Wales",
-  "Victoria",
-  "Queensland",
-  "South Australia",
-  "Western Australia",
-  "Tasmania",
-  "Northern Territory",
-  "Australian Capital Territory"
+
+# Define the sample size as an adjustable parameter
+sample_size <- 100  # Change this to any desired number of rows
+
+# Simulate data
+simulated_data <- tibble(
+  pollster = sample(c("InsiderAdvantage", "TIPP", "YouGov", "Ipsos", "Gallup"), sample_size, replace = TRUE),
+  numeric_grade = round(runif(sample_size, 1, 5), 1),
+  pollscore = round(rnorm(sample_size, mean = 0, sd = 1), 1),
+  methodology = sample(c("Online", "Phone", "Mixed"), sample_size, replace = TRUE),
+  transparency_score = round(runif(sample_size, 0, 10), 1),
+  state = sample(state.abb, sample_size, replace = TRUE),
+  start_date = sample(seq(as.Date('2024-01-01'), as.Date('2024-11-01'), by="day"), sample_size, replace = TRUE),
+  population_full = sample(c("Registered Voters", "Likely Voters"), sample_size, replace = TRUE),
+  candidate_name = sample(c("Kamala Harris", "Donald Trump"), sample_size, replace = TRUE),
+  pct = round(runif(sample_size, 45, 55), 1)
 )
 
-# Political parties
-parties <- c("Labor", "Liberal", "Greens", "National", "Other")
-
[…]

### 2024-10-27 scripts/01-test_simulated_data.R
test for simulated data

@@ -1,89 +1,78 @@
 #### Preamble ####
-# Purpose: Tests the structure and validity of the simulated Australian 
-  #electoral divisions dataset.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
+# Purpose: Tests for simulated data
+# Author: Yun Chu, Felix Li, and Wen Han Zhao 
+# Date: 22 October 2024
+# Contact: youna.zhao@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: 
-  # - The `tidyverse` package must be installed and loaded
-  # - 00-simulate_data.R must have been run
-# Any other information needed? Make sure you are in the `starter_folder` rproj
+# Pre-requisites: raw data has been downloaded from the website
+# Any other information needed? None
 
 
 #### Workspace setup ####
-library(tidyverse)
-
-analysis_data <- read_csv("data/00-simulated_data/simulated_data.csv")
-
-# Test if the data was successfully loaded
-if (exists("analysis_data")) {
-  message("Test Passed: The dataset was successfully loaded.")
-} else {
-  stop("Test Failed: The dataset could not be loaded.")
-}
+# Load libraries
+library(testthat)
+library(tibble)
+library(readr)
 
 
-#### Test data ####
+simulated_data <- read.csv('data/00-simulated_data/simulated_data.csv')
 
-# Check if the dataset has 151 rows
-if (nrow(analysis_data) == 151) {
-  message("Test Passed: The dataset has 151 rows.")
-} else {
-  stop("Test Failed: The dataset does not have 151 rows.")
-}
+# Define the test cases
+test_that("Simulated data has correct structure and values", {
+  # Check that the data frame has the expected number of rows and columns
+  expect_equal(nrow(simulated_data), sample_size)
+  expect_equal(ncol(simulated_data), 10)
+  
+  # Check column names
+  expect_equal(
+    colnames(simulated_data),
+    c("pollster", "numeric_grade", "pollscore", "methodology", 
+      "transparency_score", "state", "start_date", 
+      "population_full", "candidate_name", "pct")
+  )
+})
 
-# Check if the dataset has 3 columns
-if (ncol(analysis_data) == 3) {
-  message("Test Passed: The dataset has 3 columns.")
-} else {
-  stop("Test Failed: The dataset does not have 3 columns.")
-}
[…]

### 2024-10-27 scripts/02-download_data.R
updated downloaded data script

@@ -15,10 +15,9 @@ library(readr)
 
 #### Download data ####
 
-president_polls <- read_csv("~/Downloads/president_polls.csv")
-data <- president_polls
+#### Download data ####
+raw_data <- read_csv("https://projects.fivethirtyeight.com/polls/data/president_polls.csv")
 
 
 #### Save data ####
-
-write_csv(data, "data/01-raw_data/raw_data.csv") 
+write_csv(raw_data, "data/01-raw_data/poll_raw_data.csv") 

### 2024-10-27 scripts/01-test_simulated_data.R
updated test files

@@ -17,6 +17,8 @@ library(readr)
 
 simulated_data <- read.csv('data/00-simulated_data/simulated_data.csv')
 
+sample_size <- 100
+
 # Define the test cases
 test_that("Simulated data has correct structure and values", {
   # Check that the data frame has the expected number of rows and columns

### 2024-10-27 scripts/04-test_analysis_data.R

@@ -1,69 +1,78 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 26 September 2024 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Tests for cleaned data
+# Author: Yun Chu, Felix Li, and Wen Han Zhao 
+# Date: 22 October 2024
+# Contact: youna.zhao@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: raw data has been downloaded from the website
+# Any other information needed? None
 
 
 #### Workspace setup ####
-library(tidyverse)
+# Load libraries
 library(testthat)
+library(tibble)
+library(readr)
 
-data <- read_csv("data/02-analysis_data/analysis_data.csv")
 
+simulated_data <- read.csv('data/02-analysis_data/analysis_data.csv')
 
-#### Test data ####
-# Test that the dataset has 151 rows - there are 151 divisions in Australia
-test_that("dataset has 151 rows", {
-  expect_equal(nrow(analysis_data), 151)
+# Define the test cases
+test_that("Simulated data has correct structure and values", {
+  # Check that the data frame has the expected number of rows and columns
+  expect_equal(nrow(simulated_data), sample_size)
+  expect_equal(ncol(simulated_data), 10)
+  
+  # Check column names
+  expect_equal(
+    colnames(simulated_data),
+    c("pollster", "numeric_grade", "pollscore", "methodology", 
+      "transparency_score", "state", "start_date", 
+      "population_full", "candidate_name", "pct")
+  )
 })
 
-# Test that the dataset has 3 columns
-test_that("dataset has 3 columns", {
-  expect_equal(ncol(analysis_data), 3)
+test_that("Pollster values are from expected categories", {
+  expect_true(all(simulated_data$pollster %in% c("InsiderAdvantage", "TIPP", "YouGov", "Ipsos", "Gallup")))
 })
 
-# Test that the 'division' column is character type
-test_that("'division' is character", {
-  expect_type(analysis_data$division, "character")
+test_that("Numeric grade is within the range 1 to 5", {
+  expect_true(all(simulated_data$numeric_grade >= 1 & simulated_data$numeric_grade <= 5))
 })
 
[…]

### 2024-10-27 scripts/00-simulate_data.R
model change

@@ -1,5 +1,5 @@
 #### Preamble ####
-# Purpose: Cleans the raw data from 538 by removing missing values
+# Purpose: Simulate data of Interest
 # Author: Yun Chu, Felix Li, and Wen Han Zhao 
 # Date: 22 October 2024
 # Contact: youna.zhao@mail.utoronto.ca

### 2024-10-27 scripts/06-model_data.R

@@ -7,30 +7,99 @@
 # Pre-requisites: [...UPDATE THIS...]
 # Any other information needed? [...UPDATE THIS...]
 
-
-#### Workspace setup ####
+#### Set-up ####
+# Load libraries
 library(tidyverse)
+library(janitor)
+library(lubridate)
+library(broom)
+library(modelsummary)
 library(rstanarm)
+library(splines)
 
-#### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
 
-### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
+#### Prepare dataset ####
+# Read in the data and clean variable names
+data <- read_csv("data/01-raw_data/poll_raw_data.csv") |>
+  clean_names()
+
+# Filter data to Harris estimates based on high-quality polls after she declared
+just_harris_high_quality <- data |>
+  filter(
+    candidate_name == "Kamala Harris",
+    numeric_grade >= 2.8,
+    transparency_score >= 5,
+    pollscore <= 0 # Need to investigate this choice - come back and fix. 
+    # Also need to look at whether the pollster has multiple polls or just one or two - filter out later
+  ) |>
+  mutate(
+    state = if_else(is.na(state), "National", state), # Hacky fix for national polls - come back and check
+    end_date = mdy(end_date)
+  ) |>
+  filter(end_date >= as.Date("2024-07-21")) |> # When Harris declared
+  mutate(
+    num_harris = round((pct / 100) * sample_size, 0) # Need number not percent for some models
   )
 
 
+#### Plot data ####
+base_plot <- ggplot(just_harris_high_quality, aes(x = end_date, y = pct)) +
+  theme_classic() +
+  labs(y = "Harris percent", x = "Date")
+
+# Plots poll estimates and overall smoothing
+base_plot +
+  geom_point() +
+  geom_smooth()
+
+# Color by pollster
+# This gets messy - need to add a filter - see line 21
+base_plot +
+  geom_point(aes(color = pollster)) +
+  geom_smooth() +
+  theme(legend.position = "bottom")
+
+# Facet by pollster
[…]

### 2024-10-29 paper/paper.qmd
Update Idealized Methodology & Survey Section Update Reference

@@
 
+# Appendix 2 - Idealized Methodology and Survey
+
+## Overview
+
+This section introduces an idealized methodology and survey with $100K budget to predict the 2024 US presidential election.
+The goal is to maximize the accuracy of the prediction under the budget.
+The subsections that describe the details of the idealized methodology and survey include sampling approach, respondents recruiting method, data validation, poll aggregation and the survey questions.
+
+### Sampling Approach
+
+Cluster sampling is the sampling approach used.
+Specifically clusters are the swing states as they are the states that would affect the election result, i.e.
+Arizona, Florida, Gegorgia, Iowa, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin.
+In each swing state/clusters, units are selected based on postal code.
+Using simple random sampling, 100 distinct postal codes are randomly selected.
+People whose living address has the postal codes selected are the target respondents.
+For the postal codes that lie in the non-residential area, the postal codes would be ignored.
+
+### Recruitment Strategy
+
+The strategy to be implemented to recruit respondents is a combination of physical and online recruitment.
+
+If the building corresponds to the postal code is a condo building, a paper with QR code to the survey is going to be put in the lobby or elevator, or sent to residents through the property management.
+If the building is a residential house, then a letter with the QR code would be put in the mailbox.
+
+All respondents are rewarded with a $5 deposit to their bank account or a gift card of their choice. Each IP address is limited to answer the survey once.
+
+### Data Validation
+
+To improve accuracy of the prediction results, the following data validation approaches will be done.
+
+* Postal Code Validation
+ - All the postal codes got from respondents are going to be validated to check if it is one of the randomly selected postal codes.
+If not, the response would not be considered when the prediction is done.
+* Just-for-Rewards Prevention
+ - To identify the responses that are not seriously answered, the last question of the survey is set test if the respondents are serious and careful when answering the questions.
+* Age Validity
+ - Responses with age under 18 but answered "Yes" to "Are you registered to Vote" are discarded.
+
+### Poll Aggregation
+
+For each of the swing states, the result would be calculated based on the votes to Trump versus Harris because they are the candidates with the greatest chances of winning.
+Based on the decisiveness of the respondents and the likelihood of voting from the responses, the individual votes will be weighted when calculating the results for each swing states.
+
+To aggregate the polls for all states, the result of each states is multiplied by its electoral votes.
+After summing the electoral votes for Trump or Harris, the estimated electoral votes that Trump or Harris get is available.
+The candidate that has more than 270 electoral votes would be predicted to win the election {@nationalarchives2024electoral}.
+
+### Survey
+
+Google form of the survey is available here:  https://docs.google.com/forms/d/1obaebX3zqw7WJEd1lHrF-DVBXdZt5tbjHofjsWj9zf8/prefill
+
+1. What is the postal code of where you live?
+
+2. Are you registered to vote?
+
+* Yes
+* No
+
+3. Who would you vote for?
+
+* Donald Trump - Republican
+* Kamala Harris - Democrat
+* Other Candidates
+* Not Decided Yet
+
+4. How decisive are you to vote for the option you chose in the last question?
+
+* Very Decisive
+* Pretty Decisive
+* A Bit Indecisive
+* Very Indecisive
+
+5. How likely are you to vote for the election?
+
+* Very Likely
+* A Bit Likely
+* Not Likely
+
+6. What age group are you in?
+
+* 0 - 18
+* 19 - 30
+* 31 - 50
+* 51 - 70
+* 71+
+
+7.
+
+8.
+
+9.
+
+10. How many characters are present in the word "President"?
+
+* 6
+* 8
+* 10
+* None of the Above
+
+### Budget Specification
+
+* Physical and Online Recruitment: $30,000
+* Rewards for Respondents: $50,000
+* Data Collection & Validation: $10,000
+* Other: $10,000
+
+Total: $100,000
+
+### Conclusion
+
+The survey uses cluster sampling to sample the swing states based on postal codes.
+After data collection and validation, weighting based on the likelihood of voting and decisiveness of the respondents, the total votes for Donald Trump and Kamala Harris are calulated.
+The candidate with more than 270 votes are predicted to win the election{@nationalarchives2024electoral}.
+
 # References

### 2024-10-29 paper/paper.qmd
update bullet point format

@@
 * Postal Code Validation
- - All the postal codes got from respondents are going to be validated to check if it is one of the randomly selected postal codes.
+    + All the postal codes got from respondents are going to be validated to check if it is one of the randomly selected postal codes.
 If not, the response would not be considered when the prediction is done.
 * Just-for-Rewards Prevention
- - To identify the responses that are not seriously answered, the last question of the survey is set test if the respondents are serious and careful when answering the questions.
+    + To identify the responses that are not seriously answered, the last question of the survey is set test if the respondents are serious and careful when answering the questions.
 * Age Validity
- - Responses with age under 18 but answered "Yes" to "Are you registered to Vote" are discarded.
+    + Responses with age under 18 but answered "Yes" to "Are you registered to Vote" are discarded.
 

### 2024-10-29 scripts/06-model_data.R
model change

@@ -91,11 +91,11 @@ ggplot(just_harris_high_quality_1, aes(x = end_date)) +
   geom_point(aes(y = pct), color = "black") +
   geom_line(aes(y = fitted_date_pollster), color = "blue", linetype = "dotted") +
   theme_classic() +
-  labs(y = "Harris percent", x = "Date", title = "Linear Model: pct ~ end_date + pollster")
+  labs(y = "Harris percent", x = "Date", title = "Linear Model: pct ~ end_date + pollster + state")
 
 # All of the above would be in scripts - data cleaning scripts and modelling scripts. 
 # This is an example of how you get a results table that you could put into your Quarto doc
-modelsummary(models = list("Model 1" = model_date, "Model 2" = model_date_pollster))
+modelsummary(models = list("Model 2" = model_date_pollster))
 
 #### Save model ####
 saveRDS(

### 2024-10-30 paper/paper.qmd
updated model section

@@
 
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix @sec-model-details].
-
+As indicated in the @sec-data, we do not have a confident belief of priors of parameters of our model since Harris replace the role of Biden by indecent and we do not have enough data about her previous polling outcomes as a president candidate compared to Biden.
+Therefore, we would not use the bayesian model but instead using a more simple multilinear regression model which predicts the support rate of Trump and Harris using time towards elections date and pollsters as they are beliveved to highly corelated to the response variable in @sec-data.
+We would bulid 8 models whihc represent national trend and that of seven swing states.
 ## Model set-up
 
-Define $y_i$ as the percentage of support for a candicate in a poll.
+Define $y_{i,\text{stateA}}$ as the percentage of support for a candicate in a poll in the specific state.
 
-${y_i} = \beta_0 + \beta_1 pollscore + \beta_2 methodology + \beta_3transparency score + \beta_4 hypothetical + \epsilon$
+$$
+y_{i,national} = \beta_{0,national} + \beta_{1,time,national} + \beta_{2,pollster,national} + \epsilon_{national}
+$$ 
+$$
+y_{i,\text{AZ}} = \beta_{0,\text{AZ}} + \beta_{1,\text{time,AZ}} + \beta_{2,\text{pollster,AZ}} + \epsilon_{\text{AZ}}
+$$
+$$
+y_{i,\text{FL}} = \beta_{0,\text{FL}} + \beta_{1,\text{time,FL}} + \beta_{2,\text{pollster,FL}} + \epsilon_{\text{FL}}
+$$
+$$
+y_{i,\text{GA}} = \beta_{0,\text{GA}} + \beta_{1,\text{time,GA}} + \beta_{2,\text{pollster,GA}} + \epsilon_{\text{GA}}
+$$
+$$
+y_{i,\text{MI}} = \beta_{0,\text{MI}} + \beta_{1,\text{time,MI}} + \beta_{2,\text{pollster,MI}} + \epsilon_{\text{MI}}
+$$
+$$
+y_{i,\text{NC}} = \beta_{0,\text{NC}} + \beta_{1,\text{time,NC}} + \beta_{2,\text{pollster,NC}} + \epsilon_{\text{NC}}
+$$
+$$
+y_{i,\text{PA}} = \beta_{0,\text{PA}} + \beta_{1,\text{time,PA}} + \beta_{2,\text{pollster,PA}} + \epsilon_{\text{PA}}
+$$
+$$
+y_{i,\text{WI}} = \beta_{0,\text{WI}} + \beta_{1,\text{time,WI}} + \beta_{2,\text{pollster,WI}} + \epsilon_{\text{WI}}
+$$
 
-pollscore: A numeric value representing the score or reliability of the pollster in question (e.g., -1.1).
-"The error and bias we can attribute to a pollster.
-Negative numbers are better.
-Stands for "Predictive Optimization of Latent skill Level in Surveys, Considering Overall Record, Empirically.
+time:The date when the pollster starts polling
 
-methodology: The method used to conduct the poll (e.g., Online Panel).
-
-transparency_score: A score reflecting the pollster’s transparency about their methodology (e.g., 9.0).
-"A grade for how transparent a pollster is, calculated based on how much information it discloses about its polls and weighted by
-by recency. The highest Transparency Score is 10."
-
-hypothetical: Indicates whether the poll is about a hypothetical match-up.
+pollster: The pollster which run the polling
 
@@
 
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-# Library setup
-#install.packages("glm")
-library(glmnet)
-library(ggplot2)
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-model1
-#| tbl-cap: "model 1"
-#| warning: false
-# Get the analysis data
-analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
-
-# Model
-glm_model1 <- glm(pct ~ pollscore + methodology + transparency_score + hypothetical, data = analysis_data, family = gaussian())
-
-# Get the summary
-model_summary <- summary(glm_model1)
-
-# Convert the summary coefficients to a data frame
-summary_table <- as.data.frame(model_summary$coefficients)
-
-# Create a table with a caption and control the size
-kable(summary_table, caption = "Model 1 Summary Table", size = "small")
-```
-
-## Lasso Regularization {#sec-model-lassoregularization}
-Lasso Regularization is performed to see if the selected variable in our model would best predict the results of polls percentage of a candidate.
-If the variable do not align with our actual numbers, we should consider other variables or a interaction variable.
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-lassoregularization
-#| tbl-cap: "lasso regularization"
-#| warning: false
-# Prepare the predictors
-x <- model.matrix(pct ~ pollscore + methodology + transparency_score + hypothetical, data = analysis_data)[, -1]
-
-# Prepare the response variable
-y <- analysis_data$pct
-
-# Lasso regression (alpha = 1)
-lasso_model <- glmnet(x, y, alpha = 1, family = "gaussian")
-
-# Cross-validation for Lasso
-cv_lasso <- cv.glmnet(x, y, alpha = 1, family = "gaussian")
-
-plot(cv_lasso)
-
-# Get the best lambda
-best_lambda_lasso <- cv_lasso$lambda.min
-
-# Refit Lasso with the optimal lambda
-lasso_final_model <- glmnet(x, y, alpha = 1, lambda = best_lambda_lasso, family = "gaussian")
-
-# View Lasso coefficients
-coef.lasso <- coef(lasso_final_model)
-
-# Predict using Lasso
-pred_lasso <- predict(lasso_final_model, s = best_lambda_lasso, newx = x)
-
-```
-
-Computing the Mean Square Error would give us a sense of our lasso model prediction. 
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-msr
-#| tbl-cap: "Mean Squared Error"
-#| warning: false
-# Calculate Mean Squared Error
-mse <- mean((y - pred_lasso)^2)
-cat("Mean Squared Error:", mse, "\n")
-
-# Calculate R-squared
-sst <- sum((y - mean(y))^2)
-sse <- sum((y - pred_lasso)^2)
-r_squared <- 1 - (sse / sst)
-cat("R-squared:", r_squared, "\n")
-
-# Create a data frame for plotting
-results_df <- data.frame(
-  Metric = c("SST", "SSE", "R-squared"),
-  Value = c(sst, sse, r_squared)
-)
-
-kable(results_df, caption = "Mean Squared Error Table")
-```
-
-From this graph, we can see that the cluster points do not perfectly align on the line of the best fit so we should still investigate in other variables. 
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-actualvspredicted
-#| tbl-cap: "actual vs predicted"
-#| warning: false
-
-# Compare predicted vs actual values
-comparison <- data.frame(Actual = y, Predicted = as.vector(pred_lasso))
-
-kable(head(comparison), caption = "Actual vs Predicted")
-
-# Plot actual vs predicted values
-ggplot(comparison, aes(x = Actual, y = Predicted)) +
-  geom_point() +
-  geom_abline(slope = 1, intercept = 0, color = "red") +
-  labs(title = "Actual vs Predicted Values",
-       x = "Actual Values",
-       y = "Predicted Values") +
-  theme_minimal()
-```
+# Results {#set-result}
 

### 2024-10-30 scripts/06-model_data.R
update model script

@@ -41,66 +41,58 @@ just_harris_high_quality <- data |>
     num_harris = round((pct / 100) * sample_size, 0) # Need number not percent for some models
   )
 
+swing_states <- c("Arizona", "Florida", "Georgia", "Michigan", "Nevada", "North Carolina", "Pennsylvania")
 
-#### Plot data ####
-base_plot <- ggplot(just_harris_high_quality, aes(x = end_date, y = pct)) +
-  theme_classic() +
-  labs(y = "Harris percent", x = "Date")
+harris_models <- list()
+states <- c("National", swing_states)
+for (state in states) {
+  # Filter data for each state
+  state_data <- just_harris_high_quality %>%
+    filter(state == state)
+  
+  # Fit linear model
+  model <- lm(pct ~ end_date + pollster, data = state_data)
+  
+  # Store model in list with state name
+  harris_models[[state]] <- model
+}
 
-# Plots poll estimates and overall smoothing
-base_plot +
-  geom_point() +
-  geom_smooth()
-
-# Color by pollster
-# This gets messy - need to add a filter - see line 21
-base_plot +
-  geom_point(aes(color = pollster)) +
-  geom_smooth() +
-  theme(legend.position = "bottom")
-
-# Facet by pollster
-# Make the line 21 issue obvious
-# Also - is there duplication???? Need to go back and check
-base_plot +
-  geom_point() +
-  geom_smooth() +
-  facet_wrap(vars(pollster))
-
-# Color by pollscore
-base_plot +
-  geom_point(aes(color = factor(pollscore))) +
-  geom_smooth() +
-  theme(legend.position = "bottom")
-
-
-#### Starter models ####
-
-# Model 2: pct as a function of end_date and pollster
-model_date_pollster <- lm(pct ~ end_date + pollster+ state, data = just_harris_high_quality)
-
-# Augment data with model predictions
-just_harris_high_quality_1 <- just_harris_high_quality |>
+just_trump_high_quality <- data %>%
+  filter(
+    candidate_name == "Donald Trump",
+    numeric_grade >= 2.8,
+    transparency_score >= 5,
+    pollscore <= 0  # Placeholder for pollscore filter, adjust as necessary
+  ) %>%
   mutate(
-    fitted_date_pollster = predict(model_date_pollster),
+    state = if_else(is.na(state), "National", state), # Replace NA with "National" for national polls
+    end_date = mdy(end_date)
+  ) %>%
[…]

### 2024-10-30 paper/paper.qmd
Updated model and result

@@
 library(dplyr)
+library(rstanarm)
+library(bayesplot)
+library(ggplot2)
+library(ggpubr)
 ```
@@
 
-# Model {#set-model}
+# Model {#sec-model}
 
-As indicated in the @sec-data, we do not have a confident belief of priors of parameters of our model since Harris replace the role of Biden by indecent and we do not have enough data about her previous polling outcomes as a president candidate compared to Biden.
-Therefore, we would not use the bayesian model but instead using a more simple multilinear regression model which predicts the support rate of Trump and Harris using time towards elections date and pollsters as they are beliveved to highly corelated to the response variable in @sec-data.
-We would bulid 8 models whihc represent national trend and that of seven swing states.
-## Model set-up
+In this analysis, we aim to model the popularity trends for Kamala Harris and Donald Trump, based on high-quality polling data collected at the national level after Harris’s declaration on July 21, 2024.
+We used Bayesian linear regression models with Gaussian error structures to estimate changes in polling percentages over time for each candidate.
 
-Define $y_{i,\text{stateA}}$ as the percentage of support for a candicate in a poll in the specific state.
+## Data Filtering and Preparation
 
-$$
-y_{i,national} = \beta_{0,national} + \beta_{1,time,national} + \beta_{2,pollster,national} + \epsilon_{national}
-$$ 
-$$
-y_{i,\text{AZ}} = \beta_{0,\text{AZ}} + \beta_{1,\text{time,AZ}} + \beta_{2,\text{pollster,AZ}} + \epsilon_{\text{AZ}}
-$$
-$$
-y_{i,\text{FL}} = \beta_{0,\text{FL}} + \beta_{1,\text{time,FL}} + \beta_{2,\text{pollster,FL}} + \epsilon_{\text{FL}}
-$$
-$$
-y_{i,\text{GA}} = \beta_{0,\text{GA}} + \beta_{1,\text{time,GA}} + \beta_{2,\text{pollster,GA}} + \epsilon_{\text{GA}}
-$$
-$$
-y_{i,\text{MI}} = \beta_{0,\text{MI}} + \beta_{1,\text{time,MI}} + \beta_{2,\text{pollster,MI}} + \epsilon_{\text{MI}}
-$$
-$$
-y_{i,\text{NC}} = \beta_{0,\text{NC}} + \beta_{1,\text{time,NC}} + \beta_{2,\text{pollster,NC}} + \epsilon_{\text{NC}}
-$$
-$$
-y_{i,\text{PA}} = \beta_{0,\text{PA}} + \beta_{1,\text{time,PA}} + \beta_{2,\text{pollster,PA}} + \epsilon_{\text{PA}}
-$$
-$$
-y_{i,\text{WI}} = \beta_{0,\text{WI}} + \beta_{1,\text{time,WI}} + \beta_{2,\text{pollster,WI}} + \epsilon_{\text{WI}}
-$$
+The data were first filtered to retain only high-quality polls (with a numeric grade of 2.0 or above and a transparency score of 4 or above).
+We focused on polls where the `state` was either explicitly labeled as "National" or where the state information was missing (which we imputed as "National").
+Additionally, we filtered polls collected after July 21, 2024, when Kamala Harris announced her candidacy.
 
-time:The date when the pollster starts polling
+- **Kamala Harris Data**: After filtering for high-quality polls and setting the timeframe, we calculated the number of Harris supporters per poll by taking the product of her polling percentage and the poll’s sample size.
+- **Donald Trump Data**: We applied the same filtering and calculation methods to Trump’s data, ensuring comparability between both candidates.
 
-pollster: The pollster which run the polling
+To avoid issues with missing values in our models, we excluded any rows where key variables (e.g., polling percentage or end date) were missing.
 
-We run the model in R [@citeR].
+## Model Specification
 
-### Model justification
+Two separate Bayesian linear regression models were fitted using the `stan_glm` function. Both models specified the formula:
+
+\[
+\text{pct} \sim \text{end\_date}
+\]
+
+indicating that we are modeling the polling percentage as a function of time (end date of the polls).
+
+### Model for Kamala Harris
+
+The model for Kamala Harris was fitted as follows:
+
+- **Response Variable**: `pct` (polling percentage)
+- **Predictor**: `end_date` (date of the poll)
+- **Family**: Gaussian (normal distribution)
+- **Priors**:
+  - Normal prior for the slope: \( \text{Normal}(0, 0.1) \)
+  - Normal prior for the intercept: \( \text{Normal}(50, 5) \)
+  - Exponential prior for auxiliary parameters: \( \text{Exponential}(1) \)
+
+### Model for Donald Trump
+The model for Donald Trump was fitted using the same specifications as for Kamala Harris
 
 # Results {#set-result}
+The results of the Bayesian linear regression models for Kamala Harris and Donald Trump are summarized in @tbl-model_summary_H and @tbl-model_summary_T.
+
+```{r}
+#| label: tbl-model_summary_H
+#| echo: false
+#| tbl-cap: "Summary statistics of the Bayesian model for Harris"
+# Load required packages
+library(broom.mixed)  # For tidy model summaries from mixed models
+library(dplyr)
+library(readr)
+library(knitr)  # For nicely formatted tables
+library(modelsummary)
+
+# Load the saved models
+model1 <- readRDS(file = here::here("models/models_harris.rds"))
+
+# Extract a tidy summary of the model
+model_summary <- broom.mixed::tidy(model1, effects = "fixed", conf.int = TRUE) %>%
+  select(term, estimate, std.error, conf.low, conf.high) %>%
+  rename(
+    Term = term,
+    Estimate = estimate,
+    `Std. Error` = std.error,
+    `2.5% CI` = conf.low,
+    `97.5% CI` = conf.high
+  )
+
+# Display the summary table
+kable(model_summary, caption = "Summary of Bayesian Model for Harris", digits = 2)
+
+```
+```{r}
+#| label: tbl-model_summary_T
+#| echo: false
+#| tbl-cap: "Summary statistics of the Bayesian model for Harris"
+# Load required packages
+library(broom.mixed)  # For tidy model summaries from mixed models
+library(dplyr)
+library(readr)
+library(knitr)  # For nicely formatted tables
+
+# Load the saved models
+model2 <- readRDS(file = here::here("models/models_trump.rds"))
+
+# Extract a tidy summary of the model
+model_summary <- broom.mixed::tidy(model2, effects = "fixed", conf.int = TRUE) %>%
+  select(term, estimate, std.error, conf.low, conf.high) %>%
+  rename(
+    Term = term,
+    Estimate = estimate,
+    `Std. Error` = std.error,
+    `2.5% CI` = conf.low,
+    `97.5% CI` = conf.high
+  )
+
+# Display the summary table
+kable(model_summary, caption = "Summary of Bayesian Model for Harris", digits = 2)
+```
+## Overall trend
+The intercepts for both models are significantly negative, indicating a downward trend in polling percentages over time for both candidates.
+The positive coefficients for end_date suggest a slight increase in polling percentages as the election date approaches.
+However, the magnitude of these increases is relatively small.
+
+## Model Diagnostics
+The posterior predictive checks (Figures 1 and 2) and the credible intervals for predictors (Figures 3 and 4) indicate that the models fit the data well.
+The diagnostics for both Harris and Trump models show that the models are stable and the predictions are reliable.Further diagnostics (Figures 5 to 8) confirm the robustness of the models.
+The R-hat values are all below 1.1, indicating good convergence of the MCMC chains.
 
@@
 
+## Model check
+
+```{r}
+#| label: fig-pos_check_H
+#| echo: false
+#| fig-cap: "Posterior Predictive Check for All Models"
+
+# Load necessary libraries
+library(rstanarm)
+library(bayesplot)
+library(ggplot2)
+library(patchwork)  # For combining plots
+
+pp_check(model1) +  
+    theme_classic() +
+    theme(legend.position = "bottom")
+
+```
+```{r}
+#| label: fig-pos_check_T
+#| echo: false
+#| fig-cap: "Posterior Predictive Check for All Models"
+
+# Load necessary libraries
+library(rstanarm)
+library(bayesplot)
+library(ggplot2)
+library(patchwork)  # For combining plots
+
+pp_check(model2) +  
+    theme_classic() +
+    theme(legend.position = "bottom")
+
+```
+```{r}
+#| label: fig-CI_H
+#| echo: false
+#| fig-cap: "CI for predictors for Harris"
+
+# Load necessary libraries
+library(rstanarm)
+library(bayesplot)
+library(ggplot2)
+library(patchwork)  # For combining plots
+
+plot(
+  model1,
+  "areas")
+```
+```{r}
+#| label: fig-CI_T
+#| echo: false
+#| fig-cap: "CI for predictors for Trump"
+
+# Load necessary libraries
+library(rstanarm)
+library(bayesplot)
+library(ggplot2)
+library(patchwork)  # For combining plots
+
+plot(
+  model2,
+  "areas")
+```
+
+```{r}
+#| label: fig-DI_H
+#| echo: false
+#| fig-cap: "Model diagonostic for Harris"
+
+# Load necessary libraries
+library(rstanarm)
+library(bayesplot)
+library(ggplot2)
+library(patchwork)  # For combining plots
+
+plot(model1, "trace")
+plot(model1, "rhat")
+
+```
+```{r}
+#| label: fig-DI_T
+#| echo: false
+#| fig-cap: "Model diagonostic for Trump"
+
+# Load necessary libraries
+library(rstanarm)
+library(bayesplot)
+library(ggplot2)
+library(patchwork)  # For combining plots
+
+plot(model2, "trace")
+plot(model2, "rhat")
+
+```
 ## pollster methodology analysis

### 2024-10-30 scripts/06-model_data.R

@@ -7,92 +7,94 @@
 # Pre-requisites: [...UPDATE THIS...]
 # Any other information needed? [...UPDATE THIS...]
 
-#### Set-up ####
-# Load libraries
-library(tidyverse)
-library(janitor)
+
+library(readr)
+library(dplyr)
 library(lubridate)
-library(broom)
-library(modelsummary)
 library(rstanarm)
-library(splines)
+library(janitor) # Load the janitor package for clean_names()
 
-
-#### Prepare dataset ####
-# Read in the data and clean variable names
+# Load and clean data
 data <- read_csv("data/01-raw_data/poll_raw_data.csv") |>
   clean_names()
 
 # Filter data to Harris estimates based on high-quality polls after she declared
+# Load required packages
+library(readr)
+library(dplyr)
+library(lubridate)
+library(rstanarm)
+library(janitor)
+
+# Load and clean data
+data <- read_csv("data/01-raw_data/poll_raw_data.csv") |>
+  clean_names()
+
+# Filter and prepare data for Harris model
 just_harris_high_quality <- data |>
   filter(
     candidate_name == "Kamala Harris",
-    numeric_grade >= 2.8,
-    transparency_score >= 5,
-    pollscore <= 0 # Need to investigate this choice - come back and fix. 
-    # Also need to look at whether the pollster has multiple polls or just one or two - filter out later
+    numeric_grade >= 2.0,
+    transparency_score >= 4,
+    pollscore <= 0
   ) |>
   mutate(
-    state = if_else(is.na(state), "National", state), # Hacky fix for national polls - come back and check
+    state = if_else(is.na(state), "National", state),
     end_date = mdy(end_date)
   ) |>
-  filter(end_date >= as.Date("2024-07-21")) |> # When Harris declared
+  filter(end_date >= as.Date("2024-07-21"),
+         state == "National") |>
   mutate(
-    num_harris = round((pct / 100) * sample_size, 0) # Need number not percent for some models
-  )
+    num_harris = round((pct / 100) * sample_size, 0)
+  ) |>
+  drop_na(pct, end_date)  # Drop rows with NA in pct or end_date
 
-swing_states <- c("Arizona", "Florida", "Georgia", "Michigan", "Nevada", "North Carolina", "Pennsylvania")
-
-harris_models <- list()
-states <- c("National", swing_states)
-for (state in states) {
-  # Filter data for each state
-  state_data <- just_harris_high_quality %>%
[…]

### 2024-10-30 scripts/06-model_data.R
model change

@@ -26,10 +26,6 @@ library(lubridate)
 library(rstanarm)
 library(janitor)
 
-# Load and clean data
-data <- read_csv("data/01-raw_data/poll_raw_data.csv") |>
-  clean_names()
-
 # Filter and prepare data for Harris model
 just_harris_high_quality <- data |>
   filter(
@@ -42,12 +38,14 @@ just_harris_high_quality <- data |>
     state = if_else(is.na(state), "National", state),
     end_date = mdy(end_date)
   ) |>
-  filter(end_date >= as.Date("2024-07-21"),
-         state == "National") |>
+  filter(state == "National") |>  # Remove the filter on end_date here
+  drop_na(pct, end_date) |>  # Drop rows with NA in pct or end_date
   mutate(
-    num_harris = round((pct / 100) * sample_size, 0)
-  ) |>
-  drop_na(pct, end_date)  # Drop rows with NA in pct or end_date
+    # Get the earliest and latest end dates
+    earliest_date = min(end_date),
+    # Create a new column for days after the earliest date
+    days_after_earliest = as.numeric(end_date - earliest_date),
+  )
 
 # Filter and prepare data for Trump model
 just_trump_high_quality <- data |>
@@ -61,18 +59,20 @@ just_trump_high_quality <- data |>
     state = if_else(is.na(state), "National", state),
     end_date = mdy(end_date)
   ) |>
-  filter(end_date >= as.Date("2024-07-21"),
-         state == "National") |>
+  filter(state == "National") |>  # Remove the filter on end_date here
+  drop_na(pct, end_date) |>  # Drop rows with NA in pct or end_date
   mutate(
-    num_trump = round((pct / 100) * sample_size, 0)
-  ) |>
-  drop_na(pct, end_date)  # Drop rows with NA in pct or end_date
+    # Get the earliest and latest end dates
+    earliest_date = min(end_date),
+    # Create a new column for days after the earliest date
+    days_after_earliest = as.numeric(end_date - earliest_date),
+  )
 
 write_csv(just_harris_high_quality, "data/02-analysis_data/Harris.csv")
 write_csv(just_trump_high_quality, "data/02-analysis_data/Trump.csv")
 # Fit the model for Harris
 harris_model <- stan_glm(
-  formula = pct ~ end_date,
+  formula = pct ~ days_after_earliest,
   data = just_harris_high_quality,
   family = gaussian(),
[…]

### 2024-10-31 paper/paper.qmd
Polished result and model sec

@@
 
-In this analysis, we aim to model the popularity trends for Kamala Harris and Donald Trump, based on high-quality polling data collected at the national level after Harris’s declaration on July 21, 2024.
+In this analysis, we aim to model the popularity trends for the two 2024 president nominates, Kamala Harris and Donald Trump, based on high-quality polling data collected at the national level after Harris’s declaration on July 21, 2024.
 We used Bayesian linear regression models with Gaussian error structures to estimate changes in polling percentages over time for each candidate.
@@
 
-The data were first filtered to retain only high-quality polls (with a numeric grade of 2.0 or above and a transparency score of 4 or above).
-We focused on polls where the `state` was either explicitly labeled as "National" or where the state information was missing (which we imputed as "National").
-Additionally, we filtered polls collected after July 21, 2024, when Kamala Harris announced her candidacy.
+The data were first filtered to retain only high-quality polls, defined as those with a numeric grade of 2.0 or above and a transparency score of 4 or higher.
+We focused exclusively on polls where the *state* variable indicates national coverage, ensuring the poll represents nationwide opinions rather than a single state, which could introduce bias.
+Additionally, we filtered for polls conducted after July 21, 2024, the date when Kamala Harris announced her candidacy.
 
-- **Kamala Harris Data**: After filtering for high-quality polls and setting the timeframe, we calculated the number of Harris supporters per poll by taking the product of her polling percentage and the poll’s sample size.
-- **Donald Trump Data**: We applied the same filtering and calculation methods to Trump’s data, ensuring comparability between both candidates.
-
-To avoid issues with missing values in our models, we excluded any rows where key variables (e.g., polling percentage or end date) were missing.
+To prevent issues with missing data in our models, we excluded any rows where key variables (such as polling percentage or end date) were missing.
 
@@
 
-Two separate Bayesian linear regression models were fitted using the `stan_glm` function. Both models specified the formula:
+Two separate Bayesian linear regression models were fitted using the @rs package. Both models specified the formula:
 
-\[
-\text{pct} \sim \text{end\_date}
-\]
+\begin{align}
+Y_i | \mu_i,\sigma  & \sim \text{Normal}(\mu_i,\sigma) \\
+\mu_i & = \beta_0 + \beta_1X_{i1}\\
+\beta_0 & \sim \text{Normal}(50, 5), \\ 
+\beta_1 & \sim \text{Normal}(0, 0.1)
+\end{align}
 
-indicating that we are modeling the polling percentage as a function of time (end date of the polls).
+Where $Y_i$ is the support rate for Harris or Trump, separately, and $X_{i1}$ is the number of days passed after the latest polling sample of Trump or Harris.
 
-### Model for Kamala Harris
+## Model Justification and Limitations
+We opted to use two separate models for Trump and Harris, as their supporter demographics differ significantly in ways that are challenging to capture within a single or even a small set of variables @pew_harris_trump_2024.
+The polling dataset we obtained primarily reflects the characteristics and trends within individual polls rather than offering direct predictors of the presidential election outcome.
+To maximize the model’s reliability, we refined the dataset by selecting polls with high credibility rather than attempting to predict outcomes directly from the available data.
 
-The model for Kamala Harris was fitted as follows:
-
-- **Response Variable**: `pct` (polling percentage)
-- **Predictor**: `end_date` (date of the poll)
-- **Family**: Gaussian (normal distribution)
-- **Priors**:
-  - Normal prior for the slope: \( \text{Normal}(0, 0.1) \)
-  - Normal prior for the intercept: \( \text{Normal}(50, 5) \)
-  - Exponential prior for auxiliary parameters: \( \text{Exponential}(1) \)
-
-### Model for Donald Trump
-The model for Donald Trump was fitted using the same specifications as for Kamala Harris
+Our model uses a simple linear approach with time as the sole predictor variable.
+While this may introduce potential bias by making a simplifying assumption about the trend over time, it also has the advantage of being straightforward to fit and interpret.
+This trade-off allows us to capture general trends while minimizing the risk of overfitting to limited or potentially inconsistent data sources.
 
@@
 #| tbl-cap: "Summary statistics of the Bayesian model for Harris"
+#| warning: false
 # Load required packages
@@
 # Display the summary table
-kable(model_summary, caption = "Summary of Bayesian Model for Harris", digits = 2)
+kable(model_summary, digits = 3)
 
@@
 #| echo: false
-#| tbl-cap: "Summary statistics of the Bayesian model for Harris"
+#| tbl-cap: "Summary statistics of the Bayesian model for Trump"
 # Load required packages
@@
 # Display the summary table
-kable(model_summary, caption = "Summary of Bayesian Model for Harris", digits = 2)
+kable(model_summary, digits = 3)
 ```
 ## Overall trend
-The intercepts for both models are significantly negative, indicating a downward trend in polling percentages over time for both candidates.
-The positive coefficients for end_date suggest a slight increase in polling percentages as the election date approaches.
-However, the magnitude of these increases is relatively small.
+The intercept of both model is positive, with Harris at 40.47 and Trump slightly lower than Harris at 40.36.
+The trend of support rate for both candidates with respect of time is positive but Harris's increment (0.10) is larger than Trump's (0.06).
 
 ## Model Diagnostics
-The posterior predictive checks (Figures 1 and 2) and the credible intervals for predictors (Figures 3 and 4) indicate that the models fit the data well.
-The diagnostics for both Harris and Trump models show that the models are stable and the predictions are reliable.Further diagnostics (Figures 5 to 8) confirm the robustness of the models.
-The R-hat values are all below 1.1, indicating good convergence of the MCMC chains.
+The model diagnostics indicate a good fit and reliable predictions for both candidates detailed plots is displayed in Appendix @sec-model-details.
+The posterior predictive checks (Figures @fig-pos_check_H and @fig-pos_check_T) show that the predicted values align well with the observed data, suggesting the models capture key patterns accurately.
+Additionally, the credible intervals for predictors (Figures @fig-CI_H and @fig-CI_T) provide further evidence of a well-calibrated model, as they encompass observed support rates, indicating that the uncertainty is well-accounted for in the predictions.
+
+Further robustness checks (@fig-DI_H and @fig-DI_T) support the stability of these models.
+Specifically, the R-hat values are below 1.1 for all parameters, demonstrating good convergence of the MCMC chains and suggesting that the parameter estimates are reliable.
+This convergence implies that our sampling process reached a stable distribution, which strengthens the credibility of our inferences.
+
+Overall, these diagnostics confirm that the models are not only stable but also appropriately reflect the trends in the data, supporting their reliability for further analysis.
 
@@
 
-# Appendix {-}
+# Appendix
 
@@
 
-## Model check
+## Model check {#sec-model-details}
 
@@
 ```
+
+\newpage
 ## pollster methodology analysis

### 2024-11-02 paper/paper.qmd
update survey questions

--- 2024-10-30 paper/paper.qmd
@@
 
-As indicated in the @sec-data, we do not have a confident belief of priors of parameters of our model since Harris replace the role of Biden by indecent and we do not have enough data about her previous polling outcomes as a president candidate compared to Biden.
-Therefore, we would not use the bayesian model but instead using a more simple multilinear regression model which predicts the support rate of Trump and Harris using time towards elections date and pollsters as they are beliveved to highly corelated to the response variable in @sec-data.
-We would bulid 8 models whihc represent national trend and that of seven swing states.
+The goal of our modelling strategy is twofold. Firstly,...
+
+Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix @sec-model-details].
+
 ## Model set-up
 
-Define $y_{i,\text{stateA}}$ as the percentage of support for a candicate in a poll in the specific state.
+Define $y_i$ as the percentage of support for a candicate in a poll.
 
-$$
-y_{i,national} = \beta_{0,national} + \beta_{1,time,national} + \beta_{2,pollster,national} + \epsilon_{national}
-$$ 
-$$
-y_{i,\text{AZ}} = \beta_{0,\text{AZ}} + \beta_{1,\text{time,AZ}} + \beta_{2,\text{pollster,AZ}} + \epsilon_{\text{AZ}}
-$$
-$$
-y_{i,\text{FL}} = \beta_{0,\text{FL}} + \beta_{1,\text{time,FL}} + \beta_{2,\text{pollster,FL}} + \epsilon_{\text{FL}}
-$$
-$$
-y_{i,\text{GA}} = \beta_{0,\text{GA}} + \beta_{1,\text{time,GA}} + \beta_{2,\text{pollster,GA}} + \epsilon_{\text{GA}}
-$$
-$$
-y_{i,\text{MI}} = \beta_{0,\text{MI}} + \beta_{1,\text{time,MI}} + \beta_{2,\text{pollster,MI}} + \epsilon_{\text{MI}}
-$$
-$$
-y_{i,\text{NC}} = \beta_{0,\text{NC}} + \beta_{1,\text{time,NC}} + \beta_{2,\text{pollster,NC}} + \epsilon_{\text{NC}}
-$$
-$$
-y_{i,\text{PA}} = \beta_{0,\text{PA}} + \beta_{1,\text{time,PA}} + \beta_{2,\text{pollster,PA}} + \epsilon_{\text{PA}}
-$$
-$$
-y_{i,\text{WI}} = \beta_{0,\text{WI}} + \beta_{1,\text{time,WI}} + \beta_{2,\text{pollster,WI}} + \epsilon_{\text{WI}}
-$$
+${y_i} = \beta_0 + \beta_1 pollscore + \beta_2 methodology + \beta_3transparency score + \beta_4 hypothetical + \epsilon$
 
-time:The date when the pollster starts polling
+pollscore: A numeric value representing the score or reliability of the pollster in question (e.g., -1.1).
+"The error and bias we can attribute to a pollster.
+Negative numbers are better.
+Stands for "Predictive Optimization of Latent skill Level in Surveys, Considering Overall Record, Empirically.
 
-pollster: The pollster which run the polling
+methodology: The method used to conduct the poll (e.g., Online Panel).
+
+transparency_score: A score reflecting the pollster’s transparency about their methodology (e.g., 9.0).
+"A grade for how transparent a pollster is, calculated based on how much information it discloses about its polls and weighted by
+by recency. The highest Transparency Score is 10."
+
+hypothetical: Indicates whether the poll is about a hypothetical match-up.
 
@@
 
-# Results {#set-result}
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+# Library setup
+#install.packages("glm")
+library(glmnet)
+library(ggplot2)
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-model1
+#| tbl-cap: "model 1"
+#| warning: false
+# Get the analysis data
+analysis_data <- read_csv(here::here("data/02-analysis_data/analysis_data.csv"))
+
+# Model
+glm_model1 <- glm(pct ~ pollscore + methodology + transparency_score + hypothetical, data = analysis_data, family = gaussian())
+
+# Get the summary
+model_summary <- summary(glm_model1)
+
+# Convert the summary coefficients to a data frame
+summary_table <- as.data.frame(model_summary$coefficients)
+
+# Create a table with a caption and control the size
+kable(summary_table, caption = "Model 1 Summary Table", size = "small")
+```
+
+## Lasso Regularization {#sec-model-lassoregularization}
+Lasso Regularization is performed to see if the selected variable in our model would best predict the results of polls percentage of a candidate.
+If the variable do not align with our actual numbers, we should consider other variables or a interaction variable.
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-lassoregularization
+#| tbl-cap: "lasso regularization"
+#| warning: false
+# Prepare the predictors
+x <- model.matrix(pct ~ pollscore + methodology + transparency_score + hypothetical, data = analysis_data)[, -1]
+
+# Prepare the response variable
+y <- analysis_data$pct
+
+# Lasso regression (alpha = 1)
+lasso_model <- glmnet(x, y, alpha = 1, family = "gaussian")
+
+# Cross-validation for Lasso
+cv_lasso <- cv.glmnet(x, y, alpha = 1, family = "gaussian")
+
+plot(cv_lasso)
+
+# Get the best lambda
+best_lambda_lasso <- cv_lasso$lambda.min
+
+# Refit Lasso with the optimal lambda
+lasso_final_model <- glmnet(x, y, alpha = 1, lambda = best_lambda_lasso, family = "gaussian")
+
+# View Lasso coefficients
+coef.lasso <- coef(lasso_final_model)
+
+# Predict using Lasso
+pred_lasso <- predict(lasso_final_model, s = best_lambda_lasso, newx = x)
+
+```
+
+Computing the Mean Square Error would give us a sense of our lasso model prediction. 
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-msr
+#| tbl-cap: "Mean Squared Error"
+#| warning: false
+# Calculate Mean Squared Error
+mse <- mean((y - pred_lasso)^2)
+cat("Mean Squared Error:", mse, "\n")
+
+# Calculate R-squared
+sst <- sum((y - mean(y))^2)
+sse <- sum((y - pred_lasso)^2)
+r_squared <- 1 - (sse / sst)
+cat("R-squared:", r_squared, "\n")
+
+# Create a data frame for plotting
+results_df <- data.frame(
+  Metric = c("SST", "SSE", "R-squared"),
+  Value = c(sst, sse, r_squared)
+)
+
+kable(results_df, caption = "Mean Squared Error Table")
+```
+
+From this graph, we can see that the cluster points do not perfectly align on the line of the best fit so we should still investigate in other variables. 
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-actualvspredicted
+#| tbl-cap: "actual vs predicted"
+#| warning: false
+
+# Compare predicted vs actual values
+comparison <- data.frame(Actual = y, Predicted = as.vector(pred_lasso))
+
+kable(head(comparison), caption = "Actual vs Predicted")
+
+# Plot actual vs predicted values
+ggplot(comparison, aes(x = Actual, y = Predicted)) +
+  geom_point() +
+  geom_abline(slope = 1, intercept = 0, color = "red") +
+  labs(title = "Actual vs Predicted Values",
+       x = "Actual Values",
+       y = "Predicted Values") +
+  theme_minimal()
+```
 
@@
 
-7.
+7. sex
 
-8.
+8. race
 
-9.
+9. education
 

### 2024-11-02 scripts/00-simulate_data.py
update survey questions

@@ -1,48 +1,0 @@
-#### Preamble ####
-# Purpose: Simulates a dataset of Australian electoral divisions, including the 
-  # state and party that won each division.
-# Author: Rohan Alexander
-# Date: 26 September 2024
-# Contact: rohan.alexander@utoronto.ca
-# License: MIT
-# Pre-requisites: 
-  # - `polars` must be installed (pip install polars)
-  # - `numpy` must be installed (pip install numpy)
-
-
-#### Workspace setup ####
-import polars as pl
-import numpy as np
-np.random.seed(853)
-
-
-#### Simulate data ####
-# State names
-states = [
-    "New South Wales", "Victoria", "Queensland", "South Australia", 
-    "Western Australia", "Tasmania", "Northern Territory", 
-    "Australian Capital Territory"
-]
-
-# Political parties
-parties = ["Labor", "Liberal", "Greens", "National", "Other"]
-
-# Probabilities for state and party distribution
-state_probs = [0.25, 0.25, 0.15, 0.1, 0.1, 0.1, 0.025, 0.025]
-party_probs = [0.40, 0.40, 0.05, 0.1, 0.05]
-
-# Generate the data using numpy and polars
-divisions = [f"Division {i}" for i in range(1, 152)]
-states_sampled = np.random.choice(states, size=151, replace=True, p=state_probs)
-parties_sampled = np.random.choice(parties, size=151, replace=True, p=party_probs)
-
-# Create a polars DataFrame
-analysis_data = pl.DataFrame({
-    "division": divisions,
-    "state": states_sampled,
-    "party": parties_sampled
-})
-
-
-#### Save data ####
-analysis_data.write_csv("data/00-simulated_data/simulated_data.csv")

### 2024-11-02 scripts/06-model_data.R

@@ -7,100 +7,94 @@
 # Pre-requisites: [...UPDATE THIS...]
 # Any other information needed? [...UPDATE THIS...]
 
-#### Set-up ####
-# Load libraries
-library(tidyverse)
-library(janitor)
+
+library(readr)
+library(dplyr)
 library(lubridate)
-library(broom)
-library(modelsummary)
 library(rstanarm)
-library(splines)
+library(janitor) # Load the janitor package for clean_names()
 
-
-#### Prepare dataset ####
-# Read in the data and clean variable names
+# Load and clean data
 data <- read_csv("data/01-raw_data/poll_raw_data.csv") |>
   clean_names()
 
 # Filter data to Harris estimates based on high-quality polls after she declared
+# Load required packages
+library(readr)
+library(dplyr)
+library(lubridate)
+library(rstanarm)
+library(janitor)
+
+# Filter and prepare data for Harris model
 just_harris_high_quality <- data |>
   filter(
     candidate_name == "Kamala Harris",
-    numeric_grade >= 2.8,
-    transparency_score >= 5,
-    pollscore <= 0 # Need to investigate this choice - come back and fix. 
-    # Also need to look at whether the pollster has multiple polls or just one or two - filter out later
+    numeric_grade >= 2.0,
+    transparency_score >= 4,
+    pollscore <= 0
   ) |>
   mutate(
-    state = if_else(is.na(state), "National", state), # Hacky fix for national polls - come back and check
+    state = if_else(is.na(state), "National", state),
     end_date = mdy(end_date)
   ) |>
-  filter(end_date >= as.Date("2024-07-21")) |> # When Harris declared
+  filter(state == "National") |>  # Remove the filter on end_date here
+  drop_na(pct, end_date) |>  # Drop rows with NA in pct or end_date
   mutate(
-    num_harris = round((pct / 100) * sample_size, 0) # Need number not percent for some models
+    # Get the earliest and latest end dates
+    earliest_date = min(end_date),
+    # Create a new column for days after the earliest date
+    days_after_earliest = as.numeric(end_date - earliest_date),
   )
 
-
-#### Plot data ####
-base_plot <- ggplot(just_harris_high_quality, aes(x = end_date, y = pct)) +
-  theme_classic() +
-  labs(y = "Harris percent", x = "Date")
-
-# Plots poll estimates and overall smoothing
-base_plot +
-  geom_point() +
[…]

### 2024-11-02 scripts/06-model_data.R
update prediction graph

@@ -70,6 +70,7 @@ just_trump_high_quality <- data |>
 
 write_csv(just_harris_high_quality, "data/02-analysis_data/Harris.csv")
 write_csv(just_trump_high_quality, "data/02-analysis_data/Trump.csv")
+
 # Fit the model for Harris
 harris_model <- stan_glm(
   formula = pct ~ days_after_earliest,
@@ -91,6 +92,80 @@ trump_model <- stan_glm(
   prior_aux = exponential(rate = 1),
   seed = 853
 )
+
+# Define the target end date
+last_date <- as.Date("2024-11-05")
+
+# Calculate the number of days from earliest_date to last_date
+days_to_last_date <- as.numeric(last_date - earliest_date)
+
+# Generate prediction data frame with days_after_earliest up to days_to_last_date
+days_range <- data.frame(days_after_earliest = seq(0, days_to_last_date, by = 1))
+
+# Convert days_after_earliest to actual dates by adding them to earliest_date
+days_range$end_date <- earliest_date + days_range$days_after_earliest
+
+# Harris predictions
+harris_predictions <- predict(harris_model, newdata = days_range, se.fit = TRUE)
+days_range$harris_pct <- harris_predictions$fit
+days_range$harris_se <- harris_predictions$se.fit
+
+# Trump predictions
+trump_predictions <- predict(trump_model, newdata = days_range, se.fit = TRUE)
+days_range$trump_pct <- trump_predictions$fit
+days_range$trump_se <- trump_predictions$se.fit
+
+
+
+# Add actual poll data to plot
+just_harris_high_quality$Candidate <- "Harris"
+just_trump_high_quality$Candidate <- "Trump"
+poll_data <- rbind(just_harris_high_quality, just_trump_high_quality)
+
+# Plot
+
+library(ggplot2)
+
+ggplot(data = days_range, aes(x = end_date)) +
+  # Harris line and confidence interval
+  geom_line(aes(y = harris_pct, color = "Harris"), size = 1) +
+  geom_ribbon(aes(ymin = harris_pct - 1.96 * harris_se, ymax = harris_pct + 1.96 * harris_se, fill = "Harris"), alpha = 0.2) +
+  
+  # Trump line and confidence interval
+  geom_line(aes(y = trump_pct, color = "Trump"), size = 1) +
[…]

### 2024-11-02 scripts/06-model_data.R
update prediction graph

@@ -7,12 +7,13 @@
 # Pre-requisites: [...UPDATE THIS...]
 # Any other information needed? [...UPDATE THIS...]
 
-
+library(tidyr)
 library(readr)
 library(dplyr)
 library(lubridate)
 library(rstanarm)
 library(janitor) # Load the janitor package for clean_names()
+library(ggplot2)
 
 # Load and clean data
 data <- read_csv("data/01-raw_data/poll_raw_data.csv") |>
@@ -20,11 +21,14 @@ data <- read_csv("data/01-raw_data/poll_raw_data.csv") |>
 
 # Filter data to Harris estimates based on high-quality polls after she declared
 # Load required packages
+library(tidyr)
 library(readr)
 library(dplyr)
 library(lubridate)
 library(rstanarm)
 library(janitor)
+library(ggplot2)
+
 
 # Filter and prepare data for Harris model
 just_harris_high_quality <- data |>
@@ -41,13 +45,14 @@ just_harris_high_quality <- data |>
   filter(state == "National") |>  # Remove the filter on end_date here
   drop_na(pct, end_date) |>  # Drop rows with NA in pct or end_date
   mutate(
-    # Get the earliest and latest end dates
-    earliest_date = min(end_date),
     # Create a new column for days after the earliest date
-    days_after_earliest = as.numeric(end_date - earliest_date),
+    days_after_earliest = as.numeric(end_date - min(end_date))
   )
 
-# Filter and prepare data for Trump model
+# Define `earliest_date` globally based on `just_harris_high_quality`
+earliest_date <- min(just_harris_high_quality$end_date)
+
+# Repeat the process for Trump data
 just_trump_high_quality <- data |>
   filter(
     candidate_name == "Donald Trump",
@@ -62,12 +67,11 @@ just_trump_high_quality <- data |>
   filter(state == "National") |>  # Remove the filter on end_date here
   drop_na(pct, end_date) |>  # Drop rows with NA in pct or end_date
   mutate(
-    # Get the earliest and latest end dates
-    earliest_date = min(end_date),
     # Create a new column for days after the earliest date
-    days_after_earliest = as.numeric(end_date - earliest_date),
+    days_after_earliest = as.numeric(end_date - earliest_date)
   )
 
+# Write data to CSV
[…]

### 2024-11-02 scripts/06-model_data.R
model update

@@ -68,7 +68,7 @@ just_trump_high_quality <- data |>
     days_after_earliest = as.numeric(end_date - earliest_date),
   )
 
-write_csv(just_harris_high_quality, "data/02-analysis_data/Harris.csv")
+write_csv(just_harris_high_quality, "data/02-analysis_data/Harris.csv")ci
 write_csv(just_trump_high_quality, "data/02-analysis_data/Trump.csv")
 # Fit the model for Harris
 harris_model <- stan_glm(

### 2024-11-02 scripts/03-clean_data.R
Update clean data and model

@@ -8,13 +8,56 @@
 # Any other information needed? None
 
 #### Workspace setup ####
-library(tidyverse)
+library(readr)
+library(dplyr)
+library(lubridate)
+library(rstanarm)
+library(janitor)
+library(tidyr)
 
-#### Clean data ####
-raw_data <- read_csv("data/01-raw_data/raw_data.csv")
+data <- read_csv("data/01-raw_data/poll_raw_data.csv") |>
+  clean_names()
+# Filter and prepare data for Harris model
+just_harris_high_quality <- data |>
+  filter(
+    candidate_name == "Kamala Harris",
+    numeric_grade >= 2.0,
+    transparency_score >= 4,
+    pollscore <= 0
+  ) |>
+  mutate(
+    state = if_else(is.na(state), "National", state),
+    end_date = mdy(end_date)
+  ) |>
+  filter(state == "National") |>  # Remove the filter on end_date here
+  drop_na(pct, end_date) |>  # Drop rows with NA in pct or end_date
+  mutate(
+    # Get the earliest and latest end dates
+    earliest_date = min(end_date),
+    # Create a new column for days after the earliest date
+    days_after_earliest = as.numeric(end_date - earliest_date),
+  )
 
-# Remove rows with any missing values in the relevant columns
-data_clean <- na.omit(data[, c("pct", "pollscore", "methodology", "transparency_score", "hypothetical")])
+# Filter and prepare data for Trump model
+just_trump_high_quality <- data |>
+  filter(
+    candidate_name == "Donald Trump",
+    numeric_grade >= 2.0,
+    transparency_score >= 4,
+    pollscore <= 0
+  ) |>
+  mutate(
+    state = if_else(is.na(state), "National", state),
+    end_date = mdy(end_date)
+  ) |>
+  filter(state == "National") |>  # Remove the filter on end_date here
+  drop_na(pct, end_date) |>  # Drop rows with NA in pct or end_date
+  mutate(
+    # Get the earliest and latest end dates
+    earliest_date = min(end_date),
+    # Create a new column for days after the earliest date
+    days_after_earliest = as.numeric(end_date - earliest_date),
+  )
 
-#### Save data ####
-write_csv(data_clean, "data/02-analysis_data/analysis_data.csv")
+write_csv(just_harris_high_quality, "data/02-analysis_data/Harris.csv")
+write_csv(just_trump_high_quality, "data/02-analysis_data/Trump.csv")

### 2024-11-02 scripts/06-model_data.R
fix error

@@ -71,12 +71,8 @@ just_trump_high_quality <- data |>
     days_after_earliest = as.numeric(end_date - earliest_date)
   )
 
-<<<<<<< HEAD
 # Write data to CSV
 write_csv(just_harris_high_quality, "data/02-analysis_data/Harris.csv")
-=======
-write_csv(just_harris_high_quality, "data/02-analysis_data/Harris.csv")ci
->>>>>>> b9975540bc2e8dfa19104ee1acd5935561b0b98f
 write_csv(just_trump_high_quality, "data/02-analysis_data/Trump.csv")
 
 # Fit the model for Harris

### 2024-11-02 scripts/06-model_data.R
comment out state code

@@ -169,90 +169,90 @@ ggplot(data = days_range, aes(x = end_date)) +
 
 ### PLot the swing states Trend ###
 
-# Define a list of swing states
-swing_states <- c("Arizona", "Georgia", "Michigan", "Nevada", "North Carolina", "Pennsylvania", "Wisconsin")
-
-# Filter for swing states from the original dataset and assign them to Harris and Trump datasets
-state_poll_data <- data %>%
-  filter(state %in% swing_states) %>%
-  mutate(
-    end_date = mdy(end_date),
-    Candidate = if_else(candidate_name == "Kamala Harris", "Harris", "Trump")
-  )
-
-# Generate the prediction data frame with dates up to November 5, 2024
-days_range <- data.frame(days_after_earliest = seq(0, days_to_last_date, by = 1))
-days_range$end_date <- earliest_date + days_range$days_after_earliest
-
-# Harris predictions
-harris_predictions <- predict(harris_model, newdata = days_range, se.fit = TRUE)
-days_range$harris_pct <- harris_predictions$fit
-days_range$harris_se <- harris_predictions$se.fit
-
-# Trump predictions
-trump_predictions <- predict(trump_model, newdata = days_range, se.fit = TRUE)
-days_range$trump_pct <- trump_predictions$fit
-days_range$trump_se <- trump_predictions$se.fit
-
-# Repeat the prediction data for each state
-state_days_range <- days_range %>%
-  tidyr::crossing(state = swing_states)
-
-# Prepare for faceted plot by merging actual poll data with the predictions
-state_poll_data <- state_poll_data %>%
-  filter(Candidate %in% c("Trump", "Harris"))  # Filter for only Trump and Harris
-
-# Combine actual poll data with predictions for plotting
-plot_data <- state_days_range %>%
-  left_join(state_poll_data, by = c("end_date", "state"))
-
-# Plot with faceting by state
-ggplot(data = plot_data, aes(x = end_date)) +
-  # Harris line and confidence interval
-  geom_line(aes(y = harris_pct, color = "Harris"), linewidth = 1) +
-  geom_ribbon(aes(ymin = harris_pct - 1.96 * harris_se, ymax = harris_pct + 1.96 * harris_se, fill = "Harris"), alpha = 0.2) +
-  
-  # Trump line and confidence interval
-  geom_line(aes(y = trump_pct, color = "Trump"), linewidth = 1) +
[…]

### 2024-11-02 paper/paper.qmd
update survey questions
update prediction graph
add prior predictive check

--- 2024-10-31 paper/paper.qmd
@@
 ## Model Diagnostics
-The model diagnostics indicate a good fit and reliable predictions for both candidates detailed plots is displayed in Appendix @sec-model-details.
-The posterior predictive checks (Figures @fig-pos_check_H and @fig-pos_check_T) show that the predicted values align well with the observed data, suggesting the models capture key patterns accurately.
-Additionally, the credible intervals for predictors (Figures @fig-CI_H and @fig-CI_T) provide further evidence of a well-calibrated model, as they encompass observed support rates, indicating that the uncertainty is well-accounted for in the predictions.
+The model diagnostics indicate a strong fit and reliable predictive performance for both candidates, as detailed in the plots in Appendix @sec-model-details.
+Posterior predictive checks (Figures @fig-pos_check_H and @fig-pos_check_T) show that the predicted values align well with the observed data, suggesting that the models effectively capture key data patterns.
+Furthermore, the comparison of posterior and prior variables indicates minimal deviation, affirming an appropriate choice of priors.
+The credible intervals for the predictors (Figures @fig-CI_H and @fig-CI_T) encompass the observed support rates, demonstrating that the models adequately account for uncertainty in their predictions.
 
-Further robustness checks (@fig-DI_H and @fig-DI_T) support the stability of these models.
-Specifically, the R-hat values are below 1.1 for all parameters, demonstrating good convergence of the MCMC chains and suggesting that the parameter estimates are reliable.
-This convergence implies that our sampling process reached a stable distribution, which strengthens the credibility of our inferences.
+Additional robustness checks (Figures @fig-DI_H and @fig-DI_T) confirm the stability of these models.
+The R-hat values for all parameters are below 1.1, indicating good convergence of the MCMC chains and reliable parameter estimates.
+This convergence supports the conclusion that the sampling process reached a stable distribution, reinforcing the credibility of our inferences.
 
-Overall, these diagnostics confirm that the models are not only stable but also appropriately reflect the trends in the data, supporting their reliability for further analysis.
+In conclusion, these diagnostics verify that the models are both stable and effective in reflecting trends in the data, confirming their reliability for further analysis.
 
@@
 
+just_harris_high_quality <- read.csv(here::here("data/02-analysis_data/Harris.csv"))
+
+posterior_vs_prior(model1) +
+  theme_minimal() +
+  scale_color_brewer(palette = "Set1") +
+  theme(legend.position = "bottom") +
+  coord_flip()
+
 ```
@@
     theme(legend.position = "bottom")
+
+just_trump_high_quality <- read.csv(here::here("data/02-analysis_data/Trump.csv"))
+
+posterior_vs_prior(model2) +
+  theme_minimal() +
+  scale_color_brewer(palette = "Set1") +
+  theme(legend.position = "bottom") +
+  coord_flip()
 

### 2024-11-02 scripts/06-model_data.R
no change

@@ -1,22 +1,11 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Models the relatioship between the supprt rate of Harris and Trump with respect to time
+# Author: Yun Chu, Felix Li, and Wen Han Zhao 
+# Date: 22 October 2024
+# Contact: youna.zhao@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
-
-
-library(readr)
-library(dplyr)
-library(lubridate)
-library(rstanarm)
-library(janitor) # Load the janitor package for clean_names()
-
-# Load and clean data
-data <- read_csv("data/01-raw_data/poll_raw_data.csv") |>
-  clean_names()
+# Pre-requisites: raw data has been downloaded from the website
+# Any other information needed? None
 
 # Filter data to Harris estimates based on high-quality polls after she declared
 # Load required packages
@@ -26,50 +15,9 @@ library(lubridate)
 library(rstanarm)
 library(janitor)
 
-# Filter and prepare data for Harris model
-just_harris_high_quality <- data |>
-  filter(
-    candidate_name == "Kamala Harris",
-    numeric_grade >= 2.0,
-    transparency_score >= 4,
-    pollscore <= 0
-  ) |>
-  mutate(
-    state = if_else(is.na(state), "National", state),
-    end_date = mdy(end_date)
-  ) |>
-  filter(state == "National") |>  # Remove the filter on end_date here
-  drop_na(pct, end_date) |>  # Drop rows with NA in pct or end_date
-  mutate(
-    # Get the earliest and latest end dates
-    earliest_date = min(end_date),
-    # Create a new column for days after the earliest date
-    days_after_earliest = as.numeric(end_date - earliest_date),
-  )
+just_harris_high_quality <- read.csv("data/02-analysis_data/Harris.csv")
+just_trump_high_quality <- read.csv("data/02-analysis_data/Trump.csv")
 
-# Filter and prepare data for Trump model
-just_trump_high_quality <- data |>
-  filter(
-    candidate_name == "Donald Trump",
-    numeric_grade >= 2.0,
-    transparency_score >= 4,
[…]

### 2024-11-03 scripts/06-model_data.R
update model

@@ -4,10 +4,8 @@
 # Date: 22 October 2024
 # Contact: youna.zhao@mail.utoronto.ca
 # License: MIT
-<<<<<<< HEAD
 # Pre-requisites: raw data has been downloaded from the website
 # Any other information needed? None
-=======
 # Pre-requisites: [...UPDATE THIS...]
 # Any other information needed? [...UPDATE THIS...]
 
@@ -22,7 +20,6 @@ library(ggplot2)
 # Load and clean data
 data <- read_csv("data/01-raw_data/poll_raw_data.csv") |>
   clean_names()
->>>>>>> 76f9252b4de65f603fe61dd23d9c68092b8ed6e3
 
 # Filter data to Harris estimates based on high-quality polls after she declared
 # Load required packages
@@ -34,12 +31,9 @@ library(rstanarm)
 library(janitor)
 library(ggplot2)
 
-
-<<<<<<< HEAD
 just_harris_high_quality <- read.csv("data/02-analysis_data/Harris.csv")
 just_trump_high_quality <- read.csv("data/02-analysis_data/Trump.csv")
 
-=======
 # Filter and prepare data for Harris model
 just_harris_high_quality <- data |>
   filter(
@@ -85,7 +79,6 @@ just_trump_high_quality <- data |>
 write_csv(just_harris_high_quality, "data/02-analysis_data/Harris.csv")
 write_csv(just_trump_high_quality, "data/02-analysis_data/Trump.csv")
 
->>>>>>> 76f9252b4de65f603fe61dd23d9c68092b8ed6e3
 # Fit the model for Harris
 harris_model <- stan_glm(
   formula = pct ~ days_after_earliest,

### 2024-11-03 scripts/05-exploratory_data_analysis.R
Update Explorary Data Analysis

@@ -1,24 +1,101 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Explore the dataset 
+# Author: Yun Chu, Felix Li, and Wen Han Zhao 
+# Date: 4 November 2024
+# Contact: youna.zhao@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: raw data used from data folder
+# Any other information needed? None
 
 
 #### Workspace setup ####
 library(tidyverse)
-library(rstanarm)
+library(janitor)
+library(lubridate)
+library(dplyr)
+library(ggplot2)
 
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+analysis_data <- read_csv("data/02-analysis_data/analysis_data.csv")
+data <- read.csv("data/01-raw_data/raw_data.csv")
+
+hist(data$pct)
+
+### Filter through the data we want 
+trump <- data |> filter(
+  candidate_name == "Donald Trump",
+  numeric_grade >= 2.7,
+  pollscore <= 0,
+  transparency_score >= 5.5
+)|>
+  mutate(
+    state = if_else(is.na(state), "National", state), 
+    end_date = mdy(end_date)
+  ) |>
+  mutate(
+    num_trump = round((pct / 100) * sample_size, 0) # Need number not percent for some models
+  )
+
+### Only keeping the rows that we want to investigate
+trump <- trump |> select(sample_size, pollster, numeric_grade, pollscore, methodology, transparency_score, state, start_date, 
+                         end_date, population, party, answer, candidate_name, pct)
+### Summary of the selected dataset
+summary(trump)
+
+### Histogram ### 
+hist(trump$pct, main = "Histogram of Poll", xlab = "The pencentage of the vote for Trump in the poll", ylab = "Frequency")
+
+### Summary statistics for pct
+mean_pct <- mean(data$pct)
+sd_pct <- sd(data$pct)
+min_pct <- min(data$pct)
+max_pct <- max(data$pct)
+
[…]

### 2024-11-03 scripts/06-model_data.R
Model change

@@ -6,8 +6,8 @@
 # License: MIT
 # Pre-requisites: raw data has been downloaded from the website
 # Any other information needed? None
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: None
+# Any other information needed? None
 
 library(tidyr)
 library(readr)

### 2024-11-03 scripts/03-clean_data.R
clean data change

@@ -8,13 +8,59 @@
 # Any other information needed? None
 
 #### Workspace setup ####
-library(tidyverse)
+library(readr)
+library(dplyr)
+library(lubridate)
+library(rstanarm)
+library(janitor)
+library(tidyr)
 
 #### Clean data ####
-raw_data <- read_csv("data/01-raw_data/raw_data.csv")
+data <- read_csv("data/01-raw_data/poll_raw_data.csv") |>
+  clean_names()
+# Filter and prepare data for Harris model
+just_harris_high_quality <- data |>
+  filter(
+    candidate_name == "Kamala Harris",
+    numeric_grade >= 2.0,
+    transparency_score >= 4,
+    pollscore <= 0
+  ) |>
+  mutate(
+    state = if_else(is.na(state), "National", state),
+    end_date = mdy(end_date)
+  ) |>
+  filter(state == "National") |>  # Remove the filter on end_date here
+  drop_na(pct, end_date) |>  # Drop rows with NA in pct or end_date
+  mutate(
+    # Get the earliest and latest end dates
+    earliest_date = min(end_date),
+    # Create a new column for days after the earliest date
+    days_after_earliest = as.numeric(end_date - earliest_date),
+  )
 
-# Remove rows with any missing values in the relevant columns
-data_clean <- na.omit(data[, c("pct", "pollscore", "methodology", "transparency_score", "hypothetical")])
 
-#### Save data ####
-write_csv(data_clean, "data/02-analysis_data/analysis_data.csv")
+# Filter and prepare data for Trump model
+just_trump_high_quality <- data |>
+  filter(
+    candidate_name == "Donald Trump",
+    numeric_grade >= 2.0,
+    transparency_score >= 4,
+    pollscore <= 0
+  ) |>
+  mutate(
+    state = if_else(is.na(state), "National", state),
+    end_date = mdy(end_date)
+  ) |>
+  filter(state == "National") |>  # Remove the filter on end_date here
+  drop_na(pct, end_date) |>  # Drop rows with NA in pct or end_date
+  mutate(
+    # Get the earliest and latest end dates
+    earliest_date = min(end_date),
+    # Create a new column for days after the earliest date
+    days_after_earliest = as.numeric(end_date - earliest_date),
+  )
+
+
+write_csv(just_harris_high_quality, "data/02-analysis_data/Harris.csv")
+write_csv(just_trump_high_quality, "data/02-analysis_data/Trump.csv")

### 2024-11-03 paper/paper.qmd
update survey

@@
 
+cite R, packages, dataa
+The data used in this paper is source from FiveThirtyEight ("538presidentialpolls", 2024). T
 We use the statistical programming language R [@citeR].... Our data.... Following @tellingstories, we consider...
@@
 
-3. Who would you vote for?
+3. If the election were held today, who would you most likely vote for?
 
@@
 * Not Decided Yet
+* Prefer not to say
 
@@
 
-5. How likely are you to vote for the election?
+5. How many characters are present in the word "President"?
 
-* Very Likely
-* A Bit Likely
-* Not Likely
+* 6
+* 8
+* 10
+* None of the Above
 
@@
 
-7. sex
+7. What is your sex?
 
-8. race
+* Male
+* Female
+* Non-binary
+* Prefer not to say
 
-9. education
+8. What is your Ethnity/Race?
 
-10. How many characters are present in the word "President"?
+* White
+* Black or African American
+* Hispanic or Latino
+* Asian
+* Native American or Alaska Native
+* Native Hawaiian or Other Pacific Islander
+* Middle Eastern or North African
+* Prefer not to say
+* Other (please specify) [optional text box]
 
-* 6
-* 8
-* 10
-* None of the Above
+9. What is your Education Level?
+
+* Less than high school
+* High school diploma or equivalent
+* Some college, no degree
+* Associate’s degree
+* Bachelor’s degree
+* Master’s degree
+* Professional or doctoral degree (JD, MD, PhD, etc.)
+* Prefer not to say
+
+10.How do you plan to vote?
+
+* In-person on Election Day
+* In-person early voting
+* Mail-in/Absentee ballot
+* Undecided
+
+11. Is there anything else you’d like to share about your voting decision or concerns?
 

### 2024-11-03 scripts/03-clean_data.R
add parquet

@@ -14,6 +14,7 @@ library(lubridate)
 library(rstanarm)
 library(janitor)
 library(tidyr)
+library(arrow)
 
 #### Clean data ####
 data <- read_csv("data/01-raw_data/poll_raw_data.csv") |>
@@ -61,6 +62,5 @@ just_trump_high_quality <- data |>
     days_after_earliest = as.numeric(end_date - earliest_date),
   )
 
-
-write_csv(just_harris_high_quality, "data/02-analysis_data/Harris.csv")
-write_csv(just_trump_high_quality, "data/02-analysis_data/Trump.csv")
+write_parquet(just_harris_high_quality, "data/02-analysis_data/Harris.parquet")
+write_parquet(just_trump_high_quality, "data/02-analysis_data/Trump.parquet")

### 2024-11-03 scripts/04-test_analysis_data.R

@@ -15,7 +15,8 @@ library(tibble)
 library(readr)
 
 
-simulated_data <- read.csv('data/02-analysis_data/analysis_data.csv')
+simulated_data <- read_parquet('data/02-analysis_data/Harris.parquet')
+# For Trump: simulated_data <- read_parquet('data/02-analysis_data/Trump.parquet')
 
 # Define the test cases
 test_that("Simulated data has correct structure and values", {

### 2024-11-03 scripts/06-model_data.R

@@ -31,8 +31,8 @@ library(rstanarm)
 library(janitor)
 library(ggplot2)
 
-just_harris_high_quality <- read.csv("data/02-analysis_data/Harris.csv")
-just_trump_high_quality <- read.csv("data/02-analysis_data/Trump.csv")
+just_harris_high_quality <- read_parquet('data/02-analysis_data/Harris.parquet')
+just_trump_high_quality <- read_parquet('data/02-analysis_data/Trump.parquet')
 
 # Filter and prepare data for Harris model
 just_harris_high_quality <- data |>

### 2024-11-03 paper/paper.qmd

@@
 Cluster sampling is the sampling approach used.
-Specifically clusters are the swing states as they are the states that would affect the election result, i.e.
+Specifically clusters are states.
+In our case, the swing states as they are the states that would affect the election result, i.e.
 Arizona, Florida, Gegorgia, Iowa, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin.
-In each swing state/clusters, units are selected based on postal code.
+In each swing state, units are selected based on postal code.
 Using simple random sampling, 100 distinct postal codes are randomly selected.
-People whose living address has the postal codes selected are the target respondents.
+People whose living address have the postal codes selected are the target respondents.
 For the postal codes that lie in the non-residential area, the postal codes would be ignored.

### 2024-11-03 scripts/05-exploratory_data_analysis.R
Update for Harris

@@ -2,7 +2,6 @@
 # Purpose: Explore the dataset 
 # Author: Yun Chu, Felix Li, and Wen Han Zhao 
 # Date: 4 November 2024
-# Contact: youna.zhao@mail.utoronto.ca
 # License: MIT
 # Pre-requisites: raw data used from data folder
 # Any other information needed? None
@@ -15,15 +14,17 @@ library(lubridate)
 library(dplyr)
 library(ggplot2)
 
+
 #### Read data ####
-analysis_data <- read_csv("data/02-analysis_data/analysis_data.csv")
+analysis_data <- read_parquet("data/02-analysis_data/analysis_data.parquet")
 data <- read.csv("data/01-raw_data/raw_data.csv")
 
+### Histogram of Polls in Raw dataset
 hist(data$pct)
 
 ### Filter through the data we want 
-trump <- data |> filter(
-  candidate_name == "Donald Trump",
+harris <- data |> filter(
+  candidate_name == "Harris",
   numeric_grade >= 2.7,
   pollscore <= 0,
   transparency_score >= 5.5
@@ -33,17 +34,18 @@ trump <- data |> filter(
     end_date = mdy(end_date)
   ) |>
   mutate(
-    num_trump = round((pct / 100) * sample_size, 0) # Need number not percent for some models
+    num_harris = round((pct / 100) * sample_size, 0) # Need number not percent for some models
   )
 
 ### Only keeping the rows that we want to investigate
-trump <- trump |> select(sample_size, pollster, numeric_grade, pollscore, methodology, transparency_score, state, start_date, 
-                         end_date, population, party, answer, candidate_name, pct)
-### Summary of the selected dataset
-summary(trump)
+harris <- harris |> select(sample_size, pollster, numeric_grade, pollscore, methodology, transparency_score, state, start_date, 
+                           end_date, population, party, answer, candidate_name, pct)
 
-### Histogram ### 
-hist(trump$pct, main = "Histogram of Poll", xlab = "The pencentage of the vote for Trump in the poll", ylab = "Frequency")
+### Summary of the selected dataset
+summary(harris)
+
+### Histogram of Polls of Selected Data
+hist(harris$pct, main = "Histogram of Poll", xlab = "The pencentage of the vote for Harris in the poll", ylab = "Frequency")
 
 ### Summary statistics for pct
 mean_pct <- mean(data$pct)
[…]

### 2024-11-03 scripts/03-clean_data.R
clean data change

@@ -35,10 +35,11 @@ just_harris_high_quality <- data |>
   drop_na(pct, end_date) |>  # Drop rows with NA in pct or end_date
   mutate(
     # Get the earliest and latest end dates
-    earliest_date = min(end_date),
+    earliest_date = as.Date("July/21/2024", format = "%B/%d/%Y"),
     # Create a new column for days after the earliest date
     days_after_earliest = as.numeric(end_date - earliest_date),
-  )
+  ) |>
+  filter(days_after_earliest > 0)
 
 
 # Filter and prepare data for Trump model
@@ -57,10 +58,11 @@ just_trump_high_quality <- data |>
   drop_na(pct, end_date) |>  # Drop rows with NA in pct or end_date
   mutate(
     # Get the earliest and latest end dates
-    earliest_date = min(end_date),
+    earliest_date = as.Date("July/21/2024", format = "%B/%d/%Y"),
     # Create a new column for days after the earliest date
     days_after_earliest = as.numeric(end_date - earliest_date),
-  )
+  ) |>
+  filter(days_after_earliest > 0)
 
 write_parquet(just_harris_high_quality, "data/02-analysis_data/Harris.parquet")
 write_parquet(just_trump_high_quality, "data/02-analysis_data/Trump.parquet")

### 2025-03-12 paper/paper.qmd
addressed nathalie comments

--- 2024-11-02 paper/paper.qmd
@@
 ---
-title: "US President Prediction"
-subtitle: "Predict the results of 2024 US President Election"
+title: "Self-reported LLM usage and results on a data science project: Evidence from a Canadian undergraduate data science course"
 author: 
-  - Yun Chu
-  - Felix Li
-  - Wen Han Zhao 
-thanks: "Code and data are available at: [https://github.com/younazhao/US-President-Prediction/tree/main](https://github.com/younazhao/US-President-Prediction/tree/main)."
-date: 22 October 2024
+  - Rohan Alexander
+  - Luca Carnegie
+  - Nathalie Moon
+thanks: "Code and data are available at: https://github.com/lcarnegie/llms-achievement.
+We thank Tiffany Timbers and attendees at JSM 2024 for helpful suggestions.
+This research is currently under review by the University of Toronto's Research Ethics Board.
+Contributions: RA: Conceptualization, Data Curation, Formal Analysis, Funding Acquisition, Investigation, Methodology, Project Administration, Software, Visualization, Writing – Original Draft Preparation, Writing – Review & Editing; LC: Formal Analysis, Visualization, Writing – Original Draft Preparation, Writing – Review & Editing; NM: Conceptualization, Methodology, Writing – Original Draft Preparation, Writing – Review & Editing.
+Comments can be sent to: rohan.alexander@utoronto.ca."
+date: today
 date-format: long
-abstract: "In this study, we aim to study the 2024 U.S.
-Presidential Election with the methodology of  polls of polls across various states.
-With the insightful analyses on polls provided by Redfiled & Wilton Strategies, the conducted survey on the population highlights various key point for the prediction of US president.
-Using Generalized Linear Models (GLMs), this report analyzes the accuracy of prediction in president votes through several key variables from pollster.
-The model incorporates variables such as pollscore, methodology, transparency_score and hypothetical.
-The prediction finding highlights the significance of candicate trustworthiness among the voter decision, providing insights for swing voters.
-This analysis offers valuable information into the potential electoral outcomes driving the 2024 US president election.
-"
+abstract: "To help understand the effect of Large Language Models (LLMs) on data science practice we examine the extent to which LLM usage is correlated with the mark that a student gets on a final paper in a classroom data science setting.
+We find some very mild evidence from this observational study that LLM usage may be associated with better scores, but our main conclusion is that there is no clear relationship between more extensive LLM usage and the student's mark.
+Despite the classroom setting used for evaluation, the particular activity of interest is similar to the work done by professional data scientists.
+Our finding suggests the need for more extensive work evaluating how LLMs can be integrated into the data science workflow in a way that provides value."
 format: pdf
 number-sections: true
+table-of-contents: false
 bibliography: references.bib
@@
 
-```{r}
+```{r setup}
 #| include: false
-#| warning: false
-#| message: false
+
+library(arrow)
+library(here)
+library(knitr)
 library(tidyverse)
-library(knitr)
-library(kableExtra)
-library(readr)
-library(dplyr)
-library(rstanarm)
+library(scales)
+library(modelsummary)
 library(bayesplot)
-library(ggplot2)
-library(ggpubr)
+
+# Read in data
+all_data <-
+  read_parquet(here("data/analysis_data.parquet"))
 ```
@@
 
-Overview paragraph
+Trustworthy data science is the practice of conducting data analysis in a transparent, ethical, and reliable manner.
+These principles are upheld in various ways, through ongoing education, adherence to professional norms and, and implementation of reproducible workflows.
+To stay current, data scientists must continually update their knowledge of best practices for transparency and reproducibility, foster a culture which values different perspectives and accountability, and critically consider the broader impact of data-driven decisions.
+Recent advances, such as the wide release of user-friendly Large Language Models (LLMs), particularly OpenAI’s ChatGPT, have transformed the data science toolkit.
+These powerful tools for natural language processing and generation influence various aspects of data analysis and automation, further emphasizing the need for trustworthy practices.
 
-Estimand paragraph
+Like all new tools LLMs have created both excitement and apprehension.
+ChatGPT's public release on November 30, 2022, brought LLMs into the mainstream conversation.
+LLMs have quickly been put to use in both industry and academia.
+By now, many people, especially educators and students, have some experience with LLMs, in both personal and professional contexts.
 
-Results paragraph
+In the context of teaching statistics and data science, LLMs could be useful for many tasks.
+For instance, there is considerable interest in the potential of chatbots to act as personalized tutors for students [@fulgencio2024developing; @afzal2024tailoring].
+ChatGPT has also been the catalyst for many interesting and important conversations around academic integrity [@Eke2023], the development of critical thinking skills [@thiga2024generative], and what effective learning looks like [@BaidooAnu_2023].
 
-Why it matters paragraph
+In this paper we are interested in better understanding LLMs as a tool for producing trustworthy data science.
+We study how they were used by students in an upper-year undergraduate data science course and whether students who used LLMs tended to have higher scores than those who did not.
 
-Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....
+The tasks involved in a trustworthy data science workflow can be generally broken down into a number of key competencies [@Adhikari2021Interleaving; @Gibbs2021Building].
+One is programming, which is done when cleaning, analyzing, and visualizing data often using programming languages like R or Python.
+Another is writing, which is primarily done when communicating results.
+The potential for LLMs to positively affect students' academic performance in data science can be clearly inferred from their already-demonstrated capability in those competencies in adjacent professional fields.
+
+In terms of computer programming, @Peng_2023 found a positive impact of GitHub Copilot (an LLM-powered programming assistant) on productivity.
+Specifically, in an experiment involving 95 freelance programmers, they found that a treatment group of programmers with access to GitHub Copilot completed a standardized programming task 56% faster than the control group.
+Programmers with less experience saw the greatest improvements in productivity.
+
+@DellAcqua_2023, focusing on management consulting tasks, provide evidence that LLMs can improve writing productivity.
+In a field experiment involving consultants from Boston Consulting Group they found that the use of OpenAI's GPT-4 led to a 25% increase in delivery speed of business tasks, most of which involved some writing, as well as a 40% increase in human-rated performance on those tasks.
+Similar to computer programming, these productivity increases were most pronounced for those with below average performance, with their output increasing by 43%.
+
+On the other hand, @Valenzuela_2024 argue that LLMs lead to a loss of serendipity which may lead to less original work, and potentially de-skilling primarily with respect to programming ability, among other consequences.
+These outcomes could negatively affect students' effective learning of data science.
+
+@Ellis_Slade_2023 take a more optimistic perspective and argue that LLMs are just another technology that will impact statistics and data science education.
+Similarly, @Tu_2024 acknowledge that LLMs can streamline many parts of a data science workflow.
+With that in mind, they suggest that data-scientists-in-training should shift their self-perspectives from primarily being an "analyst" to primarily being a "product manager" responsible for strategic oversight of the analysis carried out by LLMs.
+
+At the high school level @Lazar_2023 conducted an informal survey of secondary school teachers and students on their opinions of ChatGPT, and found that while LLMs could: help creativity, provide academic support when teachers were unavailable, and model certain types of writing well.
+Teachers were also concerned about the potential of LLMs to limit students' learning in certain ways through over-reliance.
+Beyond academic integrity concerns, @Lazar_2023 found that teachers had similar concerns to @Valenzuela_2024 about de-skilling and an overall loss of agency in writing and critical thinking.
+
+@Cahill_2024 surveyed undergraduate political science students on their attitudes toward, and usage of, AI tools.
+They found that the use of ChatGPT was widespread.
+However, they also found that many students lacked confidence in using AI for academic purposes.
+In particular, only 11% "strongly agreed" that they know how to use AI to improve their writing.
+Students had nuanced views on appropriate AI use.
+Many respondents felt that using it to write whole papers was inappropriate, but using it for basic tasks like general assistance, writing feedback and basic data visualization appropriate.
+
+To understand the current state of LLMs as a tool for trustworthy data science, this paper focuses on the association between student academic performance and their LLM usage.
+Specifically, we examine the relationship between students' grades and self-reported measures of student LLM usage, as well as student attitudes toward LLMs in general.
+This is based on students' final papers and a survey, conducted in a third-year undergraduate data science course at the University of Toronto.
+By examining how students interact with and perceive LLMs as tools, and how these variables translate into student outcomes, current practice with regard to LLM integration in data science can be better understood, leading to better recommendations for their future development.
+
+The remainder of this paper is structured as follows: @sec-data visualizes and analyzes survey data and coursework from students.
+@sec-model specifies a model used to investigate the relationship.
+@sec-results describes and analyzes the model's results.
+@sec-discussion discusses the implications of the findings for data science education and future research and practice at the intersection of LLMs and trustworthy data science.
 
@@
 
-## Overview
+## Background
 
-We use the statistical programming language R [@citeR].... Our data.... Following @tellingstories, we consider...
+To investigate students' usage and attitudes towards LLMs and how they related to their academic performance, a dataset containing their usage and attitudes, coursework, and academic performance was constructed.
+This was based on three components:
 
-Overview text
+1.  an optional survey;
+2.  self-reported LLM usage; and
+3.  student marks on their final paper.
 
-## Measurement
+All data are from the cohort of students taking STA302 "Methods of Data Analysis I" in the Winter 2024 semester at the University of Toronto.
+This course had 275 students initially enrolled which, reflecting a normal rate of attrition for undergraduate statistics courses at the University of Toronto, reduced to 154 students by the end of the semester.
+Assessment was heavily based on three papers submitted over the course of the 12 week semester.
 
-Some paragraphs about how we go from a phenomena in the world to an entry in the dataset.
+The student marks that we analyze are based only on the final paper, which is done individually.
+By this stage, uninterested students have typically dropped the course, and students are familiar with course expectations.
+A typical paper submission is 10-20 pages, and requires students to conduct original research to answer a research question of interest to them.
+It reflects the skills typically used by a professional data scientist.
+Students are expected to develop a research question of interest to them, identify or collect data to answer the question, conduct statistical analysis, and write a short paper.
+Examples of final papers (shared with consent) include: @hannahyu; @emilysu; and @bennyrochwerg.
 
-## Outcome variables
+By the time they are working on their final paper, students have submitted and received feedback on two previous papers with similar requirements and rubrics to that of the final paper.
+Each paper has the same basic structure and expectations.
+Before the final paper is due, students have received feedback on all their previous work in the class (including their past papers) and there is an optional two-day period of peer review.
 
-Talk way more about it.
+The pre-requisites of this course mean that the typical student is an upper-year undergraduate.
+Coding and writing are major parts of the course.
+Students are welcome to use R or Python, but the majority code in R because that is the programming language currently mostly taught in pre-requisite courses.
+All writing must be in English.
+The primary motivation for having students write three papers as the main assessment for the course is to give them the opportunity to create a public portfolio of work they can use to apply for jobs.
 
-## Predictor variables
+Throughout the semester students were encouraged to use LLMs.
+Formal instruction was provided twice during the semester.
+The first was a masterclass taught by a computer science faculty member on the ethics of using LLMs (see @Horton2024 for details).
+The second was a masterclass taught by a TA on writing with LLMs.
 
-Add graphs, tables and text.
+Data was collected from students through an optional end-of-course survey.
+[Appendix -@sec-survey-details] details the questions asked in the survey.
+Whether or not they consented to their data being used, all respondents received a 1% increase in their final course grade for their participation.
+Consenting responses were then matched to their final paper mark, as well as the GitHub repository for their final paper.
+The responses were anonymized by removing any personal references to the students themselves including, names, emails, student numbers, and GitHub links.
 
-Use sub-sub-headings for each outcome variable and feel free to combine a few into one if they go together naturally.
+Data cleaning and analysis was done using the `R` statistical programming language [@citeR], and the `tidyverse` [@tidyverse], `janitor` [@janitor], `reshape2` [@reshape], and `readxl` [@readxl] packages.
+
+## Survey data
+
+There were 146 responses to the survey.
+Of these, 119 respondents provided authorization for their data to be collected and used.
+Four of those respondents submitted the survey twice, and after removing their second response, 115 responses remained.
+Of those, 15 respondents did not include a statement on LLM usage in the README of the GitHub repository of their final paper, leaving 100 responses.
+Finally 7 of those respondents did not provide a usable GPA response.
+This leaves 93 respondents that were of use and were merged based on student name.
+
+All but one respondent completed the survey within 45 minutes (@fig-response-time-and-gpa-1).
+That one respondent took more than 4,000 minutes to complete the survey, suggesting they took a break while filling it out.
+That respondent is included in the analysis dataset, but without them, the average time to complete the survey was 11 minutes, and the standard deviation was 8 minutes.
+
+```{r fig-response-time-and-gpa}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: Distribution of respondents' survey response times and self-reported GPA
+#| fig-subcap: ["Distribution of survey response times","\"What is your GPA?\""]
+#| layout-ncol: 1
+
+all_data |>
+  # summarize(mean = mean(mins_complete/60),
+  # standard_dev = sd(mins_complete/60))
+  ggplot(aes(x = mins_complete/60)) +
+  geom_histogram(binwidth = 1) +
+  labs(x = "Time to complete survey (minutes)", y = "Frequency") +
+  theme_minimal()
+
+all_data |>
+  # summarize(mean = mean(what_is_your_gpa, na.rm = TRUE),
+  #           standard_dev = sd(what_is_your_gpa, na.rm = TRUE))
+  ggplot(aes(x = what_is_your_gpa)) +
+  geom_histogram(binwidth = 0.1) +
+  labs(x = "Self-reported GPA", y = "Frequency") +
+  theme_minimal()
+```
+
+There is a wide distribution of self-reported GPAs (@fig-response-time-and-gpa-2).
+The majority of responses cluster around a B (3.0 out of 4.0), and the average is 3.06 with a standard deviation of 0.55.
+One factor that may affect the range is that the course is required for programs in the Statistics, Mathematics and Computer Science Departments.
+Self-reported GPAs also introduce the possibility of reporting bias.
+For instance, respondents may have provided their cumulative GPA, their most recent term's GPA, or could have misreported it entirely.
+
+Students from a range of years took the course, however the majority of respondents were in their 3rd or 4th year of study (@tbl-year-dist).
+The course's prerequisite two-course sequence is typically completed by students in their second year, which makes it difficult for students to take this course earlier than their 3rd year.
+
+```{r tbl-year-dist}
+#| message: false
+#| echo: false
+#| warning: false
+#| tbl-cap: \"What year are you?\"
+
+all_data |>
+  count(what_year_are_you) |>
+  pivot_wider(
+    names_from = what_year_are_you,
+    values_from = n,
+    values_fill = list(n = 0)
+  ) |>
+  kable()
+```
+
+Respondents had a varied self-perception of their coding and writing abilities (@fig-selfperception).
+Most respondents believe it is important to be good at writing, but many are either indifferent or do not like to write (@fig-selfperception-1).
+Respondents also do not find writing to be particularly easy, which could be associated with the reported relative antipathy toward writing.
+Most respondents were at least somewhat confident in their own writing abilities, but a substantial number felt otherwise.
+Although few respondents felt that they were confident in their writing ability, more felt that they were able to catch their mistakes, which could indicate a disconnect between how respondents perceive their work and how the work was evaluated.
+
+```{r fig-selfperception}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: Self-perception of coding and writing abilities
+#| fig-subcap: ["\"Please rate how much each statement describes you, on a scale from 'This is very different to me' to 'This is a lot like me'\"", "\"Please rate how much each statement describes you, on a scale from 'This is very different to me' to 'This is a lot like me'\""]
+
+# Writing
+writing_perceptions_long <-
+  all_data |>
+  select(
+    writing_is_easy_for_me,
+    i_like_to_write,
+    i_believe_it_is_important_to_be_a_good_writer,
+    when_i_edit_it_is_easy_for_me_to_catch_my_mistakes,
+    i_feel_confident_sharing_my_writing,
+    i_am_confident_in_my_overall_writing_ability
+  ) |>
+  pivot_longer(cols = everything(),
+               names_to = "statements",
+               values_to = "response") |>
+  mutate(
+    statements = case_when(
+      statements == "writing_is_easy_for_me" ~ "\"Writing is easy for me.\"",
+      statements == "i_like_to_write" ~ "\"I like to write.\"",
+      statements == "i_believe_it_is_important_to_be_a_good_writer" ~ "\"I believe it is important\nto be a good writer.\"",
+      statements == "when_i_edit_it_is_easy_for_me_to_catch_my_mistakes" ~ "\"When I edit it is easy for\nme to catch my mistakes.\"",
+      statements == "i_feel_confident_sharing_my_writing" ~ "\"I feel confident\nsharing my writing.\"",
+      statements == "i_am_confident_in_my_overall_writing_ability" ~ "\"I am confident in my\noverall writing ability.\"",
+      TRUE ~ statements
+    ),
+    response = case_when(
+      response == "\"This is a lot like me\"" ~ "A lot like me",
+      response == "\"This is very different to me\"" ~ "Very different\nfrom me",
+      response == "\"This somewhat describes me\"" ~ "Somewhat describes me",
+      TRUE ~ response
+    )
+  )
+
+# so that "Very different from me" is Red - better visuals
+set1_palette <- c("#4DAF4A", "#377EB8", "#E41A1C")
+
+writing_perceptions_long |>
+  ggplot(aes(x = response, fill = response)) +
+  geom_bar(position = "stack", stat = "count") +
+  facet_wrap(vars(statements)) +
+  labs(x = "Response", y = "Number of repondents", fill = "Response:") +
+  scale_fill_manual(values = set1_palette) +
+  theme_minimal() +
+  theme(axis.text.x = element_blank(), legend.position = "bottom")
+
+# Coding
+coding_perceptions_long <-
+  all_data |>
+  select(
+    coding_is_easy_for_me,
+    i_like_to_code,
+    i_believe_it_is_important_to_be_a_good_coder,
+    when_i_check_my_code_it_is_easy_for_me_to_catch_my_mistakes,
+    i_feel_confident_sharing_my_code,
+    i_am_confident_in_my_overall_coding_ability
+  ) |>
+  pivot_longer(cols = everything(),
+               names_to = "statements",
+               values_to = "response") |>
+  mutate(
+    statements = case_when(
+      statements == "coding_is_easy_for_me" ~ "\"Coding is easy for me\"",
+      statements == "i_like_to_code" ~ "\"I like to code\"",
+      statements == "i_believe_it_is_important_to_be_a_good_coder" ~ "\"I believe it is important\n to be a good coder\"",
+      statements == "when_i_check_my_code_it_is_easy_for_me_to_catch_my_mistakes" ~ "\"When I check my code, it is\n easy for me to catch my\nmistakes\"",
+      statements == "i_feel_confident_sharing_my_code" ~ "\"I feel confident\n sharing my code\"",
+      statements == "i_am_confident_in_my_overall_coding_ability" ~ "\"I am confident in\n my overall coding ability\"",
+      TRUE ~ statements
+    ),
+    response = case_when(
+      response == "\"This is a lot like me\"" ~ "A lot like me",
+      response == "\"This is very different to me\"" ~ "Very different\nfrom me",
+      response == "\"This somewhat describes me\"" ~ "Somewhat describes me",
+      TRUE ~ response
+    )
+  )
+
+coding_perceptions_long |>
+  ggplot(aes(x = response, fill = response)) +
+  geom_bar(position = "stack", stat = "count") +
+  facet_wrap(vars(statements)) +
+  labs(x = "Response", y = "Number of repondents", fill = "Response:") +
+  scale_fill_manual(values = set1_palette) +
+  theme_minimal() +
+  theme(axis.text.x = element_blank(), legend.position = "bottom")
+```
+
+Respondent self-perceptions regarding coding proficiency and importance varied (@fig-selfperception-2).
+There is strong consensus on the perceived importance of coding skills.
+However, respondents' self-assessed coding abilities and enjoyment are more heterogeneous, with a substantial proportion reporting moderate rather than high levels of ease and enjoyment in coding tasks.
+
+Overall there was a moderate level of confidence among respondents in their overall coding ability, willingness to share code, and capacity to identify errors (@fig-selfperception-2).
+Notably, respondents express slightly higher confidence in detecting their own coding mistakes compared to general coding ability or code sharing.
+These patterns suggest that while respondents have developed some coding self-efficacy, there is still considerable potential for enhancing their perceived competence and comfort across various coding-related activities.
+
+The majority of respondents were at least "somewhat familiar" with generative AI such as ChatGPT (@tbl-ai-familiarity-usage-1).
+Though respondents' self-perceptions around writing and coding are varied, there was strong consensus that the use of generative AI tools is appropriate within an academic setting (@tbl-ai-familiarity-usage-2).
+Most respondents who responded "It depends" generally found artificial intelligence tools to be appropriate, though with certain guidelines and rules governing their use.
+
+```{r tbl-ai-familiarity-usage}
+#| message: false
+#| echo: false
+#| warning: false
+#| tbl-cap: Familiarity with, and appropriateness of, generative AI
+#| tbl-subcap: ["'How familiar are you with using generative AI tools such as OpenAI's ChatGPT or equivalents?'","'To what extent do you think using generative AI tools such as ChatGPT by OpenAI (or equivalents) is ethical and appropriate for schoolwork?'"]
+
+all_data |>
+  count(ai_familiarity) |>
+  kable(col.names = c("Extent of AI familiarity", "Students"))
+
+all_data |>
+  count(ai_schoolwork_appropriate) |>
+  kable(col.names = c("Ethical & Appropriate for school?", "Students"))
+```
+
+To understand the role of LLMs in learning, students identified their usage in a more granular way by selecting various pre-defined use cases in the survey (@fig-AI-use-and-usefulness-1).
+Technical questions and explaining concepts were the two top use cases among students in the course.
+More than half of students also used LLMs for quick questions, general knowledge, writing paper code, and checking solutions.
+Just less than half used it to write paper content, which could suggest that students do not feel confident using it to improve writing; this could be similar to the attitudes expressed by students surveyed in @Cahill_2024.
+
+```{r fig-AI-use-and-usefulness}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: "Use and usefulness of generative AI"
+#| fig-subcap: ["\"If you have used generative AI tools such as OpenAI's ChatGPT or equivalents, in what ways have you used it (select all that apply)?\"", "\"How helpful did you find generative AI tools such as ChatGPT by OpenAI (or equivalents) for each component of STA302?\""]
+
+# Use
+all_data |>
+  select(how_have_you_used) |>
+  separate_rows(how_have_you_used, sep = ";") |>
+  mutate(
+    how_have_you_used = case_when(
+      how_have_you_used == "check grammar mistake" ~ "Other",
+      how_have_you_used == "debugging code" ~ "Other",
+      how_have_you_used == "mental health support" ~ "Other",
+      how_have_you_used == "Making practice questions for me with increasing difficulty" ~ "Other",
+      how_have_you_used == "Generating test questions" ~ "Other",
+      TRUE ~ how_have_you_used
+    )
+  ) |> 
+  count(how_have_you_used) |>
+  filter(how_have_you_used != "") |>
+  arrange(desc(n)) |>
+  ggplot(aes(x = n, y = reorder(how_have_you_used, n))) +
+  geom_bar(stat = "identity") +
+  labs(x = "Number of respondents", y = "Type of usage") +
+  theme_minimal()
+
+# Usefulness
+all_data |>
+  select(
+    weekly_quiz,
+    weekly_mini_essay,
+    papers_generating_ideas,
+    papers_writing_code,
+    papers_writing_content,
+    papers_improving_content
+  ) |>
+  pivot_longer(cols = everything(),
+               names_to = "task",
+               values_to = "response") |>
+  mutate(
+    # Simplify response categories and rename tasks
+    response_simplified = case_when(
+      response %in% c(
+        "\"Not helpful\"",
+        "\"I did not use generative AI for this component\""
+      ) ~ "Less Helpful",
+      response %in% c("\"Somewhat helpful\"", "\"Very helpful\"") ~ "More Helpful",
+      TRUE ~ "Unknown"  # Handle unexpected values
+    ),
+    task = case_when(
+      task == "weekly_quiz" ~ "Weekly Quiz",
+      task == "weekly_mini_essay" ~ "Weekly Mini Essay",
+      task == "papers_generating_ideas" ~ "Generating Ideas",
+      task == "papers_writing_code" ~ "Writing Code",
+      task == "papers_writing_content" ~ "Writing Content",
+      task == "papers_improving_content" ~ "Improving Content",
+      TRUE ~ "Other"  # Handle any unexpected task names
+    )
+  ) |>
+  group_by(task, response_simplified) |>
+  summarise(count = n(), .groups = 'drop') |>
+  ggplot(aes(x = response_simplified, y = count, fill = response_simplified)) +
+  geom_bar(stat = "identity") +
+  facet_wrap(~ task, scales = "free_y") +
+  labs(y = "Number of respondents", x = "", fill = "Response type: ") +
+  theme_minimal() +
+  coord_flip() +
+  scale_fill_brewer(palette = "Set1") +
+  theme(axis.text.y = element_blank(), legend.position = "bottom")
+```
+
+Respondents were also asked to rate the helpfulness of LLMs on various tasks assigned during the course on a 4-point scale of "Did not use" to "Very Helpful" (@fig-AI-use-and-usefulness-2).
+To simplify the presentation, responses were grouped into two main categories: "Less Helpful" and "More Helpful." The "Less Helpful" category combines responses where students found the AI either "Not helpful" or did not use it for the task, while the "More Helpful" category includes responses where the LLM was considered "Somewhat helpful" or "Very helpful".
+
+Respondents differed in terms of how they used LLMs in the course (@fig-AI-use-and-usefulness-2).
+While most tasks were roughly split between respondents finding LLMs helpful or not, respondents found them most helpful in generating code.
+In the context of the course, this mostly meant generating R code for transforming, analyzing, and visualizing data.
+To a lesser extent, respondents also found LLMs to be helpful in improving the existing writing they had, while at the same time not favouring it for writing content from scratch.
+
+Finally, one question asked students to elaborate on whether they thought generative AI tools such as ChatGPT were ethical and appropriate for schoolwork.
+This was an open-response question.
+To provide a sense of the responses, we used Anthropic's Claude 3.5 Sonnet model (as at 5 August 2024) to summarize the comments and it provided:
+
+> Many students view AI as a helpful supplementary tool, comparing it to resources like Google or calculators.
+They believe it can aid in understanding concepts, debugging code, brainstorming ideas, and saving time on routine tasks.
+However, there's a consensus that AI should not be used to complete entire assignments or replace original thinking.
+Students emphasize the importance of using AI ethically, citing it when appropriate, and not relying on it exclusively.
+Some argue that learning to use AI effectively is a valuable skill for future careers.
+Concerns raised include the potential for plagiarism, the risk of hindering critical thinking skills, and the possibility of receiving incorrect information.
+Overall, most students support the responsible use of AI in education, with proper guidelines and transparency, while recognizing the need to maintain academic integrity and develop independent learning skills.
+
+## LLM usage and final paper marks
+
+Two other components were merged with the survey responses: self-reported LLM usage on the final paper, and final paper mark.
+
+Students were encouraged to use LLMs to complete their papers.
+Each paper required the students to disclose their usage through a statement in the GitHub repository README for the paper.
+Even students who did not use generative AI at all were required to state this in the README.
+For students who did use generative AI, there was an additional requirement, where possible, that they save the logs of their usage in a txt file which was also included in their GitHub repository.
+
+Those README statements were gathered and parsed using OpenAI's ChatGPT 4o model (as at 26 July). The following prompt was used:
+
+> The following statement is about to what extent LLMs were used by a student.
+Please characterize it as one of: "None", "Minimal", "Somewhat", "Extensive", "Unsure".
+Respond with only one of those options.
+
+All classifications were then manually checked for reasonableness.
+
+We find a varied extent of self-reported LLM usage (@tbl-llm-usage).
+41 respondents were classified as having made extensive use of LLMs, while 28 were classified as having made somewhat use.
+31 respondents were classified as having made minimal or no use of LLMs.
+In the analysis dataset we combine those two classifications because only 8 respondents were classified as having minimal usage.
+
+```{r tbl-llm-usage}
+#| message: false
+#| echo: false
+#| warning: false
+#| tbl-cap: Self-reported LLM usage in final paper
+
+all_data |>
+  count(llm_usage) |> 
+  kable(col.names = c("Self-reported LLM usage", "Number"))
+```
+
+The third, and final, component is the mark, in percentages, on the final paper (@fig-mark-1).
+The overall mean was 78% and standard deviation was 17 percentage points.
+However there was considerable differences by the extent of LLM usage (@tbl-mark; @fig-mark-2).
+
+```{r fig-mark}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: Distribution of marks on final paper (in percentages)
+#| fig-subcap: ["Overall", "By self-reported LLM usage"]
+
+all_data |>
+  select(mark) |>
+  ggplot(aes(x = mark)) +
+  geom_histogram(binwidth = 0.05) +
+  scale_x_continuous(labels = label_percent()) +
+  labs(x = "Final paper mark (percentage)", y = "Count") +
+  theme_minimal()
+
+all_data |>
+  ggplot(mapping = aes(x = mark)) +
+  geom_histogram(binwidth = 0.01) +
+  scale_x_continuous(labels = label_percent()) +
+  labs(x = "Final paper mark (percentage)", y = "Count") +
+  facet_wrap(vars(llm_usage),
+             ncol = 2,
+             labeller = labeller(
+               llm_usage = c(
+                 "Extensive" = "Extensive LLM usage",
+                 "Minimal" = "Minimal LLM usage",
+                 "None" = "No LLM usage",
+                 "Somewhat" = "Some LLM usage"
+               )
+             )) +
+  theme_minimal()
+```
+
+```{r tbl-mark}
+#| message: false
+#| echo: false
+#| warning: false
+#| tbl-cap: Self-reported LLM usage in final paper and final paper mark
+
+all_data |>
+  group_by(llm_usage) |>
+  summarize(mean = mean(mark) |> round(2),
+            sd = sd(mark) |> round(2)) |>
+  kable(col.names = c("Self-reported LLM usage", "Mean", "Std dev"))
+```
 
@@
 
-In this analysis, we aim to model the popularity trends for the two 2024 president nominates, Kamala Harris and Donald Trump, based on high-quality polling data collected at the national level after Harris’s declaration on July 21, 2024.
-We used Bayesian linear regression models with Gaussian error structures to estimate changes in polling percentages over time for each candidate.
+The goal of our modelling strategy is to better understand how a respondents' result on their final paper associates with their self-reported LLM usage.
+Students received their result on the final paper as a proportion (a 0-100% mark), so we use a Bayesian zero-one-inflated beta regression to best fit the distribution in this interval.
+We use this type of regression given the ability of the beta distribution to effectively describe continuous and skewed  (heteroskedastic) datasets, which we can see in the asymmetrical histogram of the distribution of marks.
 
-## Data Filtering and Preparation
+Transforming the regression with a logit function as the link function ensures that the predicted values stay between zero and one, respecting the boundaries of the data.
 
-The data were first filtered to retain only high-quality polls, defined as those with a numeric grade of 2.0 or above and a transparency score of 4 or higher.
-We focused exclusively on polls where the *state* variable indicates national coverage, ensuring the poll represents nationwide opinions rather than a single state, which could introduce bias.
-Additionally, we filtered for polls conducted after July 21, 2024, the date when Kamala Harris announced her candidacy.
+We briefly describe the model that we use, which follows @kurztotherescue, which was adapted from @ospina.
+No students got zeroes, though some got full marks (100%) on the paper, so we adapt the regression to one-inflated beta model as described by @heisstotherescue to account for this.
 
-To prevent issues with missing data in our models, we excluded any rows where key variables (such as polling percentage or end date) were missing.
+Model diagnostics are included in [Appendix -@sec-model-details].
 
-## Model Specification
+The model uses a one-inflated beta regression approach to analyze the relationship between student LLM usage, GPA, and final paper marks. Let $Y_i$ represent the mark for student $i$, modeled as:
 
-Two separate Bayesian linear regression models were fitted using the @rs package. Both models specified the formula:
+$$Y_i \sim \text{Zero-One-Inflated Beta}(\mu_i, \phi_i, \pi_{zoi_i}, \pi_{coi_i})$$
 
-\begin{align}
-Y_i | \mu_i,\sigma  & \sim \text{Normal}(\mu_i,\sigma) \\
-\mu_i & = \beta_0 + \beta_1X_{i1}\\
-\beta_0 & \sim \text{Normal}(50, 5), \\ 
-\beta_1 & \sim \text{Normal}(0, 0.1)
-\end{align}
+where:
 
-Where $Y_i$ is the support rate for Harris or Trump, separately, and $X_{i1}$ is the number of days passed after the latest polling sample of Trump or Harris.
+- $\mu_i$ is the mean of the beta distribution (location parameter)
+- $\phi_i$ is the precision parameter (controlling dispersion)
+- $\pi_{zoi_i}$ is the probability of $Y_i$ being exactly 0 or 1
+- $\pi_{coi_i}$ is the probability of $Y_i$ being in the continuous part of the distribution
 
-## Model Justification and Limitations
-We opted to use two separate models for Trump and Harris, as their supporter demographics differ significantly in ways that are challenging to capture within a single or even a small set of variables @pew_harris_trump_2024.
-The polling dataset we obtained primarily reflects the characteristics and trends within individual polls rather than offering direct predictors of the presidential election outcome.
-To maximize the model’s reliability, we refined the dataset by selecting polls with high credibility rather than attempting to predict outcomes directly from the available data.
+The model components are specified as:
 
-Our model uses a simple linear approach with time as the sole predictor variable.
-While this may introduce potential bias by making a simplifying assumption about the trend over time, it also has the advantage of being straightforward to fit and interpret.
-This trade-off allows us to capture general trends while minimizing the risk of overfitting to limited or potentially inconsistent data sources.
+1. Mean component:
+   $logit(\mu_i) = \beta_{0,\mu} + \beta_{1,\mu} \cdot \text{llm\_usage}_i + \beta_{2,\mu} \cdot \text{GPA}_i$
 
-# Results {#set-result}
-The results of the Bayesian linear regression models for Kamala Harris and Donald Trump are summarized in @tbl-model_summary_H and @tbl-model_summary_T.
+2. Precision component:
+   $\log(\phi_i) = \beta_{0,\phi} + \beta_{1,\phi} \cdot \text{llm\_usage}_i + \beta_{2,\phi} \cdot \text{GPA}_i$
 
-```{r}
-#| label: tbl-model_summary_H
+3. Zero-one inflation component:
+   $logit(\pi_{zoi_i}) = \beta_{0,zoi} + \beta_{1,zoi} \cdot \text{llm\_usage}_i + \beta_{2,zoi} \cdot \text{GPA}_i$
+
+4. Continuous outcome inflation component:
+   $logit(\pi_{coi_i}) = \beta_{0,coi}$
+
+The "llm_usage" predictor is categorical, with students' self-reported LLM usage classified as "Extensive" (reference level), "Somewhat", or "None or minimal".
+GPA is a continuous variable representing students' self-reported grade point averages.
+The model is implemented using the brms package in R, which provides a Bayesian framework for fitting these regression models.
+
+The expected relationship between LLM usage and final mark is unclear, given the modelling strategy and variables collected.
+It could be that stronger students used LLMs in a more sophisticated way or that they did not need to use them.
+However, we can expect the relationship between GPA and final mark to be positive.
+
+\newpage
+
+# Results {#sec-results}
+
+Our key results are summarized in @tbl-model-results and @fig-model-results, with full output available in the Appendix. We draw on the `modelsummary` [@modelsummary] package to summarize the results.
+
+@tbl-model-results and @fig-model-results present the key results from two models estimating the association between LLM usage, GPA and final paper mark.
+The first model includes only LLM usage, while the second model also includes self-reported GPA as a predictor.
+
+```{r tbl-model-results}
+#| message: false
 #| echo: false
-#| tbl-cap: "Summary statistics of the Bayesian model for Harris"
 #| warning: false
-# Load required packages
-library(broom.mixed)  # For tidy model summaries from mixed models
-library(dplyr)
-library(readr)
-library(knitr)  # For nicely formatted tables
-library(modelsummary)
+#| tbl-cap: Selected Coefficient estimates and mean absolute deviation (MAD)
 
-# Load the saved models
-model1 <- readRDS(file = here::here("models/models_harris.rds"))
+fit1 <-
+  readRDS(file = here("models/fit_one_inflated.rds"))
 
-# Extract a tidy summary of the model
-model_summary <- broom.mixed::tidy(model1, effects = "fixed", conf.int = TRUE) %>%
-  select(term, estimate, std.error, conf.low, conf.high) %>%
-  rename(
-    Term = term,
-    Estimate = estimate,
-    `Std. Error` = std.error,
-    `2.5% CI` = conf.low,
-    `97.5% CI` = conf.high
+fit2 <-
+  readRDS(file = here("models/fit_one_inflated_gpa.rds"))
+
+modelsummary(
+  list(
+    "Base model" = fit1,
+    "Self-reported GPA Included" = fit2
+  ),
+  statistic = "mad",
+  fmt = 2,
+  output = "kableExtra",
+  kable_options = list(
+    booktabs = TRUE,
+    linesep = ""
+  ),
+  coef_map = c(
+    "b_llm_usageSomewhat" = "LLM usage: Somewhat",
+    "b_llm_usageNoneorminimal" = "LLM usage: None or minimal",
+    "b_what_is_your_gpa" = "GPA", 
+    "b_zoi_Intercept" = "Zero-one inflation intercept"
   )
+)
+```
 
-# Display the summary table
-kable(model_summary, digits = 3)
+\newpage
+
+```{r fig-model-results}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: Selected coefficient estimates and 90 per cent credibility intervals
+modelplot(
+  list(
+    "Base model" = fit1,
+    "Including self-reported GPA" = fit2
+  ),
+  conf_level = 0.9,
+  coef_map = c(
+    "b_llm_usageSomewhat" = "LLM usage: Somewhat",
+    "b_llm_usageNoneorminimal" = "LLM usage: None or minimal",
+    "b_what_is_your_gpa" = "GPA"
+  ),
+  coef_omit = NULL
+) +
+  labs(x = "") +
+  scale_color_brewer(palette = "Set1") +
+  theme(legend.position = "bottom")
 
 ```
-```{r}
-#| label: tbl-model_summary_T
-#| echo: false
-#| tbl-cap: "Summary statistics of the Bayesian model for Trump"
-# Load required packages
-library(broom.mixed)  # For tidy model summaries from mixed models
-library(dplyr)
-library(readr)
-library(knitr)  # For nicely formatted tables
 
-# Load the saved models
-model2 <- readRDS(file = here::here("models/models_trump.rds"))
+LLM usage estimates are in relation to "Extensive" LLM usage.
+In the base model, the coefficients suggest that less usage of LLMs was associated with lower scores.
+The second model retains this association, but with slightly smaller coefficients.
+For instance, the negative impact of “LLM usage: Somewhat” is slightly reduced from -0.35 to -0.29, indicating that part of the effect observed in the base model could be explained by the GPA variable.
 
-# Extract a tidy summary of the model
-model_summary <- broom.mixed::tidy(model2, effects = "fixed", conf.int = TRUE) %>%
-  select(term, estimate, std.error, conf.low, conf.high) %>%
-  rename(
-    Term = term,
-    Estimate = estimate,
-    `Std. Error` = std.error,
-    `2.5% CI` = conf.low,
-    `97.5% CI` = conf.high
-  )
+Including self-reported GPA in the second model considerably changes the  changes the interpretation of some coefficients.
+For example, the intercept value (in Zero-one inflation intercept) becomes much more negative in the GPA-included model, suggesting that the GPA significantly affects the zero-one inflation part of the model.
+At the same time, though, the essential relationship between LLM usage and final mark is retained.
 
-# Display the summary table
-kable(model_summary, digits = 3)
-```
-## Overall trend
-The intercept of both model is positive, with Harris at 40.47 and Trump slightly lower than Harris at 40.36.
-The trend of support rate for both candidates with respect of time is positive but Harris's increment (0.10) is larger than Trump's (0.06).
+The overall fit of the model improves when GPA is included.
+There is an solid increase in R-squared from 0.095 to 0.478 and improvements in the ELPD and LOOIC metrics.
+This is also suggested by the LOO comparison between the two models (@tbl-loo in [Appendix -@sec-model-details]).
 
-## Model Diagnostics
-The model diagnostics indicate a strong fit and reliable predictive performance for both candidates, as detailed in the plots in Appendix @sec-model-details.
-Posterior predictive checks (Figures @fig-pos_check_H and @fig-pos_check_T) show that the predicted values align well with the observed data, suggesting that the models effectively capture key data patterns.
-Furthermore, the comparison of posterior and prior variables indicates minimal deviation, affirming an appropriate choice of priors.
-The credible intervals for the predictors (Figures @fig-CI_H and @fig-CI_T) encompass the observed support rates, demonstrating that the models adequately account for uncertainty in their predictions.
+# Discussion {#sec-discussion}
 
-Additional robustness checks (Figures @fig-DI_H and @fig-DI_T) confirm the stability of these models.
-The R-hat values for all parameters are below 1.1, indicating good convergence of the MCMC chains and reliable parameter estimates.
-This convergence supports the conclusion that the sampling process reached a stable distribution, reinforcing the credibility of our inferences.
+There is some very mild evidence from this observational study that LLM usage may be associated with better scores.
+However, this work should mostly be treated as a direction for future research in this area.
+More broadly, our work has three major take-aways:
 
-In conclusion, these diagnostics verify that the models are both stable and effective in reflecting trends in the data, confirming their reliability for further analysis.
+1. LLMs could help enhance the trustworthiness of data science, but not without considerable further development, testing and integration in the data science process.
+
+2. Making general purpose chatbots more opinionated and knowledgeable about data science would considerably help, and
+
+3. Alternative interaction approaches, beyond a chatbot interface, could be especially useful for students learning data science in particular.
+
+## Impact of LLMs on trustworthy data science
+
+We found that many students did use LLMs to help them complete a data science project closely related to what professional data scientists do.
+A common use case to ask the LLM to write code to make a graph.
+Another was to ask for some example paragraphs.
+
+Data science code tends to be written differently than software engineering code.
+Functions and loops are used less, as there is often less reuse of code.
+The training data of these LLMs is dominated by software engineering code, and so when an LLM is asked to write code for a data science project it tends to produce code with some of those features.
+This is neither good nor bad in and of itself, however it can clash with the style of the code that surrounds it.
+
+LLMs are also being used as an alternative search engine.
+Students often asked, for instance, how to modify a graph to add a vertical line, or similar.
+Here LLMs tended to be especially useful and able to answer questions.
+In the past students may have used StackOverflow or similar to find their answers, but the nature of the chat response likely provided a faster answer.
+
+Finally, there is the question of whether the nature of the data science has fundamentally changed or improved.
+For now, there is no evidence of that.
+It was clear when a student had just used LLM-generated text without modification, and it was often clear when large amounts of code had been written by an LLM.
+But it was not obvious when small amounts of either had been used, or when LLM-generated content had just been used as an initial draft.
+
+Many of these issues were alluded to by @bommasani2021opportunities, in the context of data science, and it is striking that they remain relevant today.
+
+## General purpose compared with opinionated specific chatbots
+
+ChatGPT and equivalents are general-purpose chatbots.
+One can ask them a technical data science question and then in the same session ask for a brownie recipe!
+However this general-purpose nature means that they are not necessarily optimized toward one particular task [@thoppilan2022lamda].
+When used for data science, this means that sometimes the functions or libraries recommended may not exist, code may not run, or out-dated approaches may be used.
+Frontier models make things up and are unnervingly confident, even when wrong [@openai2023gpt4, p.
+59], which can create special issues for learners.
+
+General-purpose chatbots being used for data science are fine for experienced data scientists who have an established sense of when something might be not ideal or even wrong.
+But it is problematic for novices who may not know, can lack the confidence to push back, and possibly, may learn the wrong approach.
+
+Retrieval-augmented generation (RAG) means providing various sources, for instance books or manuals, which the LLM uses to base its response on.
+Fine-tuning means providing the previously trained LLM with a large number of example input and outputs, which again guide the response from the LLM [@raffel2020exploring].
+A chatbot-based interface, focused on the needs of data science beginners, could use RAG and/or fine-tuning to considerably improve the quality of its responses.
+The trade-off would be a reduction in the general-purpose usability of the chatbot, but in the context of this use, such a trade-off is likely acceptable.
+
+## Changed LLM-based interfaces
+
+LLMs existed before ChatGPT, but it was the launch of ChatGPT that brought them into the mainstream.
+The underlying model was not considerably different to others and the main difference that ChatGPT brought was the chat interaction.
+These are useful, but there is considerable opportunity to develop additional ways of interacting with LLMs.
+This is particularly relevant for enhancing trustworthy data science.
+
+Search-based LLM tools like Elicit and Perplexity.ai differ to ChatGPT in that they focus on search.
+In contrast to Google, which typically provides a list of websites related to a particular search, Perplexity.ai attempts to provide an answer.
+In contrast to ChatGPT, search-based LLM tools focus on answering a question rather than engaging in discussion.
+
+Students in this class were required to use GitHub to host their papers, as well as supporting code and data.
+GitHub Issues and Pull Requests were used for peer review and submission occurred based on links to GitHub repositories.
+Trustworthy data science could be considerably enhanced by establishing something analogous to continuous integration and a suite of tests, but for data science.
+The one-off nature of much data science code would mean that LLMs could be especially useful in this proposed infrastructure stack; for instance, when a data scientist commits their code or writing, a suite of static tests could run, for instance to check that the data are being read in, that classes are appropriate, etc.
+But additionally LLMs could be used to check for possible improvements in writing, and propose context-specific improvements in code.
+For instance, to make sure that what is being described in the paper is actually what is happening in the code.
+
+Finally, one exciting area of current development in LLMs is tool use.
+One early example of this was enabling LLMs to use search engines such as Google.
+As LLMs can produce writing, and writing is the input to a Google Search, enabling this tool broadened the types of queries that LLMs could respond effectively to.
+Tool use, focused on data science, exists.
+For instance, LLMs can use R and Python to respond to queries about a dataset.
+However, further development could be useful.
+For instance, when a user wrote code in an IDE to import some data, a tool-enhanced LLM could notice and automatically write code that establishes a Pydantic-based validation model for that data.
+Similarly, consider an LLM in an IDE context where a data scientist wrote out the model, in statistical notation, that they were interested in estimating, and the LLM was able write the model in Stan, run it, save the estimates, and add a summary table to the paper, all automatically and in the background.
+Similar work has occured in other contexts and found to bring substantial benefits [@chen2022evaluating].
+
+## Weaknesses and next steps
+
+There are several, substantial weaknesses of this study.
+The foundational one is that we used observational data, much of which was self-reported.
+There may be selection bias present in terms of who used LLMs, who reported their LLM usage truthfully, and even who remained in the class.
+A different design, specifically a randomized controlled trial (RCT) would deal with many of these issues, although likely at some cost.
+
+Regression reports average estimates over the full dataset.
+However, many earlier studies found distributional effects, with low-performing individuals benefiting more than high-performing individuals.
+Again, a change in design toward an RCT could enable the exploration of this question.
+Stratification of our dataset would result in small sample size, but nonetheless our data does provide some limited suggestive evidence that this effect may be present in data science.
+For instance, looking at the 28 students who received an A+ for the final paper, 13 of them had extensive LLM usage, whereas looking at the same number of worst performing students finds that only six of them had extensive LLM usage.
+
+Along these lines, we have considered LLM usage for code and writing as equivalent, but they should actually have different impacts depending on student backgrounds.
+Distinguishing between native and second-language English speakers, and then focusing on differential LLM usage, could have added a great deal of nuance to the analysis.
+
+Finally, we only considered one outcome measure, namely grade on the final paper.
+Each paper required a considerable amount of time for the student to produce.
+If an LLM was found to reduce the time taken to produce a paper, without any reduction in quality, then that would be a similarly useful outcome.
+
+Despite these shortcomings, our work clearly identifies a need for further research examining how LLMs can be used to develop a more trustworthy data science workflow, both in and outside of the classroom.
 
@@
 
-# Appendix
+# Appendix {.unnumbered}
 
-# Additional data details
+# Survey questions {#sec-survey-details}
 
-## Model check {#sec-model-details}
+1.  After carefully reading the informed consent document, please indicate below whether you consent to have your anonymized responses included in the research study?
+    -   Yes, I authorize the use of the data collected about me for the STA302 course survey to be used. I will be compensated 1% of my course grade for completing the survey.
+    -   No I do not want my data included in the research study, but I want to complete the survey. I will be compensated 1% of my course grade for completing the survey.
+    -   I do not want to complete this survey. I realize that I am forfeiting the corresponding course credit.
+2.  What is your full name on Quercus?
+3.  What is your Student ID?
+4.  What year are you?
+5.  What is your specialization?
+6.  What is/are your major/s?
+7.  What is/are your minor/s?
+8.  What is your GPA?
+9.
+Please rate how much each statement describes you, on a scale from "This is very different to me" to "This is a lot like me" \["This is very different to me"; "This somewhat describes me"; "This is a lot like me"\]
+    -   Writing is easy for me
+    -   I like to write
+    -   I believe it is important to be a good writer.
+    -   When I edit it is easy for me to catch my mistakes.
+    -   I feel confident sharing my writing.
+    -   I am confident in my overall writing ability.
+10.
+Please rate how much each statement describes you, on a scale from "This is very different to me" to "This is a lot like me".
+When answering, please consider whichever programming language you are most familiar with.
+\["This is very different to me"; "This somewhat describes me"; "This is a lot like me"\]
+    -   Coding is easy for me
+    -   I like to code
+    -   I believe it is important to be a good coder.
+    -   When I check my code it is easy for me to catch my mistakes.
+    -   I feel confident sharing my code.
+    -   I am confident in my overall coding ability.
+11. How familiar are you with using generative AI tools such as OpenAI's ChatGPT or equivalents?
+    -   Very familiar
+    -   Somewhat familiar
+    -   Not familiar
+    -   Other
+12. Have you used any generative AI tools such as OpenAI's ChatGPT or equivalents for any reason (personal or educational)?
+    -   Yes
+    -   No
+    -   Other
+13. If you have used generative AI tools such as OpenAI's ChatGPT or equivalents, in what ways have you used it (select all that apply)?
+    -   Asking technical questions
+    -   Carrying on a conversation out of curiosity
+    -   Asking general knowledge questions
+    -   Solving homework
+    -   Checking solutions
+    -   Asking quick questions when stuck
+    -   Explaining concepts
+    -   Writing essays or paragraphs
+    -   Writing code
+    -   Never used it
+    -   Other
+14. To what extent do you think using generative AI tools such as ChatGPT by OpenAI (or equivalents) is ethical and appropriate for schoolwork?
+    -   Appropriate
+    -   Inappropriate
+    -   Other
+15. Please elaborate on your answer above.
+16. Did you use any generative AI tools such as OpenAI's ChatGPT or equivalents for STA302?
+    -   Yes
+    -   No
+    -   Other
+17.
+How helpful did you find generative AI tools such as ChatGPT by OpenAI (or equivalents) for each component of STA302?
+\["Not helpful"; "Somewhat helpful"; "Very helpful"; "I did not use generative AI for this component"\]
+    -   Weekly quiz
+    -   Weekly mini-essay
+    -   Papers: Generating ideas
+    -   Papers: Writing code
+    -   Papers: Writing content
+    -   Papers: Improving content
+18. What is your recommendation for how generative AI tools such as ChatGPT by OpenAI (or equivalents) should be used in the course in future?
+19. (Optional) Any other comments?
 
-```{r}
-#| label: fig-pos_check_H
+\newpage
+
+# Model details {#sec-model-details}
+
+We use `modelsummary` to compute the full regression output.
+
+## Full Model Output
+
+```{r tbl-model-results-app}
+#| message: false
 #| echo: false
-#| fig-cap: "Posterior Predictive Check for All Models"
+#| warning: false
+#| tbl-cap: Coefficient estimates and mean absolute deviation (MAD)
 
-# Load necessary libraries
-library(rstanarm)
-library(bayesplot)
-library(ggplot2)
-library(patchwork)  # For combining plots
+fit1 <-
+  readRDS(file = here("models/fit_one_inflated.rds"))
+fit2 <-
+  readRDS(file = here("models/fit_one_inflated_gpa.rds"))
 
-pp_check(model1) +  
+modelsummary(
+  list(
+    "Base model" = fit1,
+    "Self-reported GPA Included" = fit2
+  ),
+  statistic = "mad",
+  fmt = 2
+)
+```
+
+\newpage
+
+## Full Coefficient Estimates
+
+```{r fig-model-results-app}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: Selected coefficient estimates and 90 per cent credibility intervals
+
+modelplot(
+  list(
+    "Base model" = fit1,
+    "Including self-reported GPA" = fit2
+  ),
+  conf_level = 0.9,
+  drop = c("phi", "b_zoi_Intercept")) +
+  labs(x= "") +
+  scale_color_brewer(palette = "Set1") +
+  theme(legend.position = "bottom")
+```
+
+\newpage
+
+Next, we use `bayesplot` [@bayesplot] and `loo` [@loo] conduct posterior predictive checks and evaluate model diagnostics.
+
+## Posterior predictive check
+
+```{r fig-posteriorchecks}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: Posterior predictive checking
+#| fig-subcap: ["Base model", "Including self-reported GPA"]
+#| layout-ncol: 2
+
+pp_check(fit1) +
   theme_classic() +
@@
 
-just_harris_high_quality <- read.csv(here::here("data/02-analysis_data/Harris.csv"))
-
-posterior_vs_prior(model1) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-
-```
-```{r}
-#| label: fig-pos_check_T
-#| echo: false
-#| fig-cap: "Posterior Predictive Check for All Models"
-
-# Load necessary libraries
-library(rstanarm)
-library(bayesplot)
-library(ggplot2)
-library(patchwork)  # For combining plots
-
-pp_check(model2) +  
+pp_check(fit2) +
   theme_classic() +
   theme(legend.position = "bottom")
-
-just_trump_high_quality <- read.csv(here::here("data/02-analysis_data/Trump.csv"))
-
-posterior_vs_prior(model2) +
-  theme_minimal() +
-  scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom") +
-  coord_flip()
-
-```
-```{r}
-#| label: fig-CI_H
-#| echo: false
-#| fig-cap: "CI for predictors for Harris"
-
-# Load necessary libraries
-library(rstanarm)
-library(bayesplot)
-library(ggplot2)
-library(patchwork)  # For combining plots
-
-plot(
-  model1,
-  "areas")
-```
-```{r}
-#| label: fig-CI_T
-#| echo: false
-#| fig-cap: "CI for predictors for Trump"
-
-# Load necessary libraries
-library(rstanarm)
-library(bayesplot)
-library(ggplot2)
-library(patchwork)  # For combining plots
-
-plot(
-  model2,
-  "areas")
 ```
 
-```{r}
-#| label: fig-DI_H
+```{r tbl-loo}
+#| message: false
 #| echo: false
-#| fig-cap: "Model diagonostic for Harris"
+#| warning: false
+#| tbl-cap: Model comparison
 
-# Load necessary libraries
-library(rstanarm)
-library(bayesplot)
-library(ggplot2)
-library(patchwork)  # For combining plots
-
-plot(model1, "trace")
-plot(model1, "rhat")
-
-```
-```{r}
-#| label: fig-DI_T
-#| echo: false
-#| fig-cap: "Model diagonostic for Trump"
-
-# Load necessary libraries
-library(rstanarm)
-library(bayesplot)
-library(ggplot2)
-library(patchwork)  # For combining plots
-
-plot(model2, "trace")
-plot(model2, "rhat")
-
+brms::loo_compare(loo::loo(fit1, cores = 2), loo::loo(fit2, cores = 2)) |>
+  kable(digits = 3)
 ```
@@
 \newpage
-## pollster methodology analysis
 
-This survey was conducted by Redfield & Wilton Strategies to assess the voting intentions of eligible voters in key U.S.
-swing states ahead of the 2024 Presidential Election.
-The primary goal of this poll is to provide an accurate and timely snapshot of public opinion in states where electoral outcomes are uncertain and could have a decisive impact on the overall result of the election.
-Swing states, due to their political volatility and diverse voter bases, are critical in determining the balance of power in the U.S.
-electoral system.
-Understanding voter preferences in these states is essential for political analysts, campaigns, and the general public.
+## Diagnostics
 
-### Population of Interest
-The population of interest for this survey consists of all eligible voters residing in major U.S.
-swing states, specifically Arizona, Florida, Georgia, Michigan, North Carolina, and Pennsylvania.
-These states are known for their fluctuating political alignments and are expected to play a crucial role in the upcoming election.
+```{r fig-basemodeldiagnostics}
+#| fig-height: 7
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: Base model diagnostics
 
-### Sampling Frame
-The population sampled includes eligible voters from Arizona, Florida, Georgia, Michigan, North Carolina, and Pennsylvania. Participants were selected via an online panel.
+plot(fit1)
+```
 
-### Sample
-The sample sizes for each state were as follows:
+```{r fig-gpamodeldiagnostics}
+#| fig-height: 6
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: Model including self-reported GPA diagnostics
 
-Arizona: 750 respondents
-
-Florida: 1,350 respondents
-
-Georgia: 927 respondents
-
-Michigan: 970 respondents
-
-North Carolina: 880 respondents
-
-Pennsylvania: 1,070 respondents
-
-## Weakness & Strength of the methodology
-
-In terms of strengths, Redfield & Wilton has a great reputation for producing reliable polling data.
-They have employ a mix of online and telephone survey, which add in various resources of collecting their data.
-This approach can help reduce bias.
-In addition, Redfield & Wilton often target swing states, which makes their polls results important to the US president election.
-
-The weakness of their methodology would incorporate a certain potential bias based on their political leaning of their clients or media. This could cause a certain neutrality in their dataset.
-
-Overall, Redfield & Wilton has been considered as a competent, reputable and reliable pollster with his variaty and methodology on polls.
-Even though the pollster contained a potential bias, it has been a reliable resource in prediction of US president election.
-
-# Model details {#sec-model-details}
+plot(fit2)
+```
 
@@
 
-# Appendix 2 - Idealized Methodology and Survey
-
-## Overview
-
-This section introduces an idealized methodology and survey with $100K budget to predict the 2024 US presidential election.
-The goal is to maximize the accuracy of the prediction under the budget.
-The subsections that describe the details of the idealized methodology and survey include sampling approach, respondents recruiting method, data validation, poll aggregation and the survey questions.
-
-### Sampling Approach
-
-Cluster sampling is the sampling approach used.
-Specifically clusters are the swing states as they are the states that would affect the election result, i.e.
-Arizona, Florida, Gegorgia, Iowa, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin.
-In each swing state/clusters, units are selected based on postal code.
-Using simple random sampling, 100 distinct postal codes are randomly selected.
-People whose living address has the postal codes selected are the target respondents.
-For the postal codes that lie in the non-residential area, the postal codes would be ignored.
-
-### Recruitment Strategy
-
-The strategy to be implemented to recruit respondents is a combination of physical and online recruitment.
-
-If the building corresponds to the postal code is a condo building, a paper with QR code to the survey is going to be put in the lobby or elevator, or sent to residents through the property management.
-If the building is a residential house, then a letter with the QR code would be put in the mailbox.
-
-All respondents are rewarded with a $5 deposit to their bank account or a gift card of their choice. Each IP address is limited to answer the survey once.
-
-### Data Validation
-
-To improve accuracy of the prediction results, the following data validation approaches will be done.
-
-* Postal Code Validation
-    + All the postal codes got from respondents are going to be validated to check if it is one of the randomly selected postal codes.
-If not, the response would not be considered when the prediction is done.
-* Just-for-Rewards Prevention
-    + To identify the responses that are not seriously answered, the last question of the survey is set test if the respondents are serious and careful when answering the questions.
-* Age Validity
-    + Responses with age under 18 but answered "Yes" to "Are you registered to Vote" are discarded.
-
-### Poll Aggregation
-
-For each of the swing states, the result would be calculated based on the votes to Trump versus Harris because they are the candidates with the greatest chances of winning.
-Based on the decisiveness of the respondents and the likelihood of voting from the responses, the individual votes will be weighted when calculating the results for each swing states.
-
-To aggregate the polls for all states, the result of each states is multiplied by its electoral votes.
-After summing the electoral votes for Trump or Harris, the estimated electoral votes that Trump or Harris get is available.
-The candidate that has more than 270 electoral votes would be predicted to win the election {@nationalarchives2024electoral}.
-
-### Survey
-
-Google form of the survey is available here:  https://docs.google.com/forms/d/1obaebX3zqw7WJEd1lHrF-DVBXdZt5tbjHofjsWj9zf8/prefill
-
-1. What is the postal code of where you live?
-
-2. Are you registered to vote?
-
-* Yes
-* No
-
-3. Who would you vote for?
-
-* Donald Trump - Republican
-* Kamala Harris - Democrat
-* Other Candidates
-* Not Decided Yet
-
-4. How decisive are you to vote for the option you chose in the last question?
-
-* Very Decisive
-* Pretty Decisive
-* A Bit Indecisive
-* Very Indecisive
-
-5. How likely are you to vote for the election?
-
-* Very Likely
-* A Bit Likely
-* Not Likely
-
-6. What age group are you in?
-
-* 0 - 18
-* 19 - 30
-* 31 - 50
-* 51 - 70
-* 71+
-
-7. sex
-
-8. race
-
-9. education
-
-10. How many characters are present in the word "President"?
-
-* 6
-* 8
-* 10
-* None of the Above
-
-### Budget Specification
-
-* Physical and Online Recruitment: $30,000
-* Rewards for Respondents: $50,000
-* Data Collection & Validation: $10,000
-* Other: $10,000
-
-Total: $100,000
-
-### Conclusion
-
-The survey uses cluster sampling to sample the swing states based on postal codes.
-After data collection and validation, weighting based on the likelihood of voting and decisiveness of the respondents, the total votes for Donald Trump and Kamala Harris are calulated.
-The candidate with more than 270 votes are predicted to win the election{@nationalarchives2024electoral}.
-
 # References

### 2025-03-14 paper/paper.qmd
Reconcile Lucas updates with mine

@@
 ---
-title: "Self-reported LLM usage and results on a data science project: Evidence from a Canadian undergraduate data science course"
+title: "Self-reported LLM usage and outcomes on a data science project: Evidence from a Canadian undergraduate data science course"
 author: 
@@
 We thank Tiffany Timbers and attendees at JSM 2024 for helpful suggestions.
-This research is currently under review by the University of Toronto's Research Ethics Board.
-Contributions: RA: Conceptualization, Data Curation, Formal Analysis, Funding Acquisition, Investigation, Methodology, Project Administration, Software, Visualization, Writing – Original Draft Preparation, Writing – Review & Editing; LC: Formal Analysis, Visualization, Writing – Original Draft Preparation, Writing – Review & Editing; NM: Conceptualization, Methodology, Writing – Original Draft Preparation, Writing – Review & Editing.
+This research is was approved by the University of Toronto's Research Ethics Board as Protocol 47725.
 Comments can be sent to: rohan.alexander@utoronto.ca."
@@
 date-format: long
-abstract: "To help understand the effect of Large Language Models (LLMs) on data science practice we examine the extent to which LLM usage is correlated with the mark that a student gets on a final paper in a classroom data science setting.
+abstract: "To help understand the effect of Large Language Models (LLMs) on data science practice we examine the extent to which self-reported LLM usage is correlated with the mark that a student gets on a final paper in a classroom data science setting.
 We find some very mild evidence from this observational study that LLM usage may be associated with better scores, but our main conclusion is that there is no clear relationship between more extensive LLM usage and the student's mark.
@@
 ---
+
+<!-- Contributions: RA: Conceptualization, Data Curation, Formal Analysis, Funding Acquisition, Investigation, Methodology, Project Administration, Software, Visualization, Writing – Original Draft Preparation, Writing – Review & Editing; LC: Formal Analysis, Visualization, Writing – Original Draft Preparation, Writing – Review & Editing; NM: Conceptualization, Methodology, Writing – Original Draft Preparation, Writing – Review & Editing.
+-->
 
@@
 Trustworthy data science is the practice of conducting data analysis in a transparent, ethical, and reliable manner.
-These principles are upheld in various ways, through ongoing education, adherence to professional norms and, and implementation of reproducible workflows.
+These principles are upheld in various ways, through ongoing education, adherence to professional norms and implementation of reproducible workflows.
 To stay current, data scientists must continually update their knowledge of best practices for transparency and reproducibility, foster a culture which values different perspectives and accountability, and critically consider the broader impact of data-driven decisions.
@@
 LLMs have quickly been put to use in both industry and academia.
-By now, many people, especially educators and students, have some experience with LLMs, in both personal and professional contexts.
+By now many people, especially educators and students, have some experience with LLMs in both personal and professional contexts.
 
@@
 
-In this paper we are interested in better understanding LLMs as a tool for producing trustworthy data science.
-We study how they were used by students in an upper-year undergraduate data science course and whether students who used LLMs tended to have higher scores than those who did not.
-
 The tasks involved in a trustworthy data science workflow can be generally broken down into a number of key competencies [@Adhikari2021Interleaving; @Gibbs2021Building].
@@
 Another is writing, which is primarily done when communicating results.
-The potential for LLMs to positively affect students' academic performance in data science can be clearly inferred from their already-demonstrated capability in those competencies in adjacent professional fields.
+The potential for LLMs to positively affect students' academic performance in data science follows from their already-demonstrated capability in those competencies in adjacent professional fields.
 
@@
 
-At the high school level @Lazar_2023 conducted an informal survey of secondary school teachers and students on their opinions of ChatGPT, and found that while LLMs could: help creativity, provide academic support when teachers were unavailable, and model certain types of writing well.
-Teachers were also concerned about the potential of LLMs to limit students' learning in certain ways through over-reliance.
+At the high school level @Lazar_2023 conducted an informal survey of secondary school teachers and students on their opinions of ChatGPT, and found that LLMs could help creativity, provide academic support when teachers were unavailable, and model certain types of writing well.
+But teachers were concerned about the potential of LLMs to limit students' learning in certain ways through over-reliance.
 Beyond academic integrity concerns, @Lazar_2023 found that teachers had similar concerns to @Valenzuela_2024 about de-skilling and an overall loss of agency in writing and critical thinking.
@@
 
+In this paper we are interested in better understanding LLMs as a tool for producing trustworthy data science.
+We study how they were used by students in an upper-year undergraduate data science course and whether students who used LLMs tended to have higher scores than those who did not.
+
 To understand the current state of LLMs as a tool for trustworthy data science, this paper focuses on the association between student academic performance and their LLM usage.
-Specifically, we examine the relationship between students' grades and self-reported measures of student LLM usage, as well as student attitudes toward LLMs in general.
+Specifically, we examine the relationship between the mark a student got on a final paper and self-reported measures of student LLM usage, as well as student attitudes toward LLMs in general.
 This is based on students' final papers and a survey, conducted in a third-year undergraduate data science course at the University of Toronto.
@@
 
-The remainder of this paper is structured as follows: @sec-data visualizes and analyzes survey data and coursework from students.
+The remainder of this paper is structured as follows: @sec-data visualizes and analyzes survey data, self-reported LLM usage, and final marks.
 @sec-model specifies a model used to investigate the relationship.
@@
 
-Data was collected from students through an optional end-of-course survey.
+Data were collected from students through an optional end-of-course survey.
 [Appendix -@sec-survey-details] details the questions asked in the survey.
@@
 Consenting responses were then matched to their final paper mark, as well as the GitHub repository for their final paper.
-The responses were anonymized by removing any personal references to the students themselves including, names, emails, student numbers, and GitHub links.
+The responses were then anonymized by removing any personal references to the students themselves including, names, emails, student numbers, GitHub links, and summarizing free-text responses.
 
-Data cleaning and analysis was done using the `R` statistical programming language [@citeR], and the `tidyverse` [@tidyverse], `janitor` [@janitor], `reshape2` [@reshape], and `readxl` [@readxl] packages.
+Data cleaning and analysis was done using the `R` statistical programming language [@citeR], and the 
+`arrow` [@arrow], 
+`here` [@here], 
+`janitor` [@janitor], 
+`readxl` [@readxl], and 
+`tidyverse` [@tidyverse] 
+packages.
 
@@
 Finally 7 of those respondents did not provide a usable GPA response.
-This leaves 93 respondents that were of use and were merged based on student name.
 
@@
 That one respondent took more than 4,000 minutes to complete the survey, suggesting they took a break while filling it out.
-That respondent is included in the analysis dataset, but without them, the average time to complete the survey was 11 minutes, and the standard deviation was 8 minutes.
+After that respondent is removed from the analysis the average time to complete the survey was 11 minutes, and the standard deviation was 8 minutes.
+This leaves 92 respondents that were of use and were merged based on student name.
 
@@
 One factor that may affect the range is that the course is required for programs in the Statistics, Mathematics and Computer Science Departments.
-Self-reported GPAs also introduce the possibility of reporting bias.
+While there was no reason for the respondents to not report the truth, self-reported GPAs also introduce the possibility of reporting bias.
 For instance, respondents may have provided their cumulative GPA, their most recent term's GPA, or could have misreported it entirely.
@@
 Students from a range of years took the course, however the majority of respondents were in their 3rd or 4th year of study (@tbl-year-dist).
-The course's prerequisite two-course sequence is typically completed by students in their second year, which makes it difficult for students to take this course earlier than their 3rd year.
+The course's prerequisite two-course sequence is typically completed by students in their second year, would make it difficult to take this course earlier than their 3rd year.
 
@@
 
-# so that "Very different from me" is Red - better visuals
-set1_palette <- c("#4DAF4A", "#377EB8", "#E41A1C")
-
 writing_perceptions_long |>
@@
   labs(x = "Response", y = "Number of repondents", fill = "Response:") +
-  scale_fill_manual(values = set1_palette) +
+  scale_fill_brewer(palette = "Set1") +
   theme_minimal() +
@@
 The majority of respondents were at least "somewhat familiar" with generative AI such as ChatGPT (@tbl-ai-familiarity-usage-1).
-Though respondents' self-perceptions around writing and coding are varied, there was strong consensus that the use of generative AI tools is appropriate within an academic setting (@tbl-ai-familiarity-usage-2).
-Most respondents who responded "It depends" generally found artificial intelligence tools to be appropriate, though with certain guidelines and rules governing their use.
+Though respondents' self-perceptions around writing and coding were varied, there was strong consensus that the use of generative AI tools is appropriate within an academic setting (@tbl-ai-familiarity-usage-2).
+Most respondents who selected "It depends" generally found artificial intelligence tools to be appropriate, though with certain guidelines and rules governing their use.
 
@@
 
-To understand the role of LLMs in learning, students identified their usage in a more granular way by selecting various pre-defined use cases in the survey (@fig-AI-use-and-usefulness-1).
-Technical questions and explaining concepts were the two top use cases among students in the course.
-More than half of students also used LLMs for quick questions, general knowledge, writing paper code, and checking solutions.
-Just less than half used it to write paper content, which could suggest that students do not feel confident using it to improve writing; this could be similar to the attitudes expressed by students surveyed in @Cahill_2024.
+To understand the role of LLMs in learning, respondents identified their usage in a more granular way by selecting various pre-defined use cases in the survey (@fig-AI-use-and-usefulness-1).
+Technical questions and explaining concepts were the two top use cases among respondents.
+More than half of respondents also used LLMs for quick questions, general knowledge, writing paper code, and checking solutions.
+Just less than half reported that they used it to write paper content, which could suggest that respondents do not feel confident using it to improve writing.
 
@@
 
-> Many students view AI as a helpful supplementary tool, comparing it to resources like Google or calculators.
+> Many respondents view AI as a helpful supplementary tool, comparing it to resources like Google or calculators.
 They believe it can aid in understanding concepts, debugging code, brainstorming ideas, and saving time on routine tasks.
 However, there's a consensus that AI should not be used to complete entire assignments or replace original thinking.
-Students emphasize the importance of using AI ethically, citing it when appropriate, and not relying on it exclusively.
+Respondents emphasize the importance of using AI ethically, citing it when appropriate, and not relying on it exclusively.
 Some argue that learning to use AI effectively is a valuable skill for future careers.
 Concerns raised include the potential for plagiarism, the risk of hindering critical thinking skills, and the possibility of receiving incorrect information.
-Overall, most students support the responsible use of AI in education, with proper guidelines and transparency, while recognizing the need to maintain academic integrity and develop independent learning skills.
+Overall, most respondents support the responsible use of AI in education, with proper guidelines and transparency, while recognizing the need to maintain academic integrity and develop independent learning skills.
+
+As at 14 August 2024, the default Anthropic setting is that they do not use input data for training and so these student responses should have entered future training datasets [@anthropic].
 
@@
 We find a varied extent of self-reported LLM usage (@tbl-llm-usage).
-41 respondents were classified as having made extensive use of LLMs, while 28 were classified as having made somewhat use.
-31 respondents were classified as having made minimal or no use of LLMs.
-In the analysis dataset we combine those two classifications because only 8 respondents were classified as having minimal usage.
+38 respondents were classified as having made extensive use of LLMs, while 26 were classified as having made somewhat use.
+28 respondents were classified as having made minimal or no use of LLMs.
+In the analysis dataset we combine those two classifications because only a handful of respondents were classified as having minimal usage.
 
@@
 The overall mean was 78% and standard deviation was 17 percentage points.
-However there was considerable differences by the extent of LLM usage (@tbl-mark; @fig-mark-2).
+However there were considerable differences by the extent of LLM usage (@tbl-mark; @fig-mark-2).
 
@@
 The goal of our modelling strategy is to better understand how a respondents' result on their final paper associates with their self-reported LLM usage.
-Students received their result on the final paper as a proportion (a 0-100% mark), so we use a Bayesian zero-one-inflated beta regression to best fit the distribution in this interval.
-We use this type of regression given the ability of the beta distribution to effectively describe continuous and skewed  (heteroskedastic) datasets, which we can see in the asymmetrical histogram of the distribution of marks.
+Students received their result on the final paper as a percentage between 0% and 100%.
+There were no students who got 0%, but 9 students got 100%.
+As such, we use one-inflated beta regression [@Ospina2012].
+Beta regression is commonly used when data are between 0 and 1 and is focused on estimating the mean and variance of the beta distribution.
+One approach is to consider a one-inflated distribution which is a mixture of focusing on those responses that were 1, and the usual Beta distribution aspect.
 
-Transforming the regression with a logit function as the link function ensures that the predicted values stay between zero and one, respecting the boundaries of the data.
+The main variable of interest, `llm_usage` is a categorical variable with three possible values: "Extensive", "Somewhat", "None or minimal".
+We use "None or minimal" as the reference level and so the results are in relation to that level of self-reported LLM usage.
+Self-reported GPA, `what_is_your_gpa`, is a number between 1.9 and 4 with one decimal place.
 
-We briefly describe the model that we use, which follows @kurztotherescue, which was adapted from @ospina.
-No students got zeroes, though some got full marks (100%) on the paper, so we adapt the regression to one-inflated beta model as described by @heisstotherescue to account for this.
+Define $y_i$ as the percentage that student $i$ received on the final paper:
 
-Model diagnostics are included in [Appendix -@sec-model-details].
+$$y_i \sim \mbox{Zero-One-Inflated Beta}\left(\mu_i, \phi_i, \mbox{zoi}_i, \mbox{coi}\right)$$
 
-The model uses a one-inflated beta regression approach to analyze the relationship between student LLM usage, GPA, and final paper marks. Let $Y_i$ represent the mark for student $i$, modeled as:
-
-$$Y_i \sim \text{Zero-One-Inflated Beta}(\mu_i, \phi_i, \pi_{zoi_i}, \pi_{coi_i})$$
-
-where:
+Where:
 
@@
 
-The model components are specified as:
+The model components are:
 
@@
 
-The "llm_usage" predictor is categorical, with students' self-reported LLM usage classified as "Extensive" (reference level), "Somewhat", or "None or minimal".
-GPA is a continuous variable representing students' self-reported grade point averages.
-The model is implemented using the brms package in R, which provides a Bayesian framework for fitting these regression models.
-
-The expected relationship between LLM usage and final mark is unclear, given the modelling strategy and variables collected.
-It could be that stronger students used LLMs in a more sophisticated way or that they did not need to use them.
-However, we can expect the relationship between GPA and final mark to be positive.
-
-\newpage
+We estimate the model and explore the results using the statistical programming language `R` [@citeR], and the 
+`brms` [@brms], and 
+`modelsummary` [@modelsummary] 
+packages.
+Our code to adjust for the fact that there were no students who got 0% follows @buerkner and @Heiss2021.
+Model diagnostics are included in [Appendix -@sec-model-details].
+Finally, the piecewise approach is a common one for when there may be something different about the 0s or 1s, compared with the other responses.
+Another approach for dealing with a situation in which there are exact 0s or 1s is to slightly perturb the data.
+In our case, there is not much difference between getting 99% on a paper and getting 100%.
+When we estimated that perturbed data model using Beta regression the results were similar.
 
@@
 
-Our key results are summarized in @tbl-model-results and @fig-model-results, with full output available in the Appendix. We draw on the `modelsummary` [@modelsummary] package to summarize the results.
-
-@tbl-model-results and @fig-model-results present the key results from two models estimating the association between LLM usage, GPA and final paper mark.
-The first model includes only LLM usage, while the second model also includes self-reported GPA as a predictor.
+Our results are summarized in @tbl-model-results and @fig-model-results, with full output available in the Appendix. We draw on the `modelsummary` [@modelsummary] package to summarize the results.
 
@@
 #| warning: false
-#| tbl-cap: Selected Coefficient estimates and mean absolute deviation (MAD)
+#| tbl-cap: Selected coefficient estimates and mean absolute deviation (MAD)
 
@@
   readRDS(file = here("models/fit_one_inflated.rds"))
-
 fit2 <-
@@
     "Base model" = fit1,
-    "Self-reported GPA Included" = fit2
+    "Self-reported GPA included" = fit2
   ),
@@
 
-\newpage
-
 ```{r fig-model-results}
@@
 #| fig-cap: Selected coefficient estimates and 90 per cent credibility intervals
+
 modelplot(
@@
   theme(legend.position = "bottom")
+```
 
-```
+@tbl-model-results and @fig-model-results present the results from two models estimating the association between LLM usage, GPA and final paper mark.
+The first model includes only LLM usage, while the second model also includes self-reported GPA.
 
@@
 The overall fit of the model improves when GPA is included.
-There is an solid increase in R-squared from 0.095 to 0.478 and improvements in the ELPD and LOOIC metrics.
+There is an increase in R-squared from 0.095 to 0.478 and improvements in the ELPD and LOOIC metrics.
 This is also suggested by the LOO comparison between the two models (@tbl-loo in [Appendix -@sec-model-details]).
@@
 1. LLMs could help enhance the trustworthiness of data science, but not without considerable further development, testing and integration in the data science process.
-
 2. Making general purpose chatbots more opinionated and knowledgeable about data science would considerably help, and 
-
 3. Alternative interaction approaches, beyond a chatbot interface, could be especially useful for students learning data science in particular.
@@
 
-There are several, substantial weaknesses of this study.
+There are several substantial weaknesses of this study.
 The foundational one is that we used observational data, much of which was self-reported.
@@
   conf_level = 0.9,
-  drop = c("phi", "b_zoi_Intercept")) +
-  labs(x= "") +
+  drop = c("phi", "b_zoi_Intercept")
+) +
+  labs(x = "") +
   scale_color_brewer(palette = "Set1") +
@@
 
-Next, we use `bayesplot` [@bayesplot] and `loo` [@loo] conduct posterior predictive checks and evaluate model diagnostics.
+## Posterior predictive check
 
-## Posterior predictive check
+We use `bayesplot` [@bayesplot] and `loo` [@loo] conduct posterior predictive checks and evaluate model diagnostics.
 

### 2025-03-14 scripts/01-data_cleaning.R

@@ -6,10 +6,9 @@
 
 
 #### Workspace setup ####
-library(here)
 library(tidyverse)
+library(here)
 library(janitor)
-library(stringr)
 library(readxl)
 library(arrow)
 

### 2025-03-14 scripts/03-model.R

@@ -20,6 +20,12 @@ all_data <-
   read_parquet(here::here("data/analysis_data.parquet"))
 
 
+# Change the reference level so it's in terms of none
+all_data <- 
+  all_data |> 
+  mutate(llm_usage = relevel(llm_usage, ref = "None or minimal"))
+
+
 #### Model data ####
 #### Simplified data ####
 # Change the 1s to 0.99 so we can just use Beta initially

### 2025-03-14 paper/paper.qmd
small edits

@@
 The main variable of interest, `llm_usage` is a categorical variable with three possible values: "Extensive", "Somewhat", "None or minimal".
-We use "None or minimal" as the reference level and so the results are in relation to that level of self-reported LLM usage.
-Self-reported GPA, `what_is_your_gpa`, is a number between 1.9 and 4 with one decimal place.
+We use "None or minimal" as the reference level, so the results are in relation to that level of self-reported LLM usage.
+Self-reported GPA, `what_is_your_gpa`, is a number between 1.9 and 4.0, measured with one decimal place.
 
-Define $y_i$ as the percentage that student $i$ received on the final paper:
+Define $y_i$ as the percentage (0-100%) that student $i$ received on the final paper:
 
@@
 
-We estimate the model and explore the results using the statistical programming language `R` [@citeR], and the 
-`brms` [@brms], and 
-`modelsummary` [@modelsummary] 
-packages.
+We estimate the model and explore the results using the statistical programming language `R` [@citeR],`brms` [@brms], and`modelsummary` [@modelsummary]  packages.
 Our code to adjust for the fact that there were no students who got 0% follows @buerkner and @Heiss2021.
 Model diagnostics are included in [Appendix -@sec-model-details].
-Finally, the piecewise approach is a common one for when there may be something different about the 0s or 1s, compared with the other responses.
+
+In estimating models of this type, a piecewise approach is a common one for when there may be something different about the 0s or 1s, compared with the other responses.
 Another approach for dealing with a situation in which there are exact 0s or 1s is to slightly perturb the data.
-In our case, there is not much difference between getting 99% on a paper and getting 100%.
+In our case,however, there is little difference between getting 99% on a paper and getting 100%.
 When we estimated that perturbed data model using Beta regression the results were similar.
@@
 
-LLM usage estimates are in relation to "Extensive" LLM usage.
+LLM usage estimates are in relation to "None or Minimal" LLM usage.
 In the base model, the coefficients suggest that less usage of LLMs was associated with lower scores.

### 2025-04-05 paper/paper.qmd
Add SSTA304

--- 2025-03-14 paper/paper.qmd
@@
 ---
-title: "Self-reported LLM usage and outcomes on a data science project: Evidence from a Canadian undergraduate data science course"
+title: "Self-reported LLM usage and outcomes on a data science project: Evidence from two undergraduate data science courses"
 author: 
@@
   - Nathalie Moon
-thanks: "Code and data are available at: https://github.com/lcarnegie/llms-achievement.
-We thank Tiffany Timbers and attendees at JSM 2024 for helpful suggestions.
-This research is was approved by the University of Toronto's Research Ethics Board as Protocol 47725.
-Comments can be sent to: rohan.alexander@utoronto.ca."
+thanks: "Code and data are available at: REDACTED.
+We thank attendees at JSM 2024 for helpful suggestions.
+This research was approved by the University of REDACTED's Research Ethics Board (Protocol 47725).
+Comments: REDACTED."
 date: today
@@
 We find some very mild evidence from this observational study that LLM usage may be associated with better scores, but our main conclusion is that there is no clear relationship between more extensive LLM usage and the student's mark.
-Despite the classroom setting used for evaluation, the particular activity of interest is similar to the work done by professional data scientists.
+Despite the classroom setting used for evaluation, the task of interest is similar to the work done by professional data scientists.
+Comparing self-reported usage, there was a considerable increase between the class in January-April 2024 where 41 per cent of students self-reported extensive LLM usage and the class in September-December 2024 where 69 per cent reported extensive LLM usage.
 Our finding suggests the need for more extensive work evaluating how LLMs can be integrated into the data science workflow in a way that provides value."
@@
 library(bayesplot)
+library(tidybayes)
 
@@
   read_parquet(here("data/analysis_data.parquet"))
+
+sta304_data <-
+  read_parquet(here("data/sta304_analysis_data_classified_cleaned.parquet"))
+
+# Read in models
+fit1 <-
+  readRDS(file = here("models/fit_one_inflated.rds"))
+fit2 <-
+  readRDS(file = here("models/fit_one_inflated_gpa.rds"))
+fit3 <-
+  readRDS(file = here("models/fit_one_inflated_sta304.rds"))
+fit4 <-
+  readRDS(file = here("models/fit_one_inflated_gpa_sta304.rds"))
+fit5 <-
+  readRDS(file = here("models/fit_one_inflated_gpa_english_sta304.rds"))
+fit6 <-
+  readRDS(file = here("models/fit_one_inflated_gpa_both.rds"))
 ```
@@
 To stay current, data scientists must continually update their knowledge of best practices for transparency and reproducibility, foster a culture which values different perspectives and accountability, and critically consider the broader impact of data-driven decisions.
-Recent advances, such as the wide release of user-friendly Large Language Models (LLMs), particularly OpenAI’s ChatGPT, have transformed the data science toolkit.
+Recent advances, such as the wide release of user-friendly Large Language Models (LLMs), particularly OpenAI's ChatGPT, have transformed the data science toolkit.
 These powerful tools for natural language processing and generation influence various aspects of data analysis and automation, further emphasizing the need for trustworthy practices.
@@
 ChatGPT's public release on November 30, 2022, brought LLMs into the mainstream conversation.
-LLMs have quickly been put to use in both industry and academia.
+LLMs have quickly been used in both industry and academia.
 By now many people, especially educators and students, have some experience with LLMs in both personal and professional contexts.
@@
 For instance, there is considerable interest in the potential of chatbots to act as personalized tutors for students [@fulgencio2024developing; @afzal2024tailoring].
-ChatGPT has also been the catalyst for many interesting and important conversations around academic integrity [@Eke2023], the development of critical thinking skills [@thiga2024generative], and what effective learning looks like [@BaidooAnu_2023].
-
-The tasks involved in a trustworthy data science workflow can be generally broken down into a number of key competencies [@Adhikari2021Interleaving; @Gibbs2021Building].
+ChatGPT has also been the catalyst for many interesting and important conversations around academic integrity [@Eke2023], the development of critical thinking skills [@thiga2024generative], and what effective learning looks like [@BaidooAnu_2023; @Bastani2024].
+
+The tasks involved in a trustworthy data science workflow can be generally broken down into a few key competencies [@Adhikari2021Interleaving; @Gibbs2021Building].
 One is programming, which is done when cleaning, analyzing, and visualizing data often using programming languages like R or Python.
@@
 In terms of computer programming, @Peng_2023 found a positive impact of GitHub Copilot (an LLM-powered programming assistant) on productivity.
-Specifically, in an experiment involving 95 freelance programmers, they found that a treatment group of programmers with access to GitHub Copilot completed a standardized programming task 56% faster than the control group.
+Specifically, in an experiment involving 95 freelance programmers, they found that a treatment group of programmers with access to GitHub Copilot completed a standardized programming task 56 per cent faster than the control group.
 Programmers with less experience saw the greatest improvements in productivity.
-
 @DellAcqua_2023, focusing on management consulting tasks, provide evidence that LLMs can improve writing productivity.
-In a field experiment involving consultants from Boston Consulting Group they found that the use of OpenAI's GPT-4 led to a 25% increase in delivery speed of business tasks, most of which involved some writing, as well as a 40% increase in human-rated performance on those tasks.
-Similar to computer programming, these productivity increases were most pronounced for those with below average performance, with their output increasing by 43%.
+In a field experiment involving consultants from Boston Consulting Group they found that the use of OpenAI's GPT-4 led to a 25 per cent increase in delivery speed of business tasks, most of which involved some writing, as well as a 40 per cent increase in human-rated performance on those tasks.
+Like computer programming, these productivity increases were most pronounced for those with below average performance, with their output increasing by 43 per cent.
 
@@
 These outcomes could negatively affect students' effective learning of data science.
-
 @Ellis_Slade_2023 take a more optimistic perspective and argue that LLMs are just another technology that will impact statistics and data science education.
@@
 With that in mind, they suggest that data-scientists-in-training should shift their self-perspectives from primarily being an "analyst" to primarily being a "product manager" responsible for strategic oversight of the analysis carried out by LLMs.
-
+@lehmann2025aimeetsclassroomlarge found the effect of LLMs on learning was nuanced and depended on the way the LLM was used and the level of prior knowledge.
 At the high school level @Lazar_2023 conducted an informal survey of secondary school teachers and students on their opinions of ChatGPT, and found that LLMs could help creativity, provide academic support when teachers were unavailable, and model certain types of writing well.
@@
 However, they also found that many students lacked confidence in using AI for academic purposes.
-In particular, only 11% "strongly agreed" that they know how to use AI to improve their writing.
+Only 11 per cent "strongly agreed" that they know how to use AI to improve their writing.
 Students had nuanced views on appropriate AI use.
@@
 In this paper we are interested in better understanding LLMs as a tool for producing trustworthy data science.
-We study how they were used by students in an upper-year undergraduate data science course and whether students who used LLMs tended to have higher scores than those who did not.
-
-To understand the current state of LLMs as a tool for trustworthy data science, this paper focuses on the association between student academic performance and their LLM usage.
+We study how they were used by students in two upper-year undergraduate data science courses -- one in Winter 2024 (January to April) and another in Fall 2024 (September to December) -- and whether students who used LLMs tended to have higher scores than those who did not.
+We focus on the association between student academic performance and their LLM usage.
 Specifically, we examine the relationship between the mark a student got on a final paper and self-reported measures of student LLM usage, as well as student attitudes toward LLMs in general.
-This is based on students' final papers and a survey, conducted in a third-year undergraduate data science course at the University of Toronto.
+This is based on students' final papers and a survey, conducted in third-year undergraduate data science courses at the University of REDACTED.
 By examining how students interact with and perceive LLMs as tools, and how these variables translate into student outcomes, current practice with regard to LLM integration in data science can be better understood, leading to better recommendations for their future development.
+
+We find some mild association between the mark that a respondent received and their self-reported LLM usage.
+And this is especially the case for respondents that were not native English speakers.
+More concretely, there was a considerable increase in the number of respondents who self-reported extensive LLM usage and also those who considered LLM usage to be appropriate and ethical for school, when we compared the results in the class at the start of 2024 with the class at the end of 2024.
+Taken together it is clear that considerable additional work is needed understanding the appropriate use of generative AI in statistics and data science pedagogy.
 
@@
 
-All data are from the cohort of students taking STA302 "Methods of Data Analysis I" in the Winter 2024 semester at the University of Toronto.
-This course had 275 students initially enrolled which, reflecting a normal rate of attrition for undergraduate statistics courses at the University of Toronto, reduced to 154 students by the end of the semester.
-Assessment was heavily based on three papers submitted over the course of the 12 week semester.
-
-The student marks that we analyze are based only on the final paper, which is done individually.
+All data are from students taking STA302 "Methods of Data Analysis I" in the Winter 2024 semester (January to April 2024), or students taking STA304 "Surveys, Sampling, and Observational Data" in the Fall 2024 semester (September to December 2024), at the University of REDACTED.
+STA302 had 275 students initially enrolled which reduced to 154 students by the end of the semester.
+Similarly, STA304 had 275 students initially, reduced to 196 students by the end of the semester.
+These reductions reflect a normal rate of attrition for undergraduate statistics courses at the University of REDACTED.
+Assessment in both courses was heavily based on three papers submitted over the course of the 12 week semester.
+
+The student marks that we analyze are based only on the final paper, which was done individually.
 By this stage, uninterested students have typically dropped the course, and students are familiar with course expectations.
-A typical paper submission is 10-20 pages, and requires students to conduct original research to answer a research question of interest to them.
+A typical paper submission was 10-20 pages, and required students to conduct original research to answer a research question of interest to them.
 It reflects the skills typically used by a professional data scientist.
 Students are expected to develop a research question of interest to them, identify or collect data to answer the question, conduct statistical analysis, and write a short paper.
-Examples of final papers (shared with consent) include: @hannahyu; @emilysu; and @bennyrochwerg.
+Examples of final papers (shared with consent) include: REDACTED; REDACTED; and REDACTED.
 
@@
 Each paper has the same basic structure and expectations.
-Before the final paper is due, students have received feedback on all their previous work in the class (including their past papers) and there is an optional two-day period of peer review.
+Before the final paper is due, students have received feedback on all their previous work in the class (including their past papers) and there is an optional period of peer review.
 
@@
 Throughout the semester students were encouraged to use LLMs.
-Formal instruction was provided twice during the semester.
+In STA302 formal instruction in LLM usage was provided twice during the semester.
 The first was a masterclass taught by a computer science faculty member on the ethics of using LLMs (see @Horton2024 for details).
 The second was a masterclass taught by a TA on writing with LLMs.
+In STA304 formal instruction was provided by instructor-demonstration of efficient LLM usage for coding and writing.
 
 Data were collected from students through an optional end-of-course survey.
-[Appendix -@sec-survey-details] details the questions asked in the survey.
-Whether or not they consented to their data being used, all respondents received a 1% increase in their final course grade for their participation.
+[Appendix -@sec-survey-details] details the questions asked in the survey for STA302.
+The STA304 survey was almost identical, with the one difference discussed later.
+Whether or not they consented to their data being used, all respondents received a 1 per cent increase in their final course grade for filling out the survey.
 Consenting responses were then matched to their final paper mark, as well as the GitHub repository for their final paper.
@@
 
-There were 146 responses to the survey.
+In STA302, there were 146 responses to the survey.
 Of these, 119 respondents provided authorization for their data to be collected and used.
@@
 Of those, 15 respondents did not include a statement on LLM usage in the README of the GitHub repository of their final paper, leaving 100 responses.
-Finally 7 of those respondents did not provide a usable GPA response.
-
-All but one respondent completed the survey within 45 minutes (@fig-response-time-and-gpa-1).
-That one respondent took more than 4,000 minutes to complete the survey, suggesting they took a break while filling it out.
-After that respondent is removed from the analysis the average time to complete the survey was 11 minutes, and the standard deviation was 8 minutes.
-This leaves 92 respondents that were of use and were merged based on student name.
+Finally, seven of those respondents did not provide a usable GPA response.
+
+In STA304, there were 183 responses to the survey.
+Of these, 153 respondents provided authorization for their data to be collected and used.
+Three of those respondents submitted the survey twice and their second response was removed.
+One student submitted their final paper too late to be of use, and so their survey was removed.
+Nine respondents did not include a statement on LLM usage in the README of the GitHub repository of their final paper.
+One respondent did not provide a usable GPA response.
+Finally, one respondent failed an attention check question halfway through the survey "Please select the option blue".
+
+In STA302, all but one respondent completed the survey within 60 minutes.
+After that respondent was removed from the analysis the average time to complete the survey was 11 minutes, and the standard deviation was 8 minutes (@fig-response-time-and-gpa-1).
+This left 92 respondents from STA302 that were of use and were merged based on student name.
 
@@
 #| fig-cap: Distribution of respondents' survey response times and self-reported GPA
-#| fig-subcap: ["Distribution of survey response times","\"What is your GPA?\""]
-#| layout-ncol: 1
+#| fig-subcap: ["STA302 distribution of survey response times","STA302 \"What is your GPA?\"","STA304 distribution of survey response times","STA304 \"What is your GPA?\""]
+#| layout-ncol: 2
 
@@
   # standard_dev = sd(mins_complete/60))
-  ggplot(aes(x = mins_complete/60)) +
+  ggplot(aes(x = mins_complete / 60)) +
   geom_histogram(binwidth = 1) +
@@
   theme_minimal()
-```
-
-There is a wide distribution of self-reported GPAs (@fig-response-time-and-gpa-2).
-The majority of responses cluster around a B (3.0 out of 4.0), and the average is 3.06 with a standard deviation of 0.55.
-One factor that may affect the range is that the course is required for programs in the Statistics, Mathematics and Computer Science Departments.
+
+sta304_data |>
+  # summarize(mean = mean(mins_complete/60),
+  # standard_dev = sd(mins_complete/60))
+  ggplot(aes(x = mins_complete / 60)) +
+  geom_histogram(binwidth = 1) +
+  labs(x = "Time to complete survey (minutes)", y = "Frequency") +
+  theme_minimal()
+
+sta304_data |>
+  # summarize(mean = mean(what_is_your_gpa, na.rm = TRUE),
+  #           standard_dev = sd(what_is_your_gpa, na.rm = TRUE))
+  ggplot(aes(x = what_is_your_gpa)) +
+  geom_histogram(binwidth = 0.1) +
+  labs(x = "Self-reported GPA", y = "Frequency") +
+  theme_minimal()
+```
+
+In STA304 three students took longer than one hour to complete the survey and were removed.
+After those respondents were removed the average time to complete the survey was 10 minutes and the standard deviation was 8 minutes (@fig-response-time-and-gpa-2).
+This left 131 respondents that were of use and were merged based on name.
+
+In STA302 there was a wide distribution of self-reported GPAs (@fig-response-time-and-gpa-2).
+Most responses clustered around a B (3.0 out of 4.0), and the average was 3.06 with a standard deviation of 0.55.
+One factor that may have affected the range is that the course is required for programs in the Statistics, Mathematics and Computer Science Departments.
 While there was no reason for the respondents to not report the truth, self-reported GPAs also introduce the possibility of reporting bias.
 For instance, respondents may have provided their cumulative GPA, their most recent term's GPA, or could have misreported it entirely.
-
-Students from a range of years took the course, however the majority of respondents were in their 3rd or 4th year of study (@tbl-year-dist).
-The course's prerequisite two-course sequence is typically completed by students in their second year, would make it difficult to take this course earlier than their 3rd year.
+The mean and standard deviation of self-reported GPA was the same in STA304 (@fig-response-time-and-gpa-4).
+
+Students from a range of years took both courses, but most respondents were in their 3rd or 4th year of study (@tbl-year-dist).
+The prerequisite two-course sequence is typically completed by students in their second year, would make it difficult to take either course earlier than their 3rd year.
 
@@
 #| tbl-cap: \"What year are you?\"
+#| tbl-subcap: ["STA302 responses (Winter 2024)","STA304 responses (Fall 2024)"]
 
@@
   kable()
-```
-
-Respondents had a varied self-perception of their coding and writing abilities (@fig-selfperception).
-Most respondents believe it is important to be good at writing, but many are either indifferent or do not like to write (@fig-selfperception-1).
+
+sta304_data |>
+  count(what_year_are_you) |>
+  pivot_wider(
+    names_from = what_year_are_you,
+    values_from = n,
+    values_fill = list(n = 0)
+  ) |>
+  kable()
+```
+
+Respondents from STA302 had a varied self-perception of their coding and writing abilities (@fig-selfperception).
+Most respondents believed it was important to be good at writing, but many were either indifferent or do not like to write (@fig-selfperception-1).
 Respondents also do not find writing to be particularly easy, which could be associated with the reported relative antipathy toward writing.
@@
 #| warning: false
-#| fig-cap: Self-perception of coding and writing abilities
+#| fig-cap: Self-perception of coding and writing abilities in STA302 (Winter 2024)
 #| fig-subcap: ["\"Please rate how much each statement describes you, on a scale from 'This is very different to me' to 'This is a lot like me'\"", "\"Please rate how much each statement describes you, on a scale from 'This is very different to me' to 'This is a lot like me'\""]
@@
 Respondent self-perceptions regarding coding proficiency and importance varied (@fig-selfperception-2).
-There is strong consensus on the perceived importance of coding skills.
-However, respondents' self-assessed coding abilities and enjoyment are more heterogeneous, with a substantial proportion reporting moderate rather than high levels of ease and enjoyment in coding tasks.
-
-Overall there was a moderate level of confidence among respondents in their overall coding ability, willingness to share code, and capacity to identify errors (@fig-selfperception-2).
+There was strong consensus on the perceived importance of coding skills.
+However, respondents' self-assessed coding abilities and enjoyment were more heterogeneous, with a substantial proportion reporting moderate rather than high levels of ease and enjoyment in coding tasks.
+Overall, there was a moderate level of confidence among respondents in their overall coding ability, willingness to share code, and capacity to identify errors.
 Notably, respondents express slightly higher confidence in detecting their own coding mistakes compared to general coding ability or code sharing.
@@
 
-The majority of respondents were at least "somewhat familiar" with generative AI such as ChatGPT (@tbl-ai-familiarity-usage-1).
-Though respondents' self-perceptions around writing and coding were varied, there was strong consensus that the use of generative AI tools is appropriate within an academic setting (@tbl-ai-familiarity-usage-2).
-Most respondents who selected "It depends" generally found artificial intelligence tools to be appropriate, though with certain guidelines and rules governing their use.
+Similar results were also found in STA304 (@fig-sta304-selfperception).
+
+```{r fig-sta304-selfperception}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: Self-perception of coding and writing abilities in STA304 (Fall 2024)
+#| fig-subcap: ["\"Please rate how much each statement describes you, on a scale from 'This is very different to me' to 'This is a lot like me'\"", "\"Please rate how much each statement describes you, on a scale from 'This is very different to me' to 'This is a lot like me'\""]
+
+# Writing
+writing_perceptions_long <-
+  sta304_data |>
+  select(
+    writing_is_easy_for_me,
+    i_like_to_write,
+    i_believe_it_is_important_to_be_a_good_writer,
+    when_i_edit_it_is_easy_for_me_to_catch_my_mistakes,
+    i_feel_confident_sharing_my_writing,
+    i_am_confident_in_my_overall_writing_ability
+  ) |>
+  pivot_longer(cols = everything(),
+               names_to = "statements",
+               values_to = "response") |>
+  mutate(
+    statements = case_when(
+      statements == "writing_is_easy_for_me" ~ "\"Writing is easy for me.\"",
+      statements == "i_like_to_write" ~ "\"I like to write.\"",
+      statements == "i_believe_it_is_important_to_be_a_good_writer" ~ "\"I believe it is important\nto be a good writer.\"",
+      statements == "when_i_edit_it_is_easy_for_me_to_catch_my_mistakes" ~ "\"When I edit it is easy for\nme to catch my mistakes.\"",
+      statements == "i_feel_confident_sharing_my_writing" ~ "\"I feel confident\nsharing my writing.\"",
+      statements == "i_am_confident_in_my_overall_writing_ability" ~ "\"I am confident in my\noverall writing ability.\"",
+      TRUE ~ statements
+    ),
+    response = case_when(
+      response == "\"This is a lot like me\"" ~ "A lot like me",
+      response == "\"This is very different to me\"" ~ "Very different\nfrom me",
+      response == "\"This somewhat describes me\"" ~ "Somewhat describes me",
+      TRUE ~ response
+    )
+  )
+
+writing_perceptions_long |>
+  ggplot(aes(x = response, fill = response)) +
+  geom_bar(position = "stack", stat = "count") +
+  facet_wrap(vars(statements)) +
+  labs(x = "Response", y = "Number of repondents", fill = "Response:") +
+  scale_fill_brewer(palette = "Set1") +
+  theme_minimal() +
+  theme(axis.text.x = element_blank(), legend.position = "bottom")
+
+# Coding
+coding_perceptions_long <-
+  sta304_data |>
+  select(
+    coding_is_easy_for_me,
+    i_like_to_code,
+    i_believe_it_is_important_to_be_a_good_coder,
+    when_i_check_my_code_it_is_easy_for_me_to_catch_my_mistakes,
+    i_feel_confident_sharing_my_code,
+    i_am_confident_in_my_overall_coding_ability
+  ) |>
+  pivot_longer(cols = everything(),
+               names_to = "statements",
+               values_to = "response") |>
+  mutate(
+    statements = case_when(
+      statements == "coding_is_easy_for_me" ~ "\"Coding is easy for me\"",
+      statements == "i_like_to_code" ~ "\"I like to code\"",
+      statements == "i_believe_it_is_important_to_be_a_good_coder" ~ "\"I believe it is important\n to be a good coder\"",
+      statements == "when_i_check_my_code_it_is_easy_for_me_to_catch_my_mistakes" ~ "\"When I check my code, it is\n easy for me to catch my\nmistakes\"",
+      statements == "i_feel_confident_sharing_my_code" ~ "\"I feel confident\n sharing my code\"",
+      statements == "i_am_confident_in_my_overall_coding_ability" ~ "\"I am confident in\n my overall coding ability\"",
+      TRUE ~ statements
+    ),
+    response = case_when(
+      response == "\"This is a lot like me\"" ~ "A lot like me",
+      response == "\"This is very different to me\"" ~ "Very different\nfrom me",
+      response == "\"This somewhat describes me\"" ~ "Somewhat describes me",
+      TRUE ~ response
+    )
+  )
+
+coding_perceptions_long |>
+  ggplot(aes(x = response, fill = response)) +
+  geom_bar(position = "stack", stat = "count") +
+  facet_wrap(vars(statements)) +
+  labs(x = "Response", y = "Number of repondents", fill = "Response:") +
+  scale_fill_brewer(palette = "Set1") +
+  theme_minimal() +
+  theme(axis.text.x = element_blank(), legend.position = "bottom")
+```
+
+Though STA302 respondents' self-perceptions around writing and coding were varied, there was strong consensus (79 per cent) that the use of generative AI tools was appropriate within an academic setting (@tbl-ai-familiarity-usage-1).
+Most respondents who selected "It depends" (12 per cent) generally found in comments that artificial intelligence tools to be appropriate, though with certain guidelines and rules governing their use.
+However, in STA304, almost all students (87 per cent) felt that it was ethical and appropriate to use generative AI for school and only 4 per cent felt that it was inappropriate.
+STA302 ran from January to April 2024, while STA304 ran from September to December 2024.
 
@@
 #| warning: false
-#| tbl-cap: Familiarity with, and appropriateness of, generative AI
-#| tbl-subcap: ["'How familiar are you with using generative AI tools such as OpenAI's ChatGPT or equivalents?'","'To what extent do you think using generative AI tools such as ChatGPT by OpenAI (or equivalents) is ethical and appropriate for schoolwork?'"]
-
-all_data |>
-  count(ai_familiarity) |>
-  kable(col.names = c("Extent of AI familiarity", "Students"))
+#| tbl-cap: "'To what extent do you think using generative AI tools such as ChatGPT by OpenAI (or equivalents) is ethical and appropriate for schoolwork?'"
+#| tbl-subcap: ["Responses from STA302 (Winter 2024)", "Responses from STA304 (Fall 2024)"]
 
@@
   count(ai_schoolwork_appropriate) |>
-  kable(col.names = c("Ethical & Appropriate for school?", "Students"))
-```
-
-To understand the role of LLMs in learning, respondents identified their usage in a more granular way by selecting various pre-defined use cases in the survey (@fig-AI-use-and-usefulness-1).
+  mutate(percentage = n / sum(n) * 100,
+         percentage = round(percentage, 0)) |> 
+  kable(col.names = c("Ethical & Appropriate for school?", "Number", "Percentage"))
+
+sta304_data |>
+  count(ai_schoolwork_appropriate) |>
+  mutate(percentage = n / sum(n) * 100,
+         percentage = round(percentage, 0)) |> 
+  kable(col.names = c("Ethical & Appropriate for school?", "Number", "Percentage"))
+```
+
+<!-- To understand the role of LLMs in learning, respondents identified their usage in a more granular way by selecting various pre-defined use cases in the survey (@fig-AI-use-and-usefulness-1).
 Technical questions and explaining concepts were the two top use cases among respondents.
@@
 Just less than half reported that they used it to write paper content, which could suggest that respondents do not feel confident using it to improve writing.
+-->
+
+Respondents were also asked to rate the helpfulness of LLMs on various tasks assigned during the course on a 4-point scale of "Did not use" to "Very Helpful" (@fig-AI-use-and-usefulness).
+To simplify the presentation, responses were grouped into two main categories: "Less Helpful" and "More Helpful." The "Less Helpful" category combines responses where students found the AI either "Not helpful" or did not use it for the task, while the "More Helpful" category includes responses where the LLM was considered "Somewhat helpful" or "Very helpful".
+
+Respondents differed in terms of how they used LLMs in STA302 (@fig-AI-use-and-usefulness-1).
+While most tasks were roughly split between respondents finding LLMs helpful or not, respondents found them most helpful in generating code.
+In the context of the course, this mostly meant generating R code for transforming, analyzing, and visualizing data.
+To a lesser extent, respondents also found LLMs to be helpful in improving the existing writing they had, while at the same time not favoring it for writing content from scratch.
+This was similar to the results found for STA304 (@fig-AI-use-and-usefulness-2), although there was consistently increased usage for all four aspects.
 
@@
 #| warning: false
-#| fig-cap: "Use and usefulness of generative AI"
-#| fig-subcap: ["\"If you have used generative AI tools such as OpenAI's ChatGPT or equivalents, in what ways have you used it (select all that apply)?\"", "\"How helpful did you find generative AI tools such as ChatGPT by OpenAI (or equivalents) for each component of STA302?\""]
-
-# Use
-all_data |>
-  select(how_have_you_used) |>
-  separate_rows(how_have_you_used, sep = ";") |>
-  mutate(
-    how_have_you_used = case_when(
-      how_have_you_used == "check grammar mistake" ~ "Other",
-      how_have_you_used == "debugging code" ~ "Other",
-      how_have_you_used == "mental health support" ~ "Other",
-      how_have_you_used == "Making practice questions for me with increasing difficulty" ~ "Other",
-      how_have_you_used == "Generating test questions" ~ "Other",
-      TRUE ~ how_have_you_used
-    )
-  ) |> 
-  count(how_have_you_used) |>
-  filter(how_have_you_used != "") |>
-  arrange(desc(n)) |>
-  ggplot(aes(x = n, y = reorder(how_have_you_used, n))) +
-  geom_bar(stat = "identity") +
-  labs(x = "Number of respondents", y = "Type of usage") +
-  theme_minimal()
+#| fig-cap: "\"How helpful did you find generative AI tools such as ChatGPT by OpenAI (or equivalents) for different components?\""
+#| fig-subcap: ["Responses from STA302 (Winter 2024)", "Responses from STA304 (Fall 2024)"]
+
+# # Use
+# all_data |>
+#   select(how_have_you_used) |>
+#   separate_rows(how_have_you_used, sep = ";") |>
+#   mutate(
+#     how_have_you_used = case_when(
+#       how_have_you_used == "check grammar mistake" ~ "Other",
+#       how_have_you_used == "debugging code" ~ "Other",
+#       how_have_you_used == "mental health support" ~ "Other",
+#       how_have_you_used == "Making practice questions for me with increasing difficulty" ~ "Other",
+#       how_have_you_used == "Generating test questions" ~ "Other",
+#       TRUE ~ how_have_you_used
+#     )
+#   ) |> 
+#   count(how_have_you_used) |>
+#   filter(how_have_you_used != "") |>
+#   arrange(desc(n)) |>
+#   ggplot(aes(x = n, y = reorder(how_have_you_used, n))) +
+#   geom_bar(stat = "identity") +
+#   labs(x = "Number of respondents", y = "Type of usage") +
+#   theme_minimal()
 
@@
   select(
-    weekly_quiz,
-    weekly_mini_essay,
+    # weekly_quiz,
+    # weekly_mini_essay,
     papers_generating_ideas,
@@
   theme(axis.text.y = element_blank(), legend.position = "bottom")
-```
-
-Respondents were also asked to rate the helpfulness of LLMs on various tasks assigned during the course on a 4-point scale of "Did not use" to "Very Helpful" (@fig-AI-use-and-usefulness-2).
-To simplify the presentation, responses were grouped into two main categories: "Less Helpful" and "More Helpful." The "Less Helpful" category combines responses where students found the AI either "Not helpful" or did not use it for the task, while the "More Helpful" category includes responses where the LLM was considered "Somewhat helpful" or "Very helpful".
-
-Respondents differed in terms of how they used LLMs in the course (@fig-AI-use-and-usefulness-2).
-While most tasks were roughly split between respondents finding LLMs helpful or not, respondents found them most helpful in generating code.
-In the context of the course, this mostly meant generating R code for transforming, analyzing, and visualizing data.
-To a lesser extent, respondents also found LLMs to be helpful in improving the existing writing they had, while at the same time not favouring it for writing content from scratch.
-
-Finally, one question asked students to elaborate on whether they thought generative AI tools such as ChatGPT were ethical and appropriate for schoolwork.
+
+####
+sta304_data |>
+  select(
+    generating_ideas,
+    writing_code,
+    # debugging_code,
+    writing_content,
+    improving_content
+  ) |>
+  pivot_longer(cols = everything(),
+               names_to = "task",
+               values_to = "response") |>
+  mutate(
+    # Simplify response categories and rename tasks
+    response_simplified = case_when(
+      response %in% c(
+        "\"Not helpful\"",
+        "\"I did not use generative AI for this component\""
+      ) ~ "Less Helpful",
+      response %in% c("\"Somewhat helpful\"", "\"Very helpful\"") ~ "More Helpful",
+      TRUE ~ "Unknown"  # Handle unexpected values
+    ),
+    task = case_when(
+      task == "generating_ideas" ~ "Generating Ideas",
+      task == "writing_code" ~ "Writing Code",
+      task == "debugging_code" ~ "Debugging Code",
+      task == "writing_content" ~ "Writing Content",
+      task == "improving_content" ~ "Improving Content",
+      TRUE ~ "Other"  # Handle any unexpected task names
+    )
+  ) |>
+  group_by(task, response_simplified) |>
+  summarise(count = n(), .groups = 'drop') |>
+  ggplot(aes(x = response_simplified, y = count, fill = response_simplified)) +
+  geom_bar(stat = "identity") +
+  facet_wrap(~ task, scales = "free_y") +
+  labs(y = "Number of respondents", x = "", fill = "Response type: ") +
+  theme_minimal() +
+  coord_flip() +
+  scale_fill_brewer(palette = "Set1") +
+  theme(axis.text.y = element_blank(), legend.position = "bottom")
+```
+
+One question asked students to elaborate on whether they thought generative AI tools such as ChatGPT were ethical and appropriate for schoolwork.
 This was an open-response question.
-To provide a sense of the responses, we used Anthropic's Claude 3.5 Sonnet model (as at 5 August 2024) to summarize the comments and it provided:
+To provide a sense of the responses from STA302, we used Anthropic's Claude 3.5 Sonnet model (as at 5 August 2024) to summarize the comments and it provided:
 
@@
 
-As at 14 August 2024, the default Anthropic setting is that they do not use input data for training and so these student responses should have entered future training datasets [@anthropic].
+As at 14 August 2024, the default Anthropic setting is that they do not use input data for training and so these student responses should not have entered future training datasets [@anthropic].
+
+We did a similar exercise for the responses from STA304, however using OpenAI's o3-mini model (as at 17 March 2025). It provided:
+
+> The responses indicate a general consensus that generative AI tools, such as ChatGPT, can be ethical and appropriate for schoolwork when used as a supplement rather than a substitute for genuine effort.
+Many liken their use to that of calculators or grammar checkers, highlighting benefits like idea generation, code debugging, and clarifying complex concepts.
+However, a common caution is that overreliance on these tools—especially for tasks that require critical thinking and original work—could undermine learning.
+Transparency, proper citation, and maintaining a balance between AI assistance and personal input are repeatedly emphasized, ensuring that the tools aid understanding without replacing the student’s own analytical process.
+
+Finally, in STA304 only, we asked about whether the student was an international student and whether they spoke English as their first language (@tbl-sta304-international-esl).
+It is notable that most respondents are international students (@tbl-sta304-international-esl-1) and did not speak English as their first language (@tbl-sta304-international-esl-2).
+
+```{r tbl-sta304-international-esl}
+#| message: false
+#| echo: false
+#| warning: false
+#| tbl-cap: "The background of respondents in STA304 only"
+#| tbl-subcap: ["'Are you an international student?'", "'Is English your native language (i.e. the language that you first spoke)?'"]
+
+sta304_data |> 
+  count(international_student) |> 
+  mutate(percentage = n / sum(n) * 100,
+         percentage = round(percentage, 0)) |> 
+  kable(col.names = c("International student?", "Number", "Percentage"))
+
+sta304_data |> 
+  count(native_english) |> 
+  mutate(percentage = n / sum(n) * 100,
+         percentage = round(percentage, 0)) |> 
+  kable(col.names = c("Native English speaker?", "Number", "Percentage"))
+```
 
@@
 
-Two other components were merged with the survey responses: self-reported LLM usage on the final paper, and final paper mark.
+Two other components were merged with the survey responses:
+
+1) self-reported LLM usage on the final paper, and 
+2) final paper mark.
 
 Students were encouraged to use LLMs to complete their papers.
-Each paper required the students to disclose their usage through a statement in the GitHub repository README for the paper.
+But each paper required the students to disclose their usage through a statement in the GitHub repository README for the paper.
 Even students who did not use generative AI at all were required to state this in the README.
-For students who did use generative AI, there was an additional requirement, where possible, that they save the logs of their usage in a txt file which was also included in their GitHub repository.
-
-Those README statements were gathered and parsed using OpenAI's ChatGPT 4o model (as at 26 July). The following prompt was used:
+For students who did use generative AI, there was an additional requirement, where possible, that they save the logs of their usage in a text file which was also included in their GitHub repository.
+
+Those README statements were gathered and parsed using OpenAI's ChatGPT 4o model.
+For the STA302 statements this was as at 26 July, 2024, while for the STA304 statements this was as at 16 March 2025.
+Between this time period there were underlying changes to the model despite no change in the model name.
+The following prompt was used:
 
@@
 All classifications were then manually checked for reasonableness.
+In the case of STA304, there were 31 cases where the statement provided by the student in the README was vague, and the LLM had classified it at "Somewhat".
+Manual inspection of text files provided by the students resulted in 19 of these being re-classified as "Extensive", 7 of them being re-classified as "Somewhat", and 5 of them being re-classified as "Minimal".
+None of them were re-classified as "None".
 
 We find a varied extent of self-reported LLM usage (@tbl-llm-usage).
-38 respondents were classified as having made extensive use of LLMs, while 26 were classified as having made somewhat use.
+In STA302, 38 respondents were classified as having made extensive use of LLMs, while 26 were classified as having made somewhat use.
 28 respondents were classified as having made minimal or no use of LLMs.
 In the analysis dataset we combine those two classifications because only a handful of respondents were classified as having minimal usage.
+In STA304, which occurred only six months later, we find that a vast majority (69 per cent) of students made extensive use of LLMs and only 12 per cent reported none or minimal usage.
+Partly this may be students having become more comfortable with the idea of reporting LLM usage but it also likely reflects a change in actual usage.
+Either way, the increase in the percentage of students reporting extensive LLM usage and the decrease in the percentage reporting none or minimal usage is notable and has important implications for statistics and data science pedagogy.
 
@@
 #| tbl-cap: Self-reported LLM usage in final paper
+#| tbl-subcap: ["Responses from STA302 (Winter 2024)", "Responses from STA304 (Fall 2024)"]
 
@@
   count(llm_usage) |> 
-  kable(col.names = c("Self-reported LLM usage", "Number"))
-```
-
-The third, and final, component is the mark, in percentages, on the final paper (@fig-mark-1).
-The overall mean was 78% and standard deviation was 17 percentage points.
-However there were considerable differences by the extent of LLM usage (@tbl-mark; @fig-mark-2).
+  mutate(percentage = n / sum(n) * 100,
+         percentage = round(percentage, 0)) |> 
+  kable(col.names = c("Self-reported LLM usage", "Number", "Percentage"))
+
+sta304_data |>
+  count(llm_usage) |> 
+  mutate(percentage = n / sum(n) * 100,
+         percentage = round(percentage, 0)) |> 
+  kable(col.names = c("Self-reported LLM usage", "Number", "Percentage"))
+```
+
+The third, and final, component is the mark, in percentages, on the final paper (@fig-mark).
+In STA302 the overall mean was 79 per cent and standard deviation was 16 percentage points (@fig-mark-1) while in STA304 the overall mean was 78 per cent and the standard deviation was 17 percentage points (@fig-mark-3).
+However, in both classes there were considerable differences by the extent of LLM usage (@fig-mark-2; @fig-mark-4; @tbl-mark-1; @tbl-mark-2).
 
@@
 #| fig-cap: Distribution of marks on final paper (in percentages)
-#| fig-subcap: ["Overall", "By self-reported LLM usage"]
+#| fig-subcap: ["STA302 Overall", "STA302 by self-reported LLM usage", "STA304 Overall", "STA304 by self-reported LLM usage"]
+#| layout-ncol: 2
+
+# all_data |>
+#   summarise(ave = mean(mark),
+#             std = sd(mark))
+# 
+# sta304_data |>
+#   summarise(ave = mean(mark),
+#             std = sd(mark))
 
@@
   theme_minimal()
+
+sta304_data |>
+  select(mark) |>
+  ggplot(aes(x = mark)) +
+  geom_histogram(binwidth = 0.05) +
+  scale_x_continuous(labels = label_percent()) +
+  labs(x = "Final paper mark (percentage)", y = "Count") +
+  theme_minimal()
+
+sta304_data |>
+  ggplot(mapping = aes(x = mark)) +
+  geom_histogram(binwidth = 0.01) +
+  scale_x_continuous(labels = label_percent()) +
+  labs(x = "Final paper mark (percentage)", y = "Count") +
+  facet_wrap(vars(llm_usage),
+             ncol = 2,
+             labeller = labeller(
+               llm_usage = c(
+                 "Extensive" = "Extensive LLM usage",
+                 "Minimal" = "Minimal LLM usage",
+                 "None" = "No LLM usage",
+                 "Somewhat" = "Some LLM usage"
+               )
+             )) +
+  theme_minimal()
 ```
@@
 #| tbl-cap: Self-reported LLM usage in final paper and final paper mark
+#| tbl-subcap: ["Responses from STA302 (Winter 2024)", "Responses from STA304 (Fall 2024)"]
 
@@
   kable(col.names = c("Self-reported LLM usage", "Mean", "Std dev"))
+
+sta304_data |>
+  group_by(llm_usage) |>
+  summarize(mean = mean(mark) |> round(2),
+            sd = sd(mark) |> round(2)) |>
+  kable(col.names = c("Self-reported LLM usage", "Mean", "Std dev"))
 ```
@@
 
-The goal of our modelling strategy is to better understand how a respondents' result on their final paper associates with their self-reported LLM usage.
-Students received their result on the final paper as a percentage between 0% and 100%.
-There were no students who got 0%, but 9 students got 100%.
-As such, we use one-inflated beta regression [@Ospina2012].
-Beta regression is commonly used when data are between 0 and 1 and is focused on estimating the mean and variance of the beta distribution.
-One approach is to consider a one-inflated distribution which is a mixture of focusing on those responses that were 1, and the usual Beta distribution aspect.
-
-The main variable of interest, `llm_usage` is a categorical variable with three possible values: "Extensive", "Somewhat", "None or minimal".
+The goal of our modelling is to better understand how a respondents' result on their final paper associates with their self-reported LLM usage.
+Students received their result on the final paper as a percentage between 0 per cent and 100 per cent.
+There were no students who got 0 per cent, but some students got 100 per cent.
+As such, after converting the percentage to a proportion, we use one-inflated beta regression [@Ospina2012].
+Beta regression is commonly used when data are between 0 and 1 and is focused on estimating the two shape parameters that govern the beta distribution.
+When there are some values that are 1, a variant -- one-inflated beta regression -- can be used, which is a mixture of focusing on whether a value was 1 or not, and then the usual Beta distribution aspect.
+
+The main predictor of interest, `llm_usage`, is a categorical variable with three possible values: "Extensive", "Somewhat", "None or minimal".
 We use "None or minimal" as the reference level and so the results are in relation to that level of self-reported LLM usage.
@@
 
-Define $y_i$ as the percentage that student $i$ received on the final paper:
-
-$$y_i \sim \mbox{Zero-One-Inflated Beta}\left(\mu_i, \phi_i, \mbox{zoi}_i, \mbox{coi}\right)$$
-
-Where:
-
-- $\mu_i$ is the mean of the beta distribution (location parameter)
-- $\phi_i$ is the precision parameter (controlling dispersion)
-- $\pi_{zoi_i}$ is the probability of $Y_i$ being exactly 0 or 1
-- $\pi_{coi_i}$ is the probability of $Y_i$ being in the continuous part of the distribution
-
-The model components are:
-
-1. Mean component:
-   $logit(\mu_i) = \beta_{0,\mu} + \beta_{1,\mu} \cdot \text{llm\_usage}_i + \beta_{2,\mu} \cdot \text{GPA}_i$
-
-2. Precision component:
-   $\log(\phi_i) = \beta_{0,\phi} + \beta_{1,\phi} \cdot \text{llm\_usage}_i + \beta_{2,\phi} \cdot \text{GPA}_i$
-
-3. Zero-one inflation component:
-   $logit(\pi_{zoi_i}) = \beta_{0,zoi} + \beta_{1,zoi} \cdot \text{llm\_usage}_i + \beta_{2,zoi} \cdot \text{GPA}_i$
-
-4. Continuous outcome inflation component:
-   $logit(\pi_{coi_i}) = \beta_{0,coi}$
-
-We estimate the model and explore the results using the statistical programming language `R` [@citeR], and the 
-`brms` [@brms], and 
-`modelsummary` [@modelsummary] 
-packages.
-Our code to adjust for the fact that there were no students who got 0% follows @buerkner and @Heiss2021.
-Model diagnostics are included in [Appendix -@sec-model-details].
-Finally, the piecewise approach is a common one for when there may be something different about the 0s or 1s, compared with the other responses.
-Another approach for dealing with a situation in which there are exact 0s or 1s is to slightly perturb the data.
-In our case, there is not much difference between getting 99% on a paper and getting 100%.
-When we estimated that perturbed data model using Beta regression the results were similar.
+If $y_i$ is the percentage that student $i$ received on the final paper (adjusted to a proportion), then our model is:
+$$y_i \sim \mbox{One-Inflated Beta}\left(\mu_i, \phi_i, \pi_{\mbox{zoi}_i}, \pi_{\mbox{coi}_i}\right)$$
+where:
+
+- $\mu_i$ (mean component) is the mean of the beta distribution, which accounts for marks between zero and one,
+- $\phi_i$ (precision component) is the precision of the beta distribution, which, again, accounts for marks between zero and one,
+- $\pi_{\mbox{zoi}_i}$ (zero-one inflation component) is a logistic regression model that looks at which of two groups a student is in: a) the group of students getting full or zero marks, or b) the group of students getting a mark between zero and one, and
+- $\pi_{\mbox{coi}_i}$ (continuous outcome inflation component) is used to distinguish between students that got full or zero marks (but in our case as there were no students that got zero we force this to be one).
+
+The main model specification is:
+
+1. Mean component: $\mbox{logit}(\mu_i) = \beta_{0,\mu} + \beta_{1,\mu} \cdot \text{llm\_usage}_i + \beta_{2,\mu} \cdot \text{GPA}_i$
+2. Precision component: $\log(\phi_i) = \beta_{0,\phi} + \beta_{1,\phi} \cdot \text{llm\_usage}_i + \beta_{2,\phi} \cdot \text{GPA}_i$
+3. Zero-one inflation component: $\mbox{logit}(\pi_{\mbox{zoi}_i}) = \beta_{0,\mbox{zoi}} + \beta_{1,\mbox{zoi}} \cdot \text{llm\_usage}_i + \beta_{2,\mbox{zoi}} \cdot \text{GPA}_i$
+4. Continuous outcome inflation component: $\mbox{logit}(\pi_{\mbox{coi}_i}) = \beta_{0,\mbox{coi}}$.
+
+We estimate the model and explore the results using the statistical programming language `R` [@citeR] and the 
+`brms` [@brms] package. Our code to adjust for the fact that there were no students who got zero follows @buerkner and @Heiss2021. Model diagnostics are included in [Appendix -@sec-model-details].
+
+Finally, while our approach of using zero-inflated beta regression is a common one, another approach for dealing with a situation in which there are exact zeros or ones would be to slightly perturb the data.
+In our case, that would likely have been defendable as there is not much difference between getting 99 per cent on a paper and getting 100 per cent.
+When we estimated that perturbed data model using beta regression the results were similar.
 
@@
 
-Our results are summarized in @tbl-model-results and @fig-model-results, with full output available in the Appendix. We draw on the `modelsummary` [@modelsummary] package to summarize the results.
+Our central results are summarized in @tbl-model-results and @fig-model-results, which use the `modelsummary` [@modelsummary] package.
+These focus on the main estimates i.e.
+those respondents with less than full marks.
+The full output is available in [Appendix -@sec-model-output].
 
@@
 #| tbl-cap: Selected coefficient estimates and mean absolute deviation (MAD)
-
-fit1 <-
-  readRDS(file = here("models/fit_one_inflated.rds"))
-fit2 <-
-  readRDS(file = here("models/fit_one_inflated_gpa.rds"))
 
@@
   list(
-    "Base model" = fit1,
-    "Self-reported GPA included" = fit2
+    "STA302" = fit1,
+    "STA302: GPA" = fit2,
+    "STA304" = fit3,
+    "STA304: GPA" = fit4,
+    "STA304: GPA & ESL" = fit5
   ),
@@
   output = "kableExtra",
-   kable_options = list(
-     booktabs = TRUE,
-     linesep = ""
-   ),
+  kable_options = list(booktabs = TRUE, linesep = ""),
   coef_map = c(
+    "b_Intercept" = "Intercept",
+    "b_phi_Intercept" = "Intercept (Phi)",
+    "b_zoi_Intercept" = "Intercept (ZOI)",
     "b_llm_usageSomewhat" = "LLM usage: Somewhat",
-     "b_llm_usageNoneorminimal" = "LLM usage: None or minimal",
+    "b_llm_usageExtensive" = "LLM usage: Extensive",
+    "b_phi_llm_usageSomewhat" = "LLM usage: Somewhat (Phi)",
+    "b_phi_llm_usageExtensive" = "LLM usage: Extensive (Phi)",
     "b_what_is_your_gpa" = "GPA",
-     "b_zoi_Intercept" = "Zero-one inflation intercept"
-   )
-)
+    "b_phi_what_is_your_gpa" = "GPA (Phi)",
+    "b_native_englishYes" = "English native: Yes",
+    "b_phi_native_englishYes" = "English native: Yes (Phi)"
+  )
+) |>
+  kableExtra::kable_styling(font_size = 7)
 ```
@@
 #| warning: false
+#| fig-height: 5
 #| fig-cap: Selected coefficient estimates and 90 per cent credibility intervals
@@
   list(
-    "Base model" = fit1,
-    "Including self-reported GPA" = fit2
+    "STA302" = fit1,
+    "STA302: GPA" = fit2,
+    "STA304" = fit3,
+    "STA304: GPA" = fit4,
+    "STA304: GPA & ESL" = fit5
   ),
@@
     "b_llm_usageSomewhat" = "LLM usage: Somewhat",
-     "b_llm_usageNoneorminimal" = "LLM usage: None or minimal",
-     "b_what_is_your_gpa" = "GPA"
-   ),
-   coef_omit = NULL
+    "b_llm_usageExtensive" = "LLM usage: Extensive",
+    "b_phi_llm_usageSomewhat" = "LLM usage: Somewhat (phi)",
+    "b_phi_llm_usageExtensive" = "LLM usage: Extensive (phi)",
+    "b_what_is_your_gpa" = "GPA",
+    "b_phi_what_is_your_gpa" = "GPA (phi)",
+    "b_native_englishYes" = "English native: Yes",
+    "b_phi_native_englishYes" = "English native: Yes (phi)"
+  )
 ) +
@@
   scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom")
-```
-
-@tbl-model-results and @fig-model-results present the results from two models estimating the association between LLM usage, GPA and final paper mark.
-The first model includes only LLM usage, while the second model also includes self-reported GPA.
-
-LLM usage estimates are in relation to "Extensive" LLM usage.
-In the base model, the coefficients suggest that less usage of LLMs was associated with lower scores.
-The second model retains this association, but with slightly smaller coefficients.
-For instance, the negative impact of “LLM usage: Somewhat” is slightly reduced from -0.35 to -0.29, indicating that part of the effect observed in the base model could be explained by the GPA variable.
-
-Including self-reported GPA in the second model considerably changes the  changes the interpretation of some coefficients.
-For example, the intercept value (in Zero-one inflation intercept) becomes much more negative in the GPA-included model, suggesting that the GPA significantly affects the zero-one inflation part of the model.
-At the same time, though, the essential relationship between LLM usage and final mark is retained.
-
-The overall fit of the model improves when GPA is included.
-There is an increase in R-squared from 0.095 to 0.478 and improvements in the ELPD and LOOIC metrics.
-This is also suggested by the LOO comparison between the two models (@tbl-loo in [Appendix -@sec-model-details]).
+  #theme(legend.position = "bottom") +
+  scale_y_discrete(limits=rev)
+```
+
+The initial model in @tbl-model-results is for STA302 and focuses on the association between final paper mark and LLM usage.
+The next model adds in self-reported GPA.
+Both of these have 92 observations.
+This same pattern is then repeated for STA304 and both those models have 131 observations.
+And the final model in @tbl-model-results additionally considers whether the respondent is a native English speaker.
+
+```{r}
+#| echo: false
+#| eval: false
+
+plogis(-2.1)
+plogis(-2.93)
+```
+
+To begin with, we can take advantage of the simple initial models for STA302 and STA304 to look at the ZOI intercept.
+For instance, for the STA302 model it is -2.10.
+Following @andrewagain, after undoing the logit transformation this is approximately 11%, which corresponds to the number of students in STA302 who got full marks on the final paper.
+The same holds for the STA304 model.
+
+LLM usage estimates are in relation to "None or minimal" LLM usage.
+In both the base STA302 and STA304 models, "Extensive" LLM usage was associated with slightly higher mean and precision estimates (@fig-model-results).
+But when GPA was added to the models the precision component became negative.
+That said, with 90 per cent credibility intervals there is considerable overlap with zero for many estimates.
+
+Holding all other variables constant, we can vary LLM usage from none or minimal to somewhat and then extensive (@fig-model-results-llms).
+There is an expected increase in the mark on the final paper as LLM usage increases in STA302 (@fig-model-results-llms-1), but it is less clear in STA304 (@fig-model-results-llms-2).
+It may have been that the overwhelming usage of LLMs at "extensive" levels in STA304, which occurred six months after STA302, was the reason for this seemingly different association in our data.
+
+```{r fig-model-results-llms}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-height: 5
+#| fig-cap: Comparing the effect of a hypothetical student with different LLM usage levels, all other variables constant
+#| fig-subcap: ["Based on responses from STA302 (Winter 2024)", "Based on responses from STA304 (Fall 2024)"]
+#| layout-ncol: 2
+
+fit2 |> 
+  linpred_draws(
+    newdata = data.frame(what_is_your_gpa = 3.0, llm_usage = unique(all_data$llm_usage)),
+    transform = TRUE
+  ) |> 
+  ggplot(aes(y = llm_usage, x = .linpred)) +
+  stat_halfeye(aes(fill = llm_usage), normalize = "xy") +
+  scale_fill_brewer(palette = "Set1", guide = "none") +
+  expand_limits(x = c(0.5, 0.9)) +
+  scale_x_continuous(labels = label_percent()) +
+  labs(y = NULL, x = "Final paper mark (in percentages)") +
+  theme_minimal()
+
+fit4 |> 
+  linpred_draws(
+    newdata = data.frame(what_is_your_gpa = 3.0, llm_usage = unique(all_data$llm_usage)),
+    transform = TRUE
+  ) |> 
+  ggplot(aes(y = llm_usage, x = .linpred)) +
+  stat_halfeye(aes(fill = llm_usage), normalize = "xy") +
+  scale_fill_brewer(palette = "Set1", guide = "none") +
+  expand_limits(x = c(0.5, 0.9)) +
+  scale_x_continuous(labels = label_percent()) +
+  labs(y = NULL, x = "Final paper mark (in percentages)") +
+  theme_minimal()
+```
+
+The different effect of LLM usage, by whether the students was a native English speaker, holding all other variables constant, can only be examined for STA304 (September-December 2024) as the question was not asked in the earlier STA302 (@fig-model-results-english).
+For the hypothetical student who does not speak English natively, where we changed nothing other than their LLM usage, going from none or minimal (@fig-model-results-english-1) to extensive (@fig-model-results-english-2) does appear to have been associated with considerably different marks based on our dataset.
+The distribution is smaller and the average is higher.
+
+```{r fig-model-results-english}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: Comparing the effect of a hypothetical student changing whether they were a native English speaker, all other variables constant
+#| fig-subcap: ["Holding GPA at 3.0 and LLM usage is none or minimal", "Holding GPA at 3.0 and LLM usage is extensive"]
+#| layout-ncol: 2
+
+fit5 |> 
+  linpred_draws(
+    newdata = data.frame(what_is_your_gpa = 3.0, 
+                         llm_usage = "None or minimal",
+                         native_english = unique(sta304_data$native_english)),
+    transform = TRUE
+  ) |> 
+  ggplot(aes(y = native_english, x = .linpred)) +
+  stat_halfeye(aes(fill = native_english), normalize = "xy") +
+  scale_fill_brewer(palette = "Set1", guide = "none") +
+  expand_limits(x = c(0.5, 0.9)) +
+  scale_x_continuous(labels = label_percent()) +
+  labs(y = NULL, x = "Final paper mark (in percentages)") +
+  theme_minimal()
+
+fit5 |> 
+  linpred_draws(
+    newdata = data.frame(what_is_your_gpa = 3.0, 
+                         llm_usage = "Extensive",
+                         native_english = unique(sta304_data$native_english)),
+    transform = TRUE
+  ) |> 
+  ggplot(aes(y = native_english, x = .linpred)) +
+  stat_halfeye(aes(fill = native_english), normalize = "xy") +
+  scale_fill_brewer(palette = "Set1", guide = "none") +
+  expand_limits(x = c(0.5, 0.9)) +
+  scale_x_continuous(labels = label_percent()) +
+  labs(y = NULL, x = "Final paper mark (in percentages)") +
+  theme_minimal()
+```
 
@@
 
-There is some very mild evidence from this observational study that LLM usage may be associated with better scores.
+There may be some mild evidence from this observational study that LLM usage may be associated with better scores.
 However, this work should mostly be treated as a direction for future research in this area.
@@
 1. LLMs could help enhance the trustworthiness of data science, but not without considerable further development, testing and integration in the data science process.
-2. Making general purpose chatbots more opinionated and knowledgeable about data science would considerably help, and 
-3. Alternative interaction approaches, beyond a chatbot interface, could be especially useful for students learning data science in particular.
+2. Making general purpose chatbots more opinionated and knowledgeable about data science would considerably help.
+3. Alternative interaction approaches, beyond a chatbot interface, could be especially useful for students learning data science, particularly for those who are not native English speakers.
 
@@
 Stratification of our dataset would result in small sample size, but nonetheless our data does provide some limited suggestive evidence that this effect may be present in data science.
-For instance, looking at the 28 students who received an A+ for the final paper, 13 of them had extensive LLM usage, whereas looking at the same number of worst performing students finds that only six of them had extensive LLM usage.
+For instance, considering STA302, looking at the 28 students who received an A+ for the final paper, 13 of them had extensive LLM usage, whereas looking at the same number of worst performing students finds that only six of them had extensive LLM usage.
 
 Along these lines, we have considered LLM usage for code and writing as equivalent, but they should actually have different impacts depending on student backgrounds.
-Distinguishing between native and second-language English speakers, and then focusing on differential LLM usage, could have added a great deal of nuance to the analysis.
+We were only able to distinguishing between native and second-language English speakers at a high level and a more detailed focus on differential LLM usage would add a great deal of nuance to the analysis.
 
@@
 
+In STA304, students were additionally asked about whether they were an international student "Are you an international student?", whether English was their native language "Is English your native language (i.e.
+the language that you first spoke)?", and an attention check was included "Please select the option "Blue"".
+
 \newpage
@@
 
-## Full Model Output
+## Full model output  {#sec-model-output}
 

### 2025-04-05 scripts/03-model.R

@@ -1,7 +1,7 @@
 #### Preamble ####
 # Purpose: Models
 # Author: Rohan Alexander
-# Date: 4 August 2024
+# Date: 17 March 2025
 # Contact: rohan.alexander@utoronto.ca
 # License: MIT
 # Pre-requisites: -
@@ -21,8 +21,8 @@ all_data <-
 
 
 # Change the reference level so it's in terms of none
-all_data <- 
-  all_data |> 
+all_data <-
+  all_data |>
   mutate(llm_usage = relevel(llm_usage, ref = "None or minimal"))
 
 
@@ -35,10 +35,12 @@ all_data <-
 
 # Using rstanarm
 fit1rstanarm <-
-  stan_betareg(mark_no_ones ~ llm_usage,
-               data = all_data,
-               link = "logit",
-               seed = 853)
+  stan_betareg(
+    mark_no_ones ~ llm_usage,
+    data = all_data,
+    link = "logit",
+    seed = 853
+  )
 
 fit2rstanarm <-
   stan_betareg(
@@ -76,12 +78,7 @@ fit2brms <- brm(
 # https://www.andrewheiss.com/blog/2021/11/08/beta-regression-guide/#special-case-2-one-inflated-beta-regression
 
 fit_one_inflated <- brm(
-  formula = bf(
-    mark ~ llm_usage,
-    phi ~ llm_usage,
-    zoi ~ llm_usage,
-    coi ~ 1
-  ),
+  formula = bf(mark ~ llm_usage, phi ~ llm_usage, zoi ~ llm_usage, coi ~ 1),
   family = zero_one_inflated_beta(),
   data = all_data,
   cores = 4,
@@ -114,3 +111,161 @@ saveRDS(fit2brms, file = "models/fit2brms.rds")
 saveRDS(fit_one_inflated, file = "models/fit_one_inflated.rds")
 
 saveRDS(fit_one_inflated_gpa, file = "models/fit_one_inflated_gpa.rds")
+
+
+
+
+
+#### Redo with STA304 ####
+#### Read data ####
+sta304_data <-
+  read_parquet(here("data/sta304_analysis_data_classified_cleaned.parquet"))
+
+# Change the reference level so it's in terms of none
+sta304_data <-
+  sta304_data |>
+  mutate(llm_usage = relevel(llm_usage, ref = "None or minimal"))
+
+
+#### Model data ####
+#### Simplified data ####
[…]

### 2025-04-07 paper/paper.qmd
Fix typos and update appendix

@@
 author: 
-  - Rohan Alexander
-  - Luca Carnegie
-  - Nathalie Moon
+  - REDACTED
+  - REDACTED
+  - REDACTED
 thanks: "Code and data are available at: REDACTED.
@@
 date-format: long
-abstract: "To help understand the effect of Large Language Models (LLMs) on data science practice we examine the extent to which self-reported LLM usage is correlated with the mark that a student gets on a final paper in a classroom data science setting.
-We find some very mild evidence from this observational study that LLM usage may be associated with better scores, but our main conclusion is that there is no clear relationship between more extensive LLM usage and the student's mark.
+abstract: "To help understand the effect of Large Language Models (LLMs) on data science practice we examine the extent to which self-reported LLM usage is correlated with the mark that a student received on a final paper in a classroom data science setting.
+We find some mild evidence from this observational study that LLM usage may be associated with better scores, especially for students who do not natively speak English.
+Comparing self-reported usage, there was a considerable increase between the class that occurred in January-April 2024 where 41 per cent of students self-reported extensive LLM usage and the class that occurred in September-December 2024 where 69 per cent reported extensive LLM usage.
 Despite the classroom setting used for evaluation, the task of interest is similar to the work done by professional data scientists.
-Comparing self-reported usage, there was a considerable increase between the class in January-April 2024 where 41 per cent of students self-reported extensive LLM usage and the class in September-December 2024 where 69 per cent reported extensive LLM usage.
-Our finding suggests the need for more extensive work evaluating how LLMs can be integrated into the data science workflow in a way that provides value."
+Our finding suggests the need for more extensive work evaluating how LLMs can be integrated into the data science workflow in a way that provides value in both the classroom and the workplace."
 format: pdf
@@
 Trustworthy data science is the practice of conducting data analysis in a transparent, ethical, and reliable manner.
-These principles are upheld in various ways, through ongoing education, adherence to professional norms and implementation of reproducible workflows.
+These principles are upheld in various ways including ongoing education, adherence to professional norms and implementation of reproducible workflows.
 To stay current, data scientists must continually update their knowledge of best practices for transparency and reproducibility, foster a culture which values different perspectives and accountability, and critically consider the broader impact of data-driven decisions.
@@
 @lehmann2025aimeetsclassroomlarge found the effect of LLMs on learning was nuanced and depended on the way the LLM was used and the level of prior knowledge.
-At the high school level @Lazar_2023 conducted an informal survey of secondary school teachers and students on their opinions of ChatGPT, and found that LLMs could help creativity, provide academic support when teachers were unavailable, and model certain types of writing well.
-But teachers were concerned about the potential of LLMs to limit students' learning in certain ways through over-reliance.
-Beyond academic integrity concerns, @Lazar_2023 found that teachers had similar concerns to @Valenzuela_2024 about de-skilling and an overall loss of agency in writing and critical thinking.
+At the high school level @Lazar_2023 found that LLMs could help creativity, provide academic support when teachers were unavailable, and model certain types of writing well.
+But teachers were found to have concerns about over-reliance, academic integrity, de-skilling, and an overall loss of agency in writing and critical thinking.
 
@@
 And this is especially the case for respondents that were not native English speakers.
-More concretely, there was a considerable increase in the number of respondents who self-reported extensive LLM usage and also those who considered LLM usage to be appropriate and ethical for school, when we compared the results in the class at the start of 2024 with the class at the end of 2024.
+More concretely, there was a considerable increase in the number of respondents who self-reported extensive LLM usage and also those who considered LLM usage to be appropriate and ethical for university, when we compared the results in the class at the start of 2024 with the class at the end of 2024.
 Taken together it is clear that considerable additional work is needed understanding the appropriate use of generative AI in statistics and data science pedagogy.
@@
 These reductions reflect a normal rate of attrition for undergraduate statistics courses at the University of REDACTED.
-Assessment in both courses was heavily based on three papers submitted over the course of the 12 week semester.
+Assessment in both courses was heavily based on three papers submitted over the course of the 12-week semester.
 
@@
 Though STA302 respondents' self-perceptions around writing and coding were varied, there was strong consensus (79 per cent) that the use of generative AI tools was appropriate within an academic setting (@tbl-ai-familiarity-usage-1).
-Most respondents who selected "It depends" (12 per cent) generally found in comments that artificial intelligence tools to be appropriate, though with certain guidelines and rules governing their use.
+Most respondents who selected "It depends" (12 per cent) generally found in comments that artificial intelligence tools were appropriate, though with certain guidelines and rules governing their use.
 However, in STA304, almost all students (87 per cent) felt that it was ethical and appropriate to use generative AI for school and only 4 per cent felt that it was inappropriate.
@@
 
-The goal of our modelling is to better understand how a respondents' result on their final paper associates with their self-reported LLM usage.
+The goal of our modelling is to better understand how a respondents' result on their final paper associates with their self-reported LLM usage in our dataset.
 Students received their result on the final paper as a percentage between 0 per cent and 100 per cent.
@@
 Beta regression is commonly used when data are between 0 and 1 and is focused on estimating the two shape parameters that govern the beta distribution.
-When there are some values that are 1, a variant -- one-inflated beta regression -- can be used, which is a mixture of focusing on whether a value was 1 or not, and then the usual Beta distribution aspect.
+When there are some values that are 1, a variant -- one-inflated beta regression -- can be used, which is a mixture of focusing on whether a value was 1 or not, and then the usual beta distribution aspect.
 
@@
 We use "None or minimal" as the reference level and so the results are in relation to that level of self-reported LLM usage.
-Self-reported GPA, `what_is_your_gpa`, is a number between 1.9 and 4 with one decimal place.
-
-If $y_i$ is the percentage that student $i$ received on the final paper (adjusted to a proportion), then our model is:
+Self-reported GPA, `what_is_your_gpa`, is a number between 1.5 and 4.0, with one decimal place.
+
+If $y_i$ is the proportion that student $i$ received on the final paper, then our model is:
 $$y_i \sim \mbox{One-Inflated Beta}\left(\mu_i, \phi_i, \pi_{\mbox{zoi}_i}, \pi_{\mbox{coi}_i}\right)$$
@@
 - $\phi_i$ (precision component) is the precision of the beta distribution, which, again, accounts for marks between zero and one,
-- $\pi_{\mbox{zoi}_i}$ (zero-one inflation component) is a logistic regression model that looks at which of two groups a student is in: a) the group of students getting full or zero marks, or b) the group of students getting a mark between zero and one, and
+- $\pi_{\mbox{zoi}_i}$ (zero-one inflation component) is a logistic regression model that looks at which of two groups a student is in: 1) the group of students getting full or zero marks, or 2) the group of students getting a mark between zero and one, and
 - $\pi_{\mbox{coi}_i}$ (continuous outcome inflation component) is used to distinguish between students that got full or zero marks (but in our case as there were no students that got zero we force this to be one).
@@
 This same pattern is then repeated for STA304 and both those models have 131 observations.
-And the final model in @tbl-model-results additionally considers whether the respondent is a native English speaker.
+The final model in @tbl-model-results additionally considers whether the respondent is a native English speaker.
 
@@
 For instance, for the STA302 model it is -2.10.
-Following @andrewagain, after undoing the logit transformation this is approximately 11%, which corresponds to the number of students in STA302 who got full marks on the final paper.
+After undoing the logit transformation this is approximately 11 per cent, which corresponds to the number of students in STA302 who got full marks on the final paper [@andrewagain].
 The same holds for the STA304 model.
@@
 
-Holding all other variables constant, we can vary LLM usage from none or minimal to somewhat and then extensive (@fig-model-results-llms).
+To understand the implication of the coefficient estimates we can hold all other variables constant and then vary LLM usage from "None or minimal" to "Somewhat" and then "Extensive" (@fig-model-results-llms).
 There is an expected increase in the mark on the final paper as LLM usage increases in STA302 (@fig-model-results-llms-1), but it is less clear in STA304 (@fig-model-results-llms-2).
-It may have been that the overwhelming usage of LLMs at "extensive" levels in STA304, which occurred six months after STA302, was the reason for this seemingly different association in our data.
+It may have been that the overwhelming usage of LLMs at "Extensive" levels in STA304, which occurred six months after STA302, was the reason for this seemingly different association in our data.
 
@@
 #| fig-height: 5
-#| fig-cap: Comparing the effect of a hypothetical student with different LLM usage levels, all other variables constant
+#| fig-cap: Comparing the effect of a hypothetical student with different LLM usage levels, all other variables constant, between STA302 and STA304
 #| fig-subcap: ["Based on responses from STA302 (Winter 2024)", "Based on responses from STA304 (Fall 2024)"]
@@
 
-The different effect of LLM usage, by whether the students was a native English speaker, holding all other variables constant, can only be examined for STA304 (September-December 2024) as the question was not asked in the earlier STA302 (@fig-model-results-english).
-For the hypothetical student who does not speak English natively, where we changed nothing other than their LLM usage, going from none or minimal (@fig-model-results-english-1) to extensive (@fig-model-results-english-2) does appear to have been associated with considerably different marks based on our dataset.
+The different effect of LLM usage, by whether the student was a native English speaker, holding all other variables constant, can only be examined for STA304 (September-December 2024) as the question was not asked in the earlier STA302 (@fig-model-results-english).
+For the hypothetical student who does not speak English natively, where we changed nothing other than their LLM usage, going from "None or minimal" (@fig-model-results-english-1) to "Extensive" (@fig-model-results-english-2) appears to have been associated with higher marks in our dataset.
 The distribution is smaller and the average is higher.
@@
 #| warning: false
-#| fig-cap: Comparing the effect of a hypothetical student changing whether they were a native English speaker, all other variables constant
+#| fig-cap: Comparing the effect of a hypothetical student changing whether they were a native English speaker, all other variables constant, for STA304 only
 #| fig-subcap: ["Holding GPA at 3.0 and LLM usage is none or minimal", "Holding GPA at 3.0 and LLM usage is extensive"]
@@
 We found that many students did use LLMs to help them complete a data science project closely related to what professional data scientists do.
-A common use case to ask the LLM to write code to make a graph.
-Another was to ask for some example paragraphs.
+For instance, looking at the logs, a common use case was to ask the LLM to write code to make a graph, and another was to ask for some example paragraphs of text.
 
@@
 The training data of these LLMs is dominated by software engineering code, and so when an LLM is asked to write code for a data science project it tends to produce code with some of those features.
-This is neither good nor bad in and of itself, however it can clash with the style of the code that surrounds it.
-
-LLMs are also being used as an alternative search engine.
+This is *prima facie* neither good nor bad in and of itself, however it can clash with the style of the code that surrounds it which could lead to maintainability difficulties.
+
+Again, looking at the logs, LLMs are also being used as an alternative search engine.
 Students often asked, for instance, how to modify a graph to add a vertical line, or similar.
 Here LLMs tended to be especially useful and able to answer questions.
-In the past students may have used StackOverflow or similar to find their answers, but the nature of the chat response likely provided a faster answer.
+In the past students may have used Stack Overflow or similar to find their answers, but the nature of the chat response likely provided a faster answer.
 
@@
 One can ask them a technical data science question and then in the same session ask for a brownie recipe!
-However this general-purpose nature means that they are not necessarily optimized toward one particular task [@thoppilan2022lamda].
-When used for data science, this means that sometimes the functions or libraries recommended may not exist, code may not run, or out-dated approaches may be used.
+However, this general-purpose nature means that they are not necessarily optimized toward one particular task [@thoppilan2022lamda].
+When used for data science, this means that sometimes the functions or libraries recommended may not exist, code may not run, or outdated approaches may be used.
 Frontier models make things up and are unnervingly confident, even when wrong [@openai2023gpt4, p.
@@
 Similarly, consider an LLM in an IDE context where a data scientist wrote out the model, in statistical notation, that they were interested in estimating, and the LLM was able write the model in Stan, run it, save the estimates, and add a summary table to the paper, all automatically and in the background.
-Similar work has occured in other contexts and found to bring substantial benefits [@chen2022evaluating].
+Similar work has occurred in other contexts and found to bring substantial benefits [@chen2022evaluating].
 
@@
 There may be selection bias present in terms of who used LLMs, who reported their LLM usage truthfully, and even who remained in the class.
-A different design, specifically a randomized controlled trial (RCT) would deal with many of these issues, although likely at some cost.
-
-Regression reports average estimates over the full dataset.
+A different design, specifically a randomized controlled trial (RCT) would deal with many of these issues, although would likely require financial support.
+
+Regression provides average estimates over the full dataset.
 However, many earlier studies found distributional effects, with low-performing individuals benefiting more than high-performing individuals.
@@
 Along these lines, we have considered LLM usage for code and writing as equivalent, but they should actually have different impacts depending on student backgrounds.
-We were only able to distinguishing between native and second-language English speakers at a high level and a more detailed focus on differential LLM usage would add a great deal of nuance to the analysis.
-
-Finally, we only considered one outcome measure, namely grade on the final paper.
+We were only able to distinguish between native and second-language English speakers at a high level and a more detailed focus on differential LLM usage would add a great deal of nuance to the analysis.
+
+Finally, we only considered one outcome measure, namely a student's mark on their final paper.
 Each paper required a considerable amount of time for the student to produce.
@@
 
-We use `modelsummary` to compute the full regression output.
-
 ## Full model output  {#sec-model-output}
 
-```{r tbl-model-results-app}
+```{r tbl-model-results-full}
 #| message: false
@@
 #| tbl-cap: Coefficient estimates and mean absolute deviation (MAD)
-
-fit1 <-
-  readRDS(file = here("models/fit_one_inflated.rds"))
-fit2 <-
-  readRDS(file = here("models/fit_one_inflated_gpa.rds"))
 
@@
   list(
-    "Base model" = fit1,
-    "Self-reported GPA Included" = fit2
+    "STA302" = fit1,
+    "STA302: GPA" = fit2,
+    "STA304" = fit3,
+    "STA304: GPA" = fit4,
+    "STA304: GPA & ESL" = fit5
   ),
   statistic = "mad",
-  fmt = 2
-  )
+  fmt = 2,
+  output = "kableExtra",
+  kable_options = list(booktabs = TRUE, linesep = ""),
+  # coef_map = c(
+  #   "b_Intercept" = "Intercept",
+  #   "b_phi_Intercept" = "Intercept (Phi)",
+  #   "b_zoi_Intercept" = "Intercept (ZOI)",
+  #   "b_llm_usageSomewhat" = "LLM usage: Somewhat",
+  #   "b_llm_usageExtensive" = "LLM usage: Extensive",
+  #   "b_phi_llm_usageSomewhat" = "LLM usage: Somewhat (Phi)",
+  #   "b_phi_llm_usageExtensive" = "LLM usage: Extensive (Phi)",
+  #   "b_what_is_your_gpa" = "GPA",
+  #   "b_phi_what_is_your_gpa" = "GPA (Phi)",
+  #   "b_native_englishYes" = "English native: Yes",
+  #   "b_phi_native_englishYes" = "English native: Yes (Phi)"
+  # )
+) |>
+  kableExtra::kable_styling(font_size = 7)
 ```
@@
 
-## Full Coefficient Estimates
-
-```{r fig-model-results-app}
-#| message: false
-#| echo: false
-#| warning: false
-#| fig-cap: Selected coefficient estimates and 90 per cent credibility intervals
+## Full coefficient estimates
+
+```{r fig-model-results-full}
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-height: 7
+#| fig-cap: Coefficient estimates and 90 per cent credibility intervals
 
@@
   list(
-    "Base model" = fit1,
-    "Including self-reported GPA" = fit2
+    "STA302" = fit1,
+    "STA302: GPA" = fit2,
+    "STA304" = fit3,
+    "STA304: GPA" = fit4,
+    "STA304: GPA & ESL" = fit5
   ),
   conf_level = 0.9,
-  drop = c("phi", "b_zoi_Intercept")
+  # coef_map = c(
+  #   "b_llm_usageSomewhat" = "LLM usage: Somewhat",
+  #   "b_llm_usageExtensive" = "LLM usage: Extensive",
+  #   "b_phi_llm_usageSomewhat" = "LLM usage: Somewhat (phi)",
+  #   "b_phi_llm_usageExtensive" = "LLM usage: Extensive (phi)",
+  #   "b_what_is_your_gpa" = "GPA",
+  #   "b_phi_what_is_your_gpa" = "GPA (phi)",
+  #   "b_native_englishYes" = "English native: Yes",
+  #   "b_phi_native_englishYes" = "English native: Yes (phi)"
+  # )
 ) +
@@
   scale_color_brewer(palette = "Set1") +
-  theme(legend.position = "bottom")
+  #theme(legend.position = "bottom") +
+  scale_y_discrete(limits=rev)
 ```
@@
 
+```{r fig-basemodel304diagnostics}
+#| fig-height: 7
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: Base model diagnostics (STA304)
+
+plot(fit3)
+```
+
+```{r fig-gpamodel304diagnostics}
+#| fig-height: 6
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: Model including self-reported GPA diagnostics (STA304)
+
+plot(fit4)
+```
+
+```{r fig-gpamodel304ESLdiagnostics}
+#| fig-height: 6
+#| message: false
+#| echo: false
+#| warning: false
+#| fig-cap: Model including self-reported GPA and ESL diagnostics (STA304)
+
+plot(fit5)
+```
+
 \newpage

### 2025-04-12 paper/paper.qmd
Minor changes in text of introduction

@@
 ChatGPT's public release on November 30, 2022, brought LLMs into the mainstream conversation.
-LLMs have quickly been used in both industry and academia.
+LLMs have rapidly gained traction in both industry and academia.
 By now many people, especially educators and students, have some experience with LLMs in both personal and professional contexts.
@@
 Another is writing, which is primarily done when communicating results.
-The potential for LLMs to positively affect students' academic performance in data science follows from their already-demonstrated capability in those competencies in adjacent professional fields.
+The potential for LLMs to positively affect students' academic performance in data science stems from the capabilities these models have already demonstrated in adjacent professional fields.
 
@@
 We focus on the association between student academic performance and their LLM usage.
-Specifically, we examine the relationship between the mark a student got on a final paper and self-reported measures of student LLM usage, as well as student attitudes toward LLMs in general.
+Specifically, we examine the relationship between self-reported measures of student LLM usage and the mark a student got on a final paper, as well as student attitudes toward LLMs in general.
 This is based on students' final papers and a survey, conducted in third-year undergraduate data science courses at the University of REDACTED.
@@
 
-We find some mild association between the mark that a respondent received and their self-reported LLM usage.
-And this is especially the case for respondents that were not native English speakers.
+We find some mild association between self-reported LLM usage and the mark received on the final paper, especially among students who were not native English speakers.
 More concretely, there was a considerable increase in the number of respondents who self-reported extensive LLM usage and also those who considered LLM usage to be appropriate and ethical for university, when we compared the results in the class at the start of 2024 with the class at the end of 2024.
@@
 
-To investigate students' usage and attitudes towards LLMs and how they related to their academic performance, a dataset containing their usage and attitudes, coursework, and academic performance was constructed.
+To investigate the relationship between students' LLM usage and attitudes and their academic performance, a dataset containing self-reported LLM usage, attitudes towards LLMs, final paper grades, and overall academic performance was constructed.
 This was based on three components:
 
-1. an optional survey;
-2. self-reported LLM usage; and
+1. an optional survey administered at the end of the course;
+2. self-reported LLM usage on the final paper; and
 3. student marks on their final paper.
@@
 Finally, one respondent failed an attention check question halfway through the survey "Please select the option blue".
+This left 138 usable responses.
 
@@
 Students from a range of years took both courses, but most respondents were in their 3rd or 4th year of study (@tbl-year-dist).
-The prerequisite two-course sequence is typically completed by students in their second year, would make it difficult to take either course earlier than their 3rd year.
+The prerequisite two-course sequence is typically completed by students in their second year, which would make it difficult to take either course earlier than their 3rd year.
 

### 2025-04-12 paper/paper.qmd
Added sample sizes to Table 1 (a) and (b)

@@
 #| tbl-cap: \"What year are you?\"
-#| tbl-subcap: ["STA302 responses (Winter 2024)","STA304 responses (Fall 2024)"]
+#| tbl-subcap: ["STA302 responses (Winter 2024); n=92", "STA304 responses (Fall 2024); n=131"]
 

### 2025-04-12 paper/paper.qmd
Very minor typo in the last sentence

@@
 
-Despite these shortcomings, our work clearly identifies a need for further research examining how LLMs can be used to develop a more trustworthy data science workflow, both in and outside of the classroom.
+Despite these shortcomings, our work clearly identifies a need for further research examining how LLMs can be used to develop a more trustworthy data science workflow, both in and out of the classroom.
 

### 2025-04-12 paper/paper.qmd
Reformatted table 6 to put (MAD) on same line as estimate to improve readability, and modified the caption

@@
 #| warning: false
-#| tbl-cap: Coefficient estimates and mean absolute deviation (MAD)
+#| tbl-cap: Coefficient estimates and mean absolute deviation (MAD) for five fitted models
 
@@
   ),
-  statistic = "mad",
+  statistic = NULL,
+  estimate = "{estimate} ({mad})",
   fmt = 2,

### 2025-04-12 paper/paper.qmd
Similar change to Table 6 in the results

@@
 #| warning: false
-#| tbl-cap: Selected coefficient estimates and mean absolute deviation (MAD)
+#| tbl-cap: Selected coefficient estimates and mean absolute deviation (MAD) for five fitted models
 
@@
   ),
-  statistic = "mad",
+  #statistic = "mad",
+  statistic = NULL,
+  estimate = "{estimate} ({mad})",
   fmt = 2,

### 2025-04-14 paper/paper.qmd
minor edits

@@
 
-To investigate the relationship between students' LLM usage and attitudes and their academic performance, a dataset containing self-reported LLM usage, attitudes towards LLMs, final paper grades, and overall academic performance was constructed.
+To investigate the relationship between students' LLM usage and attitudes and their academic performance, a dataset containing self-reported LLM usage, personal attitudes towards LLMs, final paper grades, and overall academic performance was constructed.
 This was based on three components:
@@
 Students from a range of years took both courses, but most respondents were in their 3rd or 4th year of study (@tbl-year-dist).
-The prerequisite two-course sequence is typically completed by students in their second year, which would make it difficult to take either course earlier than their 3rd year.
+A prerequisite two-course sequence in mathematical statistics is typically completed by students in their second year, which would make it difficult to take either course earlier than their 3rd year.
 
